### 2013-10-21 doc/tables.tex
Added latex file with compiled tables and diagram

\documentclass{article}

\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage[a4paper, top=2cm, bottom=3cm, left=4cm, right=4cm]{geometry}

\usepackage{tikz}
\usetikzlibrary{shapes, shapes.multipart, shapes.geometric}
\usetikzlibrary{positioning}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\begin{document}

\section{Method}

\begin{figure}[ht]
\begin{tikzpicture}[
    state/.style   = {draw, ellipse, aspect=2},
    action/.style  = {draw, rectangle, align=center},
    decision/.style= {draw, diamond, aspect=2, align=center},
    print/.style   = {draw, trapezium, trapezium left angle=70, trapezium right angle=-70},
    edge/.style    = {draw, ->, shorten >=2pt, shorten <=2pt},
  ]
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C:=\mathcal C(N, M)$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
  \node[action, below=of satc] (modelc)
    {$A:=\text{Model}(C \cup \{\neg P\})$ \\
     $\text{Techniques}:=\text{RefinementTechniques}$};
     \node[decision, below=of modelc] (methods) {$\text{Techniques} \neq ()$};
  \node[action, below=of methods] (ctheta)
    {$\text{Refinement} := \text{Techniques}[0]$ \\
     $\text{Techniques}:= \text{Techniques} \setminus \{\text{Refinement}\}$ \\
     $C_{\theta}:=C_{\text{Refinement}}(N, M, A)$};
  \node[decision, below=of ctheta] (satctheta) {$\text{SAT}(C_\theta)$};
  \node[action, below=of satctheta] (modelctheta)
    {$A_\theta:=\text{Model}(C_\theta)$\\
     $\delta:=\Delta(N, M, A_\theta)$\\
     $C:=C \cup \{\delta\}$};
  \node[print, right=of satc] (yes) {YES};
  \node[state, right=of yes] (end1) {END};
  \node[print, right=of methods] (dontknow) {Don't know};
  \node[state, right=of dontknow] (end2) {END};

  \draw (begin) edge[edge] (c);
  \draw (c) edge[edge] coordinate[pos=.5] (edgein) (satc);
  \draw (satc) edge[edge] node[above]{NO} (yes);
  \draw (yes) edge[edge] (end1);
  \draw (satc) edge[edge] node[right]{YES} (modelc);
  \draw (modelc) edge[edge] (methods);
  \draw (methods) edge[edge] node[right]{YES}  (ctheta);
  \draw (methods) edge[edge] node[above]{NO} (dontknow);
  \draw (ctheta) edge[edge] (satctheta);
  \draw (dontknow) edge[edge] (end2);
  \draw (satctheta) edge[edge] node[right]{YES} (modelctheta);
  \draw[edge] (satctheta.west) node[above,xshift=-1cm]{NO}
  -| ([xshift=-2.5cm] satctheta.west) |- (methods.west);
  \draw[edge] (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
  -| ([xshift=-3cm] modelctheta.west) |- (edgein);
\end{tikzpicture}
\caption{The method}
\label{fig-method}
\end{figure}

\newpage

\section{Benchmarks}

\subsection{Given by Daniel Kroening}

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & positive & negative & don't know & timeout \\
    \hline
    M1: State equation with real numbers & 23 &  0 & 54 &  0 \\
    M2: State equation with integers     & 23 &  0 & 54 &  0 \\
    \hline
    M3: M2 with TrapConditions               & 23 &  0 & 54 &  0 \\
    M4: M3 with SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
    M5: M4 with EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
    \hline
    M6: M1 with State space exploration, timeout of  1\,min & 26 & 34 & 0 & 17 \\
    M7: M1 with State space exploration, timeout of 10\,min & 26 & 39 & 0 & 12 \\
    M8: M1 with State space exploration, timeout of  8\,h   & 26 & 45 & 0 &  6 \\
    \hline
    M9: Bfc tool  & 26 & 51 & 0 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Compiled results of the benchmarks given by Daniel Kroening}
\label{table-results-compiled-dk}
\end{table}

Positive files for method 1:
\begin{lstlisting}
cprover-PN/abp
cprover-PN/basicME
cprover-PN/bingham_h150
cprover-PN/bingham_h25
cprover-PN/bingham_h250
cprover-PN/bingham_h50
cprover-PN/bingham_h500
cprover-PN/clientserver
cprover-PN/csm
cprover-PN/gsm
cprover-PN/handover
cprover-PN/keycard
cprover-PN/mesh2x2
cprover-PN/multipoll
cprover-PN/newrtp
cprover-PN/phones
cprover-PN/reaction
cprover-PN/reaction3
cprover-PN/reaction4
cprover-PN/read-write
cprover-PN/restriction
cprover_software_analysis/conditionals_vs_satabs.2/main
cprover_software_analysis/rand_cas_vs_satabs.2/main
\end{lstlisting}

Additional positive files for method 6:
\begin{lstlisting}
cprover-PN/lamport
cprover-PN/newdekker
cprover-PN/peterson
\end{lstlisting}

\newpage

\subsection{Found in mist repository}

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | p{6cm} | r | r | r | r | }
    \hline
    Method & positive & negative & don't know & timeout \\
    \hline
    M1: State equation with real numbers & 16 &  0 & 11 &  0 \\
    M2: State equation with integers     & 16 &  0 & 11 &  0 \\
    \hline
    M3: M2 with TrapConditions               & 20 &  0 &  7 &  0 \\
    M4: M3 with SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
    M5: M4 with EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
    \hline
    M10: M4 with State space exploration, timeout of 4\,h & 21 &  2 & 0 &  4 \\
    \hline
    M11: Mist tool & 23 &  4 & 0 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Compiled results of the benchmarks found in mist repository}
\label{table-results-compiled-mist}
\end{table}

Positive files for method 1:
\begin{lstlisting}
PN/MultiME
PN/basicME
PN/bingham_h150
PN/bingham_h25
PN/bingham_h250
PN/bingham_h250_attic
PN/bingham_h50
PN/csm
PN/fms
PN/fms_attic
PN/mesh2x2
PN/mesh3x2
PN/multipool
boundedPN/kanban
boundedPN/newrtp
boundedPN/read-write
\end{lstlisting}

Additional positive files for method 3:
\begin{lstlisting}
PN/pingpong
boundedPN/lamport
boundedPN/newdekker
boundedPN/peterson
\end{lstlisting}

Additional positive files for method 4:
\begin{lstlisting}
PN/manufacturing
\end{lstlisting}

Additional positive files for method 11:
\begin{lstlisting}
PN/extendedread-write
PN/extendedread-write-smallconsts
\end{lstlisting}

\end{document}

### 2013-10-22 doc/experiments-2013-10-21.tex
Experimental report.

\documentclass{article}

\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage[a4paper, top=2cm, bottom=3cm, left=4cm, right=4cm]{geometry}

\usepackage{tikz}
\usetikzlibrary{shapes, shapes.multipart, shapes.geometric}
\usetikzlibrary{positioning}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
    \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\newcommand{\bfc}{{\sc Bfc}}
\newcommand{\mist}{{\sc Mist}}
\newcommand{\pnerf}{{\sc Pnerf}}
\newcommand{\zthree}{{\sc Z3}}
\newcommand{\ttt}[1]{\texttt{#1}}

\begin{document}

\title{Experimental Results \\ \mbox{ } \\ of a Safety Checker for Petri nets}
\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
% \institute{Technische Universit\"at M\"unchen}
\date{Mon Oct 21 21:00:31 CEST 2013}

\maketitle

We report on the experimental results of a constraint based
approach for checking safety of Petri Nets.
We implemented the approach in the tool \pnerf. \pnerf\ abstracts a
given Petri Net as a set of state constraints over integers or
reals. When state constraints are not sufficient for checking safety,
\pnerf\ refines the abstraction by applying three techniques (i.e. TrapConditions,
SubnetTrapConditions, EmptyTrapConditions). Each technique constructs additional
state constraints from the spurious abstract
counterexample (i.e. the valuation corresponding state constraints) when possible.
We applied \zthree\ as the backend.
As an experiment, we substituted the refinement techniques with
concrete concrete state exploration. We also report on the
corresponding results.

\section{Benchmarks of \bfc}

We experimented on the benchmarks \ttt{cprover-PN} and
\ttt{cprover\_software\_analysis}
available at
\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
We applied 8 tool configurations to the set of benchmarks.
Table~\ref{bfc-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | r | p{7cm} | r | r | r | r | } %
    \hline
    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
    \hline
    C1 & State constraint over real numbers & 23 &  0 & 54 &  0 \\
    C2 & State constraint over integers        & 23 &  0 & 54 &  0 \\
    \hline
    C3 & C2 + TrapConditions                   & 23 &  0 & 54 &  0 \\
    C4 & C3 + SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
    C5 & C4 + EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
    \hline
    C6 & C1 + state space exploration \& TO 1\,min & 26 & 34 & 0 & 17 \\
    C7 & C1 + state space exploration \& TO 10\,min & 26 & 39 & 0 & 12 \\
    C8 & C1 + state space exploration \& TO 8\,h   & 26 & 45 & 0 &  6 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments}
\end{table}

Observations:
\begin{itemize}
\item State constraints over integers or reals correctly determine that the 23 benchmarks in Appendix~\ref{bfc-experiments-c1} are positive.
\item Composing state space constraints with concrete state space
  exploration gives the following 3 additional positives.
\begin{verbatim}
cprover-PN/lamport
cprover-PN/newdekker
cprover-PN/peterson
\end{verbatim}
\end{itemize}

\newpage

\section{Benchmarks of \mist}

We experimented on the benchmarks available at \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
and \\
\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
We applied 6 tool configurations to the set of benchmarks.
Table~\ref{mist-experiments} shows the corresponding experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | r | p{6cm} | r | r | r | r | }
    \hline
    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
    \hline
    C1 & State constraint with real numbers & 16 &  0 & 11 &  0 \\
    C2 & State constraint with integers     & 16 &  0 & 11 &  0 \\
    \hline
    C3 & C2 + TrapConditions               & 20 &  0 &  7 &  0 \\
    C4 & C3 + SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
    C5 & C4 + EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
    \hline
    C9 & C4 + state space exploration \& TO 4\,h & 21 &  2 & 0 &  4 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \mist}
\label{mist-experiments}
\end{table}

Observations:
\begin{itemize}
\item State constraints over integers or reals correctly determine that the 16 benchmarks in Appendix~\ref{mist-experiments-c1} are positive.
\item Composing state constraints with TrapConditions gives the following 4 additional positives.
\begin{verbatim}
PN/pingpong
boundedPN/lamport
boundedPN/newdekker
boundedPN/peterson
\end{verbatim}
\item Composing state constraints with TrapConditions and SubnetTrapConditions gives the additional positive \ttt{PN/manufacturing}.
  \item The remaining positive benchmarks are the following.
\begin{verbatim}
PN/extendedread-write
PN/extendedread-write-smallconsts
\end{verbatim}
\end{itemize}

\newpage

\iffalse
\section{Method}

\begin{figure}[ht]
\begin{tikzpicture}[
    state/.style   = {draw, ellipse, aspect=2},
    action/.style  = {draw, rectangle, align=center},
    decision/.style= {draw, diamond, aspect=2, align=center},
    print/.style   = {draw, trapezium, trapezium left angle=70, trapezium right angle=-70},
    edge/.style    = {draw, ->, shorten >=2pt, shorten <=2pt},
  ]
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C:=\mathcal C(N, M)$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
  \node[action, below=of satc] (modelc)
    {$A:=\text{Model}(C \cup \{\neg P\})$ \\
     $\text{Techniques}:=\text{RefinementTechniques}$};
     \node[decision, below=of modelc] (methods) {$\text{Techniques} \neq ()$};
  \node[action, below=of methods] (ctheta)
    {$\text{Refinement} := \text{Techniques}[0]$ \\
     $\text{Techniques}:= \text{Techniques} \setminus \{\text{Refinement}\}$ \\
     $C_{\theta}:=C_{\text{Refinement}}(N, M, A)$};
  \node[decision, below=of ctheta] (satctheta) {$\text{SAT}(C_\theta)$};
  \node[action, below=of satctheta] (modelctheta)
    {$A_\theta:=\text{Model}(C_\theta)$\\
     $\delta:=\Delta(N, M, A_\theta)$\\
     $C:=C \cup \{\delta\}$};
  \node[print, right=of satc] (yes) {YES};
  \node[state, right=of yes] (end1) {END};
  \node[print, right=of methods] (dontknow) {Don't know};
  \node[state, right=of dontknow] (end2) {END};

  \draw (begin) edge[edge] (c);
  \draw (c) edge[edge] coordinate[pos=.5] (edgein) (satc);
  \draw (satc) edge[edge] node[above]{NO} (yes);
  \draw (yes) edge[edge] (end1);
  \draw (satc) edge[edge] node[right]{YES} (modelc);
  \draw (modelc) edge[edge] (methods);
  \draw (methods) edge[edge] node[right]{YES}  (ctheta);
  \draw (methods) edge[edge] node[above]{NO} (dontknow);
  \draw (ctheta) edge[edge] (satctheta);
  \draw (dontknow) edge[edge] (end2);
  \draw (satctheta) edge[edge] node[right]{YES} (modelctheta);
  \draw[edge] (satctheta.west) node[above,xshift=-1cm]{NO}
  -| ([xshift=-2.5cm] satctheta.west) |- (methods.west);
  \draw[edge] (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
  -| ([xshift=-3cm] modelctheta.west) |- (edgein);
\end{tikzpicture}
\caption{The method}
\label{fig-method}
\end{figure}
\fi

\section*{Appendix}

\subsection{Benchmarks for Table~\ref{bfc-experiments}, configuration C1}
\label{bfc-experiments-c1}

\begin{verbatim}
cprover-PN/abp
cprover-PN/basicME
cprover-PN/bingham_h150
cprover-PN/bingham_h25
cprover-PN/bingham_h250
cprover-PN/bingham_h50
cprover-PN/bingham_h500
cprover-PN/clientserver
cprover-PN/csm
cprover-PN/gsm
cprover-PN/handover
cprover-PN/keycard
cprover-PN/mesh2x2
cprover-PN/multipoll
cprover-PN/newrtp
cprover-PN/phones
cprover-PN/reaction
cprover-PN/reaction3
cprover-PN/reaction4
cprover-PN/read-write
cprover-PN/restriction
cprover_software_analysis/conditionals_vs_satabs.2/main
cprover_software_analysis/rand_cas_vs_satabs.2/main
\end{verbatim}

\subsection{Benchmarks for Table~\ref{mist-experiments}, configuration C1}
\label{mist-experiments-c1}

\begin{verbatim}
PN/MultiME
PN/basicME
PN/bingham_h150
PN/bingham_h25
PN/bingham_h250
PN/bingham_h250_attic
PN/bingham_h50
PN/csm
PN/fms
PN/fms_attic
PN/mesh2x2
PN/mesh3x2
PN/multipool
boundedPN/kanban
boundedPN/newrtp
boundedPN/read-write
\end{verbatim}

\end{document}

### 2013-11-11 doc/experiments-2013-10-21.tex
Updated numbers in experiments

@@
     \hline
-    C1 & State constraint with real numbers & 16 &  0 & 11 &  0 \\
-    C2 & State constraint with integers     & 16 &  0 & 11 &  0 \\
+    C1 & State constraint with real numbers & 14 &  0 & 13 &  0 \\
+    C2 & State constraint with integers     & 14 &  0 & 13 &  0 \\
     \hline
@@
 \begin{itemize}
-\item State constraints over integers or reals correctly determine that the 16 benchmarks in Appendix~\ref{mist-experiments-c1} are positive.
-\item Composing state constraints with TrapConditions gives the following 4 additional positives.
+\item State constraints over integers or reals correctly determine that the 14 benchmarks in Appendix~\ref{mist-experiments-c1} are positive.
+\item Composing state constraints with TrapConditions gives the following 6 additional positives.
 \begin{verbatim}
 PN/pingpong
+PN/basicME
+PN/multiME
 boundedPN/lamport

### 2013-11-18 paper/workspace/main.tex
Figures for paper.

\documentclass{article}

\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}

\usepackage{tikz}
\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
	snakes,automata,backgrounds,petri,calc}
\usetikzlibrary{positioning}

\newcommand{\bfc}{{\sc Bfc}}
\newcommand{\mist}{{\sc Mist}}
\newcommand{\pnerf}{{\sc Pnerf}}
\newcommand{\zthree}{{\sc Z3}}
\newcommand{\ttt}[1]{\texttt{#1}}

\tikzstyle{place}=[circle,thick,draw=black!75,fill=white!20,
  	minimum size=6mm]
\tikzstyle{red place}=[place,draw=red!75,fill=red!20]
\tikzstyle{blue place}=[place,draw=blue!75,fill=blue!20]
\tikzstyle{transition}=[rectangle,thick,draw=black!75,
  	fill=black!20,minimum size=4mm]

\tikzstyle{state}=[draw, ellipse, aspect=2]
\tikzstyle{action}=[draw, rectangle, align=center]
\tikzstyle{decision}=[draw, diamond, aspect=2, align=center]
\tikzstyle{print}=[draw, trapezium, trapezium left angle=70, trapezium right angle=-70]
\tikzstyle{every edge}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]

\begin{document}

\title{Report on a Safety Checker for Petri nets}
\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
% \institute{Technische Universit\"at M\"unchen}
\date{Wed Nov 13 09:11:15 CEST 2013}

\section{Preliminaries}

\begin{verbatim}
* Petri net:
 N = (S, T, E, M0), where
 S : places
 T : transitions
 E : edges
 M0: initial marking

* Property:
 P, linear arithmetic formula
\end{verbatim}
\iffalse
\begin{verbatim}
 Examples:
  x + y < 0
  x + y < 0 \\and x + z > 0
  x + y < 0 \\or  x + z > 0

 ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
\end{verbatim}
\fi

\newpage
\section{Method Safety}

The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
We present the method Safety by example on a simplified mutual exlusion algorithm by Delzanno2001.

\begin{verbatim}
* Code

bool m = true;

// Thread 1:
while (true) {
  if (m) {
    m = false
    // critical section 1
    m = true
  }
  if (m) {
    m = false
    // critical section 2
    m = true
  }
}

// Thread 2:
while (true) {
  if (m) {
    m = false
    // critical section 1
    m = true
  }
  if (m) {
    m = false
    // critical section 2
    m = true
  }
}

* Property: Thread 1 and Thread 2 are never in any critical section at the same time.

* Petri net
\end{verbatim}

\begin{center}
  \begin{tikzpicture}[node distance=2cm]
    \node[blue place, tokens=2, label=above:nc] (nc) {};
    \node[place, tokens=1, label=above:m, below of=nc] (s) {};
    \node[red place, label=above:cs1, left of=s] (cs1) {};
    \node[red place, label=above:cs2, right of=s] (cs2) {};
    \node[transition, label=above:t1] (t1) at ($(s)!0.5!(cs1) + (0,1cm)$) {}
    edge [pre] (nc)
    edge [pre] (s)
    edge [post] (cs1);
    \node[transition, label=above:t2] (t2) at ($(s)!0.5!(cs2) + (0,1cm)$) {}
    edge [pre] (nc)
    edge [pre] (s)
    edge [post] (cs2);
    \node[transition, label=below:t3] (t3) at ($(s)!0.5!(cs1) - (0,1cm)$) {}
    edge [pre] (cs1)
    edge [post] (s)
    edge [post,bend left,min distance=3cm, out=105, in=105] (nc);
    \node[transition, label=below:t4] (t4) at ($(s)!0.5!(cs2) - (0,1cm)$) {}
    edge [pre] (cs2)
    edge [post] (s)
    edge [post,bend right,min distance=3cm, out=-105, in=-105] (nc);
  \end{tikzpicture}
\end{center}

\begin{verbatim}
* Property: P = ( cs1 + cs2 <= 1 )
\end{verbatim}

\newpage
\begin{verbatim}
* Method Safety:

  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
\end{verbatim}

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
    \node[print, right=of satc] (yes) {YES};
    \node[print, below=of satc] (dontknow) {Don't know};
    \node[state, right=of yes] (end1) {END};
    \node[state, below=of dontknow] (end2) {END};

    \draw (begin) edge (c);
    \draw (c) edge (satc);
    \draw (satc) edge node[above]{NO} (yes);
    \draw (yes) edge (end1);
    \draw (satc) edge node[right]{YES} (dontknow);
    \draw (dontknow) edge (end2);
  \end{tikzpicture}
\end{center}

\begin{verbatim}
* Property of state constraints C: if C U {\neg P} is unsat then N |= P

* State constraints C:

  Place equations:

  nc      = 2 - t1 - t2 + t3 + t4
  ^         ^   ^    ^    ^    ^
  |         |   |    |    |    |
  place nc  |   ------    ------
            |      |         |
            |      |       input transitions of nc
            |      |    
            |    output transitions of nc
            |
         initial number of tokens in nc

  m     = 1 - t1 - t2 + t3 + t4
  cs1   = 0 + t1 - t3
  cs2   = 0 + t2 - t4

  Non-negativity conditions:

  nc  >= 0
  m   >= 0
  cs1 >= 0
  cs2 >= 0
  t1  >= 0
  t2  >= 0
  t3  >= 0
  t4  >= 0
\end{verbatim}

\newpage
\begin{verbatim}
* Place equation:

  For a given place s the place equation is

  s = initial number of tokens of place s - output transitions of s + input transitions of s

* Non-negativity conditions:

  place      >= 0
  transition >= 0

* Subprocedure \mathcal{C}:

  Input:
    (S, T, E, I) : Petri net
  Output:
    C : State constraints

  TODO: Unroll pseudocode?
\end{verbatim}

\newpage
\section{Method Invariant}

Given Petri net \verb?N? that does not violate property \verb?P?, method Invariant constructs an invariant \verb?I? with the following properties.
\begin{itemize}
\item For each reachable marking \verb=M=, \verb?I(M)? is valid.
\item For markings that violate the property, \verb?I(M)? is unsat.
\item \verb?I? is inductive:
  \verb?\forall M. I(M) valid ^ M -> M1 implies I(M1) valid?
\end{itemize}

\begin{verbatim}
* Code, property, and Petri net: same as in section Method Safety.
\end{verbatim}

\begin{verbatim}
* Method Invariant

  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
  Subprocedure Model returns assignment A such that A satisfies C'.
  Subprocedure Inv constructs invariant I corresponding to N, P, and A.
\end{verbatim}

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
      $I := \text{Inv}(N, P, A)$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

    \draw (begin) edge (c);
    \draw (c) edge (satc);
    \draw (satc) edge node[above]{NO} (noinv);
    \draw (satc) edge node[right]{YES} (inv);
    \draw (noinv) edge (end1);
    \draw (inv) edge (printinv);
    \draw (printinv) edge (end2);
  \end{tikzpicture}
\end{center}

\newpage
\begin{verbatim}
* Dual state constraints C':

  - nc - m + cs1 <= 0
  - nc - m + cs2 <= 0
    nc + m - cs1 <= 0
    nc + m - cs2 <= 0

  2 \cdot nc + m < 2 \cdot target1

  Non-negativity conditions:

  target1 >= 0
  nc      >= 0
  m       >= 0
  cs1     >= 0
  cs2     >= 0
\end{verbatim}

\begin{align*}

2 \cdot nc + m <& 2 \cdot target1 \\
  target1 \ge& 0 \\
  nc \ge& 0 \\
  m \ge& 0 \\
  cs1 \ge& target1 \\
  cs2 \ge& target1
\end{align*}

\begin{verbatim}
* Model A:

  target1 = 1
  nc      = 0
  m       = 1
  cs1     = 0
  cs2     = 0

* Invariant:

  cs1 + cs2 + m <= 1

* Subprocedure \mathcal{C'}:

  Input:
    (S, T, E, I) : Petri net
    P            : Property
  Output:
    C'           : Dual state constraints

Input:
 N = (S, T, E, I)
   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
 P = ( cs1 + cs2 <= 1 )

Intermediate results:

Create variable p for property P.

For t1 construct inequation `target1 - nc - m + cs1 <= 0` in the following way.
    p - input_places + output_places <= 0
  We write `p` because the place cs1 appears in the property p and in the output places.

For t2 construct inequation `target1 - nc - m + cs2 <= 0` in the following way.
    p - input_places + output_places <= 0
  We write `p` because the place cs2 appears in the property p and in the output places.

For t3 construct inequation `- target1 + nc + m - cs1 <= 0` in the following way.
  - p - input_places + output_places <= 0
  We write `- p` because the place cs1 appears in the property p and in the input places.

For t4 construct inequation `- target1 + nc + m - cs2 <= 0` in the following way.
  - p - input_places + output_places <= 0
  We write `- p` because the place cs2 appears in the property p and in the input places.

The condition for writing `p` or `-p`:
  Write   `p` when output_places \cap vars(p) \neq \empty
  Write `- p` when input_places  \cap vars(p) \neq \empty

Output:
  target1 - nc - m + cs1   <= 0
  target1 - nc - m + cs2   <= 0
  - target1 + nc + m - cs1 <= 0
  - target1 + nc + m - cs2 <= 0

  2 \cdot nc + m < 2 \cdot target1

  Non-negativity conditions:

  target1 >= 0
  nc      >= 0
  m       >= 0
  cs1     >= 0
  cs2     >= 0

* Subprocedure Inv:

* Theorem Dual State Constraints:

  TODO: Unroll ?
\end{verbatim}

\newpage

\iffalse
Constraints $\mathcal C'(N, P)$ with
$Y_1 = \begin{pmatrix}target1 \end{pmatrix}$,
$Y_2 =\begin{pmatrix}nc & m & cs1 & cs2\end{pmatrix}$,
$A =\begin{pmatrix}0 & 0 & 1 & 1 \end{pmatrix}$ and
$b =\begin{pmatrix}2 \end{pmatrix}$:

\begin{align*}
  target1 - nc - m + cs1 \le& 0 \\ 
  target1 - nc - m + cs2 \le& 0 \\ 
  - target1 + nc + m - cs1 \le& 0 \\ 
  - target1 + nc + m - cs2 \le& 0 \\ 
2 \cdot nc + m <& 2 \cdot target1 \\
  target1 \ge& 0 \\
  nc \ge& 0 \\
  m \ge& 0 \\
  cs1 \ge& 0 \\
  cs2 \ge& 0
\end{align*}

Satisfying assignment is $Y_1 = \begin{pmatrix} 1 \end{pmatrix}$ and
$Y_2 = \begin{pmatrix}0 & 1 & 0 & 0 \end{pmatrix}$

Invariant $\text{Inv}(N, P, A) = (cs1 + cs2 + m \le 1)$
\fi

\iffalse

\paragraph{Intermediate results}
Constraints $\mathcal C'(N, P)$ with
$Y_1 = \begin{pmatrix}target1 \end{pmatrix}$,
$Y_2 =\begin{pmatrix}nc & m & cs1 & cs2\end{pmatrix}$,
$A =\begin{pmatrix}0 & 0 & 1 & 1 \end{pmatrix}$ and
$b =\begin{pmatrix}2 \end{pmatrix}$:

\begin{align*}
  target1 - nc - m + cs1 \le& 0 \\ 
  target1 - nc - m + cs2 \le& 0 \\ 
  - target1 + nc + m - cs1 \le& 0 \\ 
  - target1 + nc + m - cs2 \le& 0 \\ 
2 \cdot nc + m <& 2 \cdot target1 \\
  target1 \ge& 0 \\
  nc \ge& 0 \\
  m \ge& 0 \\
  cs1 \ge& 0 \\
  cs2 \ge& 0
\end{align*}

Satisfying assignment is $Y_1 = \begin{pmatrix} 1 \end{pmatrix}$ and
$Y_2 = \begin{pmatrix}0 & 1 & 0 & 0 \end{pmatrix}$

Invariant $\text{Inv}(N, P, A) = (cs1 + cs2 + m \le 1)$

\paragraph{Technique}
\begin{align*}
  \mathcal C'(N, P) =& ( ( ( Y_1 \cdot A + Y_2) \cdot C \le 0) \land \\
& ((Y_1 \cdot A + Y_2) \cdot M_0 < Y_1 \cdot b) \land \\
& (Y_1 \ge 0) \land (Y_2 \ge 0)) \\
  \text{Inv}(N, P, A) =& ( ( Y_1 \cdot A + Y_2) \cdot M \le ( Y_1 \cdot A + Y_2) \cdot M_0)
\end{align*}
where $M_0$ is the initial marking, $C$ is the incidence matrix of the net and
$A\cdot M \ge b$ is equivalent to $\neg P$.

\paragraph{Method Invariant}
\mbox{ } \\
\begin{tikzpicture}
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
  \node[print, right=of satc] (noinv) {No Invariant};
  \node[action, below=of satc] (inv) {$A_\theta:=\text{Model}(C')$\\
                                  	$I := \text{Inv}(N, P, A)$};
  \node[print, below=of inv] (printinv) {Invariant: $I$};
  \node[state, right=of noinv] (end1) {END};
  \node[state, below=of printinv] (end2) {END};

  \draw (begin) edge (c);
  \draw (c) edge (satc);
  \draw (satc) edge node[above]{NO} (noinv);
  \draw (satc) edge node[right]{YES} (inv);
  \draw (noinv) edge (end1);
  \draw (inv) edge (printinv);
  \draw (printinv) edge (end2);
\end{tikzpicture}

\section{Method Safety by Refinement}

\paragraph{Example: Mutual exclusion algorithm in Delzanno2001}

\begin{verbatim}
bool m1 = true;
bool m2 = true;

// Thread 1:
while (true) {
  if (m1 && m2) {
    m1 = false
    // critical section 1
    m1 = true
  }
  if (m1 && m2) {
    m2 = false
    // critical section 2
    m2 = true
  }
}

// Thread 2:
while (true) {
  if (m1 && m2) {
    m1 = false
    // critical section 1
    m1 = true
  }
  if (m1 && m2) {
    m2 = false
    // critical section 2
    m2 = true
  }
}

Property: The number of threads in critical sections should be at most 1
\end{verbatim}

\paragraph{Example: Petri net for mutual exclusion algorithm in Delzanno2001}
\mbox{ } \\
\begin{tikzpicture}[node distance=2cm]
  \node[blue place, tokens=2, label=above:nc] (nc) {};
  \node[place, tokens=1, label=right:m2, below of=nc, xshift=+0.75cm] (s2) {};
  \node[place, tokens=1, label=left:m1, below of=nc, xshift=-0.75cm] (s1) {};
  \node[red place, label=above:cs1, left of=s1] (cs1) {};
  \node[red place, label=above:cs2, right of=s2] (cs2) {};
  \node[transition, label=above:t1] (t1) at ($(s1)!0.5!(cs1) + (0,1cm)$) {}
	edge [pre] (nc)
	edge [pre] (s1)
	edge [pre] (s2)
	edge [post] (cs1)
	edge [post] (s2);
  \node[transition, label=above:t2] (t2) at ($(s2)!0.5!(cs2) + (0,1cm)$) {}
	edge [pre] (nc)
	edge [pre] (s2)
	edge [pre] (s1)
	edge [post] (cs2)
	edge [post] (s1);
  \node[transition, label=below:t3] (t3) at ($(s1)!0.5!(cs1) - (0,1cm)$) {}
	edge [pre] (cs1)
	edge [post] (s1)
	edge [post,bend left,min distance=3cm, out=115, in=115] (nc);
  \node[transition, label=below:t4] (t4) at ($(s2)!0.5!(cs2) - (0,1cm)$) {}
	edge [pre] (cs2)
	edge [post] (s2)
	edge [post,bend right,min distance=3cm, out=-115, in=-115] (nc);
\end{tikzpicture}
\begin{align*}
  P &= ( cs1 + cs2 \le 1 ) \\
  \neg P &= ( cs1 + cs2 \ge 2 )
\end{align*}

\paragraph{Intermediate results}
Constraints $\mathcal C(N)$ with
$M =\begin{pmatrix}nc & m1 & m2 & cs1 & cs2 \end{pmatrix}^T$,
$M_0 =\begin{pmatrix}2 & 1 & 1 & 0 & 0 \end{pmatrix}^T$,
$X =\begin{pmatrix}t1 & t2 & t3 & t4 \end{pmatrix}^T$ and
$C =\begin{pmatrix}
  -1 & -1 & 1 & 1 \\
  -1 &  0 & 1 & 0 \\
   0 & -1 & 0 & 1 \\
  1 & 0 & -1 & 0 \\
  0 & 1 & 0 & -1
\end{pmatrix}$:
\begin{align*}
  nc   =& 2 - t1 - t2 + t3 + t4 \\
  m1  =& 1 - t1 + t3 \\
  m2  =& 1 - t2 + t4 \\
  cs1 =& 0 + t1 - t3 \\
  cs2 =& 0 + t2 - t4 \\
  nc \ge& 0 \\
  m1 \ge& 0 \\
  m2 \ge& 0 \\
  cs1 \ge& 0 \\
  cs2 \ge& 0
\end{align*}

Satisfying assignment is
$M =\begin{pmatrix}0 & 0 & 0 & 1 & 1 \end{pmatrix}^T$ and
$X =\begin{pmatrix}1 & 1 & 0 & 0 \end{pmatrix}^T$.

A set satisfying the trap conditions is $S = \{m1, m2\}$.

The trap constraint is $\delta = (m1 + m2 \ge 1)$.

With the additional constraint, the set of constraints is unsatisfiable.

\paragraph{Technique trap conditions} For a petri net $N$ and an assignment $A$,
find a set $S$ that satisfies
\begin{enumerate}
  \item $S$ is a trap in the net $N$.
  \item $S$ is marked in the initial marking $M_0$.
  \item $S$ is unmarked in the assignment $A$.
\end{enumerate}
For such a set $S$, generate a constraint
$\delta = \left( \sum_{s \in S} s \ge 1 \right)$, ensuring the
trap is marked in any assignment.

\paragraph{Method Safety by Refinement}
\mbox{ } \\
\begin{tikzpicture}
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
                                    	$\delta:=\Delta(A_\theta)$\\
                                    	$C:=C \cup \{\delta\}$};
  \node[print, right=of satc] (yes) {YES};
  \node[print, right=of satctheta] (dontknow) {Don't know};
  \node[state, right=of yes] (end1) {END};
  \node[state, right=of dontknow] (end2) {END};

  \draw (begin) edge (c);
  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
  \draw (satc) edge node[above]{NO} (yes);
  \draw (yes) edge (end1);
  \draw (satc) edge node[right]{YES} (modelc);
  \draw (modelc) edge (satctheta);
  \draw (satctheta) edge node[above]{NO} (dontknow);
  \draw (dontknow) edge (end2);
  \draw (satctheta) edge node[right]{YES} (modelctheta);
  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
\end{tikzpicture}

\section{Method Invariant by Refinement}

\paragraph{Example}
See example for method Safety by Refinement.

\paragraph{Intermediate results}
Constraints $\mathcal C'(N)$ with
$Y_1 = \begin{pmatrix}target1 \end{pmatrix}$,
$Y_2 = \begin{pmatrix}trap2 \end{pmatrix}$,
$Y_3 =\begin{pmatrix}nc & m & cs1 & cs2\end{pmatrix}$,
the trap from the trap refinement as
$D =\begin{pmatrix}0 & 1 & 1 & 0 & 0 \end{pmatrix}$,
$A =\begin{pmatrix}0 & 0 & 0 & 1 & 1 \end{pmatrix}$ and
$b =\begin{pmatrix}1 \end{pmatrix}$:

\begin{align*}
  target1 - trap1 - nc - m1 + cs1 \le& 0 \\ 
  target1 - trap1 - nc - m2 + cs2 \le& 0 \\ 
  - target1 + trap1 + nc + m1 - cs1 \le& 0 \\ 
  - target1 + trap1 + nc + m2 - cs2 \le& 0 \\ 
2 \cdot trap1 + 2 \cdot nc + m1 + m2 <& 2 \cdot target1 + trap1 \\
  target1 \ge& 0 \\
  trap1 \ge& 0 \\
  nc \ge& 0 \\
  m1 \ge& 0 \\
  m2 \ge& 0 \\
  cs1 \ge& 0 \\
  cs2 \ge& 0
\end{align*}

Satisfying assignment is
$Y_1 = \begin{pmatrix} 1 \end{pmatrix}$,
$Y_2 = \begin{pmatrix} 1 \end{pmatrix}$ and
$Y_3 = \begin{pmatrix}0 & 0 & 0 & 0 & 0 \end{pmatrix}$.

Invariant $\text{Inv}(N, D, P, A) =
((cs1 + cs2 + m1 + m2 \le 2) \land (m1 + m2 \ge 1))$.

\paragraph{Technique}
\begin{align*}
  \mathcal C'(N, D, P) =& ( ( ( Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot C \le 0) \land \\
& ((Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot M_0 < Y_1 \cdot b + Y_2 \cdot \mathbf{1}) \land \\
& (Y_1 \ge 0) \land (Y_2 \ge 0) \land (Y_3 \ge 0)) \\
  \text{Inv}(N, D, P, A) =& ( ( Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot M \le ( Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot M_0) \land ( D \cdot M \ge \mathbf{1} )
\end{align*}
where $M_0$ is the initial marking, $C$ is the incidence matrix of the net, $A\cdot M \ge b$ is equivalent to $\neg P$ and $D \cdot M \ge \mathbf{1}$ is equivalent to the trap constraints.

\paragraph{Method Invariant by Refinement}
\mbox{ } \\
\begin{tikzpicture}
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
                                 	$D:=\{\}$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
                                   	$C_{\theta}:=\text{TrapConditions}(N, A)$};
  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
                                        	$\delta:=\Delta(A_\theta)$\\
                                        	$C:=C \cup \{\delta\}$\\
                                        	$D:=D \cup \{\delta\}$};
  \node[print, right=of satctheta] (noinv2) {No Invariant};
  \node[state, below=of noinv2] (end3) {END};

  \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, D, P)$};
  \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
  \node[action, below=of satcprime] (inv) {$A_\theta:=\text{Model}(C')$\\
                                      	$I := \text{Inv}(N, D, P, A)$};
  \node[print, below=of inv] (printinv) {Invariant: $I$};
  \node[print, right=of satcprime] (noinv) {No Invariant};
  \node[state, below=of noinv] (end1) {END};
  \node[state, below=of printinv] (end2) {END};

  \draw (cprime) edge (satcprime);
  \draw (satcprime) edge node[above]{NO} (noinv);
  \draw (satcprime) edge node[right]{YES} (inv);
  \draw (inv) edge (printinv);
  \draw (noinv) edge (end1);
  \draw (printinv) edge (end2);

  \draw (begin) edge (c);
  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
  \draw (satc) edge node[above]{NO} (cprime);
  \draw (satc) edge node[right]{YES} (modelc);
  \draw (modelc) edge (satctheta);
  \draw (satctheta) edge node[above]{NO} (noinv2);
  \draw (noinv2) edge (end3);
  \draw (satctheta) edge node[right]{YES} (modelctheta);
  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
\end{tikzpicture}

\fi

\end{document}

### 2013-11-18 paper/workspace/main.tex
Updated paper

@@
 
-bool m = true;
+bool m = false;
 
@@
 while (true) {
-  if (m) {
+  if (! test_and_set(m)) {
+    // critical section 1
     m = false
-    // critical section 1
-    m = true
   }
-  if (m) {
+  if (! test_and_set(m)) {
+    // critical section 2
     m = false
-    // critical section 2
-    m = true
   }
@@
 * Petri net
+
+Place m is marked in the net exactly if m is false in the algorithm.
+
 \end{verbatim}
@@
   m       >= 0
-  cs1     >= 0
-  cs2     >= 0
+  cs1     >= target1
+  cs2     >= target1
 \end{verbatim}
-
-\begin{align*}
-
-2 \cdot nc + m <& 2 \cdot target1 \\
-  target1 \ge& 0 \\
-  nc \ge& 0 \\
-  m \ge& 0 \\
-  cs1 \ge& target1 \\
-  cs2 \ge& target1
-\end{align*}
 
@@
 
-For t1 construct inequation `target1 - nc - m + cs1 <= 0` in the following way.
-    p - input_places + output_places <= 0
-  We write `p` because the place cs1 appears in the property p and in the output places.
+For t1 construct inequation `- nc - m + cs1 <= 0` in the following way.
+   - input_places + output_places <= 0
 
-For t2 construct inequation `target1 - nc - m + cs2 <= 0` in the following way.
-    p - input_places + output_places <= 0
-  We write `p` because the place cs2 appears in the property p and in the output places.
+For t2 construct inequation `- nc - m + cs2 <= 0` in the following way.
+   - input_places + output_places <= 0
 
-For t3 construct inequation `- target1 + nc + m - cs1 <= 0` in the following way.
-  - p - input_places + output_places <= 0
-  We write `- p` because the place cs1 appears in the property p and in the input places.
+For t3 construct inequation `nc + m - cs1 <= 0` in the following way.
+   - input_places + output_places <= 0
 
-For t4 construct inequation `- target1 + nc + m - cs2 <= 0` in the following way.
-  - p - input_places + output_places <= 0
-  We write `- p` because the place cs2 appears in the property p and in the input places.
-
-The condition for writing `p` or `-p`:
-  Write   `p` when output_places \cap vars(p) \neq \empty
-  Write `- p` when input_places  \cap vars(p) \neq \empty
+For t4 construct inequation `nc + m - cs2 <= 0` in the following way.
+   - input_places + output_places <= 0
 
 Output:
-  target1 - nc - m + cs1   <= 0
-  target1 - nc - m + cs2   <= 0
-  - target1 + nc + m - cs1 <= 0
-  - target1 + nc + m - cs2 <= 0
+  - nc - m + cs1   <= 0
+  - nc - m + cs2   <= 0
+  + nc + m - cs1 <= 0
+  + nc + m - cs2 <= 0
 
@@
   m       >= 0
-  cs1     >= 0
-  cs2     >= 0
+  cs1     >= target1
+  cs2     >= target1
 
@@
 \begin{verbatim}
-bool m1 = true;
-bool m2 = true;
+* Code
+
+bool m1 = false;
+bool m2 = false;
 
@@
 while (true) {
-  if (m1 && m2) {
+  if (! m2 && ! test_and_set(m1)) {
+    // critical section 1
     m1 = false
-    // critical section 1
-    m1 = true
   }
-  if (m1 && m2) {
+  if (! m1 && ! test_and_set(m2)) {
+    // critical section 2
     m2 = false
-    // critical section 2
-    m2 = true
   }
@@
 Property: The number of threads in critical sections should be at most 1
+
+* Petri net
+
+Place m1 is marked in the net exactly if m1 is false in the algorithm.
+Place m2 is marked in the net exactly if m2 is false in the algorithm.
+
 \end{verbatim}

### 2013-11-18 paper/workspace/main.tex
Updated main.

@@
 The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on a simplified mutual exlusion algorithm by Delzanno2001.
+We present the method Safety by example on a simplified mutual exclusion algorithm by Delzanno2001.
 
@@
 
-* Petri net
-
-Place m is marked in the net exactly if m is false in the algorithm.
+* Petri net:
 
@@
 \begin{verbatim}
-* Property of state constraints C: if C U {\neg P} is unsat then N |= P
+* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
 
@@
 
-  s = initial number of tokens of place s - output transitions of s + input transitions of s
+  # of tokens in s = initial number of tokens of place s - # times each output transition of s fires + # times each input transition of s fires
 
@@
 
-  place      >= 0
-  transition >= 0
+  # of tokens in place s           >= 0
+  # of times transition t is fired >= 0
 
@@
 
-Given Petri net \verb?N? that does not violate property \verb?P?, method Invariant constructs an invariant \verb?I? with the following properties.
-\begin{itemize}
-\item For each reachable marking \verb=M=, \verb?I(M)? is valid.
-\item For markings that violate the property, \verb?I(M)? is unsat.
-\item \verb?I? is inductive:
-  \verb?\forall M. I(M) valid ^ M -> M1 implies I(M1) valid?
-\end{itemize}
+The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
+We present the method Invariant by example on a simplified mutual exclusion algorithm by Delzanno2001.
 
@@
 
+\begin{verbatim}
+* Property of dual state constraints C':  If C' is sat then N |= P.
+
+* Property of invariant I:
+  - I is reachable: For each reachable marking M, I(M) is valid
+  - I is safe:      For markings that violate the property, I(M) is unsat
+  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
+\end{verbatim}
+
 \newpage
@@
   m       = 1
-  cs1     = 0
-  cs2     = 0
+  cs1     = 1
+  cs2     = 1
 
@@
 
+\end{verbatim}
+
+\newpage
+\begin{verbatim}
+* Derivation of invariant I for Petri net N that satisfies property P:
+
+  M                                    : marking          ~ places
+  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
+  X                                    : firing vector    ~ transitions
+
+  The following constraints C1 are unsat.
+  M = M0 + C*X                         : place equations
+  M >= 0                               : non-negativity conditions for places
+  X >= 0                               : non-negativity conditions for transitions
+  AM >= b                              : property P negated
+
+  Substitute M to obtain constraints C2.
+  A(M0 + CX) >= b                      : property P negated
+  M0 + CX >= 0                         : non-negativity conditions for places
+  X >= 0                               : non-negativity conditions for transitions
+
+  Rewrite each system to obtain constraints C3.
+  (-A*C)      (A*M0-b)                 : property P negated
+  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
+  (  -I)      (   0  )                 : non-negativity conditions for transitions
+
+  Apply Farkas' Lemma to obtain constraints C4.
+  yT*(-A*C)                            : 
+     (  -C) = 0                        : 
+     (  -I)                            :
+
+  yT*(A*M0-b)                          : 
+     (  M0  ) < 0                      : 
+     (   0  )                          :
+
+  y >= 0                               :
+
+  The constraints C4 are sat iff the following constraints C5 are sat.
+  y1 * A * C + y2 * C + y3 = 0         : 
+  y1 * (A*M - b) + y2 * M0 < 0         : 
+  y1 >= 0                              : 
+  y2 >= 0                              : 
+  y3 >= 0                              :
+
+  The constraints C5 are sat iff the following constraints C6 are sat.
+  (y1 * A + y2) * C  <= 0              : 
+  (y1 * A + y2) * M0 < y1 * b          : 
+  y1 >= 0                              : 
+  y2 >= 0                              :
+
+  The constraints C6 are sat iff the following constraints C' are sat.
+  \lambda * C  <= 0                    : inductivity constraint
+  \lambda * M0 < y1 * b                : safety constraint
+  \lambda >= y1 * A                    : property constraint
+  y1 >= 0                              : non-negativity constraint
+
+  For \lambda satisfying C' the invariant is the following.
+  I(M) = (\lambda * M <= \lambda * M0) : invariant
+\end{verbatim}
+
+\newpage
+\begin{verbatim}
 * Subprocedure \mathcal{C'}:
@@
 
-Intermediate results:
-
-Create variable p for property P.
-
-For t1 construct inequation `- nc - m + cs1 <= 0` in the following way.
-   - input_places + output_places <= 0
-
-For t2 construct inequation `- nc - m + cs2 <= 0` in the following way.
-   - input_places + output_places <= 0
-
-For t3 construct inequation `nc + m - cs1 <= 0` in the following way.
-   - input_places + output_places <= 0
-
-For t4 construct inequation `nc + m - cs2 <= 0` in the following way.
-   - input_places + output_places <= 0
-
-Output:
-  - nc - m + cs1   <= 0
-  - nc - m + cs2   <= 0
-  + nc + m - cs1 <= 0
-  + nc + m - cs2 <= 0
-
-  2 \cdot nc + m < 2 \cdot target1
-
-  Non-negativity conditions:
-
-  target1 >= 0
-  nc      >= 0
-  m       >= 0
-  cs1     >= target1
-  cs2     >= target1
+TODO: Write the code
 
@@
 
-* Theorem Dual State Constraints:
+  Input:
+    (S, T, E, I) : Petri net
+    A            : Satisfying assignment for C'
 
-  TODO: Unroll ?
+  Output:
+    I            : Invariant
+
+TODO: Write the code
 \end{verbatim}
@@
 \newpage
+\section{Method Safety by Refinement}
 
-\iffalse
-Constraints $\mathcal C'(N, P)$ with
-$Y_1 = \begin{pmatrix}target1 \end{pmatrix}$,
-$Y_2 =\begin{pmatrix}nc & m & cs1 & cs2\end{pmatrix}$,
-$A =\begin{pmatrix}0 & 0 & 1 & 1 \end{pmatrix}$ and
-$b =\begin{pmatrix}2 \end{pmatrix}$:
+The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
+We present the method Invariant by example on a mutual exclusion algorithm by Delzanno2001.
 
-\begin{align*}
-  target1 - nc - m + cs1 \le& 0 \\ 
-  target1 - nc - m + cs2 \le& 0 \\ 
-  - target1 + nc + m - cs1 \le& 0 \\ 
-  - target1 + nc + m - cs2 \le& 0 \\ 
-2 \cdot nc + m <& 2 \cdot target1 \\
-  target1 \ge& 0 \\
-  nc \ge& 0 \\
-  m \ge& 0 \\
-  cs1 \ge& 0 \\
-  cs2 \ge& 0
-\end{align*}
+\begin{verbatim}
 
-Satisfying assignment is $Y_1 = \begin{pmatrix} 1 \end{pmatrix}$ and
-$Y_2 = \begin{pmatrix}0 & 1 & 0 & 0 \end{pmatrix}$
+* Code
 
-Invariant $\text{Inv}(N, P, A) = (cs1 + cs2 + m \le 1)$
-\fi
+bool m1 = false;
+bool m2 = false;
+
+// Thread 1:
+while (true) {
+  if (! m2 && ! test_and_set(m1)) {
+    // critical section 1
+    m1 = false
+  }
+  if (! m1 && ! test_and_set(m2)) {
+    // critical section 2
+    m2 = false
+  }
+}
+
+// Thread 2:
+while (true) {
+  if (! m2 && ! test_and_set(m1)) {
+    // critical section 1
+    m1 = false
+  }
+  if (! m1 && ! test_and_set(m2)) {
+    // critical section 2
+    m2 = false
+  }
+}
+
+Property: Thread 1 and Thread 2 are never in any critical section at the same time.
+
+* Petri net:
+\end{verbatim}
+\begin{center}
+\begin{tikzpicture}[node distance=2cm]
+  \node[blue place, tokens=2, label=above:nc] (nc) {};
+  \node[place, tokens=1, label=right:m2, below of=nc, xshift=+0.75cm] (s2) {};
+  \node[place, tokens=1, label=left:m1, below of=nc, xshift=-0.75cm] (s1) {};
+  \node[red place, label=above:cs1, left of=s1] (cs1) {};
+  \node[red place, label=above:cs2, right of=s2] (cs2) {};
+  \node[transition, label=above:t1] (t1) at ($(s1)!0.5!(cs1) + (0,1cm)$) {}
+	edge [pre] (nc)
+	edge [pre] (s1)
+	edge [pre] (s2)
+	edge [post] (cs1)
+	edge [post] (s2);
+  \node[transition, label=above:t2] (t2) at ($(s2)!0.5!(cs2) + (0,1cm)$) {}
+	edge [pre] (nc)
+	edge [pre] (s2)
+	edge [pre] (s1)
+	edge [post] (cs2)
+	edge [post] (s1);
+  \node[transition, label=below:t3] (t3) at ($(s1)!0.5!(cs1) - (0,1cm)$) {}
+	edge [pre] (cs1)
+	edge [post] (s1)
+	edge [post,bend left,min distance=3cm, out=115, in=115] (nc);
+  \node[transition, label=below:t4] (t4) at ($(s2)!0.5!(cs2) - (0,1cm)$) {}
+	edge [pre] (cs2)
+	edge [post] (s2)
+	edge [post,bend right,min distance=3cm, out=-115, in=-115] (nc);
+\end{tikzpicture}
+\end{center}
+
+\newpage
+\begin{verbatim}
+* Method Safety:
+
+  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
+  Subprocedure \Delta constructs constraint \delta corresponding to A_\theta.
+\end{verbatim}
+\begin{center}
+\begin{tikzpicture}
+  \node[state] (begin) {BEGIN};
+  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
+  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
+  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
+                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
+  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
+  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
+                                    	$\delta:=\Delta(A_\theta)$\\
+                                    	$C:=C \cup \{\delta\}$};
+  \node[print, right=of satc] (yes) {YES};
+  \node[print, right=of satctheta] (dontknow) {Don't know};
+  \node[state, right=of yes] (end1) {END};
+  \node[state, right=of dontknow] (end2) {END};
+
+  \draw (begin) edge (c);
+  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
+  \draw (satc) edge node[above]{NO} (yes);
+  \draw (yes) edge (end1);
+  \draw (satc) edge node[right]{YES} (modelc);
+  \draw (modelc) edge (satctheta);
+  \draw (satctheta) edge node[above]{NO} (dontknow);
+  \draw (dontknow) edge (end2);
+  \draw (satctheta) edge node[right]{YES} (modelctheta);
+  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
+  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
+\end{tikzpicture}
+\end{center}
+
+\begin{verbatim}
+* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
+  1. S is a trap in the net N
+  2. S is marked in the initial marking M0
+  3. S is unmarked in the assignment A
+
+* Property of constraint \delta: Constraint \delta refines the abstraction, i.e.
+  1. A ^ \delta is unsat (\delta excludes A)
+  2. N |= \delta is sat  (\delta is a property of N)
+\end{verbatim}
+
+\begin{verbatim}
+* State constraints C:
+
+  nc  =  2 - t1 - t2 + t3 + t4
+  m1  =  1 - t1 + t3
+  m2  =  1 - t2 + t4
+  cs1 =  0 + t1 - t3
+  cs2 =  0 + t2 - t4
+
+  nc  >= 0
+  m1  >= 0
+  m2  >= 0
+  cs1 >= 0
+  cs2 >= 0
+  t1  >= 0
+  t2  >= 0
+  t3  >= 0
+  t4  >= 0
+
+* Assignment A:
+  nc  = 1
+  m1  = 0
+  m2  = 0
+  cs1 = 1
+  cs2 = 1
+  t1  = 1
+  t2  = 1
+  t3  = 0
+  t4  = 0
+
+* Trap conditions C_\theta:
+ TODO
+
+* Assignment A_\theta:
+ TODO
+
+* Constraint \delta:
+ TODO
+\end{verbatim}
 

### 2013-11-18 paper/workspace/main.tex
Updated paper
Added benchmark tables to report

@@
 * Assignment A:
-  nc  = 1
+  nc  = 0
   m1  = 0
@@
 * Trap conditions C_\theta:
- TODO
+  nc  => t1 /\ t2
+  m1  => t1 /\ t2
+  m2  => t1 /\ t2
+  cs1 => t3
+  cs2 => t4
+
+  t1  => m2 \/ cs1
+  t2  => m1 \/ cs2
+  t3  => m1 \/ nc
+  t4  => m2 \/ nc
+
+  nc \/ m1 \/ m2
+
+  ~cs1 /\ ~cs2
 
 * Assignment A_\theta:
- TODO
+  nc  = false
+  m1  = true
+  m2  = true
+  cs1 = false
+  cs2 = false
+
+  t1  = true
+  t2  = true
+  t3  = true
+  t4  = true
 
 * Constraint \delta:
- TODO
+  m1 + m2 >= 1
 \end{verbatim}
@@
 
+\section{Benchmarks of \bfc}
+
+We experimented on the benchmarks \ttt{cprover-PN} and
+\ttt{cprover\_software\_analysis}
+available at
+\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
+The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
+We applied 8 tool configurations to the set of benchmarks.
+Table~\ref{bfc-experiments} shows the corresponding experimental results.
+
+\begin{table}[h]
+\begin{center}
+  \begin{tabular}{ | r | p{7cm} | r | r | r | r | } %
+    \hline
+    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
+    \hline
+    C1 & State constraint over real numbers & 23 &  0 & 54 &  0 \\
+    C2 & State constraint over integers        & 23 &  0 & 54 &  0 \\
+    \hline
+    C3 & C2 + TrapConditions                   & 23 &  0 & 54 &  0 \\
+    C4 & C3 + SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
+    C5 & C4 + EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
+    \hline
+    C6 & C1 + state space exploration \& TO 1\,min & 26 & 34 & 0 & 17 \\
+    C7 & C1 + state space exploration \& TO 10\,min & 26 & 39 & 0 & 12 \\
+    C8 & C1 + state space exploration \& TO 8\,h   & 26 & 45 & 0 &  6 \\
+    \hline
+  \end{tabular}
+\end{center}
+\caption{Experimental results for the benchmarks of \bfc}
+\label{bfc-experiments}
+\end{table}
+
+Observations:
+\begin{itemize}
+\item State constraints over integers or reals correctly determine that the 23 benchmarks in Appendix~\ref{bfc-experiments-c1} are positive.
+\item Composing state space constraints with concrete state space
+  exploration gives the following 3 additional positives.
+\begin{verbatim}
+cprover-PN/lamport
+cprover-PN/newdekker
+cprover-PN/peterson
+\end{verbatim}
+\end{itemize}
+
+\newpage
+
+\section{Benchmarks of \mist}
+
+We experimented on the benchmarks available at \\
+\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
+and \\
+\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
+The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
+We applied 6 tool configurations to the set of benchmarks.
+Table~\ref{mist-experiments} shows the corresponding experimental results.
+
+\begin{table}[h]
+\begin{center}
+  \begin{tabular}{ | r | p{6cm} | r | r | r | r | }
+    \hline
+    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
+    \hline
+    C1 & State constraint with real numbers & 14 &  0 & 13 &  0 \\
+    C2 & State constraint with integers     & 14 &  0 & 13 &  0 \\
+    \hline
+    C3 & C2 + TrapConditions               & 20 &  0 &  7 &  0 \\
+    C4 & C3 + SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
+    C5 & C4 + EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
+    \hline
+    C9 & C4 + state space exploration \& TO 4\,h & 21 &  2 & 0 &  4 \\
+    \hline
+  \end{tabular}
+\end{center}
+\caption{Experimental results for the benchmarks of \mist}
+\label{mist-experiments}
+\end{table}
+
+Observations:
+\begin{itemize}
+\item State constraints over integers or reals correctly determine that the 14 benchmarks in Appendix~\ref{mist-experiments-c1} are positive.
+\item Composing state constraints with TrapConditions gives the following 6 additional positives.
+\begin{verbatim}
+PN/pingpong
+PN/basicME
+PN/multiME
+boundedPN/lamport
+boundedPN/newdekker
+boundedPN/peterson
+\end{verbatim}
+\item Composing state constraints with TrapConditions and SubnetTrapConditions gives the additional positive \ttt{PN/manufacturing}.
+  \item The remaining positive benchmarks are the following.
+\begin{verbatim}
+PN/extendedread-write
+PN/extendedread-write-smallconsts
+\end{verbatim}
+\end{itemize}
 \end{document}

### 2013-11-18 paper/workspace/main.tex
Updated main.
Updated paper

@@
 
-* State constraints C:
+* Example:
+  - State constraints C:
 
@@
 \begin{verbatim}
-* Dual state constraints C':
+* Example:
+  - Dual state constraints C':
 
@@
     cs2     >= target1
-\end{verbatim}
 
-\begin{verbatim}
-* Model A:
+  - Model A:
 
@@
 
-* Invariant:
+  - Invariant:
 
@@
 
+\newpage
 \begin{verbatim}
-* State constraints C:
-
+* Example
+  - State constraints C:
     nc  =  2 - t1 - t2 + t3 + t4
@@
     cs2 =  0 + t2 - t4
-
     nc  >= 0
@@
 
-* Assignment A:
+  - Assignment A:
     nc  = 0
@@
 
-* Trap conditions C_\theta:
+  - Trap conditions C_\theta:
     nc  => t1 /\ t2
@@
 
-* Assignment A_\theta:
+  - Assignment A_\theta:
     nc  = false
@@
 
-* Constraint \delta:
+  - Constraint \delta:
     m1 + m2 >= 1
 \end{verbatim}
+
+\section{Benchmarks of \bfc}
+
+We experimented on the benchmarks \ttt{cprover-PN} and
+\ttt{cprover\_software\_analysis}
+available at
+\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
+The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
+We applied 8 tool configurations to the set of benchmarks.
+Table~\ref{bfc-experiments} shows the corresponding experimental results.
+
+\begin{table}[h]
+\begin{center}
+  \begin{tabular}{ | r | p{7cm} | r | r | r | r | } %
+    \hline
+    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
+    \hline
+    C1 & State constraint over real numbers & 23 &  0 & 54 &  0 \\
+    C2 & State constraint over integers        & 23 &  0 & 54 &  0 \\
+    \hline
+    C3 & C2 + TrapConditions                   & 23 &  0 & 54 &  0 \\
+    C4 & C3 + SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
+    C5 & C4 + EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
+    \hline
+    C6 & C1 + state space exploration \& TO 1\,min & 26 & 34 & 0 & 17 \\
+    C7 & C1 + state space exploration \& TO 10\,min & 26 & 39 & 0 & 12 \\
+    C8 & C1 + state space exploration \& TO 8\,h   & 26 & 45 & 0 &  6 \\
+    \hline
+  \end{tabular}
+\end{center}
+\caption{Experimental results for the benchmarks of \bfc}
+\label{bfc-experiments}
+\end{table}
+
+Observations:
+\begin{itemize}
+\item State constraints over integers or reals correctly determine that the 23 benchmarks in Appendix~\ref{bfc-experiments-c1} are positive.
+\item Composing state space constraints with concrete state space
+  exploration gives the following 3 additional positives.
+\begin{verbatim}
+cprover-PN/lamport
+cprover-PN/newdekker
+cprover-PN/peterson
+\end{verbatim}
+\end{itemize}
+
+\newpage
+\section{Benchmarks of \mist}
+
+We experimented on the benchmarks available at \\
+\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
+and \\
+\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
+The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
+We applied 6 tool configurations to the set of benchmarks.
+Table~\ref{mist-experiments} shows the corresponding experimental results.
+
+\begin{table}[h]
+\begin{center}
+  \begin{tabular}{ | r | p{6cm} | r | r | r | r | }
+    \hline
+    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
+    \hline
+    C1 & State constraint with real numbers & 14 &  0 & 13 &  0 \\
+    C2 & State constraint with integers     & 14 &  0 & 13 &  0 \\
+    \hline
+    C3 & C2 + TrapConditions               & 20 &  0 &  7 &  0 \\
+    C4 & C3 + SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
+    C5 & C4 + EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
+    \hline
+    C9 & C4 + state space exploration \& TO 4\,h & 21 &  2 & 0 &  4 \\
+    \hline
+  \end{tabular}
+\end{center}
+\caption{Experimental results for the benchmarks of \mist}
+\label{mist-experiments}
+\end{table}
+
+Observations:
+\begin{itemize}
+\item State constraints over integers or reals correctly determine that the 14 benchmarks in Appendix~\ref{mist-experiments-c1} are positive.
+\item Composing state constraints with TrapConditions gives the following 6 additional positives.
+\begin{verbatim}
+PN/pingpong
+PN/basicME
+PN/multiME
+boundedPN/lamport
+boundedPN/newdekker
+boundedPN/peterson
+\end{verbatim}
+\item Composing state constraints with TrapConditions and SubnetTrapConditions gives the additional positive \ttt{PN/manufacturing}.
+  \item The remaining positive benchmarks are the following.
+\begin{verbatim}
+PN/extendedread-write
+PN/extendedread-write-smallconsts
+\end{verbatim}
+\end{itemize}
 
@@
 
-\section{Benchmarks of \bfc}
-
-We experimented on the benchmarks \ttt{cprover-PN} and
-\ttt{cprover\_software\_analysis}
-available at
-\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
-The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
-We applied 8 tool configurations to the set of benchmarks.
-Table~\ref{bfc-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | r | p{7cm} | r | r | r | r | } %
-    \hline
-    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
-    \hline
-    C1 & State constraint over real numbers & 23 &  0 & 54 &  0 \\
-    C2 & State constraint over integers        & 23 &  0 & 54 &  0 \\
-    \hline
-    C3 & C2 + TrapConditions                   & 23 &  0 & 54 &  0 \\
-    C4 & C3 + SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
-    C5 & C4 + EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
-    \hline
-    C6 & C1 + state space exploration \& TO 1\,min & 26 & 34 & 0 & 17 \\
-    C7 & C1 + state space exploration \& TO 10\,min & 26 & 39 & 0 & 12 \\
-    C8 & C1 + state space exploration \& TO 8\,h   & 26 & 45 & 0 &  6 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \bfc}
-\label{bfc-experiments}
-\end{table}
-
-Observations:
-\begin{itemize}
-\item State constraints over integers or reals correctly determine that the 23 benchmarks in Appendix~\ref{bfc-experiments-c1} are positive.
-\item Composing state space constraints with concrete state space
-  exploration gives the following 3 additional positives.
-\begin{verbatim}
-cprover-PN/lamport
-cprover-PN/newdekker
-cprover-PN/peterson
-\end{verbatim}
-\end{itemize}
-
-\newpage
-
-\section{Benchmarks of \mist}
-
-We experimented on the benchmarks available at \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
-and \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
-The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
-We applied 6 tool configurations to the set of benchmarks.
-Table~\ref{mist-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | r | p{6cm} | r | r | r | r | }
-    \hline
-    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
-    \hline
-    C1 & State constraint with real numbers & 14 &  0 & 13 &  0 \\
-    C2 & State constraint with integers     & 14 &  0 & 13 &  0 \\
-    \hline
-    C3 & C2 + TrapConditions               & 20 &  0 &  7 &  0 \\
-    C4 & C3 + SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
-    C5 & C4 + EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
-    \hline
-    C9 & C4 + state space exploration \& TO 4\,h & 21 &  2 & 0 &  4 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \mist}
-\label{mist-experiments}
-\end{table}
-
-Observations:
-\begin{itemize}
-\item State constraints over integers or reals correctly determine that the 14 benchmarks in Appendix~\ref{mist-experiments-c1} are positive.
-\item Composing state constraints with TrapConditions gives the following 6 additional positives.
-\begin{verbatim}
-PN/pingpong
-PN/basicME
-PN/multiME
-boundedPN/lamport
-boundedPN/newdekker
-boundedPN/peterson
-\end{verbatim}
-\item Composing state constraints with TrapConditions and SubnetTrapConditions gives the additional positive \ttt{PN/manufacturing}.
-  \item The remaining positive benchmarks are the following.
-\begin{verbatim}
-PN/extendedread-write
-PN/extendedread-write-smallconsts
-\end{verbatim}
-\end{itemize}
 \end{document}

### 2013-11-18 paper/workspace/main.tex
Updated paper
Updated main

@@
     |         |   |    |    |    |
-    place nc  |   ------    ------
+    |         |   ------    ------
+    |         |      |         |
+    |         |      |       # of tokens given to nc
     |         |      |    
-              |      |       input transitions of nc
+    |         |    # of tokens taken from nc
     |         |
-              |    output transitions of nc
+    |      initial number of tokens in nc
     |
-           initial number of tokens in nc
+    number of tokens in nc
 
@@
   Input:
-    (S, T, E, I) : Petri net
+    (S, T, E, M0) : Petri net
   Output:
@@
 
-TODO: Write the code
+TODO: Unroll pseudocode?
 
@@
 
-TODO: Write the code
+TODO: Unroll pseudocode?
 \end{verbatim}
@@
   Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs constraint \delta corresponding to A_\theta.
+  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
 \end{verbatim}
@@
 
-* Property of constraint \delta: Constraint \delta refines the abstraction, i.e.
+* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
+
+* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
   1. A ^ \delta is unsat (\delta excludes A)
@@
   - Trap conditions C_\theta:
-    nc  => t1 /\ t2
-    m1  => t1 /\ t2
-    m2  => t1 /\ t2
-    cs1 => t3
-    cs2 => t4
 
-    t1  => m2 \/ cs1
-    t2  => m1 \/ cs2
-    t3  => m1 \/ nc
-    t4  => m2 \/ nc
+    - Trap implications:
 
+    cs1 => m1 \/ nc
+    ^      ^     ^
+    |      |     |
+    |      |     nc \in S
+    |      |
+    |      m1 \in S
+    |
+    cs1 \in S
+
+    cs2 => m2 \/ nc
+    nc  => (m2 \/ cs1) /\ (m1 \/ cs2)
+    m1  => (m2 \/ cs1) /\ (m1 \/ cs2)
+    m2  => (m2 \/ cs1) /\ (m1 \/ cs2)
+
+    - At least one of the intially marked places belongs to S:
     nc \/ m1 \/ m2
 
+    - None of the marked places in A belongs to S:
     ~cs1 /\ ~cs2
@@
     cs2 = false
-
-    t1  = true
-    t2  = true
-    t3  = true
-    t4  = true
-
-  - Constraint \delta:
-    m1 + m2 >= 1
 \end{verbatim}
 
+\newpage
+\begin{verbatim}
+  - Refinement constraint \delta:
+    m1 + m2 >= 1
+    ^    ^
+    |    |
+    ------
+      |
+      S = {m1, m2}, therefore \delta excludes assignment A in the next iteration
+
+* Trap implication:
+
+  place s \in S =>    /\       \    /
+                     /  \       \  /     place p \in S
+                    /    \       \/
+                   t \in s*   p \in t*
+
+  "if s is in trap S then for each output transition t at least one successor p is in trap S"
+
+* Refinement constraint \delta:
+
+  \Sigma s >= 1  
+  A(s)
+
+  "At least one place in S is always marked"
+
+* Subprocedure TrapConditions:
+
+  Input:
+    (S, T, E, M0) : Petri net
+    A             : Satisfying assignment for C \cup { ~P }
+  Output:
+    C_\theta      : Trap conditions
+
+  TODO: Unroll pseudocode?
+
+* Subprocedure \Delta:
+
+  Input:
+    A_\theta      : Satisfying assignment for C_\theta
+  Output:
+    \delta        : Refinement constraint \delta
+
+  TODO: Unroll pseudocode?
+\end{verbatim}
+
+\newpage
 \section{Benchmarks of \bfc}

### 2013-11-21 src/minimization-constraints-to-z3.pl
Improved method for minimization of invariants

@@ -1,23 +1,0 @@
-:- dynamic place/3.        % place(Id, InTransitions, OutTransitions).
-:- dynamic assignment/2.   % assignment(Variable, Value).
-:- dynamic set_zero/1.     % set_zero(Place).
-
-:- ['load-pl-file.pl'].
-:- ['misc.pl'].
-
-z3_constraints :-
-        findall( _ , (
-                       place(P, _, _),
-                       assignment(P, 0),
-                       format('(assert (= ~p 0))\n', [P])
-                     ), _ ),
-        set_zero(P1),
-        format('(assert (= ~p 0))\n', [P1]).
-
-% Entry point
-:-      prolog_flag(argv, Argv),
-        (   foreach(F, Argv)
-        do  load_pl_file(F)
-        ),
-        z3_constraints,
-        halt.

### 2013-11-21 src/minimization-constraints.pl

@@ -1,23 +1,31 @@
 :- dynamic place/3.        % place(Id, InTransitions, OutTransitions).
 :- dynamic assignment/2.   % assignment(Variable, Value).
-:- dynamic set_zero/1.     % set_zero(Place).
 
 :- ['load-pl-file.pl'].
 :- ['misc.pl'].
 
-next_constraint :-
-        place(P, _, _),
-        assignment(P, V),
-        V > 0,
-        \+ set_zero(P),
-        portray_clause(set_zero(P)).
+minimization_constraints :-
+        findall( _ , (
+             place(P, _, _),
+             format('(declare-fun b_~p () Int)\n', [P]),
+             format('(assert (implies (> ~p 0) (= b_~p 1)))\n', [P, P]),
+             format('(assert (implies (= ~p 0) (= b_~p 0)))\n', [P, P])
+        ), _ ),
+        findall(P, place(P, _, _), Ps),
+        findall(P, (
+             place(P, _, _),
+             assignment(P, V),
+             V > 0
+        ), PsV),
+        length(PsV, Count),
+        print('(assert (< '),
+        format_sum('b_~p', Ps),
+        format(' ~p))', [Count]).
 
 % Entry point
 :-      prolog_flag(argv, Argv),
         (   foreach(F, Argv)
         do  load_pl_file(F)
         ),
-        ( next_constraint ->
-          halt(0)
-        ; halt(1)
-        ).
+        minimization_constraints,
+        halt.

### 2013-11-21 src/minimization-factor.pl

@@ -1,32 +1,0 @@
-:- dynamic place/3.        % place(Id, InTransitions, OutTransitions).
-:- dynamic assignment/2.   % assignment(Variable, Value).
-
-:- use_module(library(aggregate)).
-
-:- ['load-pl-file.pl'].
-:- ['misc.pl'].
-
-minimization_constraint :-
-        findall( _ , (
-             place(P, _, _),
-             format('(declare-fun b_~p () Int)\n', [P]),
-             format('(assert (implies (> ~p 0) (= b_~p 1)))\n', [P, P]),
-             format('(assert (implies (= ~p 0) (= b_~p 0)))\n', [P, P]),
-        ), _ ),
-        findall(P, place(P, _, _), Ps),
-        aggregate(count, (
-             place(P, _, _),
-             assignment(P, V),
-             V > 0
-        ), Sum),
-        print('(assert (< '),
-        format_sum('b_~p', Ps),
-        format(' ~p))', [Sum]).
-
-% Entry point
-:-      prolog_flag(argv, Argv),
-        (   foreach(F, Argv)
-        do  load_pl_file(F)
-        ),
-        minimization_constraint,
-        halt.

### 2013-11-21 paper/workspace/main.tex
Updated paper
Added method Invariant Minimization to paper draft

@@
 \usepackage[usenames,dvipsnames]{color}
-\usepackage{multirow}
 \usepackage{listings}
@@
 \usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
-	snakes,automata,backgrounds,petri,calc}
-\usetikzlibrary{positioning}
+	snakes,automata,backgrounds,petri,calc,positioning}
 
@@
     \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
-      $I := \text{Inv}(N, P, A)$};
+      $I := \text{Inv}(N, A)$};
     \node[print, below=of inv] (printinv) {Invariant: $I$};
@@
   Input:
-    (S, T, E, I) : Petri net
+    (S, T, E, M0) : Petri net
     P             : Property
@@
 Input:
- N = (S, T, E, I)
+ N = (S, T, E, M0)
    = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
@@
   Input:
-    (S, T, E, I) : Petri net
+    (S, T, E, M0) : Petri net
     A             : Satisfying assignment for C'
@@
 \newpage
+\section{Method Invariant Minimization}
+
+The method Invariant Minimization constructs an invariant \verb=I= that uses a minimal number places for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
+We present the method Invariant Minimization by example on a simplified mutual exclusion algorithm by Delzanno2001.
+
+\begin{verbatim}
+* Code, property, and Petri net: same as in section Method Safety.
+\end{verbatim}
+
+\begin{verbatim}
+* Method Invariant
+
+  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A
+\end{verbatim}
+
+\begin{center}
+  \begin{tikzpicture}
+    \node[state] (begin) {BEGIN};
+    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
+    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
+    \node[print, right=of satc] (noinv) {No Invariant};
+    \node[action, below=of satc] (model1) {$A:=\text{Model}(C')$};
+    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A)$};
+    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
+    \node[action, right=of satcm] (model2) {$A:=\text{Model}(C' \cup C_M)$};
+    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A)$};
+    \node[print, below=of inv] (printinv) {Invariant: $I$};
+    \node[state, right=of noinv] (end1) {END};
+    \node[state, below=of printinv] (end2) {END};
+
+    \draw (begin) edge (c);
+    \draw (c) edge (satc);
+    \draw (satc) edge node[above]{NO} (noinv);
+    \draw (noinv) edge (end1);
+    \draw (satc) edge node[right]{YES} (model1);
+    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
+    \draw (cm) edge (satcm);
+    \draw (satcm) edge node[above]{YES} (model2);
+    \draw (model2) |- (edgein);
+    \draw (satcm) edge node[right]{NO} (inv);
+    \draw (inv) edge (printinv);
+    \draw (printinv) edge (end2);
+  \end{tikzpicture}
+\end{center}
+
+\begin{verbatim}
+* Property of minimization constraints C_M generated from A: 
+    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
+\end{verbatim}
+
+\newpage
+\begin{verbatim}
+* Example:
+  - Dual state constraints C' and satisfying assignment A for C' as in example for method Invariant
+
+  - Minimization constraints:
+
+    nc  ~= 0 => m_nc = 1
+    |           --------
+    |              |
+    |           place appears in invariant
+    |
+    place coefficient in the invariant
+
+    nc   = 0 => m_nc  = 0
+    |           ---------
+    |              |
+    |           place does not appear in invariant
+    |
+    place coefficient in the invariant
+
+    m   ~= 0 => m_m   = 1
+    m    = 0 => m_m   = 0
+    cs1 ~= 0 => m_cs1 = 1
+    cs1  = 0 => m_cs1 = 0
+    cs2 ~= 0 => m_cs2 = 1
+    cs2  = 0 => m_cs2 = 0
+
+    m_nc + m_m + m_cs1 + m_cs2 < 3
+    --------------------------   |
+                |                |
+                |    number of places appearing in current invariant for A
+                |    = #{s | A(s) ~= 0} = #{m, cs1, cs2}
+                |
+                |
+  number of places appearing in new invariant
+\end{verbatim}
+
+\begin{verbatim}
+* Subprocedure \text{MinConstraints}:
+
+  Input:
+    (S, T, E, M0) : Petri net
+    A             : Assignment
+  Output:
+    C_M           : Minimization Constraints
+
+Input:
+ N   = (S, T, E, M0)
+     = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
+ A   = ( nc=0, m=1, cs1=1, cs2=1, target1=1 )
+
+TODO: Unroll pseudocode?
+
+\end{verbatim}
+
+\newpage
+
 \section{Method Safety by Refinement}
@@
     \hline
+    C10 & Method Invariant                  & 23 &  0 & 54 &  0 \\
+    \hline
   \end{tabular}
@@
     \hline
+    C10 & Method Invariant                  & 14 &  0 & 13 &  0 \\
+    \hline
   \end{tabular}
@@
 
-Invariant $\text{Inv}(N, P, A) = (cs1 + cs2 + m \le 1)$
+Invariant $\text{Inv}(N, A) = (cs1 + cs2 + m \le 1)$
 
@@
 & (Y_1 \ge 0) \land (Y_2 \ge 0)) \\
-  \text{Inv}(N, P, A) =& ( ( Y_1 \cdot A + Y_2) \cdot M \le ( Y_1 \cdot A + Y_2) \cdot M_0)
+  \text{Inv}(N, A) =& ( ( Y_1 \cdot A + Y_2) \cdot M \le ( Y_1 \cdot A + Y_2) \cdot M_0)
 \end{align*}
@@
   \node[action, below=of satc] (inv) {$A_\theta:=\text{Model}(C')$\\
-                                  	$I := \text{Inv}(N, P, A)$};
+                                  	$I := \text{Inv}(N, A)$};
   \node[print, below=of inv] (printinv) {Invariant: $I$};

### 2013-11-25 paper/workspace/main.tex
Updated table of experiments in paper draft.

@@
 % \institute{Technische Universit\"at M\"unchen}
-\date{Wed Nov 13 09:11:15 CEST 2013}
+\date{Mon Nov 25 14:24:10 CET 2013}
 
@@
     \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-    \node[print, right=of satc] (yes) {YES};
+    \node[print, right=of satc] (yes) {N satisfies P};
     \node[print, below=of satc] (dontknow) {Don't know};
@@
 
-  # of tokens in s = initial number of tokens of place s - # times each output transition of s fires + # times each input transition of s fires
+  # of tokens in s = initial number of tokens of place s
+                     - # times each output transition of s fires
+                     + # times each input transition of s fires
 
@@
                                     	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {YES};
+  \node[print, right=of satc] (yes) {N satisfies P};
   \node[print, right=of satctheta] (dontknow) {Don't know};
@@
 We experimented on the benchmarks \ttt{cprover-PN} and
-\ttt{cprover\_software\_analysis}
-available at
+\ttt{cprover\_software\_analysis} available at \\
 \ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
 The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
-We applied 8 tool configurations to the set of benchmarks.
 Table~\ref{bfc-experiments} shows the corresponding experimental results.
@@
 \begin{center}
-  \begin{tabular}{ | r | p{7cm} | r | r | r | r | } %
+  \begin{tabular}{ | p{7cm} | r | r | r | r | } %
     \hline
-    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
+    Method & pos & neg & don't know & TO \\
     \hline
-    C1 & State constraint over real numbers & 23 &  0 & 54 &  0 \\
-    C2 & State constraint over integers        & 23 &  0 & 54 &  0 \\
-    \hline
-    C3 & C2 + TrapConditions                   & 23 &  0 & 54 &  0 \\
-    C4 & C3 + SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
-    C5 & C4 + EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
-    \hline
-    C6 & C1 + state space exploration \& TO 1\,min & 26 & 34 & 0 & 17 \\
-    C7 & C1 + state space exploration \& TO 10\,min & 26 & 39 & 0 & 12 \\
-    C8 & C1 + state space exploration \& TO 8\,h   & 26 & 45 & 0 &  6 \\
-    \hline
-    C10 & Method Invariant                  & 23 &  0 & 54 &  0 \\
+    Safety (over real numbers or naturals) & 23 &  0 & 54 &  0 \\
+    Invariant                              & 23 &  0 & 54 &  0 \\
+    Invariant Minimization                 & 23 &  0 & 54 &  0 \\
+    Safety by Refinement                   & 23 &  0 & 54 &  0 \\
     \hline
@@
 
-Observations:
-\begin{itemize}
-\item State constraints over integers or reals correctly determine that the 23 benchmarks in Appendix~\ref{bfc-experiments-c1} are positive.
-\item Composing state space constraints with concrete state space
-  exploration gives the following 3 additional positives.
-\begin{verbatim}
-cprover-PN/lamport
-cprover-PN/newdekker
-cprover-PN/peterson
-\end{verbatim}
-\end{itemize}
-
-\newpage
 \section{Benchmarks of \mist}
@@
 The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
-We applied 6 tool configurations to the set of benchmarks.
 Table~\ref{mist-experiments} shows the corresponding experimental results.
@@
 \begin{center}
-  \begin{tabular}{ | r | p{6cm} | r | r | r | r | }
+  \begin{tabular}{ | p{6cm} | r | r | r | r | }
     \hline
-    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
+    Method & pos & neg & don't know & TO \\
     \hline
-    C1 & State constraint with real numbers & 14 &  0 & 13 &  0 \\
-    C2 & State constraint with integers     & 14 &  0 & 13 &  0 \\
-    \hline
-    C3 & C2 + TrapConditions               & 20 &  0 &  7 &  0 \\
-    C4 & C3 + SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
-    C5 & C4 + EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
-    \hline
-    C9 & C4 + state space exploration \& TO 4\,h & 21 &  2 & 0 &  4 \\
-    \hline
-    C10 & Method Invariant                  & 14 &  0 & 13 &  0 \\
+    Safety (over real numbers or naturals) & 14 &  0 & 13 &  0 \\
+    Invariant                              & 14 &  0 & 13 &  0 \\
+    Invariant Minimization                 & 14 &  0 & 13 &  0 \\
+    Safety by Refinement                   & 20 &  0 &  7 &  0 \\
     \hline
@@
 
-Observations:
-\begin{itemize}
-\item State constraints over integers or reals correctly determine that the 14 benchmarks in Appendix~\ref{mist-experiments-c1} are positive.
-\item Composing state constraints with TrapConditions gives the following 6 additional positives.
-\begin{verbatim}
-PN/pingpong
-PN/basicME
-PN/multiME
-boundedPN/lamport
-boundedPN/newdekker
-boundedPN/peterson
-\end{verbatim}
-\item Composing state constraints with TrapConditions and SubnetTrapConditions gives the additional positive \ttt{PN/manufacturing}.
-  \item The remaining positive benchmarks are the following.
-\begin{verbatim}
-PN/extendedread-write
-PN/extendedread-write-smallconsts
-\end{verbatim}
-\end{itemize}
-
-\iffalse
-
-\paragraph{Intermediate results}
-Constraints $\mathcal C'(N, P)$ with
-$Y_1 = \begin{pmatrix}target1 \end{pmatrix}$,
-$Y_2 =\begin{pmatrix}nc & m & cs1 & cs2\end{pmatrix}$,
-$A =\begin{pmatrix}0 & 0 & 1 & 1 \end{pmatrix}$ and
-$b =\begin{pmatrix}2 \end{pmatrix}$:
-
-\begin{align*}
-  target1 - nc - m + cs1 \le& 0 \\ 
-  target1 - nc - m + cs2 \le& 0 \\ 
-  - target1 + nc + m - cs1 \le& 0 \\ 
-  - target1 + nc + m - cs2 \le& 0 \\ 
-2 \cdot nc + m <& 2 \cdot target1 \\
-  target1 \ge& 0 \\
-  nc \ge& 0 \\
-  m \ge& 0 \\
-  cs1 \ge& 0 \\
-  cs2 \ge& 0
-\end{align*}
-
-Satisfying assignment is $Y_1 = \begin{pmatrix} 1 \end{pmatrix}$ and
-$Y_2 = \begin{pmatrix}0 & 1 & 0 & 0 \end{pmatrix}$
-
-Invariant $\text{Inv}(N, A) = (cs1 + cs2 + m \le 1)$
-
-\paragraph{Technique}
-\begin{align*}
-  \mathcal C'(N, P) =& ( ( ( Y_1 \cdot A + Y_2) \cdot C \le 0) \land \\
-& ((Y_1 \cdot A + Y_2) \cdot M_0 < Y_1 \cdot b) \land \\
-& (Y_1 \ge 0) \land (Y_2 \ge 0)) \\
-  \text{Inv}(N, A) =& ( ( Y_1 \cdot A + Y_2) \cdot M \le ( Y_1 \cdot A + Y_2) \cdot M_0)
-\end{align*}
-where $M_0$ is the initial marking, $C$ is the incidence matrix of the net and
-$A\cdot M \ge b$ is equivalent to $\neg P$.
-
-\paragraph{Method Invariant}
-\mbox{ } \\
-\begin{tikzpicture}
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-  \node[print, right=of satc] (noinv) {No Invariant};
-  \node[action, below=of satc] (inv) {$A_\theta:=\text{Model}(C')$\\
-                                  	$I := \text{Inv}(N, A)$};
-  \node[print, below=of inv] (printinv) {Invariant: $I$};
-  \node[state, right=of noinv] (end1) {END};
-  \node[state, below=of printinv] (end2) {END};
-
-  \draw (begin) edge (c);
-  \draw (c) edge (satc);
-  \draw (satc) edge node[above]{NO} (noinv);
-  \draw (satc) edge node[right]{YES} (inv);
-  \draw (noinv) edge (end1);
-  \draw (inv) edge (printinv);
-  \draw (printinv) edge (end2);
-\end{tikzpicture}
-
-\section{Method Safety by Refinement}
-
-\paragraph{Example: Mutual exclusion algorithm in Delzanno2001}
-
-\begin{verbatim}
-* Code
-
-bool m1 = false;
-bool m2 = false;
-
-// Thread 1:
-while (true) {
-  if (! m2 && ! test_and_set(m1)) {
-    // critical section 1
-    m1 = false
-  }
-  if (! m1 && ! test_and_set(m2)) {
-    // critical section 2
-    m2 = false
-  }
-}
-
-// Thread 2:
-while (true) {
-  if (! m2 && ! test_and_set(m1)) {
-    // critical section 1
-    m1 = false
-  }
-  if (! m1 && ! test_and_set(m2)) {
-    // critical section 2
-    m2 = false
-  }
-}
-
-Property: The number of threads in critical sections should be at most 1
-
-* Petri net
-
-Place m1 is marked in the net exactly if m1 is false in the algorithm.
-Place m2 is marked in the net exactly if m2 is false in the algorithm.
-
-\end{verbatim}
-
-\paragraph{Example: Petri net for mutual exclusion algorithm in Delzanno2001}
-\mbox{ } \\
-\begin{tikzpicture}[node distance=2cm]
-  \node[blue place, tokens=2, label=above:nc] (nc) {};
-  \node[place, tokens=1, label=right:m2, below of=nc, xshift=+0.75cm] (s2) {};
-  \node[place, tokens=1, label=left:m1, below of=nc, xshift=-0.75cm] (s1) {};
-  \node[red place, label=above:cs1, left of=s1] (cs1) {};
-  \node[red place, label=above:cs2, right of=s2] (cs2) {};
-  \node[transition, label=above:t1] (t1) at ($(s1)!0.5!(cs1) + (0,1cm)$) {}
-	edge [pre] (nc)
-	edge [pre] (s1)
-	edge [pre] (s2)
-	edge [post] (cs1)
-	edge [post] (s2);
-  \node[transition, label=above:t2] (t2) at ($(s2)!0.5!(cs2) + (0,1cm)$) {}
-	edge [pre] (nc)
-	edge [pre] (s2)
-	edge [pre] (s1)
-	edge [post] (cs2)
-	edge [post] (s1);
-  \node[transition, label=below:t3] (t3) at ($(s1)!0.5!(cs1) - (0,1cm)$) {}
-	edge [pre] (cs1)
-	edge [post] (s1)
-	edge [post,bend left,min distance=3cm, out=115, in=115] (nc);
-  \node[transition, label=below:t4] (t4) at ($(s2)!0.5!(cs2) - (0,1cm)$) {}
-	edge [pre] (cs2)
-	edge [post] (s2)
-	edge [post,bend right,min distance=3cm, out=-115, in=-115] (nc);
-\end{tikzpicture}
-\begin{align*}
-  P &= ( cs1 + cs2 \le 1 ) \\
-  \neg P &= ( cs1 + cs2 \ge 2 )
-\end{align*}
-
-\paragraph{Intermediate results}
-Constraints $\mathcal C(N)$ with
-$M =\begin{pmatrix}nc & m1 & m2 & cs1 & cs2 \end{pmatrix}^T$,
-$M_0 =\begin{pmatrix}2 & 1 & 1 & 0 & 0 \end{pmatrix}^T$,
-$X =\begin{pmatrix}t1 & t2 & t3 & t4 \end{pmatrix}^T$ and
-$C =\begin{pmatrix}
-  -1 & -1 & 1 & 1 \\
-  -1 &  0 & 1 & 0 \\
-   0 & -1 & 0 & 1 \\
-  1 & 0 & -1 & 0 \\
-  0 & 1 & 0 & -1
-\end{pmatrix}$:
-\begin{align*}
-  nc   =& 2 - t1 - t2 + t3 + t4 \\
-  m1  =& 1 - t1 + t3 \\
-  m2  =& 1 - t2 + t4 \\
-  cs1 =& 0 + t1 - t3 \\
-  cs2 =& 0 + t2 - t4 \\
-  nc \ge& 0 \\
-  m1 \ge& 0 \\
-  m2 \ge& 0 \\
-  cs1 \ge& 0 \\
-  cs2 \ge& 0
-\end{align*}
-
-Satisfying assignment is
-$M =\begin{pmatrix}0 & 0 & 0 & 1 & 1 \end{pmatrix}^T$ and
-$X =\begin{pmatrix}1 & 1 & 0 & 0 \end{pmatrix}^T$.
-
-A set satisfying the trap conditions is $S = \{m1, m2\}$.
-
-The trap constraint is $\delta = (m1 + m2 \ge 1)$.
-
-With the additional constraint, the set of constraints is unsatisfiable.
-
-\paragraph{Technique trap conditions} For a petri net $N$ and an assignment $A$,
-find a set $S$ that satisfies
-\begin{enumerate}
-  \item $S$ is a trap in the net $N$.
-  \item $S$ is marked in the initial marking $M_0$.
-  \item $S$ is unmarked in the assignment $A$.
-\end{enumerate}
-For such a set $S$, generate a constraint
-$\delta = \left( \sum_{s \in S} s \ge 1 \right)$, ensuring the
-trap is marked in any assignment.
-
-\paragraph{Method Safety by Refinement}
-\mbox{ } \\
-\begin{tikzpicture}
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                    	$\delta:=\Delta(A_\theta)$\\
-                                    	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {YES};
-  \node[print, right=of satctheta] (dontknow) {Don't know};
-  \node[state, right=of yes] (end1) {END};
-  \node[state, right=of dontknow] (end2) {END};
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (yes);
-  \draw (yes) edge (end1);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (dontknow);
-  \draw (dontknow) edge (end2);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-
-\section{Method Invariant by Refinement}
-
-\paragraph{Example}
-See example for method Safety by Refinement.
-
-\paragraph{Intermediate results}
-Constraints $\mathcal C'(N)$ with
-$Y_1 = \begin{pmatrix}target1 \end{pmatrix}$,
-$Y_2 = \begin{pmatrix}trap2 \end{pmatrix}$,
-$Y_3 =\begin{pmatrix}nc & m & cs1 & cs2\end{pmatrix}$,
-the trap from the trap refinement as
-$D =\begin{pmatrix}0 & 1 & 1 & 0 & 0 \end{pmatrix}$,
-$A =\begin{pmatrix}0 & 0 & 0 & 1 & 1 \end{pmatrix}$ and
-$b =\begin{pmatrix}1 \end{pmatrix}$:
-
-\begin{align*}
-  target1 - trap1 - nc - m1 + cs1 \le& 0 \\ 
-  target1 - trap1 - nc - m2 + cs2 \le& 0 \\ 
-  - target1 + trap1 + nc + m1 - cs1 \le& 0 \\ 
-  - target1 + trap1 + nc + m2 - cs2 \le& 0 \\ 
-2 \cdot trap1 + 2 \cdot nc + m1 + m2 <& 2 \cdot target1 + trap1 \\
-  target1 \ge& 0 \\
-  trap1 \ge& 0 \\
-  nc \ge& 0 \\
-  m1 \ge& 0 \\
-  m2 \ge& 0 \\
-  cs1 \ge& 0 \\
-  cs2 \ge& 0
-\end{align*}
-
-Satisfying assignment is
-$Y_1 = \begin{pmatrix} 1 \end{pmatrix}$,
-$Y_2 = \begin{pmatrix} 1 \end{pmatrix}$ and
-$Y_3 = \begin{pmatrix}0 & 0 & 0 & 0 & 0 \end{pmatrix}$.
-
-Invariant $\text{Inv}(N, D, P, A) =
-((cs1 + cs2 + m1 + m2 \le 2) \land (m1 + m2 \ge 1))$.
-
-\paragraph{Technique}
-\begin{align*}
-  \mathcal C'(N, D, P) =& ( ( ( Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot C \le 0) \land \\
-& ((Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot M_0 < Y_1 \cdot b + Y_2 \cdot \mathbf{1}) \land \\
-& (Y_1 \ge 0) \land (Y_2 \ge 0) \land (Y_3 \ge 0)) \\
-  \text{Inv}(N, D, P, A) =& ( ( Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot M \le ( Y_1 \cdot A + Y_2 \cdot D + Y_3) \cdot M_0) \land ( D \cdot M \ge \mathbf{1} )
-\end{align*}
-where $M_0$ is the initial marking, $C$ is the incidence matrix of the net, $A\cdot M \ge b$ is equivalent to $\neg P$ and $D \cdot M \ge \mathbf{1}$ is equivalent to the trap constraints.
-
-\paragraph{Method Invariant by Refinement}
-\mbox{ } \\
-\begin{tikzpicture}
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
-                                 	$D:=\{\}$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                                   	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                        	$\delta:=\Delta(A_\theta)$\\
-                                        	$C:=C \cup \{\delta\}$\\
-                                        	$D:=D \cup \{\delta\}$};
-  \node[print, right=of satctheta] (noinv2) {No Invariant};
-  \node[state, below=of noinv2] (end3) {END};
-
-  \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, D, P)$};
-  \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
-  \node[action, below=of satcprime] (inv) {$A_\theta:=\text{Model}(C')$\\
-                                      	$I := \text{Inv}(N, D, P, A)$};
-  \node[print, below=of inv] (printinv) {Invariant: $I$};
-  \node[print, right=of satcprime] (noinv) {No Invariant};
-  \node[state, below=of noinv] (end1) {END};
-  \node[state, below=of printinv] (end2) {END};
-
-  \draw (cprime) edge (satcprime);
-  \draw (satcprime) edge node[above]{NO} (noinv);
-  \draw (satcprime) edge node[right]{YES} (inv);
-  \draw (inv) edge (printinv);
-  \draw (noinv) edge (end1);
-  \draw (printinv) edge (end2);
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (cprime);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (noinv2);
-  \draw (noinv2) edge (end3);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-
-\fi
-
 \end{document}

### 2013-11-25 benchmarks/ic3-soter/concdb__single_client_writes__depth_0__fixed.spec.pl
New experiments.

@@ -1,0 +1,1290 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('s39').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
[…]

### 2013-11-25 benchmarks/ic3-soter/finite_leader__single_leader__depth_0__fixed.spec.pl

@@ -1,0 +1,608 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
[…]

### 2013-11-25 benchmarks/ic3-soter/finite_leader__single_leader__depth_2__fixed.spec.pl

@@ -1,0 +1,904 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
[…]

### 2013-11-25 benchmarks/ic3-soter/firewall__no_pred_called_with_zero__depth_0__fixed.spec.pl

@@ -1,0 +1,755 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
[…]

### 2013-11-25 benchmarks/ic3-soter/firewall__no_pred_called_with_zero__depth_1__fixed.spec.pl

@@ -1,0 +1,1807 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('s39').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
[…]

### 2013-11-25 benchmarks/ic3-soter/howait__all_workers_finished_if_wait_over__depth_0__fixed.spec.pl

@@ -1,0 +1,452 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
+place('l101').
[…]

### 2013-11-25 benchmarks/ic3-soter/parikh__should_already_be_initialized__depth_0__fixed.spec.pl

@@ -1,0 +1,150 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+transition(t1, ['l0', 's0'], ['l0', 's0', ('s0', -1), 's1', ('l0', -1), 'l1']).
+transition(t2, ['l0', 's1'], ['l0', 's1', ('s1', -1), 's2']).
+transition(t3, ['l1', 's1'], ['l1', 's1', ('l1', -1), 'l17']).
+transition(t4, ['l2', 's1'], ['l2', 's1', ('l2', -1), 'l3']).
+transition(t5, ['l3', 's1'], ['l3', 's1', ('l3', -1), 'l12']).
+transition(t6, ['l4', 's1'], ['l4', 's1', ('l4', -1), 'l5']).
+transition(t7, ['l5', 's1'], ['l5', 's1', ('l5', -1), 'l24']).
[…]

### 2013-11-25 benchmarks/ic3-soter/parikh__should_already_be_initialized__depth_1__fixed.spec.pl

@@ -1,0 +1,173 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+transition(t1, ['l0', 's0'], ['l0', 's0', ('s0', -1), 's1', ('l0', -1), 'l1']).
+transition(t2, ['l0', 's1'], ['l0', 's1', ('s1', -1), 's2']).
+transition(t3, ['l1', 's1'], ['l1', 's1', ('l1', -1), 'l17']).
+transition(t4, ['l2', 's1'], ['l2', 's1', ('l2', -1), 'l3']).
[…]

### 2013-11-25 benchmarks/ic3-soter/pipe__single_message_in_mailbox__depth_0__fixed.spec.pl

@@ -1,0 +1,451 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
+place('l101').
+place('l102').
+place('l103').
[…]

### 2013-11-25 benchmarks/ic3-soter/reslock__critical__depth_0__fixed.spec.pl

@@ -1,0 +1,762 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
+place('l101').
[…]

### 2013-11-25 benchmarks/ic3-soter/reslock__critical__depth_1__fixed.spec.pl

@@ -1,0 +1,1305 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
[…]

### 2013-11-25 benchmarks/ic3-soter/reslock__critical__depth_2__fixed.spec.pl

@@ -1,0 +1,1874 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('s39').
+place('s40').
+place('s41').
+place('s42').
+place('s43').
+place('s44').
+place('s45').
+place('s46').
+place('s47').
+place('s48').
+place('s49').
+place('s50').
+place('s51').
+place('s52').
+place('s53').
+place('s54').
+place('s55').
+place('s56').
+place('s57').
+place('s58').
+place('s59').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
[…]

### 2013-11-25 benchmarks/ic3-soter/reslockbeh__critical__depth_0__fixed.spec.pl

@@ -1,0 +1,1714 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
+place('l101').
+place('l102').
+place('l103').
+place('l104').
+place('l105').
+place('l106').
+place('l107').
+place('l108').
+place('l109').
+place('l110').
+place('l111').
+place('l112').
+place('l113').
+place('l114').
+place('l115').
+place('l116').
[…]

### 2013-11-25 benchmarks/ic3-soter/reslockbeh__critical__depth_1__fixed.spec.pl

@@ -1,0 +1,4651 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
[…]

### 2013-11-25 benchmarks/ic3-soter/ring__single_message_in_mailbox__depth_0__fixed.spec.pl

@@ -1,0 +1,843 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
+place('l101').
+place('l102').
+place('l103').
[…]

### 2013-11-25 benchmarks/ic3-soter/safe_send__sending_to_non-pid_1__depth_2__fixed.spec.pl

@@ -1,0 +1,567 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('s39').
+place('s40').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
[…]

### 2013-11-25 benchmarks/ic3-soter/safe_send__sending_to_non-pid_2__depth_2__fixed.spec.pl

@@ -1,0 +1,567 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('s39').
+place('s40').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
[…]

### 2013-11-25 benchmarks/ic3-soter/safe_send__sending_to_non-pid_3__depth_1__fixed.spec.pl

@@ -1,0 +1,567 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('s39').
+place('s40').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
[…]

### 2013-11-25 benchmarks/ic3-soter/safe_send__sending_to_non-pid_4__depth_2__fixed.spec.pl

@@ -1,0 +1,567 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('s23').
+place('s24').
+place('s25').
+place('s26').
+place('s27').
+place('s28').
+place('s29').
+place('s30').
+place('s31').
+place('s32').
+place('s33').
+place('s34').
+place('s35').
+place('s36').
+place('s37').
+place('s38').
+place('s39').
+place('s40').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
[…]

### 2013-11-25 benchmarks/ic3-soter/safe_send__sending_to_non-pid__depth_0__fixed.spec.pl

@@ -1,0 +1,244 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
[…]

### 2013-11-25 benchmarks/ic3-soter/sieve__single_message_in_counter_mailbox__depth_0__fixed.spec.pl

@@ -1,0 +1,541 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
[…]

### 2013-11-25 benchmarks/ic3-soter/sieve__single_message_in_filter_mailbox__depth_0__fixed.spec.pl

@@ -1,0 +1,542 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
[…]

### 2013-11-25 benchmarks/ic3-soter/sieve__single_message_in_sieve_mailbox__depth_0__fixed.spec.pl

@@ -1,0 +1,540 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('s13').
+place('s14').
+place('s15').
+place('s16').
+place('s17').
+place('s18').
+place('s19').
+place('s20').
+place('s21').
+place('s22').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
[…]

### 2013-11-25 benchmarks/ic3-soter/state_factory__after_receive_if_no_mail__depth_0__fixed.spec.pl

@@ -1,0 +1,1193 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
+place('l101').
+place('l102').
+place('l103').
+place('l104').
+place('l105').
+place('l106').
+place('l107').
+place('l108').
+place('l109').
+place('l110').
+place('l111').
+place('l112').
+place('l113').
+place('l114').
+place('l115').
+place('l116').
[…]

### 2013-11-25 benchmarks/ic3-soter/state_factory__single_message_in_mailbox__depth_0__fixed.spec.pl

@@ -1,0 +1,1191 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('s10').
+place('s11').
+place('s12').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
+place('l97').
+place('l98').
+place('l99').
+place('l100').
+place('l101').
+place('l102').
+place('l103').
+place('l104').
+place('l105').
+place('l106').
+place('l107').
+place('l108').
+place('l109').
+place('l110').
+place('l111').
+place('l112').
+place('l113').
+place('l114').
+place('l115').
+place('l116').
[…]

### 2013-11-25 benchmarks/ic3-soter/stutter__we_abhorr_as__depth_0__fixed.spec.pl

@@ -1,0 +1,176 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+transition(t1, ['l0', 's0'], ['l0', 's0', ('s0', -1), 's1', ('l0', -1), 'l1']).
+transition(t2, ['l0', 's1'], ['l0', 's1', ('s1', -1), 's2']).
+transition(t3, ['l1', 's1'], ['l1', 's1', ('l1', -1), 'l29']).
+transition(t4, ['l2', 's1'], ['l2', 's1', ('l2', -1), 'l3']).
+transition(t5, ['l3', 's1'], ['l3', 's1', ('l3', -1), 'l40']).
[…]

### 2013-11-25 benchmarks/ic3-soter/stutter__we_abhorr_as__depth_1__fixed.spec.pl

@@ -1,0 +1,214 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('s6').
+place('s7').
+place('s8').
+place('s9').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+place('l29').
+place('l30').
+place('l31').
+place('l32').
+place('l33').
+place('l34').
+place('l35').
+place('l36').
+place('l37').
+place('l38').
+place('l39').
+place('l40').
+place('l41').
+place('l42').
+place('l43').
+place('l44').
+place('l45').
+place('l46').
+place('l47').
+place('l48').
+place('l49').
+place('l50').
+place('l51').
+place('l52').
+place('l53').
+place('l54').
+place('l55').
+place('l56').
+place('l57').
+place('l58').
+place('l59').
+place('l60').
+place('l61').
+place('l62').
+place('l63').
+place('l64').
+place('l65').
+place('l66').
+place('l67').
+place('l68').
+place('l69').
+place('l70').
+place('l71').
+place('l72').
+place('l73').
+place('l74').
+place('l75').
+place('l76').
+place('l77').
+place('l78').
+place('l79').
+place('l80').
+place('l81').
+place('l82').
+place('l83').
+place('l84').
+place('l85').
+place('l86').
+place('l87').
+place('l88').
+place('l89').
+place('l90').
+place('l91').
+place('l92').
+place('l93').
+place('l94').
+place('l95').
+place('l96').
[…]

### 2013-11-25 benchmarks/ic3-soter/unsafe_send__sending_to_non-pid__depth_1__fixed.spec.pl

@@ -1,0 +1,68 @@
+place('s0').
+place('s1').
+place('s2').
+place('s3').
+place('s4').
+place('s5').
+place('l0').
+place('l1').
+place('l2').
+place('l3').
+place('l4').
+place('l5').
+place('l6').
+place('l7').
+place('l8').
+place('l9').
+place('l10').
+place('l11').
+place('l12').
+place('l13').
+place('l14').
+place('l15').
+place('l16').
+place('l17').
+place('l18').
+place('l19').
+place('l20').
+place('l21').
+place('l22').
+place('l23').
+place('l24').
+place('l25').
+place('l26').
+place('l27').
+place('l28').
+transition(t1, ['l0', 's0'], ['l0', 's0', ('s0', -1), 's1', ('l0', -1), 'l1']).
+transition(t2, ['l0', 's1'], ['l0', 's1', ('s1', -1), 's2']).
+transition(t3, ['l1', 's1'], ['l1', 's1', ('l1', -1), 'l8']).
+transition(t4, ['l2', 's1'], ['l2', 's1', ('l2', -1), 'l3']).
+transition(t5, ['l3', 's1'], ['l3', 's1', ('l3', -1), 'l9']).
+transition(t6, ['l4', 's1'], ['l4', 's1', ('l4', -1), 'l5']).
+transition(t7, ['l5', 's1'], ['l5', 's1', ('s1', -1), 's3', ('l5', -1), 'l15']).
+transition(t8, ['l6', 's1'], ['l6', 's1', ('l6', -1), 'l7']).
+transition(t9, ['l7', 's1'], ['l7', 's1', ('l7', -1), 'l18']).
+transition(t10, ['l8', 's1'], ['l8', 's1', ('l8', -1), 'l6']).
+transition(t11, ['l9', 's1'], ['l9', 's1', ('s1', -1), 's4', ('l9', -1), 'l11']).
+transition(t12, ['l12', 's1'], ['l12', 's1', ('l12', -1), 'l2']).
+transition(t13, ['l13', 's1'], ['l13', 's1', ('l13', -1), 'l4']).
+transition(t14, ['l14', 's1'], ['l14', 's1', ('l14', -1), 'l12']).
+transition(t15, ['l15', 's1'], ['l15', 's1', ('l15', -1), 'l21']).
+transition(t16, ['l16', 's1'], ['l16', 's1', ('l16', -1), 'l17']).
[…]

### 2013-11-25 run-benchmarks.sh
Added results of invariant benchmark on ic3-soter

@@ -1,6 +1,7 @@
 #!/bin/bash
 
-for benchmark_dir in `find benchmarks -mindepth 1 -maxdepth 1 -type d`; do
+#for benchmark_dir in `find benchmarks -mindepth 1 -maxdepth 1 -type d`; do
+for benchmark_dir in "benchmarks/ic3-soter"; do
   >$benchmark_dir/positive-pnerf.list
   >$benchmark_dir/dontknow-pnerf.list
   >$benchmark_dir/negative-pnerf.list

### 2013-11-25 paper/workspace/main.tex
Experimental setup and report for IIC.

@@
 \newcommand{\mist}{{\sc Mist}}
+\newcommand{\iic}{{\sc IIC}}
 \newcommand{\pnerf}{{\sc Pnerf}}
@@
     Safety (over real numbers or naturals) & 23 &  0 & 54 &  0 \\
-    Invariant                              & 23 &  0 & 54 &  0 \\
-    Invariant Minimization                 & 23 &  0 & 54 &  0 \\
+    Invariant \& Invariant Minimization    & 23 &  0 & 54 &  0 \\
     Safety by Refinement                   & 23 &  0 & 54 &  0 \\
@@
     Safety (over real numbers or naturals) & 14 &  0 & 13 &  0 \\
-    Invariant                              & 14 &  0 & 13 &  0 \\
-    Invariant Minimization                 & 14 &  0 & 13 &  0 \\
+    Invariant \& Invariant Minimization    & 14 &  0 & 13 &  0 \\
     Safety by Refinement                   & 20 &  0 &  7 &  0 \\
@@
 
+\section{Benchmarks of \iic}
+
+We experimented on the benchmarks available at \\
+\ttt{https://svn.rse.mpi-sws.org/RSE/IC3forWSTS/code/trunk/examples/soter}.
+The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
+Table~\ref{iic-experiments} shows the corresponding experimental results.
+
+\begin{table}[h]
+\begin{center}
+  \begin{tabular}{ | p{6cm} | r | r | r | r | }
+    \hline
+    Method & pos & neg & don't know & TO \\
+    \hline
+    Safety (over real numbers or naturals) & 2 & 0 & 36 & 0 \\
+    Invariant \& Invariant Minimization    & 2 & 0 & 26 & 0 \\
+    Safety by Refinement                   & 2 & 0 & 36 & 0 \\
+    \hline
+  \end{tabular}
+\end{center}
+\caption{Experimental results for the benchmarks of \iic}
+\label{iic-experiments}
+\end{table}
+
 \end{document}

### 2013-11-25 results-to-table.sh

@@ -16,8 +16,8 @@ results_other_tool=( positive negative error timeout )
 our_tool=pnerf
 other_tool=bfc
 
-benchmark_dirs=( 'found-in-mist-repo' 'given-by-daniel-kroening' )
-benchmark_tools=( 'mist' 'bfc' )
+benchmark_dirs=( 'found-in-mist-repo' 'given-by-daniel-kroening' 'ic3-soter' )
+benchmark_tools=( 'mist' 'bfc' 'iic' )
 for (( benchmark=0;benchmark<${#benchmark_dirs[@]};benchmark++)); do
   benchmark_dir='benchmarks/'${benchmark_dirs[$benchmark]}
   other_tool=${benchmark_tools[$benchmark]}

### 2013-11-25 run-benchmarks.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 
 #for benchmark_dir in `find benchmarks -mindepth 1 -maxdepth 1 -type d`; do
-for benchmark_dir in "benchmarks/ic3-soter"; do
+for benchmark_dir in benchmarks/ic3-soter; do
   >$benchmark_dir/positive-pnerf.list
   >$benchmark_dir/dontknow-pnerf.list
   >$benchmark_dir/negative-pnerf.list
@@ -10,13 +10,13 @@ for benchmark_dir in "benchmarks/ic3-soter"; do
   >$benchmark_dir/timing-pnerf.log
       #timeout 600 ./src/explore-states.sh -d 0 $pl_file | tee $pl_file.out
   for pl_file in `find $benchmark_dir -name "*.pl"`; do
-    T="$(date +%s%N)"
+    T="$(gdate +%s%N)"
     (
       set -o pipefail;
-      timeout 600 ./src/main $pl_file | tee $pl_file.out
+      timeout 600 ./src/main -refinement $pl_file | tee $pl_file.out
     )
     result=$?
-    T=$(($(date +%s%N)-T))
+    T=$(($(gdate +%s%N)-T))
     if [[ result -eq 0 ]]; then
         list='positive'
     elif [[ result -eq 1 ]]; then

### 2013-11-25 paper/workspace/main.tex
Updated report
Draft

@@
 
+We constructed the following two invariants by applying the method Invariant Minimization.
+
+\begin{itemize}
+\item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
+\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
+\end{itemize}
+
+\iic\ constructs the following two invariants.
+
+\begin{itemize}
+\item \verb=firewall__no_pred_called_with_zero__depth_1__fixed.spec=: \verb=(l739 < 1)=
+\item \verb=firewall__no_pred_called_with_zero__depth_2__fixed.spec=: \verb=(l3849 < 1)=
+\end{itemize}
+
 \end{document}

### 2014-01-22 paper/workspace/main.tex
Draft
Fixed typos in draft of math.

@@
   Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N, P, and A.
+  Subprocedure Inv constructs invariant I corresponding to N and A.
 \end{verbatim}
@@
 \newpage
-\section{Method Invariant Minimization}
+\section{Method Invariant with Minimization}
 
-The method Invariant Minimization constructs an invariant \verb=I= that uses a minimal number places for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-We present the method Invariant Minimization by example on a simplified mutual exclusion algorithm by Delzanno2001.
+The method Invariant with Minimization constructs an invariant \verb=I=
+that uses a minimal number of places for given Petri net \verb=N= and
+\verb=P= when \verb=N= never violates \verb=P=. 
+We present the method Invariant with Minimization by example on a
+simplified mutual exclusion algorithm by Delzanno2001.
 
@@
 \begin{verbatim}
-* Method Invariant
+* Method Invariant with Minimization
 
@@
 \begin{verbatim}
-* Method Safety:
+* Method Safety by Refinement:
 
@@
     Safety (over real numbers or naturals)      & 23 &  0 & 54 &  0 \\
-    Invariant \& Invariant Minimization    & 23 &  0 & 54 &  0 \\
+    Invariant \& Invariant with Minimization    & 23 &  0 & 54 &  0 \\
     Safety by Refinement                        & 23 &  0 & 54 &  0 \\
@@
     Safety (over real numbers or naturals)      & 14 &  0 & 13 &  0 \\
-    Invariant \& Invariant Minimization    & 14 &  0 & 13 &  0 \\
+    Invariant \& Invariant with Minimization    & 14 &  0 & 13 &  0 \\
     Safety by Refinement                        & 20 &  0 &  7 &  0 \\
@@
     Safety (over real numbers or naturals)      & 2 & 0 & 36 & 0 \\
-    Invariant \& Invariant Minimization    & 2 & 0 & 26 & 0 \\
+    Invariant \& Invariant with Minimization    & 2 & 0 & 26 & 0 \\
     Safety by Refinement                        & 2 & 0 & 36 & 0 \\
@@
 
-We constructed the following two invariants by applying the method Invariant Minimization.
+We constructed the following two invariants by applying the method
+Invariant with Minimization.
 
 \begin{itemize}
+\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
 \item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
-\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
 \end{itemize}

### 2014-01-28 doc/benchmark-tables.tex
Added scripts and output options to generate tables for benchmarks

\documentclass{article}

\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage[a4paper, landscape, top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}

\usepackage{tikz}
\usetikzlibrary{shapes, shapes.multipart, shapes.geometric}
\usetikzlibrary{positioning}

\newcommand{\bfc}{{\sc Bfc}}
\newcommand{\mist}{{\sc Mist}}
\newcommand{\iic}{{\sc Iic}}
\newcommand{\pnerf}{{\sc Pnerf}}
\newcommand{\zthree}{{\sc Z3}}
\newcommand{\ttt}[1]{\texttt{#1}}

\begin{document}

\title{Experimental Results \\ \mbox{ } \\ of a Safety Checker for Petri nets}
\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
% \institute{Technische Universit\"at M\"unchen}
\date{\today}

\section{Legend for benchmarks}

The symbols used in the headers of the tables are:

\begin{description}
  \item[P] Number of places in the petri net.
  \item[T] Number of transitions in the petri net.
  \item[R] Result of our tool. It is + if our tool can prove safety or find an invariant and - otherwise.
  \item[I] Number of iterations. This is the number of traps used for refinement or the number of times an invariant is minimized.
  \item[T] Running time of our tool in seconds.
\end{description}

The methods used by our tool are:

\begin{description}
  \item[Safety] Method Safety over reals.
  \item[Safety-Int] Method Safety over integers.
  \item[Refinement] Method Safety by Refinement over reals.
  \item[Refinement-Int] Method Safety by Refinement over integers.
  \item[Invariant] Method Invariant.
  \item[Inv-Min] Method Invariant Minimization.
  \item[Inv-Ref] Method Invariant by Refinement.
\end{description}

\section{Benchmarks of \bfc}

The set of benchmarks consists of 77 benchmarks, 26 positive and 51 negative.
Table~\ref{bfc-experiments}, \ref{bfc-experiments-2} and \ref{bfc-experiments-3}
show the experimental results.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
    \hline
           \multicolumn{3}{|c|}{Network}
         & \multicolumn{2}{|c|}{Safety}
         & \multicolumn{2}{|c|}{Safety-Int}
         & \multicolumn{3}{|c|}{Refinement}
         & \multicolumn{3}{|c|}{Refinement-Int}
         & \multicolumn{2}{|c|}{Invariant}
         & \multicolumn{3}{|c|}{Inv-Min}
         & \multicolumn{2}{|c|}{Inv-Ref} \\
    Name & P & T
        & R & T
        & R & T
        & R & I & T
        & R & I & T
        & R & T
        & R & I & T
        & R & T \\
    \hline
        \multicolumn{20}{|c|}{Positive Benchmarks} \\
    \hline
cprover-PN/abp & 42 & 43 & + & 0.19 & + & 0.18 & + & 0 & 0.13 & + & 0 & 0.17 & + & 0.26 & + & 2 & 0.56 & + & 0.42 \\
cprover-PN/basicME & 18 & 17 & + & 0.13 & + & 0.15 & + & 0 & 0.13 & + & 0 & 0.17 & + & 0.25 & + & 0 & 0.30 & + & 0.39 \\
cprover-PN/bingham\_h150 & 309 & 457 & + & 0.35 & + & 0.35 & + & 0 & 0.40 & + & 0 & 0.35 & + & 0.57 & + & 0 & 0.81 & + & 0.93 \\
cprover-PN/bingham\_h25 & 59 & 82 & + & 0.17 & + & 0.19 & + & 0 & 0.17 & + & 0 & 0.18 & + & 0.30 & + & 0 & 0.36 & + & 0.45 \\
cprover-PN/bingham\_h250 & 509 & 757 & + & 0.61 & + & 0.55 & + & 0 & 0.55 & + & 0 & 0.57 & + & 0.85 & + & 0 & 1.22 & + & 1.43 \\
cprover-PN/bingham\_h50 & 109 & 157 & + & 0.22 & + & 0.22 & + & 0 & 0.19 & + & 0 & 0.19 & + & 0.34 & + & 0 & 0.46 & + & 0.55 \\
cprover-PN/bingham\_h500 & 1009 & 1507 & + & 1.20 & + & 1.09 & + & 0 & 1.20 & + & 0 & 1.11 & + & 1.74 & + & 0 & 2.89 & + & 3.00 \\
cprover-PN/clientserver & 13 & 11 & + & 0.13 & + & 0.12 & + & 0 & 0.15 & + & 0 & 0.14 & + & 0.27 & + & 0 & 0.32 & + & 0.38 \\
cprover-PN/csm & 28 & 27 & + & 0.14 & + & 0.15 & + & 0 & 0.13 & + & 0 & 0.13 & + & 0.26 & + & 0 & 0.32 & + & 0.35 \\
cprover-PN/gsm & 73 & 156 & + & 0.26 & + & 0.29 & + & 0 & 0.21 & + & 0 & 0.20 & + & 0.64 & + & 0 & 0.68 & + & 0.78 \\
cprover-PN/handover & 126 & 255 & + & 0.23 & + & 0.23 & + & 0 & 0.25 & + & 0 & 0.26 & + & 0.37 & + & 0 & 0.50 & + & 0.60 \\
cprover-PN/keycard & 59 & 78 & + & 0.16 & + & 0.15 & + & 0 & 0.21 & + & 0 & 0.15 & + & 0.33 & + & 0 & 0.36 & + & 0.49 \\
cprover-PN/lamport & 30 & 27 & - & 0.19 & - & 0.18 & - & 0 & 0.20 & - & 0 & 0.21 & - & 0.14 & - & 0 & 0.17 & - & 0.23 \\
cprover-PN/mesh2x2 & 66 & 69 & + & 0.20 & + & 0.20 & + & 0 & 0.15 & + & 0 & 0.20 & + & 0.27 & + & 0 & 0.37 & + & 0.43 \\
cprover-PN/multipoll & 35 & 41 & + & 0.15 & + & 0.15 & + & 0 & 0.16 & + & 0 & 0.14 & + & 0.26 & + & 0 & 0.32 & + & 0.43 \\
cprover-PN/newdekker & 51 & 48 & - & 0.18 & - & 0.18 & - & 0 & 0.22 & - & 0 & 0.25 & - & 0.15 & - & 0 & 0.15 & - & 0.24 \\
cprover-PN/newrtp & 12 & 14 & + & 0.15 & + & 0.12 & + & 0 & 0.15 & + & 0 & 0.14 & + & 0.27 & + & 0 & 0.30 & + & 0.40 \\
cprover-PN/peterson & 39 & 36 & - & 0.15 & - & 0.16 & - & 0 & 0.27 & - & 0 & 0.21 & - & 0.15 & - & 0 & 0.14 & - & 0.22 \\
cprover-PN/phones & 51 & 87 & + & 0.16 & + & 0.18 & + & 0 & 0.16 & + & 0 & 0.16 & + & 0.30 & + & 0 & 0.37 & + & 0.43 \\
cprover-PN/reaction & 19 & 15 & + & 0.14 & + & 0.14 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.34 & + & 0.37 \\
cprover-PN/reaction3 & 22 & 18 & + & 0.16 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.15 & + & 0.25 & + & 0 & 0.32 & + & 0.37 \\
cprover-PN/reaction4 & 16 & 13 & + & 0.14 & + & 0.15 & + & 0 & 0.14 & + & 0 & 0.13 & + & 0.28 & + & 0 & 0.33 & + & 0.36 \\
cprover-PN/read-write & 36 & 31 & + & 0.16 & + & 0.20 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.28 & + & 0 & 0.34 & + & 0.43 \\
cprover-PN/restriction & 15 & 10 & + & 0.14 & + & 0.14 & + & 0 & 0.13 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.32 & + & 0.40 \\
cprover\_software\_analysis/conditionals\_vs\_satabs.2 & 214 & 281 & + & 0.27 & + & 0.25 & + & 0 & 0.27 & + & 0 & 0.25 & + & 0.44 & + & 0 & 0.55 & + & 0.65 \\
cprover\_software\_analysis/rand\_cas\_vs\_satabs.2 & 110 & 139 & + & 0.19 & + & 0.21 & + & 0 & 0.20 & + & 0 & 0.23 & + & 0.36 & + & 0 & 0.47 & + & 0.50 \\
    \hline
        \multicolumn{20}{|c|}{Negative Benchmarks} \\
    \hline
cprover-PN/kanban & 26 & 29 & - & 0.18 & - & 0.17 & - & 0 & 0.23 & - & 0 & 0.20 & - & 0.14 & - & 0 & 0.14 & - & 0.21 \\
cprover-PN/leabasicapproach & 30 & 27 & - & 0.16 & - & 0.16 & - & 0 & 0.24 & - & 0 & 0.23 & - & 0.16 & - & 0 & 0.15 & - & 0.22 \\
cprover-PN/platoon & 20 & 17 & - & 0.17 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.23 & - & 0.14 & - & 0 & 0.16 & - & 0.24 \\
cprover-PN/platoonFiniteHandlerPoly & 19 & 16 & - & 0.15 & - & 0.16 & - & 0 & 0.24 & - & 0 & 0.22 & - & 0.13 & - & 0 & 0.13 & - & 0.20 \\
cprover-PN/pncsacover & 64 & 68 & - & 0.19 & - & 0.18 & - & 0 & 0.28 & - & 0 & 0.25 & - & 0.16 & - & 0 & 0.17 & - & 0.28 \\
cprover-PN/reaction2 & 13 & 11 & - & 0.15 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.19 & - & 0.14 & - & 0 & 0.12 & - & 0.25 \\
cprover-PN/reaction5 & 11 & 8 & - & 0.14 & - & 0.14 & - & 1 & 0.39 & - & 1 & 0.44 & - & 0.12 & - & 0 & 0.14 & - & 0.42 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments}
\end{table}

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
    \hline
           \multicolumn{3}{|c|}{Network}
         & \multicolumn{2}{|c|}{Safety}
         & \multicolumn{2}{|c|}{Safety-Int}
         & \multicolumn{3}{|c|}{Refinement}
         & \multicolumn{3}{|c|}{Refinement-Int}
         & \multicolumn{2}{|c|}{Invariant}
         & \multicolumn{3}{|c|}{Inv-Min}
         & \multicolumn{2}{|c|}{Inv-Ref} \\
    Name & P & T
        & R & T
        & R & T
        & R & I & T
        & R & I & T
        & R & T
        & R & I & T
        & R & T \\
    \hline
        \multicolumn{20}{|c|}{Negative Benchmarks (continued)} \\
    \hline
cprover\_software\_analysis/Boop\_simple\_vf\_satabs.1 & 31 & 31 & - & 0.16 & - & 0.15 & - & 0 & 0.23 & - & 0 & 0.22 & - & 0.14 & - & 0 & 0.14 & - & 0.21 \\
cprover\_software\_analysis/Boop\_simple\_vf\_satabs.2 & 330 & 7489 & - & 3.39 & - & 3.27 & - & 0 & 5.59 & - & 0 & 5.62 & - & 2.93 & - & 0 & 3.05 & - & 5.36 \\
cprover\_software\_analysis/Function\_Pointer3\_vs\_satabs.1 & 40 & 71 & - & 0.17 & - & 0.16 & - & 1 & 0.52 & - & 1 & 0.49 & - & 0.15 & - & 0 & 0.17 & - & 0.49 \\
cprover\_software\_analysis/Function\_Pointer3\_vs\_satabs.2 & 306 & 841 & - & 0.53 & - & 0.56 & - & 0 & 0.82 & - & 1 & 1.55 & - & 0.45 & - & 0 & 0.49 & - & 0.85 \\
cprover\_software\_analysis/Function\_Pointer3\_vs\_satabs.3 & 2826 & 8961 & - & 8.12 & - & 8.86 & - & 0 & 10.21 & - & 0 & 11.29 & - & 7.49 & - & 0 & 7.56 & - & 10.19 \\
cprover\_software\_analysis/buggy\_spaghetti\_vf\_satabs.1 & 78 & 115 & - & 0.18 & - & 0.18 & - & 0 & 0.31 & - & 0 & 0.27 & - & 0.20 & - & 0 & 0.17 & - & 0.29 \\
cprover\_software\_analysis/buggy\_spaghetti\_vf\_satabs.2 & 326 & 697 & - & 0.50 & - & 0.47 & - & 0 & 0.77 & - & 0 & 0.77 & - & 0.43 & - & 0 & 0.47 & - & 0.72 \\
cprover\_software\_analysis/conditionals\_vs\_satabs.1 & 50 & 55 & - & 0.18 & - & 0.17 & - & 0 & 0.24 & - & 0 & 0.25 & - & 0.15 & - & 0 & 0.15 & - & 0.25 \\
cprover\_software\_analysis/constants\_vf\_satabs.1 & 26 & 25 & - & 0.15 & - & 0.15 & - & 0 & 0.25 & - & 0 & 0.22 & - & 0.15 & - & 0 & 0.13 & - & 0.22 \\
cprover\_software\_analysis/constants\_vf\_satabs.2 & 262 & 393 & - & 0.33 & - & 0.35 & - & 0 & 0.54 & - & 0 & 0.51 & - & 0.30 & - & 0 & 0.29 & - & 0.54 \\
cprover\_software\_analysis/dekker\_vs\_satabs.1 & 41 & 121 & - & 0.21 & - & 0.20 & - & 0 & 0.33 & - & 0 & 0.28 & - & 0.17 & - & 0 & 0.22 & - & 0.28 \\
cprover\_software\_analysis/dekker\_vs\_satabs.2 & 164 & 1345 & - & 0.67 & - & 0.72 & - & 0 & 1.08 & - & 0 & 1.08 & - & 0.62 & - & 0 & 0.61 & - & 1.08 \\
cprover\_software\_analysis/double\_lock\_p1\_vs\_satabs.1 & 64 & 205 & - & 0.22 & - & 0.24 & - & 1 & 0.79 & - & 2 & 1.59 & - & 0.36 & - & 0 & 0.36 & - & 0.79 \\
cprover\_software\_analysis/double\_lock\_p1\_vs\_satabs.2 & 570 & 7601 & - & 4.28 & - & 12.37 & - & 0 & 6.52 & - & 0 & 13.19 & - & 3.29 & - & 0 & 3.19 & - & 6.02 \\
cprover\_software\_analysis/double\_lock\_p1\_vs\_satabs.3 & 570 & 7569 & - & 3.90 & - & 11.20 & - & 0 & 5.76 & - & 0 & 13.13 & - & 3.12 & - & 0 & 3.23 & - & 5.85 \\
cprover\_software\_analysis/double\_lock\_p2\_vs\_satabs.1 & 64 & 213 & - & 0.25 & - & 0.23 & - & 0 & 0.38 & - & 1 & 0.73 & - & 0.20 & - & 0 & 0.22 & - & 0.33 \\
cprover\_software\_analysis/double\_lock\_p2\_vs\_satabs.2 & 184 & 1833 & - & 0.93 & - & 0.93 & - & 0 & 1.44 & - & 0 & 1.44 & - & 0.81 & - & 0 & 0.81 & - & 1.49 \\
cprover\_software\_analysis/double\_lock\_p3\_vs\_satabs.1 & 46 & 81 & - & 0.18 & - & 0.18 & - & 0 & 0.25 & - & 0 & 0.27 & - & 0.15 & - & 0 & 0.16 & - & 0.28 \\
cprover\_software\_analysis/double\_lock\_p3\_vs\_satabs.2 & 112 & 745 & - & 0.49 & - & 0.47 & - & 1 & 1.34 & - & 0 & 0.71 & - & 0.41 & - & 0 & 0.44 & - & 1.33 \\
cprover\_software\_analysis/double\_lock\_p3\_vs\_satabs.3 & 306 & 3137 & - & 1.56 & - & 1.54 & - & 0 & 2.35 & - & 0 & 2.34 & - & 1.32 & - & 0 & 1.32 & - & 2.34 \\
cprover\_software\_analysis/lu-fig2\_fixed\_vs\_satabs.1 & 39 & 37 & - & 0.19 & - & 0.20 & - & 0 & 0.25 & - & 0 & 0.23 & - & 0.17 & - & 0 & 0.13 & - & 0.23 \\
cprover\_software\_analysis/lu-fig2\_fixed\_vs\_satabs.2 & 61 & 185 & - & 0.23 & - & 0.21 & - & 0 & 0.33 & - & 0 & 0.35 & - & 0.20 & - & 0 & 0.23 & - & 0.32 \\
cprover\_software\_analysis/lu-fig2\_fixed\_vs\_satabs.3 & 178 & 1617 & - & 0.82 & - & 0.80 & - & 0 & 1.31 & - & 0 & 1.25 & - & 0.72 & - & 0 & 0.72 & - & 1.25 \\
cprover\_software\_analysis/peterson\_vs\_satabs.1 & 31 & 65 & - & 0.16 & - & 0.17 & - & 0 & 0.25 & - & 0 & 0.25 & - & 0.15 & - & 0 & 0.17 & - & 0.29 \\
cprover\_software\_analysis/peterson\_vs\_satabs.2 & 284 & 1985 & - & 0.95 & - & 0.98 & - & 0 & 1.52 & - & 0 & 1.54 & - & 0.91 & - & 0 & 0.93 & - & 1.54 \\
cprover\_software\_analysis/pthread5\_vs\_satabs.1 & 80 & 413 & - & 0.35 & - & 0.30 & - & 0 & 0.45 & - & 0 & 0.49 & - & 0.27 & - & 0 & 0.28 & - & 0.47 \\
cprover\_software\_analysis/pthread5\_vs\_satabs.2 & 80 & 412 & - & 0.34 & - & 0.35 & - & 0 & 0.47 & - & 0 & 0.47 & - & 0.27 & - & 0 & 0.33 & - & 0.46 \\
cprover\_software\_analysis/pthread5\_vs\_satabs.3 & 187 & 1893 & - & 1.00 & - & 0.96 & - & 0 & 1.48 & - & 0 & 1.49 & - & 0.84 & - & 0 & 0.81 & - & 1.45 \\
cprover\_software\_analysis/pthread5\_vs\_satabs.4 & 187 & 1889 & - & 1.00 & - & 1.01 & - & 0 & 1.49 & - & 0 & 1.65 & - & 0.82 & - & 0 & 0.82 & - & 1.69 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments-2}
\end{table}

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
    \hline
           \multicolumn{3}{|c|}{Network}
         & \multicolumn{2}{|c|}{Safety}
         & \multicolumn{2}{|c|}{Safety-Int}
         & \multicolumn{3}{|c|}{Refinement}
         & \multicolumn{3}{|c|}{Refinement-Int}
         & \multicolumn{2}{|c|}{Invariant}
         & \multicolumn{3}{|c|}{Inv-Min}
         & \multicolumn{2}{|c|}{Inv-Ref} \\
    Name & P & T
        & R & T
        & R & T
        & R & I & T
        & R & I & T
        & R & T
        & R & I & T
        & R & T \\
    \hline
        \multicolumn{20}{|c|}{Negative Benchmarks (continued)} \\
    \hline
cprover\_software\_analysis/rand\_cas\_vs\_satabs.1 & 48 & 53 & - & 0.16 & - & 0.20 & - & 0 & 0.25 & - & 0 & 0.24 & - & 0.15 & - & 0 & 0.16 & - & 0.25 \\
cprover\_software\_analysis/rand\_lock\_p0\_vs\_satabs.1 & 29 & 43 & - & 0.16 & - & 0.16 & - & 0 & 0.25 & - & 0 & 0.23 & - & 0.15 & - & 0 & 0.17 & - & 0.24 \\
cprover\_software\_analysis/rand\_lock\_p0\_vs\_satabs.2 & 110 & 289 & - & 0.26 & - & 0.31 & - & 0 & 0.43 & - & 0 & 0.39 & - & 0.23 & - & 0 & 0.25 & - & 0.47 \\
cprover\_software\_analysis/rand\_lock\_p0\_vs\_satabs.3 & 234 & 773 & - & 0.50 & - & 0.52 & - & 0 & 0.81 & - & 0 & 0.78 & - & 0.43 & - & 0 & 0.43 & - & 0.74 \\
cprover\_software\_analysis/simple\_loop5\_vs\_satabs.1 & 31 & 49 & - & 0.16 & - & 0.20 & - & 0 & 0.22 & - & 0 & 0.24 & - & 0.19 & - & 0 & 0.16 & - & 0.24 \\
cprover\_software\_analysis/simple\_loop5\_vs\_satabs.2 & 93 & 485 & - & 0.37 & - & 0.39 & - & 0 & 0.52 & - & 0 & 0.55 & - & 0.31 & - & 0 & 0.30 & - & 0.54 \\
cprover\_software\_analysis/spin2003\_vs\_satabs.1 & 27 & 35 & - & 0.19 & - & 0.17 & - & 0 & 0.26 & - & 0 & 0.24 & - & 0.15 & - & 0 & 0.18 & - & 0.24 \\
cprover\_software\_analysis/spin2003\_vs\_satabs.2 & 56 & 181 & - & 0.21 & - & 0.22 & - & 0 & 0.33 & - & 1 & 0.67 & - & 0.19 & - & 0 & 0.24 & - & 0.33 \\
cprover\_software\_analysis/stack\_cas\_p0\_vs\_satabs.1 & 41 & 71 & - & 0.17 & - & 0.20 & - & 0 & 0.26 & - & 0 & 0.30 & - & 0.15 & - & 0 & 0.16 & - & 0.26 \\
cprover\_software\_analysis/stack\_cas\_p0\_vs\_satabs.2 & 78 & 157 & - & 0.21 & - & 0.23 & - & 0 & 0.32 & - & 0 & 0.30 & - & 0.20 & - & 0 & 0.18 & - & 0.32 \\
cprover\_software\_analysis/stack\_cas\_p0\_vs\_satabs.3 & 158 & 345 & - & 0.31 & - & 0.33 & - & 0 & 0.47 & - & 0 & 0.51 & - & 0.31 & - & 0 & 0.25 & - & 0.44 \\
cprover\_software\_analysis/stack\_lock\_p0\_vs\_satabs.1 & 37 & 59 & - & 0.16 & - & 0.16 & - & 0 & 0.26 & - & 0 & 0.23 & - & 0.16 & - & 0 & 0.16 & - & 0.27 \\
cprover\_software\_analysis/stack\_lock\_p0\_vs\_satabs.2 & 106 & 633 & - & 0.42 & - & 0.39 & - & 0 & 0.62 & - & 0 & 0.64 & - & 0.35 & - & 0 & 0.35 & - & 0.65 \\
cprover\_software\_analysis/szymanski\_vs\_satabs.1 & 61 & 225 & - & 0.25 & - & 0.24 & - & 0 & 0.35 & - & 0 & 0.35 & - & 0.21 & - & 0 & 0.23 & - & 0.37 \\
cprover\_software\_analysis/szymanski\_vs\_satabs.2 & 568 & 8321 & - & 3.94 & - & 11.44 & - & 0 & 5.80 & - & 0 & 13.53 & - & 3.35 & - & 0 & 3.38 & - & 5.85 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments-3}
\end{table}

\section{Benchmarks of \mist}

The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
Table~\ref{mist-experiments} shows the experimental results.
In the case of \footnotemark[1], the high runtime and timeouts
for the invariant come from the special form of the property,
as the negation contains a lot of disjunctions.

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
    \hline
           \multicolumn{3}{|c|}{Network}
         & \multicolumn{2}{|c|}{Safety}
         & \multicolumn{2}{|c|}{Safety-Int}
         & \multicolumn{3}{|c|}{Refinement}
         & \multicolumn{3}{|c|}{Refinement-Int}
         & \multicolumn{2}{|c|}{Invariant}
         & \multicolumn{3}{|c|}{Inv-Min}
         & \multicolumn{2}{|c|}{Inv-Ref} \\
    Name & P & T
        & R & T
        & R & T
        & R & I & T
        & R & I & T
        & R & T
        & R & I & T
        & R & T \\
    \hline
        \multicolumn{20}{|c|}{Positive Benchmarks} \\
    \hline
PN/MultiME & 12 & 11 & - & 0.16 & - & 0.14 & + & 1 & 0.34 & + & 1 & 0.32 & - & 0.13 & - & 0 & 0.12 & + & 0.93 \\
PN/basicME & 5 & 5 & - & 0.12 & - & 0.14 & + & 1 & 0.32 & + & 1 & 0.29 & - & 0.12 & - & 0 & 0.12 & + & 0.86 \\
PN/bingham\_h150 & 153 & 302 & + & 0.22 & + & 0.22 & + & 0 & 0.24 & + & 0 & 0.22 & + & 0.41 & + & 0 & 0.50 & + & 0.61 \\
PN/bingham\_h25 & 28 & 52 & + & 0.13 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.35 & + & 0.37 \\
PN/bingham\_h250 & 253 & 502 & + & 0.32 & + & 0.31 & + & 0 & 0.30 & + & 0 & 0.32 & + & 0.50 & + & 0 & 0.70 & + & 0.83 \\
PN/bingham\_h250\_attic \footnotemark[1] & 253 & 502 & + & 11.52 & + & 18.09 & + & 0 & 11.74 & + & 0 & 18.74 & + & 6535.28 & TO & TO & TO & TO & TO \\
PN/bingham\_h50 & 53 & 102 & + & 0.17 & + & 0.15 & + & 0 & 0.17 & + & 0 & 0.17 & + & 0.30 & + & 0 & 0.38 & + & 0.40 \\
PN/csm & 14 & 14 & + & 0.14 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.14 & + & 0.22 & + & 0 & 0.29 & + & 0.36 \\
PN/extendedread-write-smallconsts & 24 & 23 & - & 0.14 & - & 0.15 & - & 0 & 0.23 & - & 0 & 0.20 & - & 0.14 & - & 0 & 0.14 & - & 0.22 \\
PN/extendedread-write & 24 & 23 & - & 0.15 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.20 & - & 0.13 & - & 0 & 0.13 & - & 0.22 \\
PN/fms & 22 & 23 & + & 0.12 & + & 0.15 & + & 0 & 0.14 & + & 0 & 0.13 & + & 0.23 & + & 0 & 0.30 & + & 0.38 \\
PN/fms\_attic & 22 & 23 & + & 0.13 & + & 0.14 & + & 0 & 0.12 & + & 0 & 0.13 & + & 0.42 & + & 0 & 0.55 & + & 0.54 \\
PN/manufacturing & 13 & 6 & - & 0.15 & - & 0.15 & - & 0 & 0.21 & - & 0 & 0.19 & - & 0.12 & - & 0 & 0.13 & - & 0.19 \\
PN/mesh2x2 & 32 & 36 & + & 0.14 & + & 0.13 & + & 0 & 0.17 & + & 0 & 0.14 & + & 0.25 & + & 0 & 0.30 & + & 0.36 \\
PN/mesh3x2 & 52 & 60 & + & 0.18 & + & 0.17 & + & 0 & 0.14 & + & 0 & 0.15 & + & 0.26 & + & 0 & 0.35 & + & 0.39 \\
PN/multipool & 18 & 25 & + & 0.15 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.13 & + & 0.25 & + & 0 & 0.31 & + & 0.38 \\
PN/pingpong & 6 & 6 & - & 0.15 & - & 0.16 & + & 1 & 0.32 & + & 1 & 0.31 & - & 0.13 & - & 0 & 0.13 & + & 0.56 \\
boundedPN/kanban & 16 & 16 & + & 0.15 & + & 0.15 & + & 0 & 0.15 & + & 0 & 0.14 & + & 0.25 & + & 0 & 0.30 & + & 0.38 \\
boundedPN/lamport & 11 & 9 & - & 0.17 & - & 0.13 & + & 1 & 0.34 & + & 1 & 0.31 & - & 0.13 & - & 0 & 0.14 & + & 0.58 \\
boundedPN/newdekker & 16 & 14 & - & 0.15 & - & 0.18 & + & 1 & 0.34 & + & 1 & 0.37 & - & 0.12 & - & 0 & 0.15 & + & 0.55 \\
boundedPN/newrtp & 9 & 12 & + & 0.14 & + & 0.12 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.23 & + & 0 & 0.30 & + & 0.38 \\
boundedPN/peterson & 14 & 12 & - & 0.13 & - & 0.17 & + & 2 & 0.57 & + & 2 & 0.55 & - & 0.13 & - & 0 & 0.15 & + & 0.75 \\
boundedPN/read-write & 13 & 9 & + & 0.14 & + & 0.13 & + & 0 & 0.16 & + & 0 & 0.13 & + & 0.22 & + & 0 & 0.33 & + & 0.38 \\
    \hline
        \multicolumn{20}{|c|}{Negative Benchmarks} \\
    \hline
PN/kanban & 16 & 20 & - & 0.14 & - & 0.15 & - & 0 & 0.21 & - & 0 & 0.22 & - & 0.13 & - & 0 & 0.13 & - & 0.19 \\
PN/leabasicapproach & 16 & 14 & - & 0.13 & - & 0.17 & - & 0 & 0.21 & - & 0 & 0.28 & - & 0.19 & - & 0 & 0.16 & - & 0.20 \\
PN/pncsacover & 31 & 36 & - & 0.17 & - & 0.18 & - & 0 & 0.23 & - & 0 & 0.22 & - & 0.15 & - & 0 & 0.13 & - & 0.25 \\
PN/pncsasemiliv & 31 & 36 & - & 0.17 & - & 0.14 & - & 0 & 0.23 & - & 0 & 0.23 & - & 0.15 & - & 0 & 0.14 & - & 0.21 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \mist}
\label{mist-experiments}
\end{table}

\section{Benchmarks of \iic}

The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
Table~\ref{iic-experiments} shows the experimental results.

\begin{verbatim}
TODO: Use fixed or non-fixed benchmark?  Source for list of positives and negatives?
\end{verbatim}

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
    \hline
           \multicolumn{3}{|c|}{Network}
         & \multicolumn{2}{|c|}{Safety}
         & \multicolumn{2}{|c|}{Safety-Int}
         & \multicolumn{3}{|c|}{Refinement}
         & \multicolumn{3}{|c|}{Refinement-Int}
         & \multicolumn{2}{|c|}{Invariant}
         & \multicolumn{3}{|c|}{Inv-Min}
         & \multicolumn{2}{|c|}{Inv-Ref} \\
    Name & P & T
        & R & T
        & R & T
        & R & I & T
        & R & I & T
        & R & T
        & R & I & T
        & R & T \\
    \hline
        \multicolumn{20}{|c|}{Positive Benchmarks} \\
    \hline
concdb\_\_single\_client\_writes\_\_depth\_0\_\_fixed & 570 & 717 & + & 0.52 & + & 0.50 & + & 0 & 0.54 & + & 0 & 0.53 & + & 0.79 & + & 0 & 1.35 & + & 1.26 \\
finite\_leader\_\_single\_leader\_\_depth\_1\_\_fixed & 454 & 447 & + & 0.43 & + & 0.40 & + & 0 & 0.45 & + & 0 & 0.37 & + & 0.60 & + & 0 & 1.52 & + & 1.00 \\
finite\_leader\_\_single\_leader\_\_depth\_2\_\_fixed & 454 & 447 & + & 0.38 & + & 0.38 & + & 0 & 0.43 & + & 0 & 0.42 & + & 0.65 & + & 0 & 1.54 & + & 1.00 \\
firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_1\_\_fixed & 780 & 1024 & + & 0.64 & + & 0.66 & + & 0 & 0.67 & + & 0 & 0.68 & + & 0.92 & + & 0 & 2.46 & + & 1.61 \\
firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_2\_\_fixed & 4014 & 6938 & + & 8.05 & + & 8.75 & + & 0 & 8.71 & + & 0 & 8.07 & + & 9.38 & + & 0 & 40.62 & + & 17.59 \\
parikh\_\_should\_already\_be\_initialized\_\_depth\_0\_\_fixed & 77 & 70 & + & 0.17 & + & 0.18 & + & 0 & 0.21 & + & 0 & 0.18 & + & 0.28 & + & 0 & 0.40 & + & 0.45 \\
parikh\_\_should\_already\_be\_initialized\_\_depth\_1\_\_fixed & 88 & 82 & + & 0.17 & + & 0.17 & + & 0 & 0.16 & + & 0 & 0.17 & + & 0.31 & + & 0 & 0.40 & + & 0.44 \\
parikh\_\_should\_already\_be\_initialized\_\_depth\_2\_\_fixed & 88 & 82 & + & 0.20 & + & 0.17 & + & 0 & 0.16 & + & 0 & 0.18 & + & 0.29 & + & 0 & 0.42 & + & 0.44 \\
pipe\_\_single\_message\_in\_mailbox\_\_depth\_0\_\_fixed & 212 & 236 & + & 0.25 & + & 0.29 & + & 0 & 0.25 & + & 0 & 0.26 & + & 0.40 & + & 0 & 0.55 & + & 0.65 \\
reslock\_\_critical\_\_depth\_0\_\_fixed & 364 & 395 & + & 0.35 & + & 0.34 & + & 0 & 0.36 & + & 0 & 0.32 & + & 0.53 & + & 0 & 1.00 & + & 0.87 \\
reslock\_\_critical\_\_depth\_1\_\_fixed & 600 & 702 & + & 0.53 & + & 0.52 & + & 0 & 0.51 & + & 0 & 0.51 & + & 0.79 & + & 0 & 2.15 & + & 1.28 \\
reslock\_\_critical\_\_depth\_2\_\_fixed & 841 & 1030 & + & 0.77 & + & 0.72 & + & 0 & 0.84 & + & 0 & 0.77 & + & 1.04 & + & 0 & 3.40 & + & 1.81 \\
reslockbeh\_\_critical\_\_depth\_0\_\_fixed & 734 & 977 & + & 0.87 & + & 0.70 & + & 0 & 0.75 & + & 0 & 0.82 & + & 1.10 & + & 0 & 2.82 & + & 2.03 \\
reslockbeh\_\_critical\_\_depth\_1\_\_fixed & 4461 & 187 & + & 0.97 & + & 0.82 & + & 0 & 0.81 & + & 0 & 0.76 & + & 1.21 & + & 0 & 46.97 & + & 1.92 \\
ring\_\_single\_message\_in\_mailbox\_\_depth\_0\_\_fixed & 387 & 453 & + & 0.38 & + & 0.36 & + & 0 & 0.37 & + & 0 & 0.40 & + & 0.57 & + & 0 & 1.06 & + & 0.94 \\
safe\_send\_\_sending\_to\_non-pid\_1\_\_depth\_1\_\_fixed & 279 & 285 & - & 0.33 & - & 0.32 & + & 1 & 0.75 & + & 1 & 0.74 & - & 0.27 & - & 0 & 0.30 & + & 1.22 \\
safe\_send\_\_sending\_to\_non-pid\_1\_\_depth\_2\_\_fixed & 279 & 285 & - & 0.34 & - & 0.29 & + & 1 & 0.82 & + & 1 & 0.72 & - & 0.27 & - & 0 & 0.28 & + & 1.21 \\
safe\_send\_\_sending\_to\_non-pid\_2\_\_depth\_1\_\_fixed & 279 & 285 & - & 0.33 & - & 0.30 & + & 1 & 0.74 & + & 1 & 0.76 & - & 0.27 & - & 0 & 0.26 & + & 1.20 \\
safe\_send\_\_sending\_to\_non-pid\_2\_\_depth\_2\_\_fixed & 279 & 285 & - & 0.30 & - & 0.32 & + & 1 & 0.74 & + & 1 & 0.72 & - & 0.27 & - & 0 & 0.27 & + & 1.17 \\
safe\_send\_\_sending\_to\_non-pid\_3\_\_depth\_1\_\_fixed & 279 & 285 & + & 0.27 & + & 0.29 & + & 0 & 0.28 & + & 0 & 0.28 & + & 0.45 & + & 0 & 0.69 & + & 0.73 \\
safe\_send\_\_sending\_to\_non-pid\_3\_\_depth\_2\_\_fixed & 279 & 285 & + & 0.27 & + & 0.27 & + & 0 & 0.32 & + & 0 & 0.25 & + & 0.42 & + & 0 & 0.72 & + & 0.69 \\
safe\_send\_\_sending\_to\_non-pid\_4\_\_depth\_1\_\_fixed & 279 & 285 & - & 0.30 & - & 0.34 & - & 5 & 2.56 & + & 3 & 1.64 & - & 0.28 & - & 0 & 0.27 & - & 2.58 \\
safe\_send\_\_sending\_to\_non-pid\_4\_\_depth\_2\_\_fixed & 279 & 285 & - & 0.31 & - & 0.30 & - & 5 & 2.65 & + & 3 & 1.61 & - & 0.27 & - & 0 & 0.27 & - & 2.58 \\
sieve\_\_single\_message\_in\_counter\_mailbox\_\_depth\_0\_\_fixed & 263 & 275 & + & 0.31 & + & 0.26 & + & 0 & 0.30 & + & 0 & 0.29 & + & 0.43 & + & 0 & 0.70 & + & 0.75 \\
sieve\_\_single\_message\_in\_filter\_mailbox\_\_depth\_0\_\_fixed & 263 & 276 & + & 0.29 & + & 0.27 & + & 0 & 0.29 & + & 0 & 0.29 & + & 0.43 & + & 0 & 0.69 & + & 0.71 \\
sieve\_\_single\_message\_in\_sieve\_mailbox\_\_depth\_0\_\_fixed & 263 & 274 & + & 0.32 & + & 0.31 & + & 0 & 0.29 & + & 0 & 0.28 & + & 0.43 & + & 0 & 0.71 & + & 0.69 \\
state\_factory\_\_after\_receive\_if\_no\_mail\_\_depth\_0\_\_fixed & 542 & 648 & + & 0.52 & + & 0.49 & + & 0 & 0.54 & + & 0 & 0.49 & + & 0.74 & + & 0 & 1.41 & + & 1.19 \\
state\_factory\_\_single\_message\_in\_mailbox\_\_depth\_0\_\_fixed & 541 & 647 & + & 0.51 & + & 0.47 & + & 0 & 0.50 & + & 0 & 0.49 & + & 0.76 & + & 0 & 1.31 & + & 1.37 \\
    \hline
        \multicolumn{20}{|c|}{Negative Benchmarks} \\
    \hline
finite\_leader\_\_single\_leader\_\_depth\_0\_\_fixed & 297 & 308 & - & 0.36 & - & 0.32 & - & 1 & 0.90 & - & 1 & 0.98 & - & 0.29 & - & 0 & 0.30 & - & 0.98 \\
firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_0\_\_fixed & 352 & 400 & - & 0.37 & - & 0.40 & - & 1 & 1.07 & - & 1 & 1.19 & - & 0.41 & - & 0 & 0.33 & - & 1.26 \\
howait\_\_all\_workers\_finished\_if\_wait\_over\_\_depth\_0\_\_fixed & 217 & 232 & - & 0.27 & - & 0.26 & - & 1 & 0.82 & - & 1 & 0.87 & - & 0.26 & - & 0 & 0.26 & - & 0.80 \\
safe\_send\_\_sending\_to\_non-pid\_\_depth\_0\_\_fixed & 121 & 120 & - & 0.20 & - & 0.22 & - & 1 & 0.63 & - & 1 & 0.59 & - & 0.18 & - & 0 & 0.19 & - & 0.62 \\
stutter\_\_we\_abhorr\_as\_\_depth\_0\_\_fixed & 87 & 86 & - & 0.17 & - & 0.18 & - & 1 & 0.58 & - & 1 & 0.60 & - & 0.17 & - & 0 & 0.19 & - & 0.55 \\
stutter\_\_we\_abhorr\_as\_\_depth\_1\_\_fixed & 107 & 104 & - & 0.20 & - & 0.20 & - & 1 & 0.54 & - & 1 & 0.58 & - & 0.19 & - & 0 & 0.17 & - & 0.56 \\
stutter\_\_we\_abhorr\_as\_\_depth\_2\_\_fixed & 107 & 104 & - & 0.19 & - & 0.22 & - & 1 & 0.62 & - & 1 & 0.59 & - & 0.18 & - & 0 & 0.17 & - & 0.62 \\
unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_0\_\_fixed & 35 & 30 & - & 0.17 & - & 0.15 & - & 1 & 0.48 & - & 1 & 0.50 & - & 0.16 & - & 0 & 0.16 & - & 0.48 \\
unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_1\_\_fixed & 35 & 30 & - & 0.16 & - & 0.16 & - & 1 & 0.45 & - & 1 & 0.45 & - & 0.14 & - & 0 & 0.17 & - & 0.47 \\
unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_2\_\_fixed & 35 & 30 & - & 0.15 & - & 0.15 & - & 1 & 0.50 & - & 1 & 0.47 & - & 0.15 & - & 0 & 0.14 & - & 0.44 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \iic}
\label{iic-experiments}
\end{table}

\end{document}

### 2014-01-28 make-benchmarks-table.sh

@@ -1,0 +1,61 @@
+#!/bin/bash
+
+benchmarks=( 'found-in-mist-repo' 'given-by-daniel-kroening' 'ic3-soter' )
+
+methods=( 'safety' 'safety-int' 'refinement' 'refinement-int' 'invariant' 'inv-min' 'inv-ref' )
+
+timeout=7200
+
+function benchmark_files {
+  benchmark=$1
+  benchmark_dir='benchmarks/'$benchmark
+  in_file=$benchmark_dir'/'$2
+  out_file=$benchmark_dir'/'$3
+  >$out_file
+  echo -n "Name;Places;Transitions;" | tee -a $out_file
+  for method in ${methods[@]}; do
+    echo -n "$method-Result;$method-Iterations;$method-Time;" | tee -a $out_file
+  done
+  echo | tee -a $out_file
+  while read pl_file; do
+    if [[ "$pl_file" =~ ^'#' ]]; then
+      echo "Skipping $pl_file"
+      continue
+    fi
+    name=$( echo -n "$pl_file" | \
+            sed -e "s/^benchmarks\\/$benchmark\\///" \
+                -e 's/\.pl$//' \
+                -e 's/\.spec$//' \
+                -e 's/\.tts$//')
+    echo -n "$name;" | tee -a $out_file
+    ./src/print-info-for-net.sh $pl_file | tee -a $out_file
+    for method in ${methods[@]}; do
+      T="$(date +%s%N)"
+      (
+        set -o pipefail;
+        timeout $timeout ./src/main "-benchmark" "-$method" $pl_file | tee -a $out_file
+      )
+      result=$?
+      T=$(($(date +%s%N)-T))
+      if [[ result -eq 124 || result -eq 137 ]]; then
+        echo -n "TO;TO;TO;"
+      else
+        time=$(printf "%4.2f" $(echo "$T / 10^9" | bc -l))
+        echo -n "$time;" | tee -a $out_file
+      fi
+    done
+    echo | tee -a $out_file
+  done <$in_file
+}
+
+for (( b=0;b<${#benchmarks[@]};b++)); do
+  benchmark=${benchmarks[$b]}
+  echo
+  echo "* $benchmark"
+  echo
+  echo "- positive"
+  benchmark_files $benchmark "positive.list" "positive-table.csv"
+  echo "- negative"
+  benchmark_files $benchmark "negative.list" "negative-table.csv"
+done
+

### 2014-01-28 run-benchmarks.sh

@@ -1,22 +1,24 @@
 #!/bin/bash
 
+benchmarks=( 'found-in-mist-repo' 'given-by-daniel-kroening' 'ic3-soter' )
+
 #for benchmark_dir in `find benchmarks -mindepth 1 -maxdepth 1 -type d`; do
-for benchmark_dir in benchmarks/ic3-soter; do
+for benchmark in $benchmarks; do
+  benchmark_dir="benchmarks/$benchmark"
   >$benchmark_dir/positive-pnerf.list
   >$benchmark_dir/dontknow-pnerf.list
   >$benchmark_dir/negative-pnerf.list
   >$benchmark_dir/timeout-pnerf.list
   >$benchmark_dir/error-pnerf.list
   >$benchmark_dir/timing-pnerf.log
-      #timeout 600 ./src/explore-states.sh -d 0 $pl_file | tee $pl_file.out
   for pl_file in `find $benchmark_dir -name "*.pl"`; do
-    T="$(gdate +%s%N)"
+    T="$(date +%s%N)"
     (
       set -o pipefail;
-      timeout 600 ./src/main -refinement $pl_file | tee $pl_file.out
+      timeout 60 ./src/main -refinement-int $pl_file | tee $pl_file.out
     )
     result=$?
-    T=$(($(gdate +%s%N)-T))
+    T=$(($(date +%s%N)-T))
     if [[ result -eq 0 ]]; then
         list='positive'
     elif [[ result -eq 1 ]]; then

### 2014-01-28 src/print-info-for-net.pl

@@ -1,0 +1,22 @@
+:- dynamic place/3.        % place(Id, InTransitions, OutTransitions).
+:- dynamic transition/3.   % transition(Id, InPlaces, OutPlaces).
+
+:- use_module(library(aggregate)).
+
+:- ['load-pl-file.pl'].
+:- ['misc.pl'].
+
+print_info :-
+        findall(P, place(P, _, _), Ps),
+        length(Ps, PlaceCount),
+        findall(T, transition(T, _, _), Ts),
+        length(Ts, TransitionCount),
+        format('~d;~d;', [PlaceCount, TransitionCount]).
+
+% Entry point
+:-      prolog_flag(argv, Argv),
+        (   foreach(F, Argv)
+        do  load_pl_file(F)
+        ),
+        print_info,
+        halt.

### 2014-01-28 src/print-info-for-net.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+function absolute_path {
+    (cd "$(dirname "$1")" && pwd)
+}
+
+sysdir=$(absolute_path $0)
+
+tmpdir=/tmp
+
+input=$1
+
+cp $(absolute_path $input)/$(basename $input) $tmpdir/input-petri-net.pl
+
+sicstus -l "$sysdir"/input-file-to-petri-net.pl -- 0 $tmpdir/input-petri-net.pl 2>/dev/null >$tmpdir/pp-petri-net.pl
+sicstus -l "$sysdir"/print-info-for-net.pl -- $tmpdir/pp-petri-net.pl 2>/dev/null
+

### 2014-01-30 paper/workspace/main.tex
Replaced example in report with Lamport's algorithm

@@
 \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
+\usepackage{graphicx}
 
@@
 The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on a simplified mutual exclusion algorithm by Delzanno2001.
+We present the method Safety by example on Lamport's 1-bit algorithm by Esparza1997.
 
@@
 
-bool m = false;
+TODO
 
-// Thread 1:
-while (true) {
-  if (! test_and_set(m)) {
-    // critical section 1
-    m = false
-  }
-  if (! test_and_set(m)) {
-    // critical section 2
-    m = false
-  }
-}
-
-// Thread 2:
-while (true) {
-  if (! test_and_set(m)) {
-    // critical section 1
-    m = false
-  }
-  if (! test_and_set(m)) {
-    // critical section 2
-    m = false
-  }
-}
-
-* Property: Thread 1 and Thread 2 are never in any critical section at the same time.
+* Property: Thread 1 and Thread 2 are never in the critical section at the same time.
 
@@
 \begin{center}
-  \begin{tikzpicture}[node distance=2cm]
-    \node[blue place, tokens=2, label=above:nc] (nc) {};
-    \node[place, tokens=1, label=above:m, below of=nc] (s) {};
-    \node[red place, label=above:cs1, left of=s] (cs1) {};
-    \node[red place, label=above:cs2, right of=s] (cs2) {};
-    \node[transition, label=above:t1] (t1) at ($(s)!0.5!(cs1) + (0,1cm)$) {}
-    edge [pre] (nc)
-    edge [pre] (s)
-    edge [post] (cs1);
-    \node[transition, label=above:t2] (t2) at ($(s)!0.5!(cs2) + (0,1cm)$) {}
-    edge [pre] (nc)
-    edge [pre] (s)
-    edge [post] (cs2);
-    \node[transition, label=below:t3] (t3) at ($(s)!0.5!(cs1) - (0,1cm)$) {}
-    edge [pre] (cs1)
-    edge [post] (s)
-    edge [post,bend left,min distance=3cm, out=105, in=105] (nc);
-    \node[transition, label=below:t4] (t4) at ($(s)!0.5!(cs2) - (0,1cm)$) {}
-    edge [pre] (cs2)
-    edge [post] (s)
-    edge [post,bend right,min distance=3cm, out=-105, in=-105] (nc);
-  \end{tikzpicture}
+  \includegraphics[width=12cm]{lamport}
 \end{center}
@@
 \begin{verbatim}
-* Property: P = ( cs1 + cs2 <= 1 )
+* Property: P = ( cs1 + cs2 >= 1 )
 \end{verbatim}
@@
 
-    nc      = 2 - t1 - t2 + t3 + t4
-    ^         ^   ^    ^    ^    ^
-    |         |   |    |    |    |
-    |         |   ------    ------
+    id1     = 1    + u1    - u2
+    ^         ^      |       |
     |         |      |       |
-    |         |      |       # of tokens given to nc
+    |         |      |       |
+    |         |      |       |
+    |         |      |     # of tokens taken from nc
     |         |      |    
-    |         |    # of tokens taken from nc
+    |         |    # of tokens given to nc
     |         |
@@
 
-    m     = 1 - t1 - t2 + t3 + t4
-    cs1   = 0 + t1 - t3
-    cs2   = 0 + t2 - t4
+    cs1 = 0 - u1 + u3
+    idle1 = 1 + u1 - u2
+    req1 = 0 + u2 - u3
+    nid1 = 0 - u1 u2
+    id2 = 1 + v2 + v5 - v6
+    cs2 = 0 + v4 - v5
+    idle2 = 1 + v3 + v5 - v6
+    req2 = 0 - v1 - v4 + v6
+    after_you2 = 0 + v1 - v2
+    await2 = 0 + v2 - v3
 
@@
 
-    nc  >= 0
-    m   >= 0
+    id1 >= 0
     cs1 >= 0
+    idle1 >= 0
+    req1 >= 0
+    nid1 >= 0
+    id2 >= 0
     cs2 >= 0
-    t1  >= 0
-    t2  >= 0
-    t3  >= 0
-    t4  >= 0
+    idle2 >= 0
+    req2 >= 0
+    after_you2 >= 0
+    await2 >= 0
+    u1 >= 0
+    u2 >= 0
+    u3 >= 0
+    v1 >= 0
+    v2 >= 0
+    v3 >= 0
+    v4 >= 0
+    v5 >= 0
+    v6 >= 0
 \end{verbatim}
@@
   # of tokens in s = initial number of tokens of place s
+                     + # times each input transition of s fires
                      - # times each output transition of s fires
-                     + # times each input transition of s fires
 
@@
 \end{verbatim}
+
+\iffalse
 
@@
 
+\fi
+
 \end{document}

### 2014-01-30 tests/lamports-alg/input-petri-net.pl

@@ -1,0 +1,30 @@
+place(cs1).
+place(idle1).
+place(nid1).
+place(req1).
+place(id1).
+
+place(cs2).
+place(idle2).
+place(req2).
+place(after_you2).
+place(await2).
+place(id2).
+
+transition(u1, [cs1, nid1], [idle1, id1]).
+transition(u2, [idle1, id1], [req1, nid1]).
+transition(u3, [req1, id2], [cs1, id2]).
+
+transition(v1, [req2, nid1], [after_you2, nid1]).
+transition(v2, [after_you2], [id2, await2]).
+transition(v3, [id1, await2], [id1, idle2]).
+transition(v4, [req2, id1], [cs2, id1]).
+transition(v5, [cs2], [idle2, id2]).
+transition(v6, [idle2, id2], [req2]).
+
+init(id1).
+init(id2).
+init(idle1).
+init(idle2).
+
+target(1, [([cs1,cs2],2)]).

### 2014-01-30 paper/workspace/main.tex
Formatted example for safety.

@@
 The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm by Esparza1997.
+We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
 
@@
 \begin{verbatim}
-* Property: P = ( cs1 + cs2 >= 1 )
+* Property P: cs1 + cs2 <= 1
 \end{verbatim}
@@
     id1        = 1 + u1 - u2
-    ^         ^      |       |
+    ^            ^   ^    ^
     |            |   |    |
@@
     req1       = 0      + u2 - u3
-    nid1 = 0 - u1 u2
+    nid1       = 0 - u1 + u2
     id2        = 1                     + v2           + v5 - v6
@@
     await2     = 0                     + v2 - v3
+\end{verbatim}
 
+\newpage
+
+\begin{verbatim}
     Non-negativity conditions:
@@
     v6         >= 0
-\end{verbatim}
 
-\newpage
-\begin{verbatim}
 * Place equation:

### 2014-01-30 paper/workspace/main.tex
Added pseudocode for method safety

@@
 
+  - Negated Property \neg P:
+
+    cs1 + cs2  >= 2
+
 * Place equation:
@@
     C : State constraints
+  Pseudocode:
 
-  TODO: Unroll pseudocode?
+  # Place equations
+  for s in S:
+    # initial number of tokens
+    print(s + " = " + M0(s)"
+    for t in T:
+      if (t, s) in E and not (s, t) in E:
+        # input transition
+        print(" + " t)
+      if (s, t) in E and not (t, s) in E:
+        # output transition
+        print(" - " t)
+    print("\n")
+
+  # Non-negativity conditions
+  for s in S:
+    print(s + " >= 0\n")
+  for t in T:
+    print(t + " >= 0\n")
 \end{verbatim}

### 2014-01-31 make-benchmarks-table.sh
Added method Invariant by Refinement with Minimization

@@ -12,11 +12,6 @@ function benchmark_files {
   in_file=$benchmark_dir'/'$2
   out_file=$benchmark_dir'/'$3
   >$out_file
-  echo -n "Name;Places;Transitions;" | tee -a $out_file
-  for method in ${methods[@]}; do
-    echo -n "$method-Result;$method-Iterations;$method-Time;" | tee -a $out_file
-  done
-  echo | tee -a $out_file
   while read pl_file; do
     if [[ "$pl_file" =~ ^'#' ]]; then
       echo "Skipping $pl_file"

### 2014-01-31 doc/benchmark-tables.tex
Updated benchmark tables with new method inv-ref-min

@@
 \usepackage{listings}
-\usepackage[a4paper, landscape, top=2cm, bottom=2cm, left=2cm, right=2cm]{geometry}
+\usepackage[a4paper, landscape, top=0.5cm, bottom=0.5cm, left=0.5cm, right=0.5cm, includefoot]{geometry}
 
@@
   \item[R] Result of our tool. It is + if our tool can prove safety or find an invariant and - otherwise.
-  \item[I] Number of iterations. This is the number of traps used for refinement or the number of times an invariant is minimized.
+  \item[I] Number of iterations. This is the number of traps used for refinement.
+  \item[M] Number of minimization steps for the invariant.
   \item[T] Running time of our tool in seconds.
@@
   \item[Safety] Method Safety over reals.
-  \item[Safety-Int] Method Safety over integers.
+  \item[Safe-Int] Method Safety over integers.
   \item[Refinement] Method Safety by Refinement over reals.
-  \item[Refinement-Int] Method Safety by Refinement over integers.
+  \item[Ref-Int] Method Safety by Refinement over integers.
   \item[Invariant] Method Invariant.
@@
   \item[Inv-Ref] Method Invariant by Refinement.
+  \item[Inv-Ref-Min] Method Invariant by Refinement with Minimization.
 \end{description}
@@
 The set of benchmarks consists of 77 benchmarks, 26 positive and 51 negative.
-Table~\ref{bfc-experiments}, \ref{bfc-experiments-2} and \ref{bfc-experiments-3}
+Table~\ref{bfc-experiments} and \ref{bfc-experiments-2}
 show the experimental results.
@@
 \begin{center}
-  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
+  \begin{tabular}{ | l | r | r | *{25}{ r | } } %
     \hline
@@
          & \multicolumn{2}{|c|}{Safety}
-         & \multicolumn{2}{|c|}{Safety-Int}
+         & \multicolumn{2}{|c|}{Safe-Int}
          & \multicolumn{3}{|c|}{Refinement}
-         & \multicolumn{3}{|c|}{Refinement-Int}
+         & \multicolumn{3}{|c|}{Ref-Int}
          & \multicolumn{2}{|c|}{Invariant}
          & \multicolumn{3}{|c|}{Inv-Min}
-         & \multicolumn{2}{|c|}{Inv-Ref} \\
+         & \multicolumn{3}{|c|}{Inv-Ref}
+         & \multicolumn{4}{|c|}{Inv-Ref-Min} \\
     Name & P & T
@@
         & R & T
+        & R & M & T
         & R & I & T
-        & R & T \\
+        & R & I & M & T \\
     \hline
-        \multicolumn{20}{|c|}{Positive Benchmarks} \\
+        \multicolumn{25}{|c|}{Positive Benchmarks} \\
     \hline
-cprover-PN/abp & 42 & 43 & + & 0.19 & + & 0.18 & + & 0 & 0.13 & + & 0 & 0.17 & + & 0.26 & + & 2 & 0.56 & + & 0.42 \\
-cprover-PN/basicME & 18 & 17 & + & 0.13 & + & 0.15 & + & 0 & 0.13 & + & 0 & 0.17 & + & 0.25 & + & 0 & 0.30 & + & 0.39 \\
-cprover-PN/bingham\_h150 & 309 & 457 & + & 0.35 & + & 0.35 & + & 0 & 0.40 & + & 0 & 0.35 & + & 0.57 & + & 0 & 0.81 & + & 0.93 \\
-cprover-PN/bingham\_h25 & 59 & 82 & + & 0.17 & + & 0.19 & + & 0 & 0.17 & + & 0 & 0.18 & + & 0.30 & + & 0 & 0.36 & + & 0.45 \\
-cprover-PN/bingham\_h250 & 509 & 757 & + & 0.61 & + & 0.55 & + & 0 & 0.55 & + & 0 & 0.57 & + & 0.85 & + & 0 & 1.22 & + & 1.43 \\
-cprover-PN/bingham\_h50 & 109 & 157 & + & 0.22 & + & 0.22 & + & 0 & 0.19 & + & 0 & 0.19 & + & 0.34 & + & 0 & 0.46 & + & 0.55 \\
-cprover-PN/bingham\_h500 & 1009 & 1507 & + & 1.20 & + & 1.09 & + & 0 & 1.20 & + & 0 & 1.11 & + & 1.74 & + & 0 & 2.89 & + & 3.00 \\
-cprover-PN/clientserver & 13 & 11 & + & 0.13 & + & 0.12 & + & 0 & 0.15 & + & 0 & 0.14 & + & 0.27 & + & 0 & 0.32 & + & 0.38 \\
-cprover-PN/csm & 28 & 27 & + & 0.14 & + & 0.15 & + & 0 & 0.13 & + & 0 & 0.13 & + & 0.26 & + & 0 & 0.32 & + & 0.35 \\
-cprover-PN/gsm & 73 & 156 & + & 0.26 & + & 0.29 & + & 0 & 0.21 & + & 0 & 0.20 & + & 0.64 & + & 0 & 0.68 & + & 0.78 \\
-cprover-PN/handover & 126 & 255 & + & 0.23 & + & 0.23 & + & 0 & 0.25 & + & 0 & 0.26 & + & 0.37 & + & 0 & 0.50 & + & 0.60 \\
-cprover-PN/keycard & 59 & 78 & + & 0.16 & + & 0.15 & + & 0 & 0.21 & + & 0 & 0.15 & + & 0.33 & + & 0 & 0.36 & + & 0.49 \\
-cprover-PN/lamport & 30 & 27 & - & 0.19 & - & 0.18 & - & 0 & 0.20 & - & 0 & 0.21 & - & 0.14 & - & 0 & 0.17 & - & 0.23 \\
-cprover-PN/mesh2x2 & 66 & 69 & + & 0.20 & + & 0.20 & + & 0 & 0.15 & + & 0 & 0.20 & + & 0.27 & + & 0 & 0.37 & + & 0.43 \\
-cprover-PN/multipoll & 35 & 41 & + & 0.15 & + & 0.15 & + & 0 & 0.16 & + & 0 & 0.14 & + & 0.26 & + & 0 & 0.32 & + & 0.43 \\
-cprover-PN/newdekker & 51 & 48 & - & 0.18 & - & 0.18 & - & 0 & 0.22 & - & 0 & 0.25 & - & 0.15 & - & 0 & 0.15 & - & 0.24 \\
-cprover-PN/newrtp & 12 & 14 & + & 0.15 & + & 0.12 & + & 0 & 0.15 & + & 0 & 0.14 & + & 0.27 & + & 0 & 0.30 & + & 0.40 \\
-cprover-PN/peterson & 39 & 36 & - & 0.15 & - & 0.16 & - & 0 & 0.27 & - & 0 & 0.21 & - & 0.15 & - & 0 & 0.14 & - & 0.22 \\
-cprover-PN/phones & 51 & 87 & + & 0.16 & + & 0.18 & + & 0 & 0.16 & + & 0 & 0.16 & + & 0.30 & + & 0 & 0.37 & + & 0.43 \\
-cprover-PN/reaction & 19 & 15 & + & 0.14 & + & 0.14 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.34 & + & 0.37 \\
-cprover-PN/reaction3 & 22 & 18 & + & 0.16 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.15 & + & 0.25 & + & 0 & 0.32 & + & 0.37 \\
-cprover-PN/reaction4 & 16 & 13 & + & 0.14 & + & 0.15 & + & 0 & 0.14 & + & 0 & 0.13 & + & 0.28 & + & 0 & 0.33 & + & 0.36 \\
-cprover-PN/read-write & 36 & 31 & + & 0.16 & + & 0.20 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.28 & + & 0 & 0.34 & + & 0.43 \\
-cprover-PN/restriction & 15 & 10 & + & 0.14 & + & 0.14 & + & 0 & 0.13 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.32 & + & 0.40 \\
-cprover\_software\_analysis/conditionals\_vs\_satabs.2 & 214 & 281 & + & 0.27 & + & 0.25 & + & 0 & 0.27 & + & 0 & 0.25 & + & 0.44 & + & 0 & 0.55 & + & 0.65 \\
-cprover\_software\_analysis/rand\_cas\_vs\_satabs.2 & 110 & 139 & + & 0.19 & + & 0.21 & + & 0 & 0.20 & + & 0 & 0.23 & + & 0.36 & + & 0 & 0.47 & + & 0.50 \\
+cprover-PN/abp & 42 & 43 & + & 0.12 & + & 0.12 & + & 0 & 0.12 & + & 0 & 0.12 & + & 0.22 & + & 2 & 0.44 & + & 0 & 0.32 & + & 0 & 2 & 0.55 \\
+cprover-PN/basicME & 18 & 17 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.21 & + & 0 & 0.27 & + & 0 & 0.31 & + & 0 & 0 & 0.36 \\
+cprover-PN/bingham\_h150 & 309 & 457 & + & 0.31 & + & 0.30 & + & 0 & 0.31 & + & 0 & 0.30 & + & 0.48 & + & 0 & 0.69 & + & 0 & 0.79 & + & 0 & 0 & 0.98 \\
+cprover-PN/bingham\_h25 & 59 & 82 & + & 0.13 & + & 0.14 & + & 0 & 0.15 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.31 & + & 0 & 0.36 & + & 0 & 0 & 0.43 \\
+cprover-PN/bingham\_h250 & 509 & 757 & + & 0.47 & + & 0.45 & + & 0 & 0.47 & + & 0 & 0.46 & + & 0.71 & + & 0 & 1.05 & + & 0 & 1.18 & + & 0 & 0 & 1.50 \\
+cprover-PN/bingham\_h50 & 109 & 157 & + & 0.17 & + & 0.17 & + & 0 & 0.17 & + & 0 & 0.17 & + & 0.29 & + & 0 & 0.38 & + & 0 & 0.44 & + & 0 & 0 & 0.53 \\
+cprover-PN/bingham\_h500 & 1009 & 1507 & + & 0.99 & + & 0.94 & + & 0 & 0.99 & + & 0 & 0.92 & + & 1.45 & + & 0 & 2.35 & + & 0 & 2.44 & + & 0 & 0 & 3.35 \\
+cprover-PN/clientserver & 13 & 11 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.22 & + & 0 & 0.25 & + & 0 & 0.29 & + & 0 & 0 & 0.35 \\
+cprover-PN/csm & 28 & 27 & + & 0.11 & + & 0.11 & + & 0 & 0.12 & + & 0 & 0.12 & + & 0.21 & + & 0 & 0.27 & + & 0 & 0.31 & + & 0 & 0 & 0.37 \\
+cprover-PN/gsm & 73 & 156 & + & 0.16 & + & 0.16 & + & 0 & 0.16 & + & 0 & 0.16 & + & 0.27 & + & 0 & 0.35 & + & 0 & 0.41 & + & 0 & 0 & 0.50 \\
+cprover-PN/handover & 126 & 255 & + & 0.20 & + & 0.20 & + & 0 & 0.20 & + & 0 & 0.20 & + & 0.31 & + & 0 & 0.43 & + & 0 & 0.50 & + & 0 & 0 & 0.61 \\
+cprover-PN/keycard & 59 & 78 & + & 0.13 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.13 & + & 0.23 & + & 0 & 0.31 & + & 0 & 0.35 & + & 0 & 0 & 0.42 \\
+cprover-PN/lamport & 30 & 27 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.18 & - & 0 & 0 & 0.18 \\
+cprover-PN/mesh2x2 & 66 & 69 & + & 0.13 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.13 & + & 0.23 & + & 0 & 0.31 & + & 0 & 0.35 & + & 0 & 0 & 0.43 \\
+cprover-PN/multipoll & 35 & 41 & + & 0.12 & + & 0.12 & + & 0 & 0.12 & + & 0 & 0.12 & + & 0.23 & + & 0 & 0.28 & + & 0 & 0.33 & + & 0 & 0 & 0.39 \\
+cprover-PN/newdekker & 51 & 48 & - & 0.13 & - & 0.13 & - & 0 & 0.19 & - & 0 & 0.19 & - & 0.13 & - & 0 & 0.14 & - & 0 & 0.19 & - & 0 & 0 & 0.19 \\
+cprover-PN/newrtp & 12 & 14 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.25 & + & 0 & 0.30 & + & 0 & 0 & 0.35 \\
+cprover-PN/peterson & 39 & 36 & - & 0.13 & - & 0.13 & - & 0 & 0.19 & - & 0 & 0.19 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.19 & - & 0 & 0 & 0.18 \\
+cprover-PN/phones & 51 & 87 & + & 0.14 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.31 & + & 0 & 0.36 & + & 0 & 0 & 0.43 \\
+cprover-PN/reaction & 19 & 15 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.26 & + & 0 & 0.30 & + & 0 & 0 & 0.35 \\
+cprover-PN/reaction3 & 22 & 18 & + & 0.12 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.26 & + & 0 & 0.30 & + & 0 & 0 & 0.36 \\
+cprover-PN/reaction4 & 16 & 13 & + & 0.11 & + & 0.11 & + & 0 & 0.12 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.26 & + & 0 & 0.29 & + & 0 & 0 & 0.35 \\
+cprover-PN/read-write & 36 & 31 & + & 0.12 & + & 0.12 & + & 0 & 0.12 & + & 0 & 0.12 & + & 0.21 & + & 0 & 0.29 & + & 0 & 0.32 & + & 0 & 0 & 0.38 \\
+cprover-PN/restriction & 15 & 10 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.26 & + & 0 & 0.29 & + & 0 & 0 & 0.35 \\
+cprover\_sa/conditionals\_vs\_satabs.2 & 214 & 281 & + & 0.21 & + & 0.22 & + & 0 & 0.21 & + & 0 & 0.21 & + & 0.34 & + & 0 & 0.47 & + & 0 & 0.54 & + & 0 & 0 & 0.68 \\
+cprover\_sa/rand\_cas\_vs\_satabs.2 & 110 & 139 & + & 0.16 & + & 0.15 & + & 0 & 0.16 & + & 0 & 0.16 & + & 0.26 & + & 0 & 0.36 & + & 0 & 0.42 & + & 0 & 0 & 0.50 \\
     \hline
-        \multicolumn{20}{|c|}{Negative Benchmarks} \\
+        \multicolumn{25}{|c|}{Negative Benchmarks} \\
     \hline
-cprover-PN/kanban & 26 & 29 & - & 0.18 & - & 0.17 & - & 0 & 0.23 & - & 0 & 0.20 & - & 0.14 & - & 0 & 0.14 & - & 0.21 \\
-cprover-PN/leabasicapproach & 30 & 27 & - & 0.16 & - & 0.16 & - & 0 & 0.24 & - & 0 & 0.23 & - & 0.16 & - & 0 & 0.15 & - & 0.22 \\
-cprover-PN/platoon & 20 & 17 & - & 0.17 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.23 & - & 0.14 & - & 0 & 0.16 & - & 0.24 \\
-cprover-PN/platoonFiniteHandlerPoly & 19 & 16 & - & 0.15 & - & 0.16 & - & 0 & 0.24 & - & 0 & 0.22 & - & 0.13 & - & 0 & 0.13 & - & 0.20 \\
-cprover-PN/pncsacover & 64 & 68 & - & 0.19 & - & 0.18 & - & 0 & 0.28 & - & 0 & 0.25 & - & 0.16 & - & 0 & 0.17 & - & 0.28 \\
-cprover-PN/reaction2 & 13 & 11 & - & 0.15 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.19 & - & 0.14 & - & 0 & 0.12 & - & 0.25 \\
-cprover-PN/reaction5 & 11 & 8 & - & 0.14 & - & 0.14 & - & 1 & 0.39 & - & 1 & 0.44 & - & 0.12 & - & 0 & 0.14 & - & 0.42 \\
+cprover-PN/kanban & 26 & 29 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.18 & - & 0 & 0 & 0.18 \\
+cprover-PN/leabasicapproach & 30 & 27 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.18 & - & 0 & 0 & 0.18 \\
+cprover-PN/platoon & 20 & 17 & - & 0.12 & - & 0.12 & - & 0 & 0.17 & - & 0 & 0.17 & - & 0.11 & - & 0 & 0.11 & - & 0 & 0.17 & - & 0 & 0 & 0.17 \\
+cprover-PN/platoonFiniteHandlerPoly & 19 & 16 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.17 & - & 0.11 & - & 0 & 0.11 & - & 0 & 0.17 & - & 0 & 0 & 0.17 \\
+cprover-PN/pncsacover & 64 & 68 & - & 0.14 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.21 & - & 0.14 & - & 0 & 0.14 & - & 0 & 0.21 & - & 0 & 0 & 0.22 \\
+cprover-PN/reaction2 & 13 & 11 & - & 0.12 & - & 0.12 & - & 0 & 0.17 & - & 0 & 0.17 & - & 0.11 & - & 0 & 0.11 & - & 0 & 0.16 & - & 0 & 0 & 0.17 \\
+cprover-PN/reaction5 & 11 & 8 & - & 0.11 & - & 0.12 & - & 1 & 0.32 & - & 1 & 0.32 & - & 0.11 & - & 0 & 0.11 & - & 1 & 0.32 & - & 1 & 0 & 0.32 \\
+cprover\_sa/Boop\_simple\_vf\_satabs.1 & 31 & 31 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.18 & - & 0 & 0 & 0.18 \\
+cprover\_sa/Boop\_simple\_vf\_satabs.2 & 330 & 7489 & - & 2.67 & - & 2.68 & - & 0 & 4.28 & - & 0 & 4.29 & - & 2.44 & - & 0 & 2.43 & - & 0 & 4.30 & - & 0 & 0 & 4.29 \\
+cprover\_sa/Function\_Pointer3\_vs\_satabs.1 & 40 & 71 & - & 0.14 & - & 0.14 & - & 1 & 0.40 & - & 1 & 0.40 & - & 0.13 & - & 0 & 0.13 & - & 1 & 0.40 & - & 1 & 0 & 0.40 \\
+cprover\_sa/Function\_Pointer3\_vs\_satabs.2 & 306 & 841 & - & 0.43 & - & 0.44 & - & 0 & 0.65 & - & 1 & 1.26 & - & 0.39 & - & 0 & 0.39 & - & 0 & 0.67 & - & 0 & 0 & 0.67 \\
+cprover\_sa/Function\_Pointer3\_vs\_satabs.3 & 2826 & 8961 & - & 6.26 & - & 7.02 & - & 0 & 8.30 & - & 0 & 8.89 & - & 5.90 & - & 0 & 5.82 & - & 0 & 8.14 & - & 0 & 0 & 8.19 \\
     \hline
@@
         & R & T
+        & R & M & T
         & R & I & T
-        & R & T \\
+        & R & I & M & T \\
     \hline
-        \multicolumn{20}{|c|}{Negative Benchmarks (continued)} \\
+        \multicolumn{25}{|c|}{Negative Benchmarks (continued)} \\
     \hline
-cprover\_software\_analysis/Boop\_simple\_vf\_satabs.1 & 31 & 31 & - & 0.16 & - & 0.15 & - & 0 & 0.23 & - & 0 & 0.22 & - & 0.14 & - & 0 & 0.14 & - & 0.21 \\
-cprover\_software\_analysis/Boop\_simple\_vf\_satabs.2 & 330 & 7489 & - & 3.39 & - & 3.27 & - & 0 & 5.59 & - & 0 & 5.62 & - & 2.93 & - & 0 & 3.05 & - & 5.36 \\
-cprover\_software\_analysis/Function\_Pointer3\_vs\_satabs.1 & 40 & 71 & - & 0.17 & - & 0.16 & - & 1 & 0.52 & - & 1 & 0.49 & - & 0.15 & - & 0 & 0.17 & - & 0.49 \\
-cprover\_software\_analysis/Function\_Pointer3\_vs\_satabs.2 & 306 & 841 & - & 0.53 & - & 0.56 & - & 0 & 0.82 & - & 1 & 1.55 & - & 0.45 & - & 0 & 0.49 & - & 0.85 \\
-cprover\_software\_analysis/Function\_Pointer3\_vs\_satabs.3 & 2826 & 8961 & - & 8.12 & - & 8.86 & - & 0 & 10.21 & - & 0 & 11.29 & - & 7.49 & - & 0 & 7.56 & - & 10.19 \\
-cprover\_software\_analysis/buggy\_spaghetti\_vf\_satabs.1 & 78 & 115 & - & 0.18 & - & 0.18 & - & 0 & 0.31 & - & 0 & 0.27 & - & 0.20 & - & 0 & 0.17 & - & 0.29 \\
-cprover\_software\_analysis/buggy\_spaghetti\_vf\_satabs.2 & 326 & 697 & - & 0.50 & - & 0.47 & - & 0 & 0.77 & - & 0 & 0.77 & - & 0.43 & - & 0 & 0.47 & - & 0.72 \\
-cprover\_software\_analysis/conditionals\_vs\_satabs.1 & 50 & 55 & - & 0.18 & - & 0.17 & - & 0 & 0.24 & - & 0 & 0.25 & - & 0.15 & - & 0 & 0.15 & - & 0.25 \\
-cprover\_software\_analysis/constants\_vf\_satabs.1 & 26 & 25 & - & 0.15 & - & 0.15 & - & 0 & 0.25 & - & 0 & 0.22 & - & 0.15 & - & 0 & 0.13 & - & 0.22 \\
-cprover\_software\_analysis/constants\_vf\_satabs.2 & 262 & 393 & - & 0.33 & - & 0.35 & - & 0 & 0.54 & - & 0 & 0.51 & - & 0.30 & - & 0 & 0.29 & - & 0.54 \\
-cprover\_software\_analysis/dekker\_vs\_satabs.1 & 41 & 121 & - & 0.21 & - & 0.20 & - & 0 & 0.33 & - & 0 & 0.28 & - & 0.17 & - & 0 & 0.22 & - & 0.28 \\
-cprover\_software\_analysis/dekker\_vs\_satabs.2 & 164 & 1345 & - & 0.67 & - & 0.72 & - & 0 & 1.08 & - & 0 & 1.08 & - & 0.62 & - & 0 & 0.61 & - & 1.08 \\
-cprover\_software\_analysis/double\_lock\_p1\_vs\_satabs.1 & 64 & 205 & - & 0.22 & - & 0.24 & - & 1 & 0.79 & - & 2 & 1.59 & - & 0.36 & - & 0 & 0.36 & - & 0.79 \\
-cprover\_software\_analysis/double\_lock\_p1\_vs\_satabs.2 & 570 & 7601 & - & 4.28 & - & 12.37 & - & 0 & 6.52 & - & 0 & 13.19 & - & 3.29 & - & 0 & 3.19 & - & 6.02 \\
-cprover\_software\_analysis/double\_lock\_p1\_vs\_satabs.3 & 570 & 7569 & - & 3.90 & - & 11.20 & - & 0 & 5.76 & - & 0 & 13.13 & - & 3.12 & - & 0 & 3.23 & - & 5.85 \\
-cprover\_software\_analysis/double\_lock\_p2\_vs\_satabs.1 & 64 & 213 & - & 0.25 & - & 0.23 & - & 0 & 0.38 & - & 1 & 0.73 & - & 0.20 & - & 0 & 0.22 & - & 0.33 \\
-cprover\_software\_analysis/double\_lock\_p2\_vs\_satabs.2 & 184 & 1833 & - & 0.93 & - & 0.93 & - & 0 & 1.44 & - & 0 & 1.44 & - & 0.81 & - & 0 & 0.81 & - & 1.49 \\
-cprover\_software\_analysis/double\_lock\_p3\_vs\_satabs.1 & 46 & 81 & - & 0.18 & - & 0.18 & - & 0 & 0.25 & - & 0 & 0.27 & - & 0.15 & - & 0 & 0.16 & - & 0.28 \\
-cprover\_software\_analysis/double\_lock\_p3\_vs\_satabs.2 & 112 & 745 & - & 0.49 & - & 0.47 & - & 1 & 1.34 & - & 0 & 0.71 & - & 0.41 & - & 0 & 0.44 & - & 1.33 \\
-cprover\_software\_analysis/double\_lock\_p3\_vs\_satabs.3 & 306 & 3137 & - & 1.56 & - & 1.54 & - & 0 & 2.35 & - & 0 & 2.34 & - & 1.32 & - & 0 & 1.32 & - & 2.34 \\
-cprover\_software\_analysis/lu-fig2\_fixed\_vs\_satabs.1 & 39 & 37 & - & 0.19 & - & 0.20 & - & 0 & 0.25 & - & 0 & 0.23 & - & 0.17 & - & 0 & 0.13 & - & 0.23 \\
-cprover\_software\_analysis/lu-fig2\_fixed\_vs\_satabs.2 & 61 & 185 & - & 0.23 & - & 0.21 & - & 0 & 0.33 & - & 0 & 0.35 & - & 0.20 & - & 0 & 0.23 & - & 0.32 \\
-cprover\_software\_analysis/lu-fig2\_fixed\_vs\_satabs.3 & 178 & 1617 & - & 0.82 & - & 0.80 & - & 0 & 1.31 & - & 0 & 1.25 & - & 0.72 & - & 0 & 0.72 & - & 1.25 \\
-cprover\_software\_analysis/peterson\_vs\_satabs.1 & 31 & 65 & - & 0.16 & - & 0.17 & - & 0 & 0.25 & - & 0 & 0.25 & - & 0.15 & - & 0 & 0.17 & - & 0.29 \\
-cprover\_software\_analysis/peterson\_vs\_satabs.2 & 284 & 1985 & - & 0.95 & - & 0.98 & - & 0 & 1.52 & - & 0 & 1.54 & - & 0.91 & - & 0 & 0.93 & - & 1.54 \\
-cprover\_software\_analysis/pthread5\_vs\_satabs.1 & 80 & 413 & - & 0.35 & - & 0.30 & - & 0 & 0.45 & - & 0 & 0.49 & - & 0.27 & - & 0 & 0.28 & - & 0.47 \\
-cprover\_software\_analysis/pthread5\_vs\_satabs.2 & 80 & 412 & - & 0.34 & - & 0.35 & - & 0 & 0.47 & - & 0 & 0.47 & - & 0.27 & - & 0 & 0.33 & - & 0.46 \\
-cprover\_software\_analysis/pthread5\_vs\_satabs.3 & 187 & 1893 & - & 1.00 & - & 0.96 & - & 0 & 1.48 & - & 0 & 1.49 & - & 0.84 & - & 0 & 0.81 & - & 1.45 \\
-cprover\_software\_analysis/pthread5\_vs\_satabs.4 & 187 & 1889 & - & 1.00 & - & 1.01 & - & 0 & 1.49 & - & 0 & 1.65 & - & 0.82 & - & 0 & 0.82 & - & 1.69 \\
+cprover\_sa/buggy\_spaghetti\_vf\_satabs.1 & 78 & 115 & - & 0.16 & - & 0.16 & - & 0 & 0.23 & - & 0 & 0.23 & - & 0.14 & - & 0 & 0.14 & - & 0 & 0.23 & - & 0 & 0 & 0.23 \\
+cprover\_sa/buggy\_spaghetti\_vf\_satabs.2 & 326 & 697 & - & 0.41 & - & 0.40 & - & 0 & 0.61 & - & 0 & 0.61 & - & 0.36 & - & 0 & 0.36 & - & 0 & 0.60 & - & 0 & 0 & 0.60 \\
+cprover\_sa/conditionals\_vs\_satabs.1 & 50 & 55 & - & 0.14 & - & 0.14 & - & 0 & 0.20 & - & 0 & 0.20 & - & 0.13 & - & 0 & 0.12 & - & 0 & 0.19 & - & 0 & 0 & 0.19 \\
+cprover\_sa/constants\_vf\_satabs.1 & 26 & 25 & - & 0.12 & - & 0.12 & - & 0 & 0.17 & - & 0 & 0.17 & - & 0.11 & - & 0 & 0.12 & - & 0 & 0.17 & - & 0 & 0 & 0.17 \\
+cprover\_sa/constants\_vf\_satabs.2 & 262 & 393 & - & 0.27 & - & 0.27 & - & 0 & 0.41 & - & 0 & 0.42 & - & 0.25 & - & 0 & 0.24 & - & 0 & 0.41 & - & 0 & 0 & 0.42 \\
+cprover\_sa/dekker\_vs\_satabs.1 & 41 & 121 & - & 0.15 & - & 0.16 & - & 0 & 0.23 & - & 0 & 0.24 & - & 0.14 & - & 0 & 0.14 & - & 0 & 0.23 & - & 0 & 0 & 0.23 \\
+cprover\_sa/dekker\_vs\_satabs.2 & 164 & 1345 & - & 0.56 & - & 0.57 & - & 0 & 0.88 & - & 0 & 0.88 & - & 0.50 & - & 0 & 0.50 & - & 0 & 0.88 & - & 0 & 0 & 0.89 \\
+cprover\_sa/double\_lock\_p1\_vs\_satabs.1 & 64 & 205 & - & 0.19 & - & 0.19 & - & 1 & 0.54 & - & 2 & 0.81 & - & 0.17 & - & 0 & 0.17 & - & 1 & 0.54 & - & 1 & 0 & 0.54 \\
+cprover\_sa/double\_lock\_p1\_vs\_satabs.2 & 570 & 7601 & - & 3.29 & - & 10.49 & - & 0 & 4.91 & - & 0 & 12.05 & - & 2.59 & - & 0 & 2.65 & - & 0 & 4.89 & - & 0 & 0 & 4.87 \\
+cprover\_sa/double\_lock\_p1\_vs\_satabs.3 & 570 & 7569 & - & 3.15 & - & 10.34 & - & 0 & 4.70 & - & 0 & 11.86 & - & 2.59 & - & 0 & 2.60 & - & 0 & 4.69 & - & 0 & 0 & 4.68 \\
+cprover\_sa/double\_lock\_p2\_vs\_satabs.1 & 64 & 213 & - & 0.19 & - & 0.20 & - & 0 & 0.28 & - & 1 & 0.57 & - & 0.18 & - & 0 & 0.17 & - & 0 & 0.28 & - & 0 & 0 & 0.28 \\
+cprover\_sa/double\_lock\_p2\_vs\_satabs.2 & 184 & 1833 & - & 0.78 & - & 0.77 & - & 0 & 1.18 & - & 0 & 1.17 & - & 0.67 & - & 0 & 0.66 & - & 0 & 1.18 & - & 0 & 0 & 1.18 \\
+cprover\_sa/double\_lock\_p3\_vs\_satabs.1 & 46 & 81 & - & 0.14 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.21 & - & 0.13 & - & 0 & 0.13 & - & 0 & 0.21 & - & 0 & 0 & 0.21 \\
+cprover\_sa/double\_lock\_p3\_vs\_satabs.2 & 112 & 745 & - & 0.37 & - & 0.38 & - & 1 & 1.06 & - & 0 & 0.58 & - & 0.33 & - & 0 & 0.34 & - & 1 & 1.07 & - & 1 & 0 & 1.07 \\
+cprover\_sa/double\_lock\_p3\_vs\_satabs.3 & 306 & 3137 & - & 1.28 & - & 1.28 & - & 0 & 1.93 & - & 0 & 1.92 & - & 1.08 & - & 0 & 1.09 & - & 0 & 1.92 & - & 0 & 0 & 1.94 \\
+cprover\_sa/lu-fig2\_vs\_satabs.1 & 39 & 37 & - & 0.14 & - & 0.13 & - & 0 & 0.19 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.19 & - & 0 & 0 & 0.19 \\
+cprover\_sa/lu-fig2\_vs\_satabs.2 & 61 & 185 & - & 0.18 & - & 0.18 & - & 0 & 0.27 & - & 0 & 0.27 & - & 0.17 & - & 0 & 0.17 & - & 0 & 0.27 & - & 0 & 0 & 0.27 \\
+cprover\_sa/lu-fig2\_vs\_satabs.3 & 178 & 1617 & - & 0.69 & - & 0.71 & - & 0 & 1.03 & - & 0 & 1.04 & - & 0.59 & - & 0 & 0.60 & - & 0 & 1.03 & - & 0 & 0 & 1.03 \\
+cprover\_sa/peterson\_vs\_satabs.1 & 31 & 65 & - & 0.14 & - & 0.14 & - & 0 & 0.20 & - & 0 & 0.20 & - & 0.13 & - & 0 & 0.13 & - & 0 & 0.21 & - & 0 & 0 & 0.20 \\
+cprover\_sa/peterson\_vs\_satabs.2 & 284 & 1985 & - & 0.81 & - & 0.83 & - & 0 & 1.24 & - & 0 & 1.27 & - & 0.73 & - & 0 & 0.81 & - & 0 & 1.26 & - & 0 & 0 & 1.24 \\
+cprover\_sa/pthread5\_vs\_satabs.1 & 80 & 413 & - & 0.26 & - & 0.26 & - & 0 & 0.38 & - & 0 & 0.41 & - & 0.24 & - & 0 & 0.23 & - & 0 & 0.39 & - & 0 & 0 & 0.39 \\
+cprover\_sa/pthread5\_vs\_satabs.2 & 80 & 412 & - & 0.25 & - & 0.26 & - & 0 & 0.39 & - & 0 & 0.39 & - & 0.24 & - & 0 & 0.23 & - & 0 & 0.38 & - & 0 & 0 & 0.39 \\
+cprover\_sa/pthread5\_vs\_satabs.3 & 187 & 1893 & - & 0.79 & - & 0.79 & - & 0 & 1.21 & - & 0 & 1.20 & - & 0.68 & - & 0 & 0.67 & - & 0 & 1.21 & - & 0 & 0 & 1.22 \\
+cprover\_sa/pthread5\_vs\_satabs.4 & 187 & 1889 & - & 0.78 & - & 0.78 & - & 0 & 1.20 & - & 0 & 1.21 & - & 0.67 & - & 0 & 0.67 & - & 0 & 1.19 & - & 0 & 0 & 1.22 \\
+cprover\_sa/rand\_cas\_vs\_satabs.1 & 48 & 53 & - & 0.13 & - & 0.13 & - & 0 & 0.20 & - & 0 & 0.19 & - & 0.13 & - & 0 & 0.12 & - & 0 & 0.19 & - & 0 & 0 & 0.19 \\
+cprover\_sa/rand\_lock\_p0\_vs\_satabs.1 & 29 & 43 & - & 0.13 & - & 0.14 & - & 0 & 0.19 & - & 0 & 0.19 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.19 & - & 0 & 0 & 0.19 \\
+cprover\_sa/rand\_lock\_p0\_vs\_satabs.2 & 110 & 289 & - & 0.22 & - & 0.22 & - & 0 & 0.33 & - & 0 & 0.34 & - & 0.20 & - & 0 & 0.20 & - & 0 & 0.33 & - & 0 & 0 & 0.33 \\
+cprover\_sa/rand\_lock\_p0\_vs\_satabs.3 & 234 & 773 & - & 0.40 & - & 0.44 & - & 0 & 0.62 & - & 0 & 0.62 & - & 0.36 & - & 0 & 0.36 & - & 0 & 0.61 & - & 0 & 0 & 0.62 \\
+cprover\_sa/simple\_loop5\_vs\_satabs.1 & 31 & 49 & - & 0.13 & - & 0.13 & - & 0 & 0.19 & - & 0 & 0.19 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.19 & - & 0 & 0 & 0.19 \\
+cprover\_sa/simple\_loop5\_vs\_satabs.2 & 93 & 485 & - & 0.28 & - & 0.28 & - & 0 & 0.42 & - & 0 & 0.42 & - & 0.25 & - & 0 & 0.25 & - & 0 & 0.44 & - & 0 & 0 & 0.43 \\
+cprover\_sa/spin2003\_vs\_satabs.1 & 27 & 35 & - & 0.13 & - & 0.13 & - & 0 & 0.19 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.18 & - & 0 & 0 & 0.18 \\
+cprover\_sa/spin2003\_vs\_satabs.2 & 56 & 181 & - & 0.18 & - & 0.18 & - & 0 & 0.26 & - & 1 & 0.51 & - & 0.16 & - & 0 & 0.17 & - & 0 & 0.27 & - & 0 & 0 & 0.26 \\
+cprover\_sa/stack\_cas\_p0\_vs\_satabs.1 & 41 & 71 & - & 0.14 & - & 0.14 & - & 0 & 0.20 & - & 0 & 0.20 & - & 0.13 & - & 0 & 0.14 & - & 0 & 0.20 & - & 0 & 0 & 0.20 \\
+cprover\_sa/stack\_cas\_p0\_vs\_satabs.2 & 78 & 157 & - & 0.17 & - & 0.18 & - & 0 & 0.25 & - & 0 & 0.25 & - & 0.16 & - & 0 & 0.16 & - & 0 & 0.25 & - & 0 & 0 & 0.25 \\
+cprover\_sa/stack\_cas\_p0\_vs\_satabs.3 & 158 & 345 & - & 0.24 & - & 0.24 & - & 0 & 0.37 & - & 0 & 0.37 & - & 0.22 & - & 0 & 0.22 & - & 0 & 0.37 & - & 0 & 0 & 0.37 \\
+cprover\_sa/stack\_lock\_p0\_vs\_satabs.1 & 37 & 59 & - & 0.15 & - & 0.14 & - & 0 & 0.20 & - & 0 & 0.20 & - & 0.12 & - & 0 & 0.13 & - & 0 & 0.21 & - & 0 & 0 & 0.20 \\
+cprover\_sa/stack\_lock\_p0\_vs\_satabs.2 & 106 & 633 & - & 0.33 & - & 0.34 & - & 0 & 0.50 & - & 0 & 0.50 & - & 0.30 & - & 0 & 0.30 & - & 0 & 0.50 & - & 0 & 0 & 0.50 \\
+cprover\_sa/szymanski\_vs\_satabs.1 & 61 & 225 & - & 0.19 & - & 0.19 & - & 0 & 0.28 & - & 0 & 0.29 & - & 0.17 & - & 0 & 0.17 & - & 0 & 0.28 & - & 0 & 0 & 0.29 \\
+cprover\_sa/szymanski\_vs\_satabs.2 & 568 & 8321 & - & 3.09 & - & 10.58 & - & 0 & 4.73 & - & 0 & 12.24 & - & 2.74 & - & 0 & 2.77 & - & 0 & 4.76 & - & 0 & 0 & 4.73 \\
     \hline
@@
 \end{center}
-\caption{Experimental results for the benchmarks of \bfc}
+\caption{Experimental results for the benchmarks of \bfc (continued)}
 \label{bfc-experiments-2}
-\end{table}
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
-    \hline
-           \multicolumn{3}{|c|}{Network}
-         & \multicolumn{2}{|c|}{Safety}
-         & \multicolumn{2}{|c|}{Safety-Int}
-         & \multicolumn{3}{|c|}{Refinement}
-         & \multicolumn{3}{|c|}{Refinement-Int}
-         & \multicolumn{2}{|c|}{Invariant}
-         & \multicolumn{3}{|c|}{Inv-Min}
-         & \multicolumn{2}{|c|}{Inv-Ref} \\
-    Name & P & T
-        & R & T
-        & R & T
-        & R & I & T
-        & R & I & T
-        & R & T
-        & R & I & T
-        & R & T \\
-    \hline
-        \multicolumn{20}{|c|}{Negative Benchmarks (continued)} \\
-    \hline
-cprover\_software\_analysis/rand\_cas\_vs\_satabs.1 & 48 & 53 & - & 0.16 & - & 0.20 & - & 0 & 0.25 & - & 0 & 0.24 & - & 0.15 & - & 0 & 0.16 & - & 0.25 \\
-cprover\_software\_analysis/rand\_lock\_p0\_vs\_satabs.1 & 29 & 43 & - & 0.16 & - & 0.16 & - & 0 & 0.25 & - & 0 & 0.23 & - & 0.15 & - & 0 & 0.17 & - & 0.24 \\
-cprover\_software\_analysis/rand\_lock\_p0\_vs\_satabs.2 & 110 & 289 & - & 0.26 & - & 0.31 & - & 0 & 0.43 & - & 0 & 0.39 & - & 0.23 & - & 0 & 0.25 & - & 0.47 \\
-cprover\_software\_analysis/rand\_lock\_p0\_vs\_satabs.3 & 234 & 773 & - & 0.50 & - & 0.52 & - & 0 & 0.81 & - & 0 & 0.78 & - & 0.43 & - & 0 & 0.43 & - & 0.74 \\
-cprover\_software\_analysis/simple\_loop5\_vs\_satabs.1 & 31 & 49 & - & 0.16 & - & 0.20 & - & 0 & 0.22 & - & 0 & 0.24 & - & 0.19 & - & 0 & 0.16 & - & 0.24 \\
-cprover\_software\_analysis/simple\_loop5\_vs\_satabs.2 & 93 & 485 & - & 0.37 & - & 0.39 & - & 0 & 0.52 & - & 0 & 0.55 & - & 0.31 & - & 0 & 0.30 & - & 0.54 \\
-cprover\_software\_analysis/spin2003\_vs\_satabs.1 & 27 & 35 & - & 0.19 & - & 0.17 & - & 0 & 0.26 & - & 0 & 0.24 & - & 0.15 & - & 0 & 0.18 & - & 0.24 \\
-cprover\_software\_analysis/spin2003\_vs\_satabs.2 & 56 & 181 & - & 0.21 & - & 0.22 & - & 0 & 0.33 & - & 1 & 0.67 & - & 0.19 & - & 0 & 0.24 & - & 0.33 \\
-cprover\_software\_analysis/stack\_cas\_p0\_vs\_satabs.1 & 41 & 71 & - & 0.17 & - & 0.20 & - & 0 & 0.26 & - & 0 & 0.30 & - & 0.15 & - & 0 & 0.16 & - & 0.26 \\
-cprover\_software\_analysis/stack\_cas\_p0\_vs\_satabs.2 & 78 & 157 & - & 0.21 & - & 0.23 & - & 0 & 0.32 & - & 0 & 0.30 & - & 0.20 & - & 0 & 0.18 & - & 0.32 \\
-cprover\_software\_analysis/stack\_cas\_p0\_vs\_satabs.3 & 158 & 345 & - & 0.31 & - & 0.33 & - & 0 & 0.47 & - & 0 & 0.51 & - & 0.31 & - & 0 & 0.25 & - & 0.44 \\
-cprover\_software\_analysis/stack\_lock\_p0\_vs\_satabs.1 & 37 & 59 & - & 0.16 & - & 0.16 & - & 0 & 0.26 & - & 0 & 0.23 & - & 0.16 & - & 0 & 0.16 & - & 0.27 \\
-cprover\_software\_analysis/stack\_lock\_p0\_vs\_satabs.2 & 106 & 633 & - & 0.42 & - & 0.39 & - & 0 & 0.62 & - & 0 & 0.64 & - & 0.35 & - & 0 & 0.35 & - & 0.65 \\
-cprover\_software\_analysis/szymanski\_vs\_satabs.1 & 61 & 225 & - & 0.25 & - & 0.24 & - & 0 & 0.35 & - & 0 & 0.35 & - & 0.21 & - & 0 & 0.23 & - & 0.37 \\
-cprover\_software\_analysis/szymanski\_vs\_satabs.2 & 568 & 8321 & - & 3.94 & - & 11.44 & - & 0 & 5.80 & - & 0 & 13.53 & - & 3.35 & - & 0 & 3.38 & - & 5.85 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \bfc}
-\label{bfc-experiments-3}
 \end{table}
@@
 In the case of \footnotemark[1], the high runtime and timeouts
-for the invariant come from the special form of the property,
-as the negation contains a lot of disjunctions.
+for the invariant come from the special form of the property.
+The negation contains 8989 disjunctions, which have to be solved separately.
 
@@
 \begin{center}
-  \begin{tabular}{ | l | r | r | *{17}{ r | } } %
+  \begin{tabular}{ | l | r | r | *{22}{ r | } } %
     \hline
@@
         & R & T
+        & R & M & T
         & R & I & T
-        & R & T \\
+        & R & I & M & T \\
     \hline
-        \multicolumn{20}{|c|}{Positive Benchmarks} \\
+        \multicolumn{25}{|c|}{Positive Benchmarks} \\
     \hline
-PN/MultiME & 12 & 11 & - & 0.16 & - & 0.14 & + & 1 & 0.34 & + & 1 & 0.32 & - & 0.13 & - & 0 & 0.12 & + & 0.93 \\
-PN/basicME & 5 & 5 & - & 0.12 & - & 0.14 & + & 1 & 0.32 & + & 1 & 0.29 & - & 0.12 & - & 0 & 0.12 & + & 0.86 \\
-PN/bingham\_h150 & 153 & 302 & + & 0.22 & + & 0.22 & + & 0 & 0.24 & + & 0 & 0.22 & + & 0.41 & + & 0 & 0.50 & + & 0.61 \\
-PN/bingham\_h25 & 28 & 52 & + & 0.13 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.35 & + & 0.37 \\
-PN/bingham\_h250 & 253 & 502 & + & 0.32 & + & 0.31 & + & 0 & 0.30 & + & 0 & 0.32 & + & 0.50 & + & 0 & 0.70 & + & 0.83 \\
-PN/bingham\_h250\_attic \footnotemark[1] & 253 & 502 & + & 11.52 & + & 18.09 & + & 0 & 11.74 & + & 0 & 18.74 & + & 6535.28 & TO & TO & TO & TO & TO \\
-PN/bingham\_h50 & 53 & 102 & + & 0.17 & + & 0.15 & + & 0 & 0.17 & + & 0 & 0.17 & + & 0.30 & + & 0 & 0.38 & + & 0.40 \\
-PN/csm & 14 & 14 & + & 0.14 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.14 & + & 0.22 & + & 0 & 0.29 & + & 0.36 \\
-PN/extendedread-write-smallconsts & 24 & 23 & - & 0.14 & - & 0.15 & - & 0 & 0.23 & - & 0 & 0.20 & - & 0.14 & - & 0 & 0.14 & - & 0.22 \\
-PN/extendedread-write & 24 & 23 & - & 0.15 & - & 0.14 & - & 0 & 0.21 & - & 0 & 0.20 & - & 0.13 & - & 0 & 0.13 & - & 0.22 \\
-PN/fms & 22 & 23 & + & 0.12 & + & 0.15 & + & 0 & 0.14 & + & 0 & 0.13 & + & 0.23 & + & 0 & 0.30 & + & 0.38 \\
-PN/fms\_attic & 22 & 23 & + & 0.13 & + & 0.14 & + & 0 & 0.12 & + & 0 & 0.13 & + & 0.42 & + & 0 & 0.55 & + & 0.54 \\
-PN/manufacturing & 13 & 6 & - & 0.15 & - & 0.15 & - & 0 & 0.21 & - & 0 & 0.19 & - & 0.12 & - & 0 & 0.13 & - & 0.19 \\
-PN/mesh2x2 & 32 & 36 & + & 0.14 & + & 0.13 & + & 0 & 0.17 & + & 0 & 0.14 & + & 0.25 & + & 0 & 0.30 & + & 0.36 \\
-PN/mesh3x2 & 52 & 60 & + & 0.18 & + & 0.17 & + & 0 & 0.14 & + & 0 & 0.15 & + & 0.26 & + & 0 & 0.35 & + & 0.39 \\
-PN/multipool & 18 & 25 & + & 0.15 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.13 & + & 0.25 & + & 0 & 0.31 & + & 0.38 \\
-PN/pingpong & 6 & 6 & - & 0.15 & - & 0.16 & + & 1 & 0.32 & + & 1 & 0.31 & - & 0.13 & - & 0 & 0.13 & + & 0.56 \\
-boundedPN/kanban & 16 & 16 & + & 0.15 & + & 0.15 & + & 0 & 0.15 & + & 0 & 0.14 & + & 0.25 & + & 0 & 0.30 & + & 0.38 \\
-boundedPN/lamport & 11 & 9 & - & 0.17 & - & 0.13 & + & 1 & 0.34 & + & 1 & 0.31 & - & 0.13 & - & 0 & 0.14 & + & 0.58 \\
-boundedPN/newdekker & 16 & 14 & - & 0.15 & - & 0.18 & + & 1 & 0.34 & + & 1 & 0.37 & - & 0.12 & - & 0 & 0.15 & + & 0.55 \\
-boundedPN/newrtp & 9 & 12 & + & 0.14 & + & 0.12 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.23 & + & 0 & 0.30 & + & 0.38 \\
-boundedPN/peterson & 14 & 12 & - & 0.13 & - & 0.17 & + & 2 & 0.57 & + & 2 & 0.55 & - & 0.13 & - & 0 & 0.15 & + & 0.75 \\
-boundedPN/read-write & 13 & 9 & + & 0.14 & + & 0.13 & + & 0 & 0.16 & + & 0 & 0.13 & + & 0.22 & + & 0 & 0.33 & + & 0.38 \\
+PN/MultiME & 12 & 11 & - & 0.12 & - & 0.12 & + & 1 & 0.27 & + & 1 & 0.27 & - & 0.11 & - & 0 & 0.11 & + & 1 & 0.76 & + & 1 & 0 & 0.92 \\
+PN/basicME & 5 & 5 & - & 0.12 & - & 0.11 & + & 1 & 0.26 & + & 1 & 0.26 & - & 0.11 & - & 0 & 0.12 & + & 1 & 0.73 & + & 1 & 0 & 0.90 \\
+PN/bingham\_h150 & 153 & 302 & + & 0.21 & + & 0.21 & + & 0 & 0.21 & + & 0 & 0.21 & + & 0.34 & + & 0 & 0.46 & + & 0 & 0.53 & + & 0 & 0 & 0.65 \\
+PN/bingham\_h25 & 28 & 52 & + & 0.12 & + & 0.12 & + & 0 & 0.12 & + & 0 & 0.12 & + & 0.22 & + & 0 & 0.28 & + & 0 & 0.32 & + & 0 & 0 & 0.39 \\
+PN/bingham\_h250 & 253 & 502 & + & 0.28 & + & 0.29 & + & 0 & 0.28 & + & 0 & 0.28 & + & 0.46 & + & 0 & 0.61 & + & 0 & 0.71 & + & 0 & 0 & 0.87 \\
+PN/bingham\_h250\_attic \footnotemark[1] & 253 & 502 & + & 11.52 & + & 18.09 & + & 0 & 11.74 & + & 0 & 18.74 & + & 6535.28 & TO & TO & TO & TO & TO & TO & TO & TO & TO & TO \\
+PN/bingham\_h50 & 53 & 102 & + & 0.14 & + & 0.14 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.31 & + & 0 & 0.36 & + & 0 & 0 & 0.44 \\
+PN/csm & 14 & 14 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.26 & + & 0 & 0.30 & + & 0 & 0 & 0.35 \\
+PN/extendedread-write-smallconsts & 24 & 23 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.19 & - & 0 & 0 & 0.18 \\
+PN/extendedread-write & 24 & 23 & - & 0.13 & - & 0.14 & - & 0 & 0.18 & - & 0 & 0.20 & - & 0.12 & - & 0 & 0.13 & - & 0 & 0.20 & - & 0 & 0 & 0.18 \\
+PN/fms & 22 & 23 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.13 & + & 0.22 & + & 0 & 0.26 & + & 0 & 0.30 & + & 0 & 0 & 0.36 \\
+PN/fms\_attic & 22 & 23 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.35 & + & 0 & 0.47 & + & 0 & 0.45 & + & 0 & 0 & 0.58 \\
+PN/manufacturing & 13 & 6 & - & 0.12 & - & 0.12 & - & 0 & 0.16 & - & 0 & 0.16 & - & 0.11 & - & 0 & 0.11 & - & 0 & 0.16 & - & 0 & 0 & 0.17 \\
+PN/mesh2x2 & 32 & 36 & + & 0.12 & + & 0.12 & + & 0 & 0.12 & + & 0 & 0.12 & + & 0.21 & + & 0 & 0.27 & + & 0 & 0.31 & + & 0 & 0 & 0.38 \\
+PN/mesh3x2 & 52 & 60 & + & 0.13 & + & 0.13 & + & 0 & 0.12 & + & 0 & 0.13 & + & 0.22 & + & 0 & 0.30 & + & 0 & 0.41 & + & 0 & 0 & 0.47 \\
+PN/multipool & 18 & 25 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.26 & + & 0 & 0.30 & + & 0 & 0 & 0.36 \\
+PN/pingpong & 6 & 6 & - & 0.11 & - & 0.11 & + & 1 & 0.26 & + & 1 & 0.26 & - & 0.11 & - & 0 & 0.11 & + & 1 & 0.44 & + & 1 & 0 & 0.50 \\
+boundedPN/kanban & 16 & 16 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.26 & + & 0 & 0.29 & + & 0 & 0 & 0.35 \\
+boundedPN/lamport & 11 & 9 & - & 0.11 & - & 0.12 & + & 1 & 0.28 & + & 1 & 0.28 & - & 0.11 & - & 0 & 0.11 & + & 1 & 0.46 & + & 1 & 0 & 0.51 \\
+boundedPN/newdekker & 16 & 14 & - & 0.12 & - & 0.12 & + & 1 & 0.27 & + & 1 & 0.28 & - & 0.11 & - & 0 & 0.11 & + & 1 & 0.48 & + & 1 & 0 & 0.51 \\
+boundedPN/newrtp & 9 & 12 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.25 & + & 0 & 0.29 & + & 0 & 0 & 0.34 \\
+boundedPN/peterson & 14 & 12 & - & 0.12 & - & 0.12 & + & 2 & 0.45 & + & 2 & 0.45 & - & 0.11 & - & 0 & 0.11 & + & 2 & 0.62 & + & 2 & 0 & 0.67 \\
+boundedPN/read-write & 13 & 9 & + & 0.11 & + & 0.11 & + & 0 & 0.11 & + & 0 & 0.11 & + & 0.20 & + & 0 & 0.25 & + & 0 & 0.29 & + & 0 & 0 & 0.35 \\
     \hline
-        \multicolumn{20}{|c|}{Negative Benchmarks} \\
+        \multicolumn{25}{|c|}{Negative Benchmarks} \\
     \hline
-PN/kanban & 16 & 20 & - & 0.14 & - & 0.15 & - & 0 & 0.21 & - & 0 & 0.22 & - & 0.13 & - & 0 & 0.13 & - & 0.19 \\
-PN/leabasicapproach & 16 & 14 & - & 0.13 & - & 0.17 & - & 0 & 0.21 & - & 0 & 0.28 & - & 0.19 & - & 0 & 0.16 & - & 0.20 \\
-PN/pncsacover & 31 & 36 & - & 0.17 & - & 0.18 & - & 0 & 0.23 & - & 0 & 0.22 & - & 0.15 & - & 0 & 0.13 & - & 0.25 \\
-PN/pncsasemiliv & 31 & 36 & - & 0.17 & - & 0.14 & - & 0 & 0.23 & - & 0 & 0.23 & - & 0.15 & - & 0 & 0.14 & - & 0.21 \\
+PN/kanban & 16 & 20 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.17 & - & 0.11 & - & 0 & 0.11 & - & 0 & 0.17 & - & 0 & 0 & 0.17 \\
+PN/leabasicapproach & 16 & 14 & - & 0.12 & - & 0.12 & - & 0 & 0.17 & - & 0 & 0.17 & - & 0.12 & - & 0 & 0.11 & - & 0 & 0.17 & - & 0 & 0 & 0.18 \\
+PN/pncsacover & 31 & 36 & - & 0.13 & - & 0.13 & - & 0 & 0.18 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.18 & - & 0 & 0 & 0.18 \\
+PN/pncsasemiliv & 31 & 36 & - & 0.12 & - & 0.12 & - & 0 & 0.18 & - & 0 & 0.18 & - & 0.12 & - & 0 & 0.12 & - & 0 & 0.18 & - & 0 & 0 & 0.18 \\
     \hline
@@
 
-\begin{verbatim}
+\iffalse
 TODO: Use fixed or non-fixed benchmark?  Source for list of positives and negatives?
-\end{verbatim}
+\fi
 
@@
         & R & T
+        & R & M & T
         & R & I & T
-        & R & T \\
+        & R & I & M & T \\
     \hline
-        \multicolumn{20}{|c|}{Positive Benchmarks} \\
+        \multicolumn{25}{|c|}{Positive Benchmarks} \\
     \hline
-concdb\_\_single\_client\_writes\_\_depth\_0\_\_fixed & 570 & 717 & + & 0.52 & + & 0.50 & + & 0 & 0.54 & + & 0 & 0.53 & + & 0.79 & + & 0 & 1.35 & + & 1.26 \\
-finite\_leader\_\_single\_leader\_\_depth\_1\_\_fixed & 454 & 447 & + & 0.43 & + & 0.40 & + & 0 & 0.45 & + & 0 & 0.37 & + & 0.60 & + & 0 & 1.52 & + & 1.00 \\
-finite\_leader\_\_single\_leader\_\_depth\_2\_\_fixed & 454 & 447 & + & 0.38 & + & 0.38 & + & 0 & 0.43 & + & 0 & 0.42 & + & 0.65 & + & 0 & 1.54 & + & 1.00 \\
-firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_1\_\_fixed & 780 & 1024 & + & 0.64 & + & 0.66 & + & 0 & 0.67 & + & 0 & 0.68 & + & 0.92 & + & 0 & 2.46 & + & 1.61 \\
-firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_2\_\_fixed & 4014 & 6938 & + & 8.05 & + & 8.75 & + & 0 & 8.71 & + & 0 & 8.07 & + & 9.38 & + & 0 & 40.62 & + & 17.59 \\
-parikh\_\_should\_already\_be\_initialized\_\_depth\_0\_\_fixed & 77 & 70 & + & 0.17 & + & 0.18 & + & 0 & 0.21 & + & 0 & 0.18 & + & 0.28 & + & 0 & 0.40 & + & 0.45 \\
-parikh\_\_should\_already\_be\_initialized\_\_depth\_1\_\_fixed & 88 & 82 & + & 0.17 & + & 0.17 & + & 0 & 0.16 & + & 0 & 0.17 & + & 0.31 & + & 0 & 0.40 & + & 0.44 \\
-parikh\_\_should\_already\_be\_initialized\_\_depth\_2\_\_fixed & 88 & 82 & + & 0.20 & + & 0.17 & + & 0 & 0.16 & + & 0 & 0.18 & + & 0.29 & + & 0 & 0.42 & + & 0.44 \\
-pipe\_\_single\_message\_in\_mailbox\_\_depth\_0\_\_fixed & 212 & 236 & + & 0.25 & + & 0.29 & + & 0 & 0.25 & + & 0 & 0.26 & + & 0.40 & + & 0 & 0.55 & + & 0.65 \\
-reslock\_\_critical\_\_depth\_0\_\_fixed & 364 & 395 & + & 0.35 & + & 0.34 & + & 0 & 0.36 & + & 0 & 0.32 & + & 0.53 & + & 0 & 1.00 & + & 0.87 \\
-reslock\_\_critical\_\_depth\_1\_\_fixed & 600 & 702 & + & 0.53 & + & 0.52 & + & 0 & 0.51 & + & 0 & 0.51 & + & 0.79 & + & 0 & 2.15 & + & 1.28 \\
-reslock\_\_critical\_\_depth\_2\_\_fixed & 841 & 1030 & + & 0.77 & + & 0.72 & + & 0 & 0.84 & + & 0 & 0.77 & + & 1.04 & + & 0 & 3.40 & + & 1.81 \\
-reslockbeh\_\_critical\_\_depth\_0\_\_fixed & 734 & 977 & + & 0.87 & + & 0.70 & + & 0 & 0.75 & + & 0 & 0.82 & + & 1.10 & + & 0 & 2.82 & + & 2.03 \\
-reslockbeh\_\_critical\_\_depth\_1\_\_fixed & 4461 & 187 & + & 0.97 & + & 0.82 & + & 0 & 0.81 & + & 0 & 0.76 & + & 1.21 & + & 0 & 46.97 & + & 1.92 \\
-ring\_\_single\_message\_in\_mailbox\_\_depth\_0\_\_fixed & 387 & 453 & + & 0.38 & + & 0.36 & + & 0 & 0.37 & + & 0 & 0.40 & + & 0.57 & + & 0 & 1.06 & + & 0.94 \\
-safe\_send\_\_sending\_to\_non-pid\_1\_\_depth\_1\_\_fixed & 279 & 285 & - & 0.33 & - & 0.32 & + & 1 & 0.75 & + & 1 & 0.74 & - & 0.27 & - & 0 & 0.30 & + & 1.22 \\
-safe\_send\_\_sending\_to\_non-pid\_1\_\_depth\_2\_\_fixed & 279 & 285 & - & 0.34 & - & 0.29 & + & 1 & 0.82 & + & 1 & 0.72 & - & 0.27 & - & 0 & 0.28 & + & 1.21 \\
-safe\_send\_\_sending\_to\_non-pid\_2\_\_depth\_1\_\_fixed & 279 & 285 & - & 0.33 & - & 0.30 & + & 1 & 0.74 & + & 1 & 0.76 & - & 0.27 & - & 0 & 0.26 & + & 1.20 \\
-safe\_send\_\_sending\_to\_non-pid\_2\_\_depth\_2\_\_fixed & 279 & 285 & - & 0.30 & - & 0.32 & + & 1 & 0.74 & + & 1 & 0.72 & - & 0.27 & - & 0 & 0.27 & + & 1.17 \\
-safe\_send\_\_sending\_to\_non-pid\_3\_\_depth\_1\_\_fixed & 279 & 285 & + & 0.27 & + & 0.29 & + & 0 & 0.28 & + & 0 & 0.28 & + & 0.45 & + & 0 & 0.69 & + & 0.73 \\
-safe\_send\_\_sending\_to\_non-pid\_3\_\_depth\_2\_\_fixed & 279 & 285 & + & 0.27 & + & 0.27 & + & 0 & 0.32 & + & 0 & 0.25 & + & 0.42 & + & 0 & 0.72 & + & 0.69 \\
-safe\_send\_\_sending\_to\_non-pid\_4\_\_depth\_1\_\_fixed & 279 & 285 & - & 0.30 & - & 0.34 & - & 5 & 2.56 & + & 3 & 1.64 & - & 0.28 & - & 0 & 0.27 & - & 2.58 \\
-safe\_send\_\_sending\_to\_non-pid\_4\_\_depth\_2\_\_fixed & 279 & 285 & - & 0.31 & - & 0.30 & - & 5 & 2.65 & + & 3 & 1.61 & - & 0.27 & - & 0 & 0.27 & - & 2.58 \\
-sieve\_\_single\_message\_in\_counter\_mailbox\_\_depth\_0\_\_fixed & 263 & 275 & + & 0.31 & + & 0.26 & + & 0 & 0.30 & + & 0 & 0.29 & + & 0.43 & + & 0 & 0.70 & + & 0.75 \\
-sieve\_\_single\_message\_in\_filter\_mailbox\_\_depth\_0\_\_fixed & 263 & 276 & + & 0.29 & + & 0.27 & + & 0 & 0.29 & + & 0 & 0.29 & + & 0.43 & + & 0 & 0.69 & + & 0.71 \\
-sieve\_\_single\_message\_in\_sieve\_mailbox\_\_depth\_0\_\_fixed & 263 & 274 & + & 0.32 & + & 0.31 & + & 0 & 0.29 & + & 0 & 0.28 & + & 0.43 & + & 0 & 0.71 & + & 0.69 \\
-state\_factory\_\_after\_receive\_if\_no\_mail\_\_depth\_0\_\_fixed & 542 & 648 & + & 0.52 & + & 0.49 & + & 0 & 0.54 & + & 0 & 0.49 & + & 0.74 & + & 0 & 1.41 & + & 1.19 \\
-state\_factory\_\_single\_message\_in\_mailbox\_\_depth\_0\_\_fixed & 541 & 647 & + & 0.51 & + & 0.47 & + & 0 & 0.50 & + & 0 & 0.49 & + & 0.76 & + & 0 & 1.31 & + & 1.37 \\
+concdb\_\_single\_client\_writes\_\_depth\_0\ & 570 & 717 & + & 0.44 & + & 0.42 & + & 0 & 0.44 & + & 0 & 0.42 & + & 0.64 & + & 0 & 1.11 & + & 0 & 1.05 & + & 0 & 0 & 1.52 \\
+finite\_leader\_\_single\_leader\_\_depth\_1\ & 454 & 447 & + & 0.33 & + & 0.31 & + & 0 & 0.33 & + & 0 & 0.31 & + & 0.50 & + & 0 & 1.22 & + & 0 & 0.81 & + & 0 & 0 & 1.53 \\
+finite\_leader\_\_single\_leader\_\_depth\_2\ & 454 & 447 & + & 0.34 & + & 0.31 & + & 0 & 0.33 & + & 0 & 0.31 & + & 0.52 & + & 0 & 1.22 & + & 0 & 0.81 & + & 0 & 0 & 1.54 \\
+firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_1\ & 780 & 1024 & + & 0.54 & + & 0.54 & + & 0 & 0.54 & + & 0 & 0.52 & + & 0.78 & + & 0 & 2.02 & + & 0 & 1.29 & + & 0 & 0 & 2.54 \\
+firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_2\ & 4014 & 6938 & + & 6.56 & + & 6.53 & + & 0 & 6.59 & + & 0 & 6.60 & + & 7.66 & + & 0 & 34.03 & + & 0 & 14.22 & + & 0 & 0 & 40.53 \\
+parikh\_\_should\_already\_be\_initialized\_\_depth\_0\ & 77 & 70 & + & 0.13 & + & 0.13 & + & 0 & 0.13 & + & 0 & 0.13 & + & 0.23 & + & 0 & 0.32 & + & 0 & 0.36 & + & 0 & 0 & 0.44 \\
+parikh\_\_should\_already\_be\_initialized\_\_depth\_1\ & 88 & 82 & + & 0.14 & + & 0.14 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.32 & + & 0 & 0.38 & + & 0 & 0 & 0.45 \\
+parikh\_\_should\_already\_be\_initialized\_\_depth\_2\ & 88 & 82 & + & 0.14 & + & 0.14 & + & 0 & 0.14 & + & 0 & 0.14 & + & 0.24 & + & 0 & 0.32 & + & 0 & 0.38 & + & 0 & 0 & 0.45 \\
+pipe\_\_single\_message\_in\_mailbox\_\_depth\_0\ & 212 & 236 & + & 0.21 & + & 0.20 & + & 0 & 0.20 & + & 0 & 0.20 & + & 0.33 & + & 0 & 0.46 & + & 0 & 0.51 & + & 0 & 0 & 0.65 \\
+reslock\_\_critical\_\_depth\_0\ & 364 & 395 & + & 0.28 & + & 0.27 & + & 0 & 0.28 & + & 0 & 0.27 & + & 0.43 & + & 0 & 0.75 & + & 0 & 0.69 & + & 0 & 0 & 1.03 \\
+reslock\_\_critical\_\_depth\_1\ & 600 & 702 & + & 0.43 & + & 0.41 & + & 0 & 0.43 & + & 0 & 0.42 & + & 0.61 & + & 0 & 1.67 & + & 0 & 1.02 & + & 0 & 0 & 2.10 \\
+reslock\_\_critical\_\_depth\_2\ & 841 & 1030 & + & 0.63 & + & 0.58 & + & 0 & 0.63 & + & 0 & 0.59 & + & 0.81 & + & 0 & 2.58 & + & 0 & 1.44 & + & 0 & 0 & 3.21 \\
+reslockbeh\_\_critical\_\_depth\_0\ & 734 & 977 & + & 0.60 & + & 0.56 & + & 0 & 0.60 & + & 0 & 0.57 & + & 0.77 & + & 0 & 1.97 & + & 0 & 1.35 & + & 0 & 0 & 2.59 \\
+reslockbeh\_\_critical\_\_depth\_1\ & 4461 & 187 & + & 0.63 & + & 0.60 & + & 0 & 0.63 & + & 0 & 0.60 & + & 0.90 & + & 0 & 37.72 & + & 0 & 1.54 & + & 0 & 0 & 38.54 \\
+ring\_\_single\_message\_in\_mailbox\_\_depth\_0\ & 387 & 453 & + & 0.30 & + & 0.30 & + & 0 & 0.30 & + & 0 & 0.31 & + & 0.45 & + & 0 & 0.81 & + & 0 & 0.74 & + & 0 & 0 & 1.10 \\
+safe\_send\_\_sending\_to\_non-pid\_1\_\_depth\_1\ & 279 & 285 & - & 0.26 & - & 0.25 & + & 1 & 0.60 & + & 1 & 0.57 & - & 0.23 & - & 0 & 0.23 & + & 1 & 0.96 & + & 1 & 2 & 2.45 \\
+safe\_send\_\_sending\_to\_non-pid\_1\_\_depth\_2\ & 279 & 285 & - & 0.26 & - & 0.25 & + & 1 & 0.60 & + & 1 & 0.57 & - & 0.23 & - & 0 & 0.23 & + & 1 & 0.97 & + & 1 & 2 & 2.45 \\
+safe\_send\_\_sending\_to\_non-pid\_2\_\_depth\_1\ & 279 & 285 & - & 0.25 & - & 0.25 & + & 1 & 0.59 & + & 1 & 0.57 & - & 0.22 & - & 0 & 0.22 & + & 1 & 0.97 & + & 1 & 1 & 1.67 \\
+safe\_send\_\_sending\_to\_non-pid\_2\_\_depth\_2\ & 279 & 285 & - & 0.25 & - & 0.24 & + & 1 & 0.60 & + & 1 & 0.57 & - & 0.22 & - & 0 & 0.23 & + & 1 & 0.96 & + & 1 & 1 & 1.66 \\
+safe\_send\_\_sending\_to\_non-pid\_3\_\_depth\_1\ & 279 & 285 & + & 0.23 & + & 0.22 & + & 0 & 0.22 & + & 0 & 0.22 & + & 0.35 & + & 0 & 0.57 & + & 0 & 0.58 & + & 0 & 0 & 0.79 \\
+safe\_send\_\_sending\_to\_non-pid\_3\_\_depth\_2\ & 279 & 285 & + & 0.22 & + & 0.22 & + & 0 & 0.23 & + & 0 & 0.22 & + & 0.36 & + & 0 & 0.56 & + & 0 & 0.56 & + & 0 & 0 & 0.77 \\
+safe\_send\_\_sending\_to\_non-pid\_4\_\_depth\_1\ & 279 & 285 & - & 0.24 & - & 0.25 & - & 5 & 2.14 & + & 3 & 1.26 & - & 0.22 & - & 0 & 0.22 & - & 5 & 2.14 & - & 5 & 0 & 2.13 \\
+safe\_send\_\_sending\_to\_non-pid\_4\_\_depth\_2\ & 279 & 285 & - & 0.25 & - & 0.25 & - & 5 & 2.13 & + & 3 & 1.25 & - & 0.22 & - & 0 & 0.22 & - & 5 & 2.12 & - & 5 & 0 & 2.15 \\
+sieve\_\_single\_message\_in\_counter\_mailbox\_\_depth\_0\ & 263 & 275 & + & 0.23 & + & 0.22 & + & 0 & 0.22 & + & 0 & 0.22 & + & 0.36 & + & 0 & 0.57 & + & 0 & 0.57 & + & 0 & 0 & 0.78 \\
+sieve\_\_single\_message\_in\_filter\_mailbox\_\_depth\_0\ & 263 & 276 & + & 0.23 & + & 0.22 & + & 0 & 0.23 & + & 0 & 0.22 & + & 0.35 & + & 0 & 0.56 & + & 0 & 0.57 & + & 0 & 0 & 0.77 \\
+sieve\_\_single\_message\_in\_sieve\_mailbox\_\_depth\_0\ & 263 & 274 & + & 0.22 & + & 0.22 & + & 0 & 0.23 & + & 0 & 0.22 & + & 0.36 & + & 0 & 0.57 & + & 0 & 0.58 & + & 0 & 0 & 0.78 \\
+state\_factory\_\_after\_receive\_if\_no\_mail\_\_depth\_0\ & 542 & 648 & + & 0.42 & + & 0.40 & + & 0 & 0.42 & + & 0 & 0.40 & + & 0.58 & + & 0 & 1.14 & + & 0 & 0.99 & + & 0 & 0 & 1.57 \\
+state\_factory\_\_single\_message\_in\_mailbox\_\_depth\_0\ & 541 & 647 & + & 0.42 & + & 0.40 & + & 0 & 0.42 & + & 0 & 0.39 & + & 0.58 & + & 0 & 0.96 & + & 0 & 0.99 & + & 0 & 0 & 1.37 \\
     \hline
-        \multicolumn{20}{|c|}{Negative Benchmarks} \\
+        \multicolumn{25}{|c|}{Negative Benchmarks} \\
     \hline
-finite\_leader\_\_single\_leader\_\_depth\_0\_\_fixed & 297 & 308 & - & 0.36 & - & 0.32 & - & 1 & 0.90 & - & 1 & 0.98 & - & 0.29 & - & 0 & 0.30 & - & 0.98 \\
-firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_0\_\_fixed & 352 & 400 & - & 0.37 & - & 0.40 & - & 1 & 1.07 & - & 1 & 1.19 & - & 0.41 & - & 0 & 0.33 & - & 1.26 \\
-howait\_\_all\_workers\_finished\_if\_wait\_over\_\_depth\_0\_\_fixed & 217 & 232 & - & 0.27 & - & 0.26 & - & 1 & 0.82 & - & 1 & 0.87 & - & 0.26 & - & 0 & 0.26 & - & 0.80 \\
-safe\_send\_\_sending\_to\_non-pid\_\_depth\_0\_\_fixed & 121 & 120 & - & 0.20 & - & 0.22 & - & 1 & 0.63 & - & 1 & 0.59 & - & 0.18 & - & 0 & 0.19 & - & 0.62 \\
-stutter\_\_we\_abhorr\_as\_\_depth\_0\_\_fixed & 87 & 86 & - & 0.17 & - & 0.18 & - & 1 & 0.58 & - & 1 & 0.60 & - & 0.17 & - & 0 & 0.19 & - & 0.55 \\
-stutter\_\_we\_abhorr\_as\_\_depth\_1\_\_fixed & 107 & 104 & - & 0.20 & - & 0.20 & - & 1 & 0.54 & - & 1 & 0.58 & - & 0.19 & - & 0 & 0.17 & - & 0.56 \\
-stutter\_\_we\_abhorr\_as\_\_depth\_2\_\_fixed & 107 & 104 & - & 0.19 & - & 0.22 & - & 1 & 0.62 & - & 1 & 0.59 & - & 0.18 & - & 0 & 0.17 & - & 0.62 \\
-unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_0\_\_fixed & 35 & 30 & - & 0.17 & - & 0.15 & - & 1 & 0.48 & - & 1 & 0.50 & - & 0.16 & - & 0 & 0.16 & - & 0.48 \\
-unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_1\_\_fixed & 35 & 30 & - & 0.16 & - & 0.16 & - & 1 & 0.45 & - & 1 & 0.45 & - & 0.14 & - & 0 & 0.17 & - & 0.47 \\
-unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_2\_\_fixed & 35 & 30 & - & 0.15 & - & 0.15 & - & 1 & 0.50 & - & 1 & 0.47 & - & 0.15 & - & 0 & 0.14 & - & 0.44 \\
+finite\_leader\_\_single\_leader\_\_depth\_0\ & 297 & 308 & - & 0.26 & - & 0.25 & - & 1 & 0.74 & - & 1 & 0.74 & - & 0.25 & - & 0 & 0.24 & - & 1 & 0.74 & - & 1 & 0 & 0.74 \\
+firewall\_\_no\_pred\_called\_with\_zero\_\_depth\_0\ & 352 & 400 & - & 0.30 & - & 0.29 & - & 1 & 0.87 & - & 1 & 0.86 & - & 0.27 & - & 0 & 0.27 & - & 1 & 0.87 & - & 1 & 0 & 0.86 \\
+howait\_\_all\_workers\_finished\_if\_wait\_over\_\_depth\_0\ & 217 & 232 & - & 0.21 & - & 0.23 & - & 1 & 0.63 & - & 1 & 0.63 & - & 0.20 & - & 0 & 0.20 & - & 1 & 0.63 & - & 1 & 0 & 0.63 \\
+safe\_send\_\_sending\_to\_non-pid\_\_depth\_0\ & 121 & 120 & - & 0.17 & - & 0.17 & - & 1 & 0.48 & - & 1 & 0.48 & - & 0.16 & - & 0 & 0.16 & - & 1 & 0.48 & - & 1 & 0 & 0.48 \\
+stutter\_\_we\_abhorr\_as\_\_depth\_0\ & 87 & 86 & - & 0.15 & - & 0.15 & - & 1 & 0.43 & - & 1 & 0.43 & - & 0.14 & - & 0 & 0.14 & - & 1 & 0.44 & - & 1 & 0 & 0.43 \\
+stutter\_\_we\_abhorr\_as\_\_depth\_1\ & 107 & 104 & - & 0.16 & - & 0.16 & - & 1 & 0.46 & - & 1 & 0.45 & - & 0.15 & - & 0 & 0.15 & - & 1 & 0.47 & - & 1 & 0 & 0.45 \\
+stutter\_\_we\_abhorr\_as\_\_depth\_2\ & 107 & 104 & - & 0.16 & - & 0.17 & - & 1 & 0.46 & - & 1 & 0.45 & - & 0.15 & - & 0 & 0.15 & - & 1 & 0.46 & - & 1 & 0 & 0.46 \\
+unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_0\ & 35 & 30 & - & 0.12 & - & 0.12 & - & 1 & 0.35 & - & 1 & 0.35 & - & 0.12 & - & 0 & 0.12 & - & 1 & 0.36 & - & 1 & 0 & 0.35 \\
+unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_1\ & 35 & 30 & - & 0.12 & - & 0.13 & - & 1 & 0.35 & - & 1 & 0.36 & - & 0.12 & - & 0 & 0.12 & - & 1 & 0.35 & - & 1 & 0 & 0.35 \\
+unsafe\_send\_\_sending\_to\_non-pid\_\_depth\_2\ & 35 & 30 & - & 0.12 & - & 0.13 & - & 1 & 0.35 & - & 1 & 0.37 & - & 0.12 & - & 0 & 0.12 & - & 1 & 0.35 & - & 1 & 0 & 0.35 \\
     \hline

### 2014-01-31 make-benchmarks-table.sh

@@ -2,7 +2,7 @@
 
 benchmarks=( 'found-in-mist-repo' 'given-by-daniel-kroening' 'ic3-soter' )
 
-methods=( 'safety' 'safety-int' 'refinement' 'refinement-int' 'invariant' 'inv-min' 'inv-ref' )
+methods=( 'safety' 'safety-int' 'refinement' 'refinement-int' 'invariant' 'inv-min' 'inv-ref' 'inv-ref-min' )
 
 timeout=7200
 

### 2014-01-31 paper/workspace/main.tex
Fixed typo in report

@@
         # input transition
-        print(" + " t)
+        print(" + " + t)
       if (s, t) in E and not (t, s) in E:
         # output transition
-        print(" - " t)
+        print(" - " + t)
     print("\n")

### 2014-02-01 paper/workspace/main.tex
Added method Safety by Refinement with Lamport example

--- 2014-01-30 paper/workspace/main.tex
@@
   # of times transition t is fired >= 0
+\end{verbatim}
 
+\newpage
+
+\begin{verbatim}
 * Subprocedure \mathcal{C}:
@@
 
+  Pseudocode:
+
+  # Place equations
+  for s in S:
+    # initial number of tokens
+    print(s + " = " + M0(s))
+    for t in T:
+      if (t, s) in E and not (s, t) in E:
+        # input transition
+        print(" + " + t)
+      if (s, t) in E and not (t, s) in E:
+        # output transition
+        print(" - " + t)
+    print("\n")
+
+  # Non-negativity conditions
+  for s in S:
+    print(s + " >= 0\n")
+  for t in T:
+    print(t + " >= 0\n")
+\end{verbatim}
+
+\newpage
+
+\section{Method Safety by Refinement}
+
+The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
+
+\begin{verbatim}
+* Method Safety by Refinement:
+
+  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
+  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
+\end{verbatim}
+\begin{center}
+\begin{tikzpicture}
+  \node[state] (begin) {BEGIN};
+  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
+  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
+  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
+                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
+  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
+  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
+                                    	$\delta:=\Delta(A_\theta)$\\
+                                    	$C:=C \cup \{\delta\}$};
+  \node[print, right=of satc] (yes) {N satisfies P};
+  \node[print, right=of satctheta] (dontknow) {Don't know};
+  \node[state, right=of yes] (end1) {END};
+  \node[state, right=of dontknow] (end2) {END};
+
+  \draw (begin) edge (c);
+  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
+  \draw (satc) edge node[above]{NO} (yes);
+  \draw (yes) edge (end1);
+  \draw (satc) edge node[right]{YES} (modelc);
+  \draw (modelc) edge (satctheta);
+  \draw (satctheta) edge node[above]{NO} (dontknow);
+  \draw (dontknow) edge (end2);
+  \draw (satctheta) edge node[right]{YES} (modelctheta);
+  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
+  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
+\end{tikzpicture}
+\end{center}
+
+\begin{verbatim}
+* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
+  1. S is a trap in the net N
+  2. S is marked in the initial marking M0
+  3. S is unmarked in the assignment A
+
+* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
+
+* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
+  1. A ^ \delta is unsat (\delta excludes A)
+  2. N |= \delta is sat  (\delta is a property of N)
+\end{verbatim}
+
+\newpage
+\begin{verbatim}
+* Example
+  - State constraints C and \neg P are the same as in the example for method Safety.
+
+  - Assignment A:
+    id1        = 0
+    cs1        = 1
+    idle1      = 0
+    req1       = 0
+    nid1       = 1
+    id2        = 0
+    cs2        = 1
+    idle2      = 0
+    req2       = 0
+    after_you2 = 0
+    await2     = 0
+    u1         = 0
+    u2         = 1
+    u3         = 1
+    v1         = 0
+    v2         = 0
+    v3         = 0
+    v4         = 1
+    v5         = 0
+    v6         = 1
+
+  - Trap conditions C_\theta:
+
+    - Trap implications:
+                       v4              v3                 u2
+                  ------------    --------------    --------------     
+    id1        => (cs2 \/ id1) /\ (id1 \/ idle2) /\ (req1 \/ nid1)
+    ^              ^      ^        ^      ^          ^       ^
+    |              |      |        |      |          |       |
+    |              |     id1 \in S |     idle2 \in S |      nid1 \in S
+    |              |               |                 |
+    |             cs2 \in S       id1 \in S         req1 \in S
+    |
+    id1 \in S
+
+    cs1        => idle1 \/ id1
+    idle1      => req1 \/ nid1
+    req1       => cs1 \/ id2
+    nid1       => (after_you2 \/ nid1) /\ (idle1 \/ id1)
+    id2        => req2 /\ (cs1 \/ id2)
+    cs2        => (idle2 \/ id2)
+    idle2      => req2
+    req2       => (cs2 \/ id1) /\ (cs2 \/ id1)
+    after_you2 => id2 \/ await2
+    await2     => id1 \/ idle2
+
+    - At least one of the initially marked places belongs to S:
+    id1 \/ id2 \/ idle1 \/ idle2
+
+    - None of the marked places in A belongs to S:
+    ~cs1 /\ ~cs2 /\ ~nid1
+\end{verbatim}
+
+\newpage
+
+\begin{verbatim}
+  - Assignment A_\theta:
+    id1        = true
+    cs1        = false
+    idle1      = false
+    req1       = true
+    nid1       = false
+    id2        = true
+    cs2        = false
+    idle2      = false
+    req2       = true
+    after_you2 = true
+    await2     = false
+  - Refinement constraint \delta:
+    id1 + id2 + req1 + req2 + after_you2 >= 1
+    ^     ^     ^      ^      ^
+    |     |     |      |      |
+    ---------------------------
+      |
+      S = {iq1, iq2, req1, req2, after_you2}, therefore \delta excludes
+      assignment A in the next iteration
+
+* Trap implication:
+
+  place s \in S =>    /\       \    /
+                     /  \       \  /     place p \in S
+                    /    \       \/
+                   t \in s*   p \in t*
+
+  "if s is in trap S then for each output transition t at least one successor p is in trap S"
+
+* Refinement constraint \delta:
+
+  \Sigma s >= 1  
+  A(s)
+
+  "At least one place in S is always marked"
+
+* Subprocedure TrapConditions:
+
+  Input:
+    (S, T, E, M0) : Petri net
+    A             : Satisfying assignment for C \cup { ~P }
+  Output:
+    C_\theta      : Trap conditions
+
   TODO: Unroll pseudocode?
+
+* Subprocedure \Delta:
+
+  Input:
+    A_\theta      : Satisfying assignment for C_\theta
+  Output:
+    \delta        : Refinement constraint \delta
+
+  Pseudocode:
+
+  # Trap implications
+  for s in S:
+    print(s + " => true")
+    for t in s*:
+      print(" /\ ( false")
+      for p in t*:
+        print(" \/ " + p)
+     print(")")
+    print("\n")
+
+  # Initially marked
+  print("false")
+  for s in S:
+    if M0(s) > 0:
+      print(" \/ " + s)
+  print("\n")
+
+  # Unmarked in A
+  for s in S:
+    if A(s) = 0:
+      print("~" + p + "\n")
 \end{verbatim}
@@
 \newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Invariant by example on a mutual exclusion algorithm by Delzanno2001.
-
-\begin{verbatim}
-
-* Code
-
-bool m1 = false;
-bool m2 = false;
-
-// Thread 1:
-while (true) {
-  if (! m2 && ! test_and_set(m1)) {
-    // critical section 1
-    m1 = false
-  }
-  if (! m1 && ! test_and_set(m2)) {
-    // critical section 2
-    m2 = false
-  }
-}
-
-// Thread 2:
-while (true) {
-  if (! m2 && ! test_and_set(m1)) {
-    // critical section 1
-    m1 = false
-  }
-  if (! m1 && ! test_and_set(m2)) {
-    // critical section 2
-    m2 = false
-  }
-}
-
-Property: Thread 1 and Thread 2 are never in any critical section at the same time.
-
-* Petri net:
-\end{verbatim}
-\begin{center}
-\begin{tikzpicture}[node distance=2cm]
-  \node[blue place, tokens=2, label=above:nc] (nc) {};
-  \node[place, tokens=1, label=right:m2, below of=nc, xshift=+0.75cm] (s2) {};
-  \node[place, tokens=1, label=left:m1, below of=nc, xshift=-0.75cm] (s1) {};
-  \node[red place, label=above:cs1, left of=s1] (cs1) {};
-  \node[red place, label=above:cs2, right of=s2] (cs2) {};
-  \node[transition, label=above:t1] (t1) at ($(s1)!0.5!(cs1) + (0,1cm)$) {}
-	edge [pre] (nc)
-	edge [pre] (s1)
-	edge [pre] (s2)
-	edge [post] (cs1)
-	edge [post] (s2);
-  \node[transition, label=above:t2] (t2) at ($(s2)!0.5!(cs2) + (0,1cm)$) {}
-	edge [pre] (nc)
-	edge [pre] (s2)
-	edge [pre] (s1)
-	edge [post] (cs2)
-	edge [post] (s1);
-  \node[transition, label=below:t3] (t3) at ($(s1)!0.5!(cs1) - (0,1cm)$) {}
-	edge [pre] (cs1)
-	edge [post] (s1)
-	edge [post,bend left,min distance=3cm, out=115, in=115] (nc);
-  \node[transition, label=below:t4] (t4) at ($(s2)!0.5!(cs2) - (0,1cm)$) {}
-	edge [pre] (cs2)
-	edge [post] (s2)
-	edge [post,bend right,min distance=3cm, out=-115, in=-115] (nc);
-\end{tikzpicture}
-\end{center}
-
-\newpage
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-\begin{center}
-\begin{tikzpicture}
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                    	$\delta:=\Delta(A_\theta)$\\
-                                    	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {N satisfies P};
-  \node[print, right=of satctheta] (dontknow) {Don't know};
-  \node[state, right=of yes] (end1) {END};
-  \node[state, right=of dontknow] (end2) {END};
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (yes);
-  \draw (yes) edge (end1);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (dontknow);
-  \draw (dontknow) edge (end2);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example
-  - State constraints C:
-    nc  =  2 - t1 - t2 + t3 + t4
-    m1  =  1 - t1 + t3
-    m2  =  1 - t2 + t4
-    cs1 =  0 + t1 - t3
-    cs2 =  0 + t2 - t4
-    nc  >= 0
-    m1  >= 0
-    m2  >= 0
-    cs1 >= 0
-    cs2 >= 0
-    t1  >= 0
-    t2  >= 0
-    t3  >= 0
-    t4  >= 0
-
-  - Assignment A:
-    nc  = 0
-    m1  = 0
-    m2  = 0
-    cs1 = 1
-    cs2 = 1
-    t1  = 1
-    t2  = 1
-    t3  = 0
-    t4  = 0
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-
-    cs1 => m1 \/ nc
-    ^      ^     ^
-    |      |     |
-    |      |     nc \in S
-    |      |
-    |      m1 \in S
-    |
-    cs1 \in S
-
-    cs2 => m2 \/ nc
-    nc  => (m2 \/ cs1) /\ (m1 \/ cs2)
-    m1  => (m2 \/ cs1) /\ (m1 \/ cs2)
-    m2  => (m2 \/ cs1) /\ (m1 \/ cs2)
-
-    - At least one of the intially marked places belongs to S:
-    nc \/ m1 \/ m2
-
-    - None of the marked places in A belongs to S:
-    ~cs1 /\ ~cs2
-
-  - Assignment A_\theta:
-    nc  = false
-    m1  = true
-    m2  = true
-    cs1 = false
-    cs2 = false
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-  - Refinement constraint \delta:
-    m1 + m2 >= 1
-    ^    ^
-    |    |
-    ------
-      |
-      S = {m1, m2}, therefore \delta excludes assignment A in the next iteration
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  TODO: Unroll pseudocode?
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  TODO: Unroll pseudocode?
-\end{verbatim}
-
-\newpage
 \section{Benchmarks of \bfc}

### 2014-02-01 paper/workspace/main.tex
bla
Added pseudocode for subprocedure \Delta
Updated paper draft

@@
   # of times transition t is fired >= 0
-\end{verbatim}
 
-\newpage
-
-\begin{verbatim}
 * Subprocedure \mathcal{C}:
@@
 
-  # Place equations
-  for s in S:
-    # initial number of tokens
-    print(s + " = " + M0(s))
-    for t in T:
-      if (t, s) in E and not (s, t) in E:
-        # input transition
-        print(" + " + t)
-      if (s, t) in E and not (t, s) in E:
-        # output transition
-        print(" - " + t)
-    print("\n")
-
-  # Non-negativity conditions
-  for s in S:
-    print(s + " >= 0\n")
-  for t in T:
-    print(t + " >= 0\n")
   \end{verbatim}
+\begin{align*}
+  C =& \left( \bigwedge_{s \in S}
+    s = M_0(s) + \sum_{t \in ^\bullet s} t - \sum_{t \in s^\bullet} t
+  \right) \land \\
+  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
+    \left( \bigwedge_{t \in T} t \ge 0 \right)
+\end{align*}
 
@@
 
+\end{verbatim}
+
+\newpage
+
+\begin{verbatim}
+
 * Subprocedure TrapConditions:
@@
 
-  TODO: Unroll pseudocode?
+  Pseudocode:
+
+\end{verbatim}
+\begin{align*}
+  C_\theta =& \left( \bigwedge_{s \in S} \left( s \Rightarrow
+        \bigwedge_{t \in s^\bullet} \bigvee_{p \in t^\bullet} p
+    \right) \right) \land \\
+  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
+  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
+\end{align*}
+\begin{verbatim}
 
@@
 
-  # Trap implications
-  for s in S:
-    print(s + " => true")
-    for t in s*:
-      print(" /\ ( false")
-      for p in t*:
-        print(" \/ " + p)
-     print(")")
-    print("\n")
-
-  # Initially marked
-  print("false")
-  for s in S:
-    if M0(s) > 0:
-      print(" \/ " + s)
-  print("\n")
-
-  # Unmarked in A
-  for s in S:
-    if A(s) = 0:
-      print("~" + p + "\n")
 \end{verbatim}
-
-\iffalse
+\begin{align*}
+  & \delta = \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
+\end{align*}
 
@@
 The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-We present the method Invariant by example on a simplified mutual exclusion algorithm by Delzanno2001.
 
@@
 * Example:
+
+  - With weakened property P \/ (\neg \delta)
+
+  - Negated property is \neg P /\ \delta
+
   - Dual state constraints C':
+    0 >= (- cs1) + id1 + idle1 + (- nid1)
+    0 >= (- id1) + (- idle1) + nid1 + req1
+    0 >= cs1 + (- req1)
+    0 >= after_you2 + (- req2)
+    0 >= (- after_you2) + await2 + id2
+    0 >= (- await2) + idle2
+    0 >= cs2 + (- req2)
+    0 >= (- cs2) + id2 + idle2
+    0 >= (- id2) + (- idle2) + req2
 
-    - nc - m + cs1 <= 0
-    - nc - m + cs2 <= 0
-      nc + m - cs1 <= 0
-      nc + m - cs2 <= 0
+    id1 + id2 + idle1 + idle2 < 2 * target_1 + trap_1
 
-    2 \cdot nc + m < 2 \cdot target1
-
-    Non-negativity conditions:
-
-    target1 >= 0
-    nc      >= 0
-    m       >= 0
-    cs1     >= target1
-    cs2     >= target1
+    id1        >= trap_1
+    cs1        >= target_1
+    idle1      >= 0
+    req1       >= trap_1
+    nid1       >= 0
+    id2        >= trap_1
+    cs2        >= target_1
+    idle2      >= 0
+    req2       >= trap_1
+    after_you2 >= trap_1
+    await2     >= 0
+    target_1   >= 0
+    trap_1     >= 0
 
@@
 
-    target1 = 1
-    nc      = 0
-    m       = 1
+    id1        = 1
     cs1        = 1
+    idle1      = 0
+    req1       = 1
+    nid1       = 0
+    id2        = 1
     cs2        = 1
+    idle2      = 0
+    req2       = 1
+    await2     = 0
+    after_you2 = 1
+    target_1   = 1
+    trap_1     = 1
 
@@
 
-    cs1 + cs2 + m <= 1
+    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
+
+  - Invariant with trap constraint \delta added:
+
+    id1 + id2 + req1 + req2 + after_you2             >= 1
+    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
 

### 2014-02-01 paper/workspace/main-example-filip.tex
Added version of paper draft with Filip's example

--- 2014-02-01 paper/workspace/main.tex
@@
 The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
+We present the method Safety by example on a modified Petri net from [Esparza1997].
 
@@
 
-* Property: Thread 1 and Thread 2 are never in the critical section at the same time.
+* Property: z is never marked
 
@@
 \begin{center}
-  \includegraphics[width=12cm]{lamport}
+  \includegraphics[width=6cm]{example-filip}
 \end{center}
@@
 \begin{verbatim}
-* Property P: cs1 + cs2 <= 1
+* Property P: z = 0
 \end{verbatim}
@@
 
-    id1        = 1 + u1 - u2
-    ^            ^   ^    ^
-    |            |   |    |
-    |            |   |    |
-    |            |   |    |
-    |            |   |   # of tokens taken from nc
+    x = 1 - t
+    ^   ^   ^
     |   |   |
-    |            |  # of tokens given to nc
+    |   |   |
+    |   |   |
+    |   |   |
+    |   |   |
+    |   |  # of tokens taken from x
     |   |
-    |          initial number of tokens in nc
+    |   initial number of tokens in x
     |
-    number of tokens in nc
+    number of tokens in x
 
-    cs1        = 0 - u1      + u3
-    idle1      = 1 + u1 - u2
-    req1       = 0      + u2 - u3
-    nid1       = 0 - u1 + u2
-    id2        = 1                     + v2           + v5 - v6
-    cs2        = 0                               + v4 - v5
-    idle2      = 1                          + v3 + v5      - v6
-    req2       = 0                - v1           - v4      + v6
-    after_you2 = 0                + v1 - v2
-    await2     = 0                     + v2 - v3
+    y = 0
+    z = 0 + t
+    w = 0 + t
 \end{verbatim}
@@
 
-    id1        >= 0
-    cs1        >= 0
-    idle1      >= 0
-    req1       >= 0
-    nid1       >= 0
-    id2        >= 0
-    cs2        >= 0
-    idle2      >= 0
-    req2       >= 0
-    after_you2 >= 0
-    await2     >= 0
-    u1         >= 0
-    u2         >= 0
-    u3         >= 0
-    v1         >= 0
-    v2         >= 0
-    v3         >= 0
-    v4         >= 0
-    v5         >= 0
-    v6         >= 0
+    x >= 0
+    y >= 0
+    z >= 0
+    w >= 0
+    t >= 0
 
@@
 
-    cs1 + cs2  >= 2
+    z >= 1
 
@@
 \begin{align*}
-  C =& \left( \bigwedge_{s \in S}
-    s = M_0(s) + \sum_{t \in ^\bullet s} t - \sum_{t \in s^\bullet} t
-  \right) \land \\
+  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
+    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
+  \right) \right) \land \\
   & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
@@
   - Assignment A:
-    id1        = 0
-    cs1        = 1
-    idle1      = 0
-    req1       = 0
-    nid1       = 1
-    id2        = 0
-    cs2        = 1
-    idle2      = 0
-    req2       = 0
-    after_you2 = 0
-    await2     = 0
-    u1         = 0
-    u2         = 1
-    u3         = 1
-    v1         = 0
-    v2         = 0
-    v3         = 0
-    v4         = 1
-    v5         = 0
-    v6         = 1
+    x = 0
+    y = 0
+    z = 1
+    w = 1
+    t = 1
 
@@
     - Trap implications:
-                       v4              v3                 u2
-                  ------------    --------------    --------------     
-    id1        => (cs2 \/ id1) /\ (id1 \/ idle2) /\ (req1 \/ nid1)
-    ^              ^      ^        ^      ^          ^       ^
-    |              |      |        |      |          |       |
-    |              |     id1 \in S |     idle2 \in S |      nid1 \in S
-    |              |               |                 |
-    |             cs2 \in S       id1 \in S         req1 \in S
+              t
+         -----------
+    x => y \/ z \/ w
+    ^    ^    |    |
+    |    |    |    w \in S
+    |    |    |
+    |    |    z \in S
+    |    |
+    |    y \in S
     |
-    id1 \in S
+    x \in S
 
-    cs1        => idle1 \/ id1
-    idle1      => req1 \/ nid1
-    req1       => cs1 \/ id2
-    nid1       => (after_you2 \/ nid1) /\ (idle1 \/ id1)
-    id2        => req2 /\ (cs1 \/ id2)
-    cs2        => (idle2 \/ id2)
-    idle2      => req2
-    req2       => (cs2 \/ id1) /\ (cs2 \/ id1)
-    after_you2 => id2 \/ await2
-    await2     => id1 \/ idle2
+    y => y \/ z \/ w
+    z => true
+    w => true
 
     - At least one of the initially marked places belongs to S:
-    id1 \/ id2 \/ idle1 \/ idle2
+    x
 
     - None of the marked places in A belongs to S:
-    ~cs1 /\ ~cs2 /\ ~nid1
+    ~z /\ ~w
 \end{verbatim}
@@
   - Assignment A_\theta:
-    id1        = true
-    cs1        = false
-    idle1      = false
-    req1       = true
-    nid1       = false
-    id2        = true
-    cs2        = false
-    idle2      = false
-    req2       = true
-    after_you2 = true
-    await2     = false
+    x = true
+    y = true
+    z = false
+    w = false
+
   - Refinement constraint \delta:
-    id1 + id2 + req1 + req2 + after_you2 >= 1
-    ^     ^     ^      ^      ^
-    |     |     |      |      |
-    ---------------------------
+    x + y >= 1
+    ^   ^
+    |   |
+    -----
       |
-      S = {iq1, iq2, req1, req2, after_you2}, therefore \delta excludes
-      assignment A in the next iteration
+      S = {x, y}, therefore \delta excludes assignment A in the next iteration
 
@@
 \begin{align*}
-  C_\theta =& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-        \bigwedge_{t \in s^\bullet} \bigvee_{p \in t^\bullet} p
+  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
+      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
     \right) \right) \land \\
@@
 \begin{align*}
-  & \delta = \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
+  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
 \end{align*}
+
+\iffalse
 
@@
 \end{verbatim}
-
-\iffalse
 

### 2014-02-01 src/petri-net-to-dot.sh

@@ -8,6 +8,6 @@ sysdir=$(absolute_path "$0")
 
 input=$1
 
-sicstus -l "$sysdir"/input-file-to-petri-net.pl -- `pwd`/$input 2>/dev/null >/tmp/petri-net.pl
+sicstus -l "$sysdir"/input-file-to-petri-net.pl -- 0 `pwd`/$input 2>/dev/null >/tmp/petri-net.pl
 
 sicstus -l "$sysdir"/petri-net-to-dot.pl -- /tmp/petri-net.pl 2>/dev/null

### 2014-02-01 paper/workspace/main.tex
Updated paper drafts

@@
 
-TODO
+\end{verbatim}
 
+\includegraphics[scale=0.75]{lamport-code}
+
+\begin{verbatim}
 * Property: Thread 1 and Thread 2 are never in the critical section at the same time.
@@
 \begin{align*}
-  C =& \left( \bigwedge_{s \in S}
-    s = M_0(s) + \sum_{t \in ^\bullet s} t - \sum_{t \in s^\bullet} t
-  \right) \land \\
-  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
+  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
+    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
+  \right) \right) \land
+    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
     \left( \bigwedge_{t \in T} t \ge 0 \right)
@@
     await2     = false
+
   - Refinement constraint \delta:
@@
 \begin{align*}
-  C_\theta =& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-        \bigwedge_{t \in s^\bullet} \bigvee_{p \in t^\bullet} p
-    \right) \right) \land \\
-  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
-  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
+  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
+      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
+    \right) \right) \land
+    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
+    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
 \end{align*}
@@
 \begin{align*}
-  & \delta = \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
+  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
 \end{align*}
+
+\iffalse
 

### 2014-02-01 paper/workspace/main.tex
bla
Changed example in paper draft to subnet of Lamport's algorithm

@@
 
-\end{verbatim}
+Process 1:
 
-\includegraphics[scale=0.75]{lamport-code}
+bit1 := false
+while true do
+  bit1 := true
+  a: while bit2 do skip od
+  b: <critical section>
+  bit1 := false
+od
 
-\begin{verbatim}
-* Property: Thread 1 and Thread 2 are never in the critical section at the same time.
+Process 2:
+
+bit2 := false
+while true do
+  l: bit2 := true
+  if bit1 then
+    bit2 := false
+    while bit1 do skip od
+    goto l
+  fi
+  <critical section>
+  bit2 := false
+od
+
+* Property: Process 1 and Process 2 are never in their respective critical section at the same time.
 
@@
 \begin{center}
-  \includegraphics[width=12cm]{lamport}
+  \includegraphics[width=10cm]{lamport}
 \end{center}
@@
 \begin{verbatim}
-* Property P: cs1 + cs2 <= 1
+
+* Subnet of Petri net:
+
 \end{verbatim}
 
-\newpage
+\begin{center}
+\begin{tikzpicture}[node distance=2cm]
+  \node[place, tokens=1, label=left:a] (a) {};
+  \node[place, tokens=0, label=left:b, below of=a] (b) {};
+  \node[place, tokens=0, label=right:nbit2, right of=b, xshift=-1cm, yshift=+1cm] (nbit2) {};
+  \node[transition, label=left:t1, above of=a, yshift=-1cm] (t1) {}
+    edge [pre, bend left, min distance=2cm, out=225, in=-45] (b)
+    edge [post] (a);
+  \node[transition, label=left:t2] (t2) at ($(a)!0.5!(b)$) {}
+    edge [pre] (a)
+    edge [pre] (nbit2)
+    edge [post] (b)
+    edge [post] (nbit2);
+\end{tikzpicture}
+\end{center}
+
+\begin{verbatim}
+* Property P: a + b <= 1
+\end{verbatim}
+
 \begin{verbatim}
@@
 * Property of state constraints C: If C U {\neg P} is unsat then N |= P.
+\end{verbatim}
 
+\newpage
+
+\begin{verbatim}
 * Example:
@@
 
-    id1        = 1 + u1 - u2
+    a     = 1 + t1 - t2
     ^       ^   ^    ^
@@
     |       |   |    |
-    |            |   |   # of tokens taken from nc
+    |       |   |   # of tokens taken from a
     |       |   |    
-    |            |  # of tokens given to nc
+    |       |  # of tokens given to a
     |       |
-    |          initial number of tokens in nc
+    |      initial number of tokens in a
     |
-    number of tokens in nc
+    number of tokens in a
 
-    cs1        = 0 - u1      + u3
-    idle1      = 1 + u1 - u2
-    req1       = 0      + u2 - u3
-    nid1       = 0 - u1 + u2
-    id2        = 1                     + v2           + v5 - v6
-    cs2        = 0                               + v4 - v5
-    idle2      = 1                          + v3 + v5      - v6
-    req2       = 0                - v1           - v4      + v6
-    after_you2 = 0                + v1 - v2
-    await2     = 0                     + v2 - v3
-\end{verbatim}
+    b     = 0 - t1 + t2
+    nbit2 = 0
 
-\newpage
-
-\begin{verbatim}
 Non-negativity conditions:
 
-    id1        >= 0
-    cs1        >= 0
-    idle1      >= 0
-    req1       >= 0
-    nid1       >= 0
-    id2        >= 0
-    cs2        >= 0
-    idle2      >= 0
-    req2       >= 0
-    after_you2 >= 0
-    await2     >= 0
-    u1         >= 0
-    u2         >= 0
-    u3         >= 0
-    v1         >= 0
-    v2         >= 0
-    v3         >= 0
-    v4         >= 0
-    v5         >= 0
-    v6         >= 0
+    a     >= 0
+    b     >= 0
+    nbit2 >= 0
 
@@
 
-    cs1 + cs2  >= 2
+    a + b >= 2
 
@@
 * Example
-  - State constraints C and \neg P are the same as in the example for method Safety.
+  - State constraints C same as in the example for method Safety.
+
+  - Property P:
+
+    b = 0
+
+  - Negated Property \neg P:
+
+    b >= 1
 
   - Assignment A:
-    id1        = 0
-    cs1        = 1
-    idle1      = 0
-    req1       = 0
-    nid1       = 1
-    id2        = 0
-    cs2        = 1
-    idle2      = 0
-    req2       = 0
-    after_you2 = 0
-    await2     = 0
-    u1         = 0
-    u2         = 1
-    u3         = 1
-    v1         = 0
-    v2         = 0
-    v3         = 0
-    v4         = 1
-    v5         = 0
-    v6         = 1
+    a     = 0
+    b     = 1
+    nbit2 = 0
+    t1    = 0
+    t2    = 1
 
@@
     - Trap implications:
-                       v4              v3                 u2
-                  ------------    --------------    --------------     
-    id1        => (cs2 \/ id1) /\ (id1 \/ idle2) /\ (req1 \/ nid1)
-    ^              ^      ^        ^      ^          ^       ^
-    |              |      |        |      |          |       |
-    |              |     id1 \in S |     idle2 \in S |      nid1 \in S
-    |              |               |                 |
-    |             cs2 \in S       id1 \in S         req1 \in S
+               t2
+             ----------
+    a     => b \/ nbit2
+    ^           ^      ^
+    |           |      |
+    |           |     nbit2 \in S
+    |           |
+    |          b \in S
     |
-    id1 \in S
+    a \in S
 
-    cs1        => idle1 \/ id1
-    idle1      => req1 \/ nid1
-    req1       => cs1 \/ id2
-    nid1       => (after_you2 \/ nid1) /\ (idle1 \/ id1)
-    id2        => req2 /\ (cs1 \/ id2)
-    cs2        => (idle2 \/ id2)
-    idle2      => req2
-    req2       => (cs2 \/ id1) /\ (cs2 \/ id1)
-    after_you2 => id2 \/ await2
-    await2     => id1 \/ idle2
+    nbit2 => b \/ nbit2
+    b     => a
 
     - At least one of the initially marked places belongs to S:
-    id1 \/ id2 \/ idle1 \/ idle2
+    a
 
     - None of the marked places in A belongs to S:
-    ~cs1 /\ ~cs2 /\ ~nid1
+    ~b
+
+  - Assignment A_\theta:
+    a     = true
+    b     = false
+    nbit2 = true
+
+  - Refinement constraint \delta:
+    a + b >= 1
+    ^   ^
+    |   |
+    -----
+      |
+      S = {a, b}, therefore \delta excludes assignment A in the next iteration
 \end{verbatim}
@@
 \begin{verbatim}
-  - Assignment A_\theta:
-    id1        = true
-    cs1        = false
-    idle1      = false
-    req1       = true
-    nid1       = false
-    id2        = true
-    cs2        = false
-    idle2      = false
-    req2       = true
-    after_you2 = true
-    await2     = false
-
-  - Refinement constraint \delta:
-    id1 + id2 + req1 + req2 + after_you2 >= 1
-    ^     ^     ^      ^      ^
-    |     |     |      |      |
-    ---------------------------
-      |
-      S = {iq1, iq2, req1, req2, after_you2}, therefore \delta excludes
-      assignment A in the next iteration
 
@@
   "At least one place in S is always marked"
-
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
 

### 2014-02-01 paper/workspace/main.tex
Changed labels of code and net

@@
     while true do
-  bit1 := true
-  a: while bit2 do skip od
-  b: <critical section>
+p1:   bit1 := true
+p2:   while bit2 do skip od
+p3:   <critical section>
       bit1 := false
@@
     while true do
-  l: bit2 := true
-  if bit1 then
-    bit2 := false
-    while bit1 do skip od
-    goto l
+q1:   bit2 := true
+q2:   if bit1 then
+q3:     bit2 := false
+q4:     while bit1 do skip od
+        goto q1
       fi
-  <critical section>
+q5:   <critical section>
       bit2 := false
@@
   \node[place, tokens=0, label=left:b, below of=a] (b) {};
-  \node[place, tokens=0, label=right:nbit2, right of=b, xshift=-1cm, yshift=+1cm] (nbit2) {};
+  \node[place, tokens=0, label=right:$\neg bit2$, right of=b, xshift=-1cm, yshift=+1cm] (nbit2) {};
   \node[transition, label=left:t1, above of=a, yshift=-1cm] (t1) {}

### 2014-02-03 paper/example-latex/rebuttal-reviews.txt
Added example paper for latex guidelines.

----------------------- REVIEW 1 ---------------------
PAPER: 114
TITLE: Binary Reachability Analysis of Functional Programs
AUTHORS: (anonymous)

OVERALL RATING: -2 (reject)
REVIEWER'S CONFIDENCE: 3 (high)
IMPORTANCE: 2 (Okay: The paper addresses a problem that has good, or many, solutions.)
NOVELTY: 2 (Incremental. A straightforward next step to an existing idea.)
CLARITY: 2 (Needs Improvement)

A program transformation that makes it possible to reduce (an aspect of) the problem of termination checking to the problem of reachability (in the normal sense of model checking of safety properties), for higher-order functional programs.  In particular, the functionality in this paper takes as input a program and a proposed termination relation for one of the functions, and the tool verifies that this relation really proves termination.

My higher-level take: This paper presents one piece of a Terminator-like checker for (a subset of?) OCaml programs.  A full tool for this termination checking problem would use abstract interpretation, abstraction refinement, etc., to learn appropriate relations explaining termination of all recursive definitions.  The technique in this paper just helps check a single such relation, and the contribution is not to show how the checking may be done, but rather to show how to reduce it to a different verification problem.  Thus, it's important to demonstrate that the reduced programs fall into a set that can be verified in practice.  The experiments provide some evidence of such tractability, but it's not overwhelmingly convincing.  I found the exposition to be disorganized and confusing, and would expect it to be even trickier for readers not used to thinking about such concepts as functional programming, monads, and metaprogramming with quotations and antiquotations.

A significant amount of verbiage and OCaml code is used to explain a program instrumentation strategy.  In the first pages of this discussion, I was very confused about how what I was reading related to the termination-checking problem.  By the end, I thought it was a simple enough idea to deserve no more than a page of prose.  Basically, we want to prove that some binary relation over procedure argument lists characterizes every recursive call that may be made in the course of evaluating some original call.  We instrument a program to nondeterministically record the arguments used in some call to the function we are focusing on.  While we continue to evaluate that call, every further argument list is checked against the saved list with the relation.  There are certainly details to be gotten right in implementing the instrumentation, but they don't strike me as "research contribution" level.  This reduction seems like one that most CAV folks would be ready to believe as sound
just from the high-level description, and it's not surprising that it can be implemented for a Turing-complete language.

I don't know if anyone has considered this idea before.  The practicality rests on the tractability of the new verification problem that we reduce to.  (Actually, it's not clear to me that it wouldn't be simpler to treat the original termination problem directly.)  The set of experiment programs used in this paper seems mildly misleading.  As mentioned in the paper, it's true that traditional techniques like size change termination work on inductive types like natural numbers, while OCaml's int's are a different story.  However, most of the experiment programs here either (a) work with simple restriction to primitive recursion (like in Coq) or (b) use int's as if they were naturals, with recursive functions whose base cases are "n <= 0" rather than just "n = 0".  These programs are trivially handled by (perhaps minor tweaks to) techniques lightweight enough that they are already incorporated into the type checkers of several production-quality functional language implementati
ons.

Of the 15 experiment programs, I identify 5 that don't fit within the above criteria.
- The usual suspects of the Ackermann function and McCarthy's 91.  These are reasonable examples where fancy termination checking is called for, though I question whether such reasoning is common enough to deserve tool support.
- Sorted list merging, which is well-founded under a relation comparing sums of lengths of argument lists.  Another reasonable example.
- A strange recursive function to manipulate an algebraic datatype seemingly inspired by red-black trees, but using code that does not seem to bear any relationship to useful tree manipulations.  I don't think there's anything practical to be learned from this one.
- A program with an intentional definition of a diverging recursive function, which appears as an argument within a non-trivial tree of combinator calls, such that the diverger is never called.  This strikes me as undermotivated, since the program can simply be run to see that it terminates, in few enough steps that one can imagine a heuristic to try exactly this for any input program!  This would be similar to the partial evaluation that goes on during termination checking in Agda or Coq.

So, in my opinion, there are 3 programs among the 15 experiments that provide any sort of interesting exercise for termination checking, and all 3 are purely first-order.  As a result, I don't believe this paper makes a persuasive case for advancing the state of the art in termination checking for higher-order programs.

Fig. 3's Augment function is not total if argument expressions to "f" may diverge.  Is that a problem?

Minor points:

In the abstract, the close repetition of the word "imperative" with two different meanings seems a bit like a bad joke. ;)

Page 2: Missing word "is" in "variables responsible for"
Page 2: Missing "order" in "higher functional"
Page 5: "defived"?  More spelling errors appear later.  Please use spell-check.
Page 6: Missing "we" before "will refer to"
Page 7: Accidential inclusion of text "dots" instead of a LaTeX operator?
Page 7: Something wrong in "Once the obtained the"
Page 8: Fig. 3: What is the use of OCaml code like "_ as sigma"?  Why not just write "sigma"?
Page 8: Fig. 3: Line 8 seems to pass wrong argument in middle recursive call.
Page 9: Missing "a" between "is Mini-OCaml"
Page 9: "and return" should be "and returns"
Page 9: Fig. 5: Why make "monadic" mutually recursive with "pre_m"?
Page 10: The monadic example at the top of the page doesn't make sense.  The notation doesn't explain the connection between the two lines of the example, and OCaml parsing rules would imply that you are using a type family "int" with a single argument "pre_m" (just as you could use "list" with "int", for instance).
Page 10: Notation like "dot" (where the quotes are taken literally from the paper) is confusing for introducing an operator.  I assumed you were talking about an infix operator "dot", where actually it is the quotes you are introducing as an operator!
Page 14: Missing "on" between "transformation sereni163"?  (Not sure what's going on here, since the benchmark ID changes within the paragraph.)
Page 14: Fig. 9: What is "dros"?
Page 14: "infinite decent" should be "infinite descent"
Page 15: "a iterative" should be "an iterative"

----------------------- REVIEW 2 ---------------------
PAPER: 114
TITLE: Binary Reachability Analysis of Functional Programs
AUTHORS: (anonymous)

OVERALL RATING: 1 (weak accept)
REVIEWER'S CONFIDENCE: 1 (low)
IMPORTANCE: 2 (Okay: The paper addresses a problem that has good, or many, solutions.)
NOVELTY: 2 (Incremental. A straightforward next step to an existing idea.)
CLARITY: 3 (Well written)

This paper, which is perhaps closer to the area of ICFP than CAV,
describes a method for proving termination of higher order functional
programs written in a subset of OCaml.

The paper is written for termination experts and is likely to be hard
to understand by non-experts. Its contribution is to successfully
adapt and extend ideas used to establish termination of imperative
programs, by reduction to reachability analysis, to functional programs.

The method is quite complex, involving monads, program transformation,
binary reachability analysis and other things this reviewer admits to
not being completely familiar with. It is hard for me to judge how
well presented the material is for suitably expert readers. There seem
to be quite complex interlocking ideas going into the algorithm: the
paper appears to be a substantial contribution, but one wonders how
incremental it is.

Termination of higher order recursive functions is a major concern in
theorem proving when using systems like Isabelle/HOL or Coq. It would
be interesting to know whether the work described has potential
relevance to this. There is no comparison, for example, with recent
work like that of Alexander Krauss at TUM. Is there any connection
between the termination problem Krauss solves, or are they completely
separate.

I think this is a substantial paper, though from a low confidence
perspective. Possible criticisms are that it is incremental and too
narrow to be of general interest.

Typos spotted: "candiate", "An program", "$v_1 dots, v_N$"

----------------------- REVIEW 3 ---------------------
PAPER: 114
TITLE: Binary Reachability Analysis of Functional Programs
AUTHORS: (anonymous)

OVERALL RATING: 2 (accept)
REVIEWER'S CONFIDENCE: 3 (high)
IMPORTANCE: 4 (Critical: The paper addresses a problem that is important and is in dire need of solutions.)
NOVELTY: 3 (Solid. A new approach in an established field.)
CLARITY: 4 (Lucid)

This paper presents a "binary reachability analysis" -- a core
component of termination analysis using transition invariants -- for
higher-order functional programs. The main technical idea is to reduce
the binary reachability analysis problem to an ordinary reachability
problem, using a sort of a product construction. The reachability
problem is then solved using an existing higher-order safety checker
(DSolve). Some empirical results about the method's utility are
presented.

This is a good paper that should be accepted for the following reasons:

* The paper is trying to solve a problem that is interesting
as well as important. Research on software verification has
traditionally neglected higher-order features, and given the
importance of such features in modern languages, it is important that
this state of affairs changes. Also, higher-order programs pose a lot
of interesting challenges for verification.

* The technical approach is elegant and utilizes the existing
capabilities of DSolve. Of course, a lot of this ground has previously
been trod in the context of Terminator. However, some intriguing new
issues (e.g., partial application) arise from the higher-order aspects
of the problem, and the paper resolves them nicely.

However, I wonder: how easy would be it for DSolve to verify the kind 
of conditions generated by your reduction on larger programs? These
monadic programs seem quite complex, and I can imagine DSolve
stumbling on them.

An evaluation on a large, more "real-world" example would alleviate 
this concern.

* The paper is well-written. I really liked the
examples in Sec. 2, which bring out the additional challenges of
higher-order verification nicely.

Questions/comments:

* If you had references and/or imperative features in your language,
would things change too much?

* The paper would benefit from a discussion of the limitations of
the approach. Can you give some examples of programs that you were
NOT able to verify?

* Table 9 should probably be augmented with runtimes for your method.

Minor:

Pg. 5, last line: "the the"

----------------------- REVIEW 4 ---------------------
PAPER: 114
TITLE: Binary Reachability Analysis of Functional Programs
AUTHORS: (anonymous)

OVERALL RATING: -2 (reject)
REVIEWER'S CONFIDENCE: 2 (medium)
IMPORTANCE: 2 (Okay: The paper addresses a problem that has good, or many, solutions.)
NOVELTY: 2 (Incremental. A straightforward next step to an existing idea.)
CLARITY: 2 (Needs Improvement)

Pro: reduction to reachability
Con: doesn't demonstrate scaling

The authors consider the construction of "transition invariants" for
higher order programs written in a mini dialect of Ocaml. The key insight
of the authors is that the construction of such invariants can be
reduced to reachability questions. This allows leveraging previous
work on deciding reachability for higher order programs.

On the positive side the reduction is technically interesting. The 
authors clearly lay out how evaluation trees can be "augmented" with
a state vector, then how information is passed during evaluation.

My major concern with the paper, and why I do not support acceptance,
is that the paper claims to demonstrate scaling but then the experiments
do not support the claim. Section 6 lists the programs that were 
considered, but nothing about the speed of verification, the size 
of state required, the size of the programs, 
or any other performance metric. It is not possible to conclude much
from this without prior detailed knowledge of the evaluation programs.

### 2014-02-03 paper/example-latex/sec_introduction.tex
Added example paper for latex guidelines.

\section{Introduction}

Tools and techniques for proving program termination are important for
increasing software quality~\cite{CACM}.
System routines written in imperative programming languages received a
significant amount of attention recently,
e.g.,~\cite{copytrick,Julia,KroeningCAV10,GieslBytecode,YangESOP08,SepLogTerm}. 
A number of the proposed approaches rely on transition invariants -- a
termination argument that can be constructed incrementally using
abstract interpretation~\cite{transitioninvariants}.
Transition invariants are binary relations over program states.
Checking if an incrementally constructed candidate is in fact a
transition invariant of the program is called binary reachability
analysis.
For imperative programs, its efficient implementation can be obtained
by a reduction to the reachability analysis, for which practical tools
are available, e.g.,~\cite{Slam,Blast,fsoft,Impact}.
The reduction is based on a program transformation that stores one
component of the pair of states under consideration in auxiliary program
variables, and then checks if the pair is in the transition
invariant~\cite{copytrick}. 
The transformed program is verified using an existing safety checker. 
If the safety checker succeeds then the original program terminates on all inputs.  
% The set of reachable program states of the transformed program can be
% directly used in the binary reachability analysis.

For functional programs, recent approaches for proving termination
apply the size change termination (SCT) argument~\cite{SCTPOPL2001}.
This argument requires checking the presence of an infinite descent within
data values passed to application sites of the program on any infinite
traversal of the call graph.
This check can be realized in two steps.
First, every program function is translated into a set of so-called
size-change graphs that keep track of decrease in values between the
actual arguments and values at the application sites in the
function.
Second, the presence of a descent is checked by computing a transitive
closure of the size-change graphs. 
Originally, the SCT analysis was formulated for first order functional
programs manipulating well-founded data, yet using an appropriate
control-flow analysis, it can be extended to higher order programs,
see
e.g.,~\cite{SereniICFP07,Sereni05terminationanalysis,SereniAPLAS05}.
Alternatively, an encoding into term rewriting can be used to make
sophisticated decidable well-founded orderings on terms applicable to proving
termination of higher order programs~\cite{GieslHaskell}.

The SCT analysis is a decision problem (that checks if there is an infinite descent in the abstract program defined by size change graphs), however it is an incomplete method for proving termination. 
SCT can return ``don't know'' for terminating programs that
manipulate non-well-founded data, e.g., integers, or when an interplay
of several variables witnesses program termination. 
In such cases, a termination prover needs to apply a more general termination argument.
Usually, such termination arguments require proving that certain expressions over program variables decrease as the computation progresses and yet the decrease cannot happen beyond a certain bound.

In this paper, we present a general approach for proving termination of higher order functional programs that goes beyond the SCT analysis.
Our approach explores the applicability of transition invariants to
this task by proposing an extension (wrt.\ imperative case) that deals
with partial applications, a programming construct that is particular for functional programs.
Partial applications of curried functions, i.e., functions that return other functions, represent a major obstacle for the binary reachability analysis. 
For a curried function, the set of variables whose values need to be stored in auxiliary variables keeps increasing as the function is subsequently applied to its arguments. 
However these arguments are not necessarily supplied simultaneously,
which requires intermediate storage of the argument values given so
far.
In this paper, we address such complications.

We develop the binary reachability analysis for higher order programs in two steps.
First, we show how intermediate nodes of program evaluation trees, \mbox{so-called} judgements, can be augmented with auxiliary values needed for tracking binary reachability. 
The auxiliary values store arguments provided at application sites.
Then, we show how this augmentation can be performed on the program source code such that the evaluation trees of the augmented program correspond to the result of augmenting the evaluation trees of the original program.
The source code transformation introduces additional parameters to functions occurring in the program. For curried functions, these additional parameters are interleaved with the original parameters, which allows us to deal with partial applications.

Our binary reachability analysis for higher order programs opens up an approach for termination proving in the presence of higher order functions that exploits a highly optimized safety checker, e.g.,~\cite{TerauchiPOPL10,RupakRanjit,KobayashiPLDI11,Dsolve,HMC},
for checking the validity of a candidate termination argument.
Hence, we can directly benefit from sophisticated abstraction
techniques and algorithmic improvements offered by these tools, as inspired by~\cite{copytrick}.

In summary, this paper makes the following contributions.
%
\begin{itemize}
\item A notion of binary reachability analysis for 
  higher order functional programs.
\item A program transformation that reduces the binary
  reachability analysis to the reachability analysis.
  % code   transformation. 
\item An implementation of our approach and its evaluation on micro
  benchmarks from the literature.
\end{itemize}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_preliminaries.tex
Added example paper for latex guidelines.

\section{Preliminaries}
\label{sec-preliminaries}

\input{fig_syn_and_sem}

In this section we describe \miniocaml, the programming language that
we use to represent programs. We also present the logger
monad~\cite{freetheorems} extended with an update operation.

\paragraph{\bf \miniocaml syntax}

Let \vars\ be a set of \emph{variables}, e.g., \ttt{x},
\ttt{List.map}, and~\ttt{myVar}.  Let \csts\ be a set of
\emph{constants}, e.g., \ttt{+}, \ttt{1}, \ttt{2.5}, and \ttt{"h"}.
Let \cstrs\ be a set of \emph{constructors}, e.g., \ttt{::},
\ttt{Some}, \ttt{[]}, and \ttt{true}.  We assume $\vst, \var \in
\vars$, $\cst\in \csts$, and $\cstr\in\cstrs$.
Figure~\ref{fig-syn-and-sem}(a) presents the syntax of \miniocaml
expressions. 
We encode \ttt{if-then-else} expressions using match expressions. 
Function applications are left associative. 
We assume that tuples of values are encoded using tuple constructors.
We assume that text in \ttt{type writer font} is \miniocaml code.

\paragraph{\bf \miniocaml types}

We use the \ocaml\ type system~\cite{LeroyPhD92} to type \miniocaml\
expressions. Let \basetypes\ be a set of base types, e.g., \ttt{int},
\ttt{string}, and \ttt{in\_channel}. Let \typevars\ be a set of type
variables, e.g., \ttt{'a} and \ttt{'b}. Let \typecstrs\ be a set of
type constructors, e.g., \ttt{list}, \ttt{option}, and \ttt{bool}.  We assume
$\tybase \in \basetypes$, $\tyvar \in \typevars$, and $\tycstr \in
\typecstrs$.  Figure~\ref{fig-syn-and-sem}(b) presents the set of \miniocaml\
types.  We write the typing proposition $\expr~:~\type$ if expression
\expr\ is of type \type\ under some typing context. We say that expression \expr\ is well-typed
if there exists a type \type\ such that $\expr~:~\type$. Examples of
valid propositions are $ \ttt{1}~:~\ttt{int} $ and $
\ttt{+}~:~\ttt{int~->~int~->~int} $.

\paragraph{\bf \miniocaml semantics}

Figure~\ref{fig-syn-and-sem}(c) presents  \emph{values} computed by \miniocaml
programs using judgements $\evjvar$ of the form
%
\begin{equation*}
  \evjs  \ni \evjvar ::=  \evj{\ectx}{\expr}{\val}
\end{equation*} 
%
Every judgement is derived by applying rules shown in
Figure~\ref{fig-syn-and-sem}(d).
A judgement \evjvar\ is \emph{valid} if there exists an evaluation
tree with \evjvar\ as the root.
Each evaluation tree is given by the set of its edges.
Each edge is a sequence of judgements $\evjvar[1], \dots, \evjvar[n],
\evjvar$, where $\evjvar[1], \dots, \evjvar[n]$ are the predecessor
nodes and \evjvar\ is the successor node.
If $n=0$ then the edge represents a leaf node.

For example, we consider the evaluation tree \evt\ for \ttt{1 + 2} as
shown below.
\begin{center}
  \small
  \[
  \inference{ \inference{}{ \evj{\emptyectx}{\ttt{2}}{\ttt{2}} } &
    \inference{ \inference{}{ \evj{\emptyectx}{\ttt{1}}{\ttt{1}} } &
      \inference{}{ \evj{\emptyectx}{\ttt{+}}{\ttt{+}} } }{
      \evj{\emptyectx}{\ttt{(+) 1}}{\ttt{+}_1} } }{
    \evj{\emptyectx}{\ttt{1 + 2}}{\ttt{3}} }
  \]
\end{center}
Let
$\evjvar[1]~=~(\evj{\emptyectx}{\ttt{1+2}}{\ttt{3}})$, 
$\evjvar[2]~=~(\evj{\emptyectx}{\ttt{2}}{\ttt{2}})$, 
$\evjvar[3]~=~(\evj{\emptyectx}{\ttt{(+)~1}}{\ttt{+}_1})$, 
$\evjvar[4]~=~(\evj{\emptyectx}{\ttt{1}}{\ttt{1}})$, 
$\evjvar[5]~=~(\evj{\emptyectx}{\ttt{+}}{\ttt{+}})$.
The evaluation tree \evt\ is given by the set of five edges below.
%
\[
\evt~=~\{ \evjvar[2], \evjvar[4], \evjvar[5], (\evjvar[4],
\evjvar[5], \evjvar[3]), (\evjvar[2], \evjvar[3], \evjvar[1]) \}
\]
%
Let $\tsf{eval}\ \ectx\ \expr$ be the value of expression \expr\ in
the environment \ectx, i.e., $v = \tsf{eval}\ \ectx\ \expr$
if there is an evaluation tree for \evj{\ectx}{\expr}{\val}.

\paragraph{\bf Recursion relations and binary reachability}

We are interested in keeping track of (possibly partial) applications
of a function defined in a program.
Let $f$ be a function identifier of type $\type[1] \rightarrow \ldots
\rightarrow \type[m] \rightarrow \type$ that is bound using a let-rec
binding, and $\theArity$ be a number between $1$ and $m$.  
An \emph{$\theFunction/\theArity$-application judgement} describes evaluation of
an application of $f$ to $\theArity$-many actual parameters, i.e., it
is a judgement of the form~\evj{\ectx}{f\ \expr[1]\ \ldots\
\expr[n]}{\val}.
A \emph{$\theFunction/\theArity$-recursion relation} consists of pairs
of value tuples $(v_1, \dots, v_\theArity)$ and~$(u_1, \dots,
u_\theArity)$ that satisfy the following condition.
For each $(v_1, \dots, v_\theArity)$ and~$(u_1, \dots, u_\theArity)$
in the relation we require existence of a pair of valid
$\theFunction/\theArity$-application judgements $\evjvar[1] =
\evj{\ectx[1]}{f\ \expr[1]\ \ldots\ \expr[\theArity]}{\val[]}$ and
$\evjvar[2] = \evj{\ectx[2]}{f\ \expr[1]\ \ldots\
\expr[\theArity]}{u}$ such that $\evjvar[2]$ appears in the evaluation
tree of $\evjvar[1]$ and for each $i \in 1..\theArity$ we have
$\tsf{eval}\ \ectx[1]\ \expr[i] = \val[i]$ and $\tsf{eval}\ \ectx[2]\
\expr[i] = u_i$.

The goal of \emph{binary reachability analysis} is to check if the
$\theFunction/\theArity$-recursion relation of the program is
contained in a given binary relation.

We fix $\theFunction$, $\theArity$, and a binary relation~$\theTI$ for
the rest of this paper.
Furthermore, we assume that $\theTI$ is represented as an assertion
over tuples of variables $(\mathtt{a}_1, \dots, \mathtt{a}_\theArity)$
and $(\mathtt{m}_1, \dots, \mathtt{m}_\theArity)$.
This assumption will be used in Figure~\ref{fig-tsel}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_relwork.tex
Added example paper for latex guidelines.

\section{Related work}
\label{sec:relwork}

% \todo{TODO: WHAT THEY DO, WHAT WE CAN DO THAT THEY CAN, WHAT WE CAN DO
%   THAT IS AN IMPROVEMENT OVER THEM}

\newcommand{\oldRelWork}{false}
\ifthenelse{\equal{\oldRelWork}{true}}{

In this section we describe existing techniques for the
analysis and verification of functional programs.

\newcommand{\lfts}{\ensuremath{\tsf{LF}^\diamond}\xspace}
\paragraph{Resource bounds analysis.}

Hofmann et. al.~\cite{Hofmann11,Hofmann03} present type systems and
corresponding type inference algorithms for the analysis of
quantitative resource bounds.  The computed bounds are represented by
linear or polynomial expressions obtained by solving a linear
program based encoding of the type constraints.  The type system
keeps track of the resource consumption using an augmentation of type
judgements by resource counters.  These counters can be viewed as
monitor states that are embedded into type judgements.  In
contrast, our monitor states are embedded into the program.

The type inference algorithm of \cite{Hofmann11} can be used in combination
with our product construction to infer quantitative properties of
monitored programs. 
On the other hand, extending \cite{Hofmann11} to keep track of other
properties requires a redesign of the type system and the invention of
the corresponding inference algorithm.

\paragraph{Resource usage verification.}
The resource usage verification problem consists in checking that the
resource access traces generated by a program satisfy a
specification~\cite{Kobayashi09}.  Several works address the problem
of resource usage verification under different names,
e.g.~\cite{Terauchi02,Kobayashi05,Kobayashi09,Kobayashi11}.

Kobayashi~\cite{Kobayashi09} presents an algorithm for checking
properties given as finite state tree automata over a finite alphabet.
In constrast, we do not give a checking algorithm since we deal with
infinite state systems. Instead, we rely on the
reachability/termination tool \dsolve~\cite{Dsolve}. Our monitors can
check non-regular properties with the help of a
reachability/termination backend.

Foster~et.~al.~\cite{Terauchi02} presents a type inference system
for checking properties given as assume/assert program
annotations. Their annotations are over atomic predicates on program
values called \emph{type qualifiers}. Their type inference system is
similar to the liquid type inference system of \dsolve. In contrast to
our work, they can handle functional programs with references.

% addresses the resource usage verification
% problem for higher-order functional programs.
% % The resource usage
% % verification problem consists in verifying that the resource access
% % traces given by a program satisfy a specification. A specification is
% % a language of infinite words on a finite alphabet.
% Kobayashi gives a novel model checking technique for finite-state
% programs. Their technique applies a type checking approach to the
% model checking of higher-order recursion schemes corresponding to user
% programs. The technique is applied on $\omega$-regular safety
% properties. In contrast, our work enables reasoning about temporal
% properties of infinite-state programs using existing
% techniques. Combining their technique with approach is an interesting
% research goal.

% Our tool can verify the examples given in
% Section~3.1 of~\cite{Kobayashi09} (see experiments 15-18,43,44 in Figure~\ref{fig-experiments}) but we failed to verify the 2-place
% buffer in Example~6.1 of~\cite{Kobayashi09} because constructing the
% corresponding monitor specification requires ingenuity. Our work and
% Kobayashi's are different in that our we provide a framework for
% expressing temporal properties that can be verified with existing
% tools and Kobayashi developed a full verification framework.

% Kobayashi \cite{Kobayashi09} applied HORS as an abstract model of
% higher order functional programs. Kobayashi presented a type based
% algorithm for checking that HORS satisfy finite state machine-based
% specifications. There are two main differences to our work. First, we
% target \ocaml\ programs, while Kobayashi's considers a prototype
% functional language. Second, we deal with properties represented by
% infinite state monitors.

\paragraph{Contract checking.}

Contracts are pre- and post-condition specifications for
functions. Xu created a verification tool \cite{DanaPhd} for Haskell that is
based on contracts and partial evaluation. 
Pre- and post-conditions can only refer to the program execution
before and after procedure calls. In contrast, our monitor
specifications have the ability of observing program executions at all times.

\paragraph{Liquid typing.}

\dsolve\ \cite{Dsolve} is a reachability checker based on refinement
type inference.  The type inference algorithm consists of two
parts. First, a set of constraints over refinement predicates is
generated from program code. Second, an iterative algorithm tests
candidate solutions constructed from a set of predicate schemes in the
theory of linear arithmetic and uninterpreted functions. While \dsolve\
is effective for the verification of assertion validity, our product
construction allows \dsolve\ to verify arbitrary safety properties.

}{

\paragraph{Termination and control-flow analysis}

Traditionally, termination analysis of higher-order programs is
developed on top of a control-flow
analysis~\cite{Sereni05terminationanalysis,SereniAPLAS05,SereniICFP07}.
Our approach relies on the applied safety checker to keep track of
control flow. 
In principle, our transformation could benefit from the results of
control-flow analysis, as discussed in Section~\ref{sec:conclusion}. 
Practically, such additional information was not necessary when
proving termination of all examples presented
in~\cite{Sereni05terminationanalysis,SereniAPLAS05,SereniICFP07}.
Adding a control-flow analysis
pass~\cite{Shivers88,MightJFP08,MightPOPL11,MightSCHEME2010} before the transformation
would be akin to the application of (function) pointer analysis for
imperative programs. 
We leave a study of such an integration as future work.

  \paragraph{Abstraction for termination.} The size change termination
  argument~\cite{SCTPOPL2001} can be extended to higher-order functional
  programs, see e.g.~\cite{JonesBohr04,SereniAPLAS05}. This argument requires checking the presence of an
  infinite descent in values passed to application sites of the program
  on any infinite traversal of the call graph. In contrast, our
  approach can keep track of a rank descent in arbitrary expressions
  over program values.
  In principle, SCT can be seen a specific abstract domain that yields
  termination arguments related to disjunctive
  well-foundedness~\cite{PodelskiSAS10}.
  We leave the question if general abstraction techniques for
termination~\cite{CousotHigherOrder94,CousotPOPL12} can be reduced by
an appropriate source to source transformation as furture work.

\paragraph{Contract checking.}

Contracts are pre- and post-condition specifications for functions. Xu
created a verification tool \cite{DanaPhd} for Haskell that is based
on contracts and partial evaluation.  Their approach works well for
checking safety properties, and can even detect divergence of
programs. In contrast, our approach is specialized to the verification
of termination, and is a step towards the automated verification of
termination through counterexample guided abstraction refinement.

\paragraph{Liquid typing.}

\dsolve\ \cite{Dsolve} is a reachability checker based on refinement
type inference.  The type inference algorithm consists of two
parts. First, a set of constraints over refinement predicates is
generated from program code. Second, an iterative algorithm tests
candidate solutions constructed from a set of predicate schemes in the
theory of linear arithmetic and uninterpreted functions. As our
experiments show, the composition of our transformation and \dsolve\ 
yields a binary reachability analysis tool.

}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_semantics.tex
Added example paper for latex guidelines.

\section{Binary reachability on evaluation trees}
\label{sec-semantics}

In this section we make the first step towards our program
transformation.
We present an augmentation of evaluation trees that allows us to
reduce the binary reachability analysis to the validity analysis of
annotated judgement.
Each judgement is augmented with a Boolean and an $\theArity$-ary
tuple of values, which we will refer to as a \emph{state}.

\input{fig_ex_monitoring}

Before presenting the augmentation procedure, we consider examples of
tree augmentation shown in Figure~\ref{fig-monitoring}.
The root of the tree is augmented with a state $\vstf = (\lfalse, 0)$,
where $\lfalse$ indicates that no argument has been used for the
augmentation yet.
We use $\vst$ to augment judgements in the subtree for the branches
that do not correspond to the evaluation of the body of~\texttt{f}.
When augmenting the subtree that deals with the body, we can
nondeterministically decide to start augmenting with a state that
records the argument of the current application.
That is, in the body subtree we augment with the state $\vstone =
\stone$.
Here, $\ltrue$ indicates that we took a snapshot of the current
application argument, and $1$ is the argument value.
The remaining judgements are augmented with $\vstone$, since we will
not change the snapshot if it was taken, i.e., if the first component
of the augmenting state is~$\ltrue$.

We proceed with an algorithm \monitor that takes as input an initial
state and an evaluation tree and produces an augmented tree.
Each augmented judgement is of the form
\mevj{\evj{\ectx}{\expr}{\val}}{\stupa}{\st}, where \st\ is a state.
As an \emph{initial} state we take a pair $(\lfalse, (v_1 \dots,
v_\theArity))$ where $v_1\ \dots, v_\theArity$ are some values.

See Figure~\ref{fig-monitor-alg}.
\monitor traverses the input tree recursively, by starting from the
root.
Whenever the current judgement is a
$\theFunction/\theArity$-application judgement, then we choose whether
to create a snapshot of the arguments and store them in the state that
is used to augment the subtree that evaluates the body.
We only create a snapshot if the Boolean component of the current
state is~$\lfalse$.
In case we currently do not deal with a
$\theFunction/\theArity$-application judgement, no state change
happens and we proceed with the subtrees.
Once we obtain the augmented versions of the subtrees, we put them
together by creating a node that connects the roots of the subtrees.

  % \begin{tabular}{r@{\ }l l}
  %   $\mevjs \ni$ & \mevjvar \;\; ::=
  %   \mevj{\evj{\ectx}{\expr}{\val}}{\stupa}{\stdown}
  %   & monitored judgement \\[\jot]
  %   $2^{{\mevjs}^+} \ni$ & \mevt  & monitored evaluation tree
  % \end{tabular}

\input{fig_monitor_alg}

We establish a formal relationship between the
$\theFunction/\theArity$-recursion relation with the augmented
judgements obtained by applying \monitor using the following theorem.
%
\begin{theorem}[\monitor keeps track of
$\theFunction/\theArity$-recursion relation]
\label{thm-monitor}
%
A pair $(v_1, \dots, v_\theArity)$ and~$(u_1, \dots, u_\theArity)$ is
in the $\theFunction/\theArity$-recursion relation if and only if the
result of applying \monitor wrt.\ some sequence of nondeterministic
choices on \evjvar and an initial state contains an augmented
judgement of the form
%
\begin{center}
  \mevj{\evj{\ectx}{$\theFunction$\ \expr[1] \dots
      \expr[\theArity]}{\val}}{\stupa}{
    (\ltrue, (v_1, \dots, v_\theArity))}
\end{center}
%
such that for each $i \in 1..\theArity$ we have $\tsf{eval}\ \ectx\
\expr[i] = u_i$
%
\end{theorem}
%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/workspace/main.tex

@@
 \tikzstyle{print}=[draw, trapezium, trapezium left angle=70, trapezium right angle=-70]
-\tikzstyle{every edge}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]
+\tikzstyle{every path}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]
 
@@
 \begin{center}
-\begin{tikzpicture}[node distance=2cm]
-  \node[place, tokens=1, label=left:a] (a) {};
-  \node[place, tokens=0, label=left:b, below of=a] (b) {};
-  \node[place, tokens=0, label=right:$\neg bit2$, right of=b, xshift=-1cm, yshift=+1cm] (nbit2) {};
-  \node[transition, label=left:t1, above of=a, yshift=-1cm] (t1) {}
-    edge [pre, bend left, min distance=2cm, out=225, in=-45] (b)
-    edge [post] (a);
-  \node[transition, label=left:t2] (t2) at ($(a)!0.5!(b)$) {}
-    edge [pre] (a)
-    edge [pre] (nbit2)
-    edge [post] (b)
-    edge [post] (nbit2);
+\begin{tikzpicture}[node distance=1cm]
+  \node[transition, label=left:t1] (t1) {};
+  \node[place, tokens=1, label=left:a, below of=t1] (a) {};
+  \node[transition, label=left:t2, below of=a] (t2) {};
+  \node[place, tokens=0, label=left:b, below of=t2] (b) {};
+  \node[place, tokens=0, label=right:$\neg bit2$, right of=t2] (nbit2) {};
+  \draw (t1) edge[pre, bend left, min distance=2cm, out=225, in=-45] (b);
+  \draw (t1) edge[post] (a);
+  \draw (t2) edge[pre] (nbit2);
+  \draw (t2) edge[post] (nbit2);
+  \draw (t2) edge[pre] (a);
+  \draw (t2) edge[post] (b);
 \end{tikzpicture}
@@
 
-\iffalse
-
 \newpage
@@
   Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A.
+  Subprocedure Inv constructs invariant I corresponding to N and A'.
 \end{verbatim}
@@
     \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
-      $I := \text{Inv}(N, A)$};
+    \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
+      $I := \text{Inv}(N, A')$};
     \node[print, below=of inv] (printinv) {Invariant: $I$};
@@
 
-  - With weakened property P \/ (\neg \delta)
+  - Dual state constraints C':
 
-  - Negated property is \neg P /\ \delta
+    0        >= - b
+    0        >= (- a) + b
 
-  - Dual state constraints C':
-    0 >= (- cs1) + id1 + idle1 + (- nid1)
-    0 >= (- id1) + (- idle1) + nid1 + req1
-    0 >= cs1 + (- req1)
-    0 >= after_you2 + (- req2)
-    0 >= (- after_you2) + await2 + id2
-    0 >= (- await2) + idle2
-    0 >= cs2 + (- req2)
-    0 >= (- cs2) + id2 + idle2
-    0 >= (- id2) + (- idle2) + req2
+    a        <  2 * target_1
 
-    id1 + id2 + idle1 + idle2 < 2 * target_1 + trap_1
+    a        >= target_1
+    b        >= target_1
+    nbit2    >= 0
+    target_1 >= 0
 
-    id1        >= trap_1
-    cs1        >= target_1
-    idle1      >= 0
-    req1       >= trap_1
-    nid1       >= 0
-    id2        >= trap_1
-    cs2        >= target_1
-    idle2      >= 0
-    req2       >= trap_1
-    after_you2 >= trap_1
-    await2     >= 0
-    target_1   >= 0
-    trap_1     >= 0
+  - Model A':
 
-  - Model A:
-
-    id1        = 1
-    cs1        = 1
-    idle1      = 0
-    req1       = 1
-    nid1       = 0
-    id2        = 1
-    cs2        = 1
-    idle2      = 0
-    req2       = 1
-    await2     = 0
-    after_you2 = 1
+    a        = 1
+    b        = 1
+    nbit2    = 1
     target_1 = 1
-    trap_1     = 1
 
@@
 
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
-
-  - Invariant with trap constraint \delta added:
-
-    id1 + id2 + req1 + req2 + after_you2             >= 1
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
+    a + b + nbit2 <= 1
 
@@
 \verb=P= when \verb=N= never violates \verb=P=.
-We present the method Invariant with Minimization by example on a
-simplified mutual exclusion algorithm by Delzanno2001.
 
@@
 
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A
+  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A'
 \end{verbatim}
@@
     \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (model1) {$A:=\text{Model}(C')$};
-    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A)$};
+    \node[action, below=of satc] (model1) {$A':=\text{Model}(C')$};
+    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A')$};
     \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
-    \node[action, right=of satcm] (model2) {$A:=\text{Model}(C' \cup C_M)$};
-    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A)$};
+    \node[action, right=of satcm] (model2) {$A':=\text{Model}(C' \cup C_M)$};
+    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A')$};
     \node[print, below=of inv] (printinv) {Invariant: $I$};
@@
 * Example:
-  - Dual state constraints C' and satisfying assignment A for C' as in example for method Invariant
+  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
 
-  - Minimization constraints:
+  - Minimization constraints C_M:
 
-    nc  ~= 0 => m_nc = 1
-    |           --------
+    a     > 0 => b_a     = 1
+    |            -----------
     |               |
@@
 
-    nc   = 0 => m_nc  = 0
-    |           ---------
+    a     = 0 => b_a     = 0
+    |           ------------
     |               |
@@
 
-    m   ~= 0 => m_m   = 1
-    m    = 0 => m_m   = 0
-    cs1 ~= 0 => m_cs1 = 1
-    cs1  = 0 => m_cs1 = 0
-    cs2 ~= 0 => m_cs2 = 1
-    cs2  = 0 => m_cs2 = 0
+    b     > 0 => b_b     = 1
+    b     = 0 => b_b     = 0
+    nbit2 > 0 => b_nbit2 = 1
+    nbit2 = 0 => b_nbit2 = 0
 
-    m_nc + m_m + m_cs1 + m_cs2 < 3
-    --------------------------   |
+    b_a + b_b + b_nbit2 < 3
+    -------------------   |
                 |         |
-                |    number of places appearing in current invariant for A
-                |    = #{s | A(s) ~= 0} = #{m, cs1, cs2}
+                |    number of places appearing in current invariant for A'
+                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
                 |
@@
   number of places appearing in new invariant
+
+  - Model A' for C' \cup C_M:
+
+    a        = 1
+    b        = 1
+    nbit2    = 0
+    target_1 = 1
+
+  - Minimized Invariant:
+
+    a + b <= 1
+
 \end{verbatim}
@@
  N   = (S, T, E, M0)
-     = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- A   = ( nc=0, m=1, cs1=1, cs2=1, target1=1 )
+ A   : Satisfying assignment for C'
 
 TODO: Unroll pseudocode?
+
+\end{verbatim}
+
+\newpage
+
+\section{Method Invariant by Refinement}
+
+\begin{verbatim}
+* Code, property, and Petri net: same as in section Method Safety by Refinement.
+
+* Example:
+
+  - Trap constraints as in example from section Method Safety by Refinement.
+
+  - Dual state constraints C':
+
+    0        >= - b
+    0        >= (- a) + b
+
+    a        <  target_1 + trap_1
+
+    a        >= trap_1
+    b        >= target_1
+    nbit2    >= trap_1
+    target_1 >= 0
+    trap_1   >= 0
+
+  - Model A':
+
+    a        = 1
+    b        = 1
+    nbit2    = 1
+    target_1 = 1
+    trap_1   = 1
+
+  - Invariant:
+
+    a     + nbit2 >= 1
+    a + b + nbit2 <= 1
 
@@
 
-\fi
-
 \end{document}

### 2014-02-03 paper/workspace/main.tex
Pushing the preamble.

@@
-\documentclass{article}
+\documentclass{llncs}
 
-\usepackage{amsmath}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{listings}
-\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-\usepackage{graphicx}
-
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
-	snakes,automata,backgrounds,petri,calc,positioning}
-
-\newcommand{\bfc}{{\sc Bfc}}
-\newcommand{\mist}{{\sc Mist}}
-\newcommand{\iic}{{\sc IIC}}
-\newcommand{\pnerf}{{\sc Pnerf}}
-\newcommand{\zthree}{{\sc Z3}}
-\newcommand{\ttt}[1]{\texttt{#1}}
-
-\tikzstyle{place}=[circle,thick,draw=black!75,fill=white!20,
-  	minimum size=6mm]
-\tikzstyle{red place}=[place,draw=red!75,fill=red!20]
-\tikzstyle{blue place}=[place,draw=blue!75,fill=blue!20]
-\tikzstyle{transition}=[rectangle,thick,draw=black!75,
-  	fill=black!20,minimum size=4mm]
-
-\tikzstyle{state}=[draw, ellipse, aspect=2]
-\tikzstyle{action}=[draw, rectangle, align=center]
-\tikzstyle{decision}=[draw, diamond, aspect=2, align=center]
-\tikzstyle{print}=[draw, trapezium, trapezium left angle=70, trapezium right angle=-70]
-\tikzstyle{every path}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]
+\input{preamble}
 
@@
 
-\title{Report on a Safety Checker for Petri nets}
-\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
-% \institute{Technische Universit\"at M\"unchen}
-\date{Mon Nov 25 14:24:10 CET 2013}
+\ifthenelse{\equal{\isDraft}{false}}{
+  \title{Verification of Petri Nets Using Linear Programming}
+  \author{
+    Javier Esparza\inst{1}
+    \and
+    Ruslán Ledesma-Garza\inst{1}
+    \and
+    Rupak Majumdar\inst{2}
+    \and
+    Philipp Meyer\inst{1}
+    \and
+    Filip Niksic\inst{2}
+  }
+  \institute{
+    Technische Universität München
+    \and
+    MPI-SWS, Kaiserslautern and Saarbrücken
+  }
+  \maketitle
+}{}
 

### 2014-02-03 paper/workspace/main.tex
Title and abstract.

@@
 \ifthenelse{\equal{\isDraft}{false}}{
-  \title{Verification of Petri Nets Using Linear Programming}
+  \title{An SMT-based Approach to Coverability Analysis}
   \author{
@@
     \and
-    Ruslán Ledesma-Garza\inst{1}
+    Rusl\'{a}n Ledesma-Garza\inst{1}
     \and
@@
   \institute{
-    Technische Universität München
+    Technische Universit\"at M\"unchen
     \and
-    MPI-SWS, Kaiserslautern and Saarbrücken
+    MPI-SWS, Kaiserslautern and Saarbr\"ucken
   }
+
   \maketitle
-}{}
+
+  \begin{abstract}
+    Model checkers based on Petri net coverability have been used
+    successfully in recent years to verify safety properties of concurrent
+    shared-memory or asynchronous message-passing software. The bottleneck
+    in the verification process is the iterative computation of the
+    (backward) reachable or (forward) coverable set, and most recent
+    research has focused on non-trivial heuristics to speed it up.
+
+    In this paper, we revisit a constraint approach to the problem based on
+    classical Petri net analysis techniques: marking equation, place
+    invariants, and traps. Given a marking M to be covered, the approach
+    constructs constraints corresponding to increasingly precise
+    overapproximations of the reachable markings covering M. If one of
+    these constraints is unsatisfiable, then M is not coverable.
+
+    Previous work on this approach suffered from the absence of efficient
+    decision procedures for linear arithmetic, which nowadays exist as a
+    part of SMT solvers. We show how to utilize an SMT solver to implement
+    the constraint approach, and additionally, to generate a minimized
+    inductive invariant from a safety proof.
+
+    We empirically evaluate our procedure on a large set of existing Petri
+    net benchmarks arising out of software verification. Surprisingly, even
+    though our technique is incomplete, we show that it can quickly
+    discharge all safe instances, usually competitive with, and occasionally
+    much faster than, state-of-the-art techniques. Additionally, the
+    inductive invariants computed are usually orders of magnitude smaller
+    than those produced by existing solvers.
+  \end{abstract}
+}{
+  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
+}
+
+\section{Introduction}
+
+This paper makes the following contributions.
+\begin{itemize}
+\item An application of \zthree\ to the constraint approach and its
+  corresponding evaluation.
+\item A method that given a network and a property constructs a
+  minimal invariant.
+\end{itemize}
 

### 2014-02-03 paper/workspace/main.tex
Added pseudocode for Invariant methods

--- 2014-02-03 paper/workspace/main.tex
@@
     \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
+    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
     \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
@@
 
-    0        >= - b
+    0        >=    a  - b
     0        >= (- a) + b
@@
   Input:
-    (S, T, E, M0) : Petri net
-    P             : Property
+    N = (S, T, E, M0)  : Petri net
+    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
+              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
   Output:
@@
  N      = (S, T, E, M0)
-   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- P = ( cs1 + cs2 <= 1 )
+        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
+\neg P  = ( a + b >= 2 )
 
-TODO: Unroll pseudocode?
+Pseudocode:
+\end{verbatim}
 
+\begin{align*}
+  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
+                      \sum_{(t, s) \in E} s
+                    - \sum_{(s, t) \in E} s \right) \right) \land
+     \left( \sum_{s \in S} M_0(s) \cdot s <
+       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
+     & \left( \bigwedge_{s \in S} \left ( s \ge 
+       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
+     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
+\end{align*}
+
+\begin{verbatim}
 * Subprocedure Inv:
@@
   Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C'
+    N = (S, T, E, M0) : Petri net
+    A'                : Satisfying assignment for C'
 
@@
 
-TODO: Unroll pseudocode?
+  Pseudocode:
 \end{verbatim}
+
+\begin{align*}
+  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
+                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
+\end{align*}
 
@@
 \begin{verbatim}
-* Property of minimization constraints C_M generated from A: 
-    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
+* Property of minimization constraints C_M generated from A': 
+    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
 \end{verbatim}
@@
                 |
-                |
   number of places appearing in new invariant
@@
 
-\end{verbatim}
-
-\begin{verbatim}
 * Subprocedure \text{MinConstraints}:
@@
   Input:
-    (S, T, E, M0) : Petri net
-    A             : Assignment
+    N = (S, T, E, M0)  : Petri net
+    A'                 : Satisfying assignment for C'
   Output:
@@
 
-Input:
- N   = (S, T, E, M0)
- A   : Satisfying assignment for C'
-
-TODO: Unroll pseudocode?
+  Pseudocode:
 
 \end{verbatim}
+
+\begin{align*}
+  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
+      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
+    \right) \right) \land
+    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
+\end{align*}
 
@@
 
+* Method Invariant by Refinement
+\end{verbatim}
+
+\begin{tikzpicture}
+  \node[state] (begin) {BEGIN};
+  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
+                                        $D:=\{\}$};
+  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
+  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
+                                        $C_{\theta}:=\text{TrapConditions}(N, A)$};
+  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
+  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
+                                                $\delta:=\Delta(A_\theta)$\\
+                                                $C:=C \cup \{\delta\}$\\
+                                                $D:=D \cup \{\delta\}$};
+  \node[print, right=of satctheta] (noinv2) {No Invariant};
+  \node[state, below=of noinv2] (end3) {END};
+
+  \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, \neg P \land D)$};
+  \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
+  \node[action, below=of satcprime] (inv) {$A':=\text{Model}(C')$\\
+                                        $I:=D \land \text{Inv}(N, A')$};
+  \node[print, below=of inv] (printinv) {Invariant: $I$};
+  \node[print, right=of satcprime] (noinv) {No Invariant};
+  \node[state, below=of noinv] (end1) {END};
+  \node[state, below=of printinv] (end2) {END};
+
+  \draw (cprime) edge (satcprime);
+  \draw (satcprime) edge node[above]{NO} (noinv);
+  \draw (satcprime) edge node[right]{YES} (inv);
+  \draw (inv) edge (printinv);
+  \draw (noinv) edge (end1);
+  \draw (printinv) edge (end2);
+
+  \draw (begin) edge (c);
+  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
+  \draw (satc) edge node[above]{NO} (cprime);
+  \draw (satc) edge node[right]{YES} (modelc);
+  \draw (modelc) edge (satctheta);
+  \draw (satctheta) edge node[above]{NO} (noinv2);
+  \draw (noinv2) edge (end3);
+  \draw (satctheta) edge node[right]{YES} (modelctheta);
+  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
+  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
+\end{tikzpicture}
+
+\newpage
+
+\begin{verbatim}
 * Example:

### 2014-02-03 paper/example-latex/rebuttal-response.txt

Thank you for the comments and suggestions!

Review 1:

The current set of benchmarks was suggested to us by experts on checking termination of higher order programs and comes from the suggested literature.
We realize that a larger set of benchmarks provides a better insight into pros and cons of the approach, we are working on it.

Our previous experience with the Terminator tool for proving termination of C routines shows that while a collection of simple tricks can deal with a large variety of simple examples. Nevertheless, for more complicated programs it is important to have a tool that can deal with general cases as well, which is the goal of our approach.
Due to its generality, our approach allows to implement the size change termination algorithm by relying on the correspondence between SCT and transition invariants [Jones et al, SAS 2010].

Regarding Fig 3, we will formalize augment as a tree transducer thus making it applicable to evaluation trees for diverging expressions.

Review 2:

Kraus et al focus on finding ranking functions by composing, e.g., via lexicographic composition or SCT approach, predefined well-founded measures manually given to datatypes (which are usually recorded in Isabelle theories). 
Usually predefined measures work well, but when more complex measures, e.g., some weighted combination thereof needs to be discovered, an automatic method is very desirable [Nipkow, personal communication].

Review 3:

Transformed programs contain a significant portion of expressions that pass state around without changing it. To deal with larger examples we are implementing a beta-reduction based program simplification that removes such expressions.
DSolve relies on logical qualifiers, akin to a set of predicates used in predicate abstraction of imperative programs. When dealing with complex termination arguments the default set of qualifiers used by DSolve is not expressive enough. An abstraction refinement technique for termination, e.g., Cook et al SAS 2005, could be applied here.
For a language with references, our notion of binary reachability and instrumentation technique would not change much. The biggest limitation is the ability of the reachability checker to handle imperative features when references are passed around in closures, which is a research challenge. For references that act as “globals”, a transformation that passes them in a state monad can be useful.
Regarding current limitations, here is an example:

let rec f g = if g () < 0 then 0 else f (fun () -> g () - 1);;

While in theory our approach can prove termination of f/1, our implementation cannot show this program terminating for all g's because:
1) we cannot look inside at the closure corresponding to g using instrumentation code.
2) we do not want to apply g inside a transition invariant check, because g may modify the state of the program (e.g. may modify a reference).

Review 4:

The current set of benchmarks was suggested to us by experts on checking termination of higher order programs and comes from the suggested literature.
We realize that a larger set of benchmarks provides a better insight into pros and cons of the approach, we are working on it.

### 2014-02-03 paper/example-latex/abstract.tex

%

A number of recent approaches for proving program termination rely on
transition invariants -- a termination argument that can be
constructed incrementally using abstract interpretation.
These approaches use binary reachability analysis to check if a
candidate transition invariant holds for a given program.
For imperative programs, its efficient implementation can be obtained
by a reduction to reachability analysis, for which practical tools
are available.
In this paper, we show how a binary reachability analysis can be put
to work for proving termination of higher order functional programs.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/main.tex

% \documentclass[nocopyrightspace]{sigplanconf}
%\documentclass{llncs}
\documentclass{llncs}

\sloppy

\input{preamble}

\begin{document}

% \conferenceinfo{WXYZ '05}{date, City.} 
% \copyrightyear{2005} 
% \copyrightdata{[to be supplied]}

% \titlebanner{}
% \preprintfooter{}

% \pagestyle{headings}

\title{Binary Reachability Analysis of Higher Order Functional Programs}
%\title{Temporal Monitoring and Verification of Higher Order Functional Programs}
\author{Rusl\'{a}n Ledesma-Garza \and Andrey Rybalchenko}
\institute{Technische Universit\"at M\"unchen}

\maketitle

\begin{abstract}
%

\input{abstract}
\iffalse
The automata-theoretic approach to program verification is an
effective and uniform way of dealing with complex temporal properties.
The crux of the approach is to represent the property by a monitor
that inspects program computations.
The monitor keeps record of the (intermediate) inspection results, and
if the property violation is detected then an appropriate note is
recorded.
Then, the monitor is composed with the program, which produces a
product program that monitors its own computation and can be analyzed
using tools for proving reachability and (fair) termination (i.e.,
sophisticated temporal reasoning is no longer necessary).
While already ubiquitous for dealing with imperative programs, monitor
based approaches were recently applied for termporal verification of
higher order functional programs by using higher order recursion
schemes (HORS), e.g.~\cite{Ong06,Kobayashi09}.

In this paper, we present an alternative approach to monitoring and
verifying higher order functional programs that applies the monitor
composition directly on the program source code (instead of composing
with a HORS derived from the program by applying finite state
abstraction on data types).
We formulate our approach for a \miniocaml language as a program
transformation algorithm and present its implementation.
Our experimental evaluation in combination with the DSolve verifier
for OCaml~\cite{DSolveCAV} shows that the scalability and
effectiveness of existing abstraction-based assertion checking tools
can be effectively leveraged for dealing with temporal safety and
liveness properties of higher order functional programs.

The automata-theoretic approach is a standard method for dealing with
complex temporal properties of imperative programs and only requires
program verifiers that deal reachability and (fair) termination
properties.
The automata-theoretic semantics of imperative programs directly
supports monitoring and product construction.
Until now, the temporal reasoning about functional programs does not
rely on monitoring and reduction to reachability and (fair)
termination.
Instead, the reasoning is executed using type systems and type
inference algorithms that are developed/customized for the specific
property at hand.
\fi

\iffalse

In this paper, we present an application of the automata-theoretic
approach to the temporal verification, i.e., safety and liveness, of
\miniocaml\ programs with higher-order procedures.  
First, we present the notion of monitoring for \miniocaml\ computations
given as evaluation trees, and define monitor specifications that can be
composed with \miniocaml\ programs.
Second, we present a product construction algorithm that composes an
\miniocaml\ program with a monitor specification in the presence of
higher order procedures.
Third, we present an implementation of the proposed approach that uses
the \dsolve\ reachability checker, together with its experimental
evaluation on examples from the literature.
The examples include resource analysis and termination checking tasks.
\fi
\end{abstract}

% \mbox{ }
% \\ \\
%\nocite{Hofmann00,Hofmann02,Hofmann03,Kobayashi09}

\input{sec_introduction}
\input{sec_illustration}
\input{sec_preliminaries}
\input{sec_semantics}
%\input{sec_logger}
\input{sec_transformation}
%\input{sec_trans_assert}
\input{sec_experiments}
\input{sec_relwork.tex}
\input{sec_conclusions}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2014-02-03 paper/example-latex/sec_conclusions.tex

\section{A limitation and future work}
\label{sec:conclusion}

Compared to imperative programs, higher order functions impose
additional complication on the binary reachability analysis.
The major current limitation of our approach lies in the treatment of
partial applications when arguments are provided at different program
points, as illustrated by the following example. 
%
\begin{small}
\begin{verbatim}
let rec f x = if x > 0 then f (x - 1) else fun y -> y + 1 in
let g = f 1 in
g 2
\end{verbatim}
\end{small}
%
Proving termination of $\mathtt{f}/2$ applications is not possible,
since the second argument to \texttt{f} is given indirectly through an
application of~\texttt{g}. 
Without a control-flow analysis, we cannot store \texttt{2} in the
auxiliary state since it is not known syntactically that \texttt{2} is
the second argument for~\texttt{f}.
Removing the above limitation is an important step for future work,
which can be accomplished either by relying on results of a
control-flow analysis, e.g.~\cite{MightPOPL11}, that is performed
before the transformation takes place.
In simple cases, by using the following transformation that is based
on $\beta$ reduction.
%
\begin{small}
\begin{verbatim}
let rec f x = if x > 0 then f (x - 1) else fun y -> y + 1 in
(* let g = f 1 in *)
(f 1) 2
\end{verbatim}
\end{small}
%
In the transformed program, both arguments are given to \texttt{f}
directly.

% In the future we would like to integrate our binary reachability
% analysis in a counter example guided abstraction refinement loop,
% following the steps of~\cite{copytrick}.

% Our monitorability condition on expressions interferes with modularity
% in the following two ways. Consider a user program \expr\ with a free
% variable representing a library value, say \ttt{List.map~:~('a->'b)->'a~list->'b~list}. 
% \product\ can construct a product \exprprod\ for \expr. 
% First, \exprprod\ can only be used for runtime verification in the presence
% of a monadic value \ttt{List.map\_m} of type
%  \ttt{\monadic~('a->'b)->'a~list->'b~list}. Since \product\ can deal
%  with arbitrary higher-order function declarations, a solution is to
%  prepend \expr\ with the definition of \ttt{List.map}. Second, the
%  static reasoning about \exprprod\ can only be done under given
%  assumptions about \ttt{List.map}. A solution is to use tool-specific
%  codifications of assumptions on free variables. As shown in the
%  illustration section and our experiments, \dsolve\ provides such a mechanism.

% % \begin{remark}[TODO: Higher-order functions]
% %   \product\ can construct products for expressions featuring
% %   higher-order functions. Given a higher-order abstraction, \product\
% %   constructs a higher-order abstraction.  Higher-order abstractions
% %   given by \product\ can only be applied to product parameters. Thus,
% %   runtime verification of higher-order products can only be done in
% %   the presence of product parameters. Static verification of
% %   higher-order products can be done by making assumptions on given
% %   product parameters. \dsolve\ provides a ``value specifications'' for
% %   making these assumptions~\cite{Dsolve}.
% % \end{remark}

% We leave as future work the development of a user-friendly language
% specification based on temporal logics.

% \paragraph{Future work}
% \begin{enumerate}
% \item User friendliness of monitor specifications.
% \item Cannot assign blame to original locations of the program neither
%   on runtime nor statically. \dsolve\ only provides 
% \end{enumerate}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_experiments.tex

\section{Experimental evaluation}
\label{sec-experiments}

In this section we describe our implementation and the corresponding
experimental evaluation.

\paragraph{Implementation.}
We implemented \product\ as an extension to the \camlp\
parser~\cite{Camlp4-dist}. Our implementation takes as input a user
program and a specification consisting of a function name, an arity,
and candidate transition invariant. Our implementation produces a
transformed program following the procedure depicted in
Figure~\ref{fig-ex-prod}.

\paragraph{Experiments.}
\input{fig_experiments}
Our experiments consisted of two steps. First we applied our
transformation to the set of benchmarks summarized in
Figure~\ref{fig-experiments}. Then we analyzed the transformed
benchmarks using the reachability checker \dsolve~\cite{Dsolve, DsolveCAV}.
\ifthenelse{\equal{\isTechReport}{true}}{
The set of benchmarks is available at~\ttt{http://www7.in.tum.de/\~{}ruslan/binreach/}.
}{
The set of benchmarks is available at~\cite{FunVTechReport}.
}
Our benchmarks feature higher order functions and algebraic
data types (lists).  We summarize our verified benchmarks in
Figure~\ref{fig-experiments}.
Benchmarks 11-15 correspond to the higher-order programs
in~\cite{Sereni05terminationanalysis} that are strict and
\mbox{type-check}.
The experiments show that our transformation can be used together with a state of
the art static analyzer to prove termination of higher-order programs found
in the literature.

% \paragraph{Evaluation.}
% The experiments show that our transformation can be used together with a state of
% the art static analyzer to prove termination of higher-order programs found
% in the literature.

% For example, consider the higher-order benchmark \ttt{sereni29} shown below.
% \begin{verbatim}
% let rec length l =
%   match l with
%     | [] -> 0
%     | _ :: t -> 1 + length t
% let rec map f xs =
%   match xs with
%     | [] -> []
%     | y :: ys -> f y :: map f ys in
% let compose f g x = f (g x) in
% let add x y = x + y in
% map (compose (add 1) (add 2)) [0;1]
% \end{verbatim}
% We applied our program transformation on \ttt{sereni29} using the function
% identifier \ttt{map}, the number of parameters $2$, and the transition
% invariant candidate $\ttt{length}(a_2) \geq 0 \wedge
% \ttt{length}(m_2) > \ttt{length}(a_2)$. Then we obtained a
% transformed program that we verified correct using \dsolve.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_illustration.old.tex

\section{Examples}

In this section we show several functional programs that illustrate
complications that arise when proving termination.  We also highlight
how our implementation applies binary reachability analysis to address
such complications.

We write our examples in OCaml, and use \texttt{nondet ()} to model
non-deterministic input, which is useful for abstracting complex
expressions that are irrelevant for our reasoning process.

\paragraph{\bf Non-well-founded data}

The program \texttt{choice} is shown below.
%
\begin{center}
\begin{minipage}[h]{.5\linewidth}
\begin{small}
\begin{verbatim}
let rec choice (x, y) = 
  if x > 0 && y > 0 then
    if nondet () then 
      choice (x-1, x)
    else 
      choice (y-2, x+1)
  else 
    ()
in
choice (read_int (), read_int ())
\end{verbatim}
\end{small}
\end{minipage}
\end{center}
%
\texttt{choice} requires complex termination reasoning, since the else
branch may lead to the increase of the value that is passed as the
second component of the tuple. 
For this reason, for example, a size-change based method is no longer
applicable.

Using transition invariants, we can prove termination by showing that
between every pair of applications of \texttt{choice} either the value
of the first component of the parameter, the second component, or
their sum is decreasing.

\paragraph{\bf Partial applications}

We consider the following curried function \texttt{f} that has a type 
\texttt{int -> int -> int}.
%
\begin{center}
\begin{minipage}[h]{.8\linewidth}
\begin{small}
\begin{verbatim}
let rec f x = if x > 0 then f (x-1) else fun y -> f x y
\end{verbatim}
\end{small}
\end{minipage}
\end{center}
%
This function shows that -- in contrast to proving termination of
recursive procedures in imperative programs -- it is important to
differentiate between partial and complete applications when dealing
with curried functions.
First, we observe that any partial application of \texttt{f}
terminates.
For example, \texttt{f 10} stops after ten recursive calls and returns
a function \texttt{fun y -> f 10 y} where \texttt{f} is bound to a
closure.
That is, there is no infinite sequence of \texttt{f} applications that
are passed only one argument.
In contrast, any complete application of \texttt{f} does not terminate.
For example, \texttt{f 1 1} will lead to an infinite sequence of
\texttt{f} applications such that each of them is given two arguments.

Our binary reachability analysis takes as input a specification that
determines which kind of applications we want to keep track of. 
The specification consists of a function identifier, e.g., \texttt{f},
a number of parameters, e.g., one, and a transition invariant candidate. 
Then, such a specification requires that applications of \texttt{f} to
one argument satisfy the transition invariant candidate.
Alternatively, we may focus on applications of \texttt{f} to two
arguments.

When dealing with one-ary applications, the binary reachability
analysis can show that every pair of one-ary applications is included
in a transition invariant.

Note that if complete applications of a function terminate, then every
partial application of the function terminates.
For example, the following function \texttt{g} does not have any
infinite application sequences neither for complete nor for partial
applications.
%
\begin{center}
\begin{minipage}[h]{.8\linewidth}
\begin{small}
\begin{verbatim}
let rec g x = if x > 0 then g (x-1) else fun y -> x+y
\end{verbatim}
\end{small}
\end{minipage}
\end{center}

\paragraph{\bf Functions as arguments}

The following example contains functions that take functions as
arguments.
%
\begin{center}
\begin{minipage}[h]{.8\linewidth}
\begin{small}
\begin{verbatim}
let rec map_skip (f, skip) = function
  | x :: xr -> f x :: map_skip (f, skip) (skip xr)
  | [] -> []
in
let add x y = x+y in
let tail = function
  | _ :: tl -> tl
  | [] -> []
in
map_skip (add 1, tail) [0;1]
\end{verbatim}
\end{small}
\end{minipage}
\end{center}
%
While such arguments may not participate in the expressions that
exhibit decent, it is important to keep track of their effect on the
arguments of the recursive calls.
For example, it is important that no application of \texttt{skip}
increases the length of its input list.

Our binary reachability analysis keep track of the above condition by
relying on a program transformation that augments functions with
additional inputs.
By storing arguments of a function application in such inputs, they
become available, i.e., they persist through applications of functions
that are passed as parameter, for the comparison with the transition
invariant when necessary.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_illustration.tex

\section{Illustration}

In this section we illustrate what our transformation adds to the
program in order to keep track of pairs of argument valuations for
checking a transition invariant candidate.

We consider the following curried function \texttt{f} that has a type
\texttt{x:int -> y:int -> ret:int}.
Here, we annotated the parameter and return value types with
identifiers to improve readability.
%
\begin{center}
\begin{minipage}[h]{.8\linewidth}
\begin{small}
\begin{verbatim}
let rec f x = if x > 0 then f (x-1) else fun y -> f x y
\end{verbatim}
\end{small}
\end{minipage}
\end{center}
%
This function shows that -- in contrast to proving termination of
recursive procedures in imperative programs -- it is important to
differentiate between partial and complete applications when dealing
with curried functions.
First, we observe that any partial application of \texttt{f}
terminates.
For example, \texttt{f 10} stops after ten recursive calls and returns
a function \texttt{fun y -> f 10 y} where \texttt{f} is bound to a
closure.
That is, there is no infinite sequence of \texttt{f} applications that
are passed only one argument.
In contrast, any complete application of \texttt{f} does not terminate.
For example, \texttt{f 1 1} will lead to an infinite sequence of
\texttt{f} applications such that each of them is given two arguments.

Our binary reachability analysis takes as input a specification that
determines which kind of applications we want to keep track of. 
The specification consists of a function identifier, e.g., \texttt{f},
a number of parameters, e.g., one, and a transition invariant candidate. 
Then, such a specification requires that applications of \texttt{f} to
one argument satisfy the transition invariant candidate.
Alternatively, we may focus on applications of \texttt{f} to two
arguments.

Once the specification is given, we transform \texttt{f} into a
function \texttt{f\_m} that keeps track of arguments on which \texttt{f} was
applied using additional parameters \texttt{old\_x} and~\texttt{old\_y}.
As a result, \texttt{f\_m} fulfills two requirements. 
First, it computes a result value \texttt{res} such that $\mathtt{res}
= \mathtt{f\ x\ y}$. 
Second, it computes new values of additional parameters.
If \texttt{f\_m} was an imperative program, we would obtain the type
%
\begin{small}
\begin{verbatim}
        x:int * y:int * copied:bool * old_x:int * old_y:int ->
          ret:int * new_copied:bool * new_x:int * new_y:int
\end{verbatim}
\end{small}
% 
where \texttt{new\_x} and \texttt{new\_y} are computed as follows. 
If \texttt{old\_x} already stores a value that was given to \texttt{x} in the past, i.e.,
if $\mathtt{copied} = \mathtt{true}$, then $\mathtt{new\_x} = \mathtt{old\_x}$. 
Otherwise, \texttt{f\_m} can nondeterministically either store
\texttt{x} in \texttt{new\_x} and set $\mathtt{new\_copied} =
\mathtt{true}$, or leave $\mathtt{new\_x} = \mathtt{old\_x}$
and~$\mathtt{new\_copied} = \mathtt{false}$.
The computation of \texttt{new\_y} is similar.
Given a transformed program, checking binary reachability amounts to
checking that at each application site the pair of tuples
$(\mathtt{old\_x}, \mathtt{old\_y})$ and $(\mathtt{x}, \mathtt{y})$
satisfies the transition invariant whenever \texttt{copied}
is~\texttt{true}.

Due to partial applications we cannot expect that values of \texttt{x}
and \texttt{y} are provided simultaneously, which complicates both
computation of \texttt{new\_x} and \texttt{new\_y} and checking if
$(\mathtt{old\_x}, \mathtt{old\_y})$ together with $(\mathtt{x},
\mathtt{y})$ satisfy the transition invariant. 
Hence, we need to keep track of arguments as they are provided, which
requires ``waiting'' for missing arguments.
We implement this waiting process by introducing additional parameters
\texttt{old\_state\_x} and \texttt{old\_state\_y} for each partial
application, together with their updated versions
\texttt{new\_state\_x} and~\texttt{new\_state\_y}.
Each additional parameter accumulates arguments in its first
component, and it keeps a tuple of previously provided arguments in
its second component. 
We obtain the following type for \texttt{f\_m}.
%
\begin{small}
\begin{verbatim}
x:int 
-> old_state_x:((int * int) *                (* accumulate x and y     *)
                (bool * int * int))          (* store copied, x, and y *)
-> (y:int 
    -> old_state_y:((int * int) *            (* accumulate x and y     *)    
                    (bool * int * int))      (* store copied, x, and y *)
    -> ret:int * new_state_y:((int * int) * 
                              (bool * int * int))) *
   new_state_x:((int * int) * 
                (bool * int * int))
\end{verbatim}
\end{small}
% 
We refer to \texttt{(int * int) * (bool * int * int)} as~\texttt{state}. 
Then \texttt{f\_m} has the type:
%
\begin{small}
\begin{verbatim}
      x:int -> old_state_x:state -> 
        (y:int -> old_state_y:state -> ret:int * new_state_y:state) *
        new_state_x:state
\end{verbatim}
\end{small}
% 
We formalize the above transformation in
Section~\ref{sec-transformation}. 
Figure~\ref{fig-ex-prod} presents a detailed execution protocol of
applying our transformation on the above program.

\iffalse
When dealing with unary applications, the binary reachability
analysis can show that every pair of unary applications is included
in a transition invariant using the following check.
\fi

Note that if complete applications of a function terminate, then every
partial application of the function terminates.
For example, consider the following function~\texttt{g}.
%
\begin{center}
\begin{verbatim}
let rec g x = if x > 0 then g (x-1) else fun y -> x+y
\end{verbatim}
\end{center}
%
This function does not have any infinite application sequences neither
for complete nor for partial applications.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_logger.tex

\paragraph{\bf Logger monad}
A \emph{monad} consists of a type constructor \ttt{m} of arity~1 and two
operations
%
\begin{small}
\begin{align*}
  \ttt{unit~}&\ttt{:~'a~->~'a~m} \\
  \ttt{(~>>=~)~}&\ttt{:~'a~m~->~('a~->~'b~m)~->~'b~m}
\end{align*}
\end{small} 
%
These operations need to satisfy three conditions called \emph{left
unit}, \emph{right unit}, and \emph{associative}~\cite{Wadler95}.  
We assume that \ttt{state} is a given type.  
Figure~\ref{fig-state-monad} presents a variant of the \emph{logger
monad}~\cite{freetheorems}. The state
update operator \ttt{update} is of type
\ttt{(state~->~state)~->~unit~m}. A monadic expression (resp. value)
is an expression (resp. value) of the logger monad type.

For example, the monadic expression \ttt{unit 1} evaluates to a
function that takes a state \st\ and returns a pair (\ttt{1}, \st). As
another example, the following monadic expression evaluates to a
function that takes a state \st\ and returns a pair (\ttt{1}, \st\ $+$
\ttt{1}).
\begin{center}
  \ttt{update (fun s -> s + 1) >>= fun () -> unit \ttt{1}}
\end{center}

\begin{figure}[t]
  \linespread{1}
  \begin{minipage}[t]{\columnwidth}
    \small
   \begin{tabular}{l@{\ }l}
     1 & \ttt{(* logger monad type *)} \\[\jot]
     2 & \ttt{type 'a m = state -> 'a * state } \\[\jot]
     3 & \ttt{(* unit operator *)} \\[\jot]
     4 & \ttt{let unit a = fun s -> (a, s) } \\[\jot]
     5 & \ttt{(* bind operator *)} \\[\jot]
     6 & \ttt{let ( >>= ) m k = fun s0 -> } \\[\jot]
     7 & \ttt{\tabTTTTTTTT let v1, s1 = m s0 in } \\[\jot]
     8 & \ttt{\tabTTTTTTTT k v1 s1 } \\[\jot]
     9 & \ttt{(* state transform operator *)} \\[\jot]
     10 & \ttt{let update f = fun s -> ( (), f s ) }
    \end{tabular}
  \end{minipage}
  \caption{Logger monad with state transform operator
    \ttt{update}. The unit operator takes a value and constructs a
    monadic value. The bind operator takes a monadic value and a
    function returning a monadic value, and constructs a new monadic
    value. The state transform operator creates a new monadic value by
    applying the state transformer \ttt{f}.}
  \label{fig-state-monad}
\end{figure}

\iffalse
We clarify the product expressions above by expressing them in
\haskell's \emph{do notation}~\cite{DoNotation}. For
the product of expression \ttt{1} and \msplus, the corresponding
\haskell\ program is the following.
\begin{center}
  \begin{minipage}{.4\columnwidth}
   \small
    \ttt{do \\
      \tabT update (fun s -> s) \\
      \tabT update (fun s -> s) \\
      \tabT unit \ttt{1}}
 \end{minipage}
\end{center}
For the product of expression \ttt{+} and \msplus, the corresponding
\haskell\ program is the following.
\begin{center}
  \begin{minipage}{.9\columnwidth}
    \small
    \ttt{do \\
      \tabT update (fun s -> s) \\
      \tabT update (fun s -> s + 1) \\
      \tabT unit (fun x -> unit (fun y -> unit ((+)\ x\ y)))} 
 \end{minipage}
\end{center}
\fi
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/example-latex/sec_transformation.tex

\section{Program transformation}
\label{sec-transformation}

In this section we present a program transformation that realizes the
function \monitor presented in Section~\ref{sec-semantics}.
To implement the state passing between judgemets we apply a so-called
logger monad.

\input{sec_logger}

\paragraph{\bf Transformation of types}

We transform each program expression into a monadic expression that
keeps track of the state that results in the judgement augmentation.
Figure~\ref{fig-types-prods} presents the function \monadic\ that maps
types of expressions in the original program to types of the
transformed program.
Function \monadic\ indicates that a tranformed program is a \miniocaml\
function that takes an initial state and returns a \mbox{pre-monadic}
program value together with a final state.

\begin{figure}[t]
  \linespread{1}
  \begin{minipage}[t]{\columnwidth}
    \small
    \begin{tabular}{l@{\ }l}
      1 & \algLet \algRec \monadic\ \type\ = (\premonadic\ \type) \moncstr \\[\jot]
      & \\[\jot]
      2 & \algAnd \premonadic\ = \algFunction \\[\jot]
      3 & \tabT $|$ \ttt{\type[1] -> \type[2]}\
      \algArrow \ttt{$(\premonadic\ \type[1])$ -> $(\premonadic\
        \type[2])$ m} \\[\jot]
      % 6 & \tabT $|$ \ttt{\type[1]\ *\ \ldots\ *\ \type[n]}\
      % \algArrow \ttt{$(\premonadic\ \type[1])$\ *\ \ldots\ *\
      %   $(\premonadic\ \type[n])$} \\[\jot] 
      4 & \tabT $|$ \ttt{(\type[1],\ \ldots,\ \type[n])\ \tycstr}\ 
      \algArrow \ttt{($(\premonadic\ \type[1])$,\ \ldots,\ $(\premonadic\
      \type[n])$)\ \tycstr} \\[\jot]
      5 & \tabT $|$ \type\ \algArrow \type
    \end{tabular}
  \end{minipage}
  \caption{Type transformation function \monadic.}
  \label{fig-types-prods}
\end{figure}

For example, consider the following applications of \monadic.
%
\begin{align*}
  \monadic\ \texttt{(int -> (int -> int))} = \;
  \begin{array}[t]{@{}l@{}}
    \texttt{(int -> ((\premonadic\ int -> int) m)) m}\\[\jot]
    \texttt{(int -> ((int -> int m) m)) m}
    % (int -> ((prem int -> int) m)) m
    % (int -> ((int -> int m) m)) m
  \end{array}
\end{align*}
%
\input{fig_tsel}

\input{fig_prod_constr}

\paragraph{\bf Transformation of expressions}

We present the transformation function \product\ in
Figure~\ref{fig-prod-alg}. 
\product uses two auxiliary functions $\tselEnter$ and $\tselExit$
shown in Figure~\ref{fig-tsel}.

For a expression \expr, \product\ traverses the abstract syntax tree
of \expr\ and gives a core monadic expression that evaluates the user
program together with two state transform operations.
\product generates \miniocaml expressions using the $\quo{\cdot}$
function.
For example, $\quo{\ttt{let x = 1 in 1}}$ emits the expression
\ttt{let x = 1 in 1}.
Within $\quo{\cdot}$ we can perform an evaluation by
applying~$\aq{\cdot}$.
For example, $\quo{\ttt{let x = \aq{\quo{1+2}} in 1}}$ emits $\ttt{let
x = 1+2 in 1}$.

The important case is the tranformation of
$\theFunction/\theArity$-applications.
Such applications are recognized in $\tselEnter$ and $\tselExit$.
The emitted code either saves the argument values into the state, or
propagates further the current state.
Furthermore, $\tselEnter$ performs a check if the snapshot stored in a
state together with the arguments of a
$\theFunction/\theArity$-application satisfy the transition invariant
candidate~$\theTI$.
This check is guarded by the condition that the snapshot must have
been stored previously.

We show an example application of \product in Figure~\ref{fig-ex-prod}
for analysing $\texttt{f}/1$-applications.
First, we present subexpressions of the program and then show the
result of the application of \product on them (we have partially
simplified the transformed expressions to improve readability).
%
\begin{figure}[!t]
  \centering
{
\small
\begin{verbatim}
e = let rec f x = if x > 0 then f (x - 1) else fun y -> f x y in f 1
e1 = if x > 0 then f (x - 1) else fun y -> f x y
e2 = x > 0
e3 = 0
e4 = (>) x
e5 = x
e6 = (>)
e7 = f (x - 1)
...

Transform e = let rec f = fun x -> 'Transform e1' in 'Transform e20'
Transform e1 = 'Transform e2' >>= fun x2 -> (if x2 then 'Transform e7'
                                             else 'Transform e14')
Transform e2 = fun s_full -> ('Transform e3' >>= fun x3 ->
                              'Transform e4' >>= fun x4 ->
                              update (fun s -> s) >>= fun () s_partial ->
                             (x4 x3 >>= fun xapp2 ->
                              update (fun s -> s) >>= fun () ->
                              unit xapp2) s_partial) s_full
Transform e3 = unit 0
Transform e4 = fun s_full -> ('Transform e5' >>= fun x5 ->
                              'Transform e6' >>= fun x6 ->
                              update (fun s -> s) >>= fun () s_partial ->
                             (x6 x5 >>= fun xapp4 ->
                              update (fun s -> s) >>= fun () ->
                              unit xapp4) s_partial) s_full
Transform e5 = unit x
Transform e6 = unit (fun z1 -> unit (fun z2 -> unit (z1 > z2)))
Transform e7 = fun s_full -> ('Transform e8' >>= fun x8 ->
                              'Transform e13' >>= fun x13 ->
                              update (fun ( _, m ) ->
                                let m_c, m_1 = m in
                                if m_c then assert ( x8 > 0 && m_1 > x8 );
                                x8, if not m_c && nondet () then true, x8 else m
                              ) >>= fun () s_partial ->
                             (x13 x8 >>= fun xapp7 ->
                              update (fun _ -> s_full) >>= fun () ->
                              unit xapp7) s_partial) s_full
...
\end{verbatim}
}
  \caption{Example application of \product\ with $N = 1$ and \mbox{$\theTI = (\
      a_1 > 0 \wedge m_1 > a_1 \ )$}. Given a stored snapshot, the
    transformed application \ttt{f (x - 1)} checks that the current
    actual satisfies $\theTI$.}
  \label{fig-ex-prod}
\end{figure}
%

We establish a relationship between augmented evaluation trees and
evaluation trees of the transformed program in the following theorem.
%
\begin{theorem}[\product implements \monitor]
\label{thm-product}
%
A pair $(v_1, \dots, v_\theArity)$ and~$(u_1, \dots, u_\theArity)$ is
obtained from the augmented evaluation tree as described in
Theorem~\ref{thm-monitor}
if and only if a judgement of the following form appears in the
evaluation tree of the
program obtained by applying \product:
%
\begin{center}
  \mevj{\evj{\ectx}{$\theFunction$\ \expr[1]\ \_\!\_\ \dots\
      \expr[\theArity]\ s}{\val}}{\stupa}{}
\end{center}
%
such that $\tsf{eval}\ \ectx\ s = (true, (v_1, \dots, v_\theArity))$
and for each $i \in 1..\theArity$ we have \mbox{$\tsf{eval}\ \ectx\ \expr[i]
= u_i$}.
%
\end{theorem}
%

The following corollary of Theorem~\ref{thm-product} allows one to
rely on the assertion validity in the transformed program to implement
the binary reachability analysis of the original program.
%
\begin{theorem}[Binary reachability analysis as assertion checking]
\label{thm-assert}
%
Each pair $(v_1, \dots, v_\theArity)$ and~$(u_1, \dots, u_\theArity)$
in the $\theFunction/\theArity$-recursion relation of the program 
satisfies $\theTI$ if and only if the assertion inserted by
$\tselEnter$ is valid in the transformed program.
%
\end{theorem}
%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-03 paper/workspace/main.tex
Added references to other subprocedures
Page numbers.

@@
 
+\newcommand{\isDraft}{true}
+
 \input{preamble}
+
+\ifthenelse{\equal{\isDraft}{true}}{
+  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
+  \pagestyle{plain}
+}{}
 
@@
     \and
-    Rusl\'{a}n Ledesma-Garza\inst{1}
-    \and
     Rupak Majumdar\inst{2}
@@
     Filip Niksic\inst{2}
+    \and
+    Rusl\'{a}n Ledesma-Garza\inst{1}
   }
@@
 }{
-  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
 }
@@
 * Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
 
-\begin{verbatim}
 * Method Invariant with Minimization
@@
     a + b <= 1
+
+* Subprocedures \mathcal{C'} and Inv are the same as
+  in Method Invariant
 
@@
 
+* Subprocedures TrapConditions and \Delta are the same as
+  in Method Safety by Refinement
+
+* Subprocedures \mathcal{C'} and Inv are the same as
+  in Method Invariant
+
 \end{verbatim}

### 2014-02-04 paper/workspace/main.tex
Split in sections, added references.

@@
 \begin{document}
+
+\nocite{DOsualdoSAS13}
 
@@
 
-\section{Preliminaries}
-
-\begin{verbatim}
-* Petri net:
- N = (S, T, E, M0), where
- S : places
- T : transitions
- E : edges
- M0: initial marking
-
-* Property:
- P, linear arithmetic formula
-\end{verbatim}
-\iffalse
-\begin{verbatim}
- Examples:
-  x + y < 0
-  x + y < 0 \\and x + z > 0
-  x + y < 0 \\or  x + z > 0
-
- ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
-  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
-\end{verbatim}
-\fi
+\input{sec_illustration}
+\input{sec_preliminaries}
 
@@
 We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
-
-\begin{verbatim}
-* Code
-
-Process 1:
-
-    bit1 := false
-    while true do
-p1:   bit1 := true
-p2:   while bit2 do skip od
-p3:   <critical section>
-      bit1 := false
-    od
-
-Process 2:
-
-    bit2 := false
-    while true do
-q1:   bit2 := true
-q2:   if bit1 then
-q3:     bit2 := false
-q4:     while bit1 do skip od
-        goto q1
-      fi
-q5:   <critical section>
-      bit2 := false
-od
-
-* Property: Process 1 and Process 2 are never in their respective critical section at the same time.
-
-* Petri net:
-
-\end{verbatim}
-
-\begin{center}
-  \includegraphics[width=10cm]{lamport}
-\end{center}
-
-\begin{verbatim}
-
-* Subnet of Petri net:
-
-\end{verbatim}
-
-\begin{center}
-\begin{tikzpicture}[node distance=1cm]
-  \node[transition, label=left:t1] (t1) {};
-  \node[place, tokens=1, label=left:a, below of=t1] (a) {};
-  \node[transition, label=left:t2, below of=a] (t2) {};
-  \node[place, tokens=0, label=left:b, below of=t2] (b) {};
-  \node[place, tokens=0, label=right:$\neg bit2$, right of=t2] (nbit2) {};
-  \draw (t1) edge[pre, bend left, min distance=2cm, out=225, in=-45] (b);
-  \draw (t1) edge[post] (a);
-  \draw (t2) edge[pre] (nbit2);
-  \draw (t2) edge[post] (nbit2);
-  \draw (t2) edge[pre] (a);
-  \draw (t2) edge[post] (b);
-\end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property P: a + b <= 1
-\end{verbatim}
 
@@
 * Property of state constraints C: If C U {\neg P} is unsat then N |= P.
+
+* Place equation:
+
+  For a given place s the place equation is
+
+  # of tokens in s = initial number of tokens of place s
+                     + # times each input transition of s fires
+                     - # times each output transition of s fires
+
+* Non-negativity conditions:
+
+  # of tokens in place s           >= 0
+  # of times transition t is fired >= 0
+
+* Subprocedure \mathcal{C}:
+
+  Input:
+    (S, T, E, M0) : Petri net
+  Output:
+    C : State constraints
+
+  Pseudocode:
+
 \end{verbatim}
+
+\begin{align*}
+  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
+    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
+  \right) \right) \land
+    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
+    \left( \bigwedge_{t \in T} t \ge 0 \right)
+\end{align*}
 
@@
 \begin{verbatim}
-* Example:
+* Example
+
+  - Code:
+
+  Process 1:                     |      Process 2:
+                                 |
+      bit1 := false              |      bit2 := false
+      while true do              |      while true do
+  p1:   bit1 := true             |  q1:   bit2 := true
+  p2:   while bit2 do skip od    |  q2:   if bit1 then
+  p3:   <critical section>       |  q3:     bit2 := true
+        bit1 := false            |  q4:     while bit1 do skip od
+      od                         |          goto q1
+                                 |        fi
+                                 |  q5:   <critical section>
+                                 |        bit2 := false
+                                 |      od
+
+  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.
+
+  - Petri net:
+
+\end{verbatim}
+
+\begin{center}
+  \includegraphics[width=10cm]{lamport}
+\end{center}
+
+\newpage
+
+\begin{verbatim}
   - State constraints C:
@@
     a + b >= 2
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
 \end{verbatim}
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land
-    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
 
@@
 \newpage
+
+\begin{verbatim}
+* Subprocedure TrapConditions:
+
+  Input:
+    (S, T, E, M0) : Petri net
+    A             : Satisfying assignment for C \cup { ~P }
+  Output:
+    C_\theta      : Trap conditions
+
+  Pseudocode:
+
+\end{verbatim}
+\begin{align*}
+  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
+      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
+    \right) \right) \land
+    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
+    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
+\end{align*}
+\begin{verbatim}
+
+* Subprocedure \Delta:
+
+  Input:
+    A_\theta      : Satisfying assignment for C_\theta
+  Output:
+    \delta        : Refinement constraint \delta
+
+  Pseudocode:
+
+\end{verbatim}
+\begin{align*}
+  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
+\end{align*}
+
+\newpage
+
 \begin{verbatim}
 * Example
+
   - State constraints C same as in the example for method Safety.
@@
   "At least one place in S is always marked"
-
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
 \end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land
-    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
-    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
-\end{align*}
 
 \newpage
+
 \section{Method Invariant}
@@
 \newpage
+
 \begin{verbatim}
-* Example:
+* Subprocedure \mathcal{C'}:
 
-  - Dual state constraints C':
+  Input:
+    N = (S, T, E, M0)  : Petri net
+    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
+              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
+  Output:
+    C'            : Dual state constraints
 
-    0        >=    a  - b
-    0        >= (- a) + b
+Input:
+ N      = (S, T, E, M0)
+        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
+\neg P  = ( a + b >= 2 )
 
-    a        <  2 * target_1
-
-    a        >= target_1
-    b        >= target_1
-    nbit2    >= 0
-    target_1 >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-
-  - Invariant:
-
-    a + b + nbit2 <= 1
-
+Pseudocode:
 \end{verbatim}
 
+\begin{align*}
+  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
+                      \sum_{(t, s) \in E} s
+                    - \sum_{(s, t) \in E} s \right) \right) \land
+     \left( \sum_{s \in S} M_0(s) \cdot s <
+       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
+     & \left( \bigwedge_{s \in S} \left ( s \ge 
+       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
+     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
+\end{align*}
+
+\begin{verbatim}
+* Subprocedure Inv:
+
+  Input:
+    N = (S, T, E, M0) : Petri net
+    A'                : Satisfying assignment for C'
+
+  Output:
+    I                 : Invariant
+
+  Pseudocode:
+\end{verbatim}
+
+\begin{align*}
+  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
+                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
+\end{align*}
+
 \newpage
+
 \begin{verbatim}
@@
 \newpage
+
 \begin{verbatim}
-* Subprocedure \mathcal{C'}:
+* Example:
 
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
-              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
-  Output:
-    C'            : Dual state constraints
+  - Dual state constraints C':
 
-Input:
- N      = (S, T, E, M0)
-        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
-\neg P  = ( a + b >= 2 )
+    0        >=    a  - b
+    0        >= (- a) + b
 
-Pseudocode:
+    a        <  2 * target_1
+
+    a        >= target_1
+    b        >= target_1
+    nbit2    >= 0
+    target_1 >= 0
+
+  - Model A':
+
+    a        = 1
+    b        = 1
+    nbit2    = 1
+    target_1 = 1
+
+  - Invariant:
+
+    a + b + nbit2 <= 1
+
 \end{verbatim}
 
-\begin{align*}
-  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
-                      \sum_{(t, s) \in E} s
-                    - \sum_{(s, t) \in E} s \right) \right) \land
-     \left( \sum_{s \in S} M_0(s) \cdot s <
-       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
-     & \left( \bigwedge_{s \in S} \left ( s \ge 
-       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
-     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
-\end{align*}
+\newpage
 
-\begin{verbatim}
-* Subprocedure Inv:
-
-  Input:
-    N = (S, T, E, M0) : Petri net
-    A'                : Satisfying assignment for C'
-
-  Output:
-    I                 : Invariant
-
-  Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
-                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
-\end{align*}
-
-\newpage
 \section{Method Invariant with Minimization}
@@
 \newpage
+
+\begin{verbatim}
+* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant
+
+* Subprocedure \text{MinConstraints}:
+
+  Input:
+    N = (S, T, E, M0)  : Petri net
+    A'                 : Satisfying assignment for C'
+  Output:
+    C_M                : Minimization Constraints
+
+  Pseudocode:
+
+\end{verbatim}
+
+\begin{align*}
+  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
+      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
+    \right) \right) \land
+    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
+\end{align*}
+
+\newpage
+
 \begin{verbatim}
 * Example:
+
   - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
@@
     a + b <= 1
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    A'                 : Satisfying assignment for C'
-  Output:
-    C_M                : Minimization Constraints
-
-  Pseudocode:
-
 \end{verbatim}
-
-\begin{align*}
-  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
-      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
-    \right) \right) \land
-    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
-\end{align*}
 
@@
 
+\begin{verbatim}
+* Subprocedures TrapConditions and \Delta are the same as
+  in Method Safety by Refinement
+
+* Subprocedures \mathcal{C'} and Inv are the same as
+  in Method Invariant
+\end{verbatim}
+
 \newpage
@@
     a + b + nbit2 <= 1
-
-* Subprocedures TrapConditions and \Delta are the same as
-  in Method Safety by Refinement
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-
 \end{verbatim}
 
-\iffalse
-
-\newpage
-\section{Benchmarks of \bfc}
-
-We experimented on the benchmarks \ttt{cprover-PN} and
-\ttt{cprover\_software\_analysis} available at \\
-\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
-The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
-Table~\ref{bfc-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{7cm} | r | r | r | r | } %
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 23 &  0 & 54 &  0 \\
-    Invariant \& Invariant with Minimization    & 23 &  0 & 54 &  0 \\
-    Safety by Refinement                        & 23 &  0 & 54 &  0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \bfc}
-\label{bfc-experiments}
-\end{table}
-
-\section{Benchmarks of \mist}
-
-We experimented on the benchmarks available at \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
-and \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
-The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
-Table~\ref{mist-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{6cm} | r | r | r | r | }
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 14 &  0 & 13 &  0 \\
-    Invariant \& Invariant with Minimization    & 14 &  0 & 13 &  0 \\
-    Safety by Refinement                        & 20 &  0 &  7 &  0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \mist}
-\label{mist-experiments}
-\end{table}
-
-\section{Benchmarks of \iic}
-
-We experimented on the benchmarks available at \\
-\ttt{https://svn.rse.mpi-sws.org/RSE/IC3forWSTS/code/trunk/examples/soter}.
-The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
-Table~\ref{iic-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{6cm} | r | r | r | r | }
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 2 & 0 & 36 & 0 \\
-    Invariant \& Invariant with Minimization    & 2 & 0 & 26 & 0 \\
-    Safety by Refinement                        & 2 & 0 & 36 & 0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \iic}
-\label{iic-experiments}
-\end{table}
-
-We constructed the following two invariants by applying the method
-Invariant with Minimization.
-
-\begin{itemize}
-\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
-\item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
-\end{itemize}
-
-\iic\ constructs the following two invariants.
-
-\begin{itemize}
-\item \verb=firewall__no_pred_called_with_zero__depth_1__fixed.spec=: \verb=(l739 < 1)=
-\item \verb=firewall__no_pred_called_with_zero__depth_2__fixed.spec=: \verb=(l3849 < 1)=
-\end{itemize}
-
-\fi
+\input{sec_experiments}
+\bibliographystyle{abbrv}
+\bibliography{references}
 

### 2014-02-05 paper/workspace/main.tex
Added pseudocode for methods in paper draft

--- 2014-02-03 paper/workspace/main.tex
@@
 
+\begin{figure}
 \begin{center}
@@
 \end{center}
+\caption{Diagram for Method Safety}
+\label{fig:method-safety-pseudocode}
+\end{figure}
+
+\begin{figure}
+\begin{algorithmic}[1]
+\State $C := \mathcal C(N)$
+\If {SAT$(C \cup \{\neg P\})$}
+  \State \Return 'The petri net may not satisfy the property'
+\Else
+  \State \Return 'The petri net satisfies the property!'
+\EndIf
+\end{algorithmic}
+\caption{Pseudocode for Method Safety}
+\label{fig:method-safety-pseudocode}
+\end{figure}
 
@@
 \end{verbatim}
+
+\begin{figure}
 \begin{center}
@@
 \end{center}
+\caption{Diagram for Method Safety by Refinement}
+\label{fig:method-safety-by-refinement-diagram}
+\end{figure}
+
+\begin{figure}
+\begin{algorithmic}[1]
+\State $C := \mathcal C(N)$
+\While {SAT$(C \cup \{\neg P\})$}
+  \State $A := $Model$(C \cup \{\neg P\})$
+  \State $C_\theta := $TrapConditions$(N, A)$
+  \If {SAT$(C_\theta)$}
+    \State $A_\theta := $Model$(C_\theta)$
+    \State $\delta := \Delta(A_\theta)$
+    \State $C := C \cup \{\delta\}$
+  \Else
+    \State \Return 'The petri net may not satisfy the property'
+  \EndIf
+\EndWhile
+\State \Return 'The petri net satisfies the property!'
+\end{algorithmic}
+\caption{Pseudocode for Method Safety by Refinement}
+\label{fig:method-safety-by-refinement-pseudocode}
+\end{figure}
 
@@
 \end{center}
+\caption{Diagram for Method Invariant}
+\label{fig:method-invariant-diagram}
+\end{figure}
+
+\begin{figure}
+\begin{algorithmic}[1]
+\State $C' := \mathcal C'(N, \neg P)$
+\If {SAT$(C')$}
+  \State $A' := $Model$(C')$
+  \State $I := $Inv$(N, A')$
+  \State \Return 'Invariant I for the petri net: $I$'
+\Else
+  \State \Return 'Failed at finding an invariant'
+\EndIf
+\end{algorithmic}
+\caption{Pseudocode for Method Invariant}
+\label{fig:method-invariant-pseudocode}
+\end{figure}
 
@@
 \end{center}
+\caption{Diagram for Method Invariant with Minimization}
+\label{fig:method-invariant-with-minimization-diagram}
+\end{figure}
+
+\begin{figure}
+\begin{algorithmic}[1]
+\State $C' := \mathcal C'(N, \neg P)$
+\If {SAT$(C')$}
+  \State $A' := $Model$(C')$
+  \State $C_M := $MinConstraints$(N, A')$
+  \While {SAT$(C' \cup C_M)$}
+    \State $A' := $Model$(C' \cup C_M)$
+    \State $C_M := $MinConstraints$(N, A')$
+  \EndWhile
+  \State $I := $Inv$(N, A')$
+  \State \Return 'Invariant I for the petri net: $I$'
+\Else
+  \State \Return 'Failed at finding an invariant'
+\EndIf
+\end{algorithmic}
+\caption{Pseudocode for Method Invariant with Minimization}
+\label{fig:method-invariant-with-minimization-pseudocode}
+\end{figure}
 
@@
 
+\begin{figure}
 \begin{tikzpicture}
@@
 \end{tikzpicture}
+\caption{Diagram for Method Invariant by Refinement}
+\label{fig:method-invariant-by-refinement-diagram}
+\end{figure}
+
+\begin{figure}
+\begin{algorithmic}[1]
+\State $C := \mathcal C(N)$
+\State $D := $true
+\While {SAT$(C \cup \{\neg P\})$}
+  \State $A := $Model$(C \cup \{\neg P\})$
+  \State $C_\theta := $TrapConditions$(N, A)$
+  \If {SAT$(C_\theta)$}
+    \State $A_\theta := $Model$(C_\theta)$
+    \State $\delta := \Delta(A_\theta)$
+    \State $C := C \cup \{\delta\}$
+    \State $D := D \land \delta$
+  \Else
+    \State \Return 'Failed at finding an invariant'
+  \EndIf
+\EndWhile
+\State $C' := \mathcal C'(N, \neg P \land D)$
+\If {SAT$(C')$}
+  \State $I := D \land $Inv$(N, A')$
+  \State \Return 'Invariant I for the petri net: $I$'
+\Else
+  \State \Return 'Failed at finding an invariant'
+\EndIf
+\end{algorithmic}
+\caption{Pseudocode for Method Invariant by Refinement}
+\label{fig:method-invariant-by-refinement-pseudocode}
+\end{figure}
 
@@
   in Method Invariant
-
 \end{verbatim}

### 2014-02-05 tests/lamports-alg/input-petri-net.pl
Unrolling lamport's 1-bit algorithm.

@@ -1,30 +1,30 @@
-place(cs1).
-place(idle1).
-place(nid1).
-place(req1).
-place(id1).
+place(p1).
+place(p2).
+place(p3).
+place(bit1).
+place(nbit1).
 
-place(cs2).
-place(idle2).
-place(req2).
-place(after_you2).
-place(await2).
-place(id2).
+place(q1).
+place(q2).
+place(q3).
+place(q4).
+place(q5).
+place(nbit2).
 
-transition(u1, [cs1, nid1], [idle1, id1]).
-transition(u2, [idle1, id1], [req1, nid1]).
-transition(u3, [req1, id2], [cs1, id2]).
+transition(s1, [p1,nbit1], [p2,bit1]).
+transition(s2, [p2,nbit2], [p3,nbit2]).
+transition(s3, [p3,bit1], [p1,nbit1]).
 
-transition(v1, [req2, nid1], [after_you2, nid1]).
-transition(v2, [after_you2], [id2, await2]).
-transition(v3, [id1, await2], [id1, idle2]).
-transition(v4, [req2, id1], [cs2, id1]).
-transition(v5, [cs2], [idle2, id2]).
-transition(v6, [idle2, id2], [req2]).
+transition(t1, [q1, nbit2], [q2]).
+transition(t2, [q2, bit1], [q3, bit1]).
+transition(t3, [q3], [q4, nbit2]).
+transition(t4, [q4, nbit1], [q1, nbit1]).
+transition(t5, [q2, nbit1], [q5, nbit1]).
+transition(t6, [q5], [q1, nbit2]).
 
-init(id1).
-init(id2).
-init(idle1).
-init(idle2).
+init(p1).
+init(q1).
+init(nbit1).
+init(nbit2).
 
-target(1, [([cs1,cs2],2)]).
+target(1, [([p3,q5],2)]).

### 2014-02-05 paper/workspace/sec_experiments.tex
Included experiments by Filip.

\section{Experiments}

To evaluate the described methods, we implemented them in a tool called
\emph{Petrinizer}. Petrinizer is essentially a set of Bash and Prolog scripts
that mediate between the input, specified in the input format of
MIST2\footnote{\url{https://github.com/pierreganty/mist}}, and the SMT
solver Z3\footnote{\url{http://z3.codeplex.com/}} \cite{DeMouraTACAS08}. With
the choice of Bash and Prolog over languages with bindings for Z3 API, we
restricted Z3's full power. Indeed, by using Z3 as an external tool, we
were not able to fully exploit the incremental nature of its linear arithmetic
solver (TODO: cite appropriate papers). However, even the non-optimal
implementation turned out to be robust and scalable, with the main advantage of
being able to tackle problem instances that are out of reach for
state-of-the-art coverability checkers.

The evaluation of Petrinizer had three main goals. First, we wanted to compare
its performance against state-of-the-art tools like MIST2, BFC
and IIC (cite!). Second, as the method it
implements is incomplete, we wanted to measure its rate of success on safe problem instances. As a
subgoal, we wanted to investigate the usefulness and necessity of traps, as
well as to what extent real arithmetic suffices over integer arithmetic in
proving safety. And last, since the language of linear arithmetic, used by
Petrinizer, allows for a more succint representation of formulas than the
language used by IIC, we wanted to compare sizes of invariants the two tools produce.

Inputs that were used in tests come from various sources (TODO: write a
sentence to justify the variety).
One source is the collection of Petri nets from the literature that is part of the MIST2 toolkit.
The second source are Petri nets originating from the analysis of concurrent C
programs that were used in the evaluation of BFC \cite{KaiserCONCUR12}. Then
there are inputs originating from the provenance analysis of a medical and a
bug-tracking system \cite{MajumdarSAS13}. Finally, we generated a set of Petri
nets from Erlang programs, using an Erlang verification tool Soter
\cite{DOsualdoSAS13}. (TODO: Acknowledge Emanuele D'Osualdo for help.)

All experiments were performed on the identical machines, equipped with
Intel Xeon 2.66 GHz CPUs and 48 GB of memory, running Linux 3.2.48.1 in 64-bit
mode. Execution time was limited to 100,000 seconds (27 hours, 46 minutes and
40 seconds), and memory to 2 GB.

\subsection{Performance}

Explain what we observe:
\begin{itemize}
  \item On small examples IIC mostly outperforms Petrinizer. Probably due
    to a slower parser, setting up and invoking Z3.
  \item On examples of a similar size, coming from the same source,
    unlike with IIC, resource consumption is fairly uniform. (Medical examples.)
  \item Petrinizer is able to handle some huge examples that are out of
    scope for IIC.
  \item There were no cases when IIC was able to finish, but Petrinizer ran
    out of resources. As a matter of fact, Petrinizer finishes in all cases,
    whereas other tools somethimes fail.
\end{itemize}

\emph{Invariant sizes.} Again, explain how we measure it. For IIC, we measure
the number of non-null literals, the number of clauses, total length of
the invariant in characters. For Petrinizer, we measure the number of
non-null coefficients in the linear expressions and the total length of the
invariant in characters. Argue why it makes sense to compare any of these
quantities. Point out the difference in size with and without the invariant
minimization.

\emph{Rate of success.} Point out that Petrinizer mostly successfully proves
safety. Specifically, it proves safety in all of the safe Soter examples.
Point out the cases where traps were actually useful, and potentially progress
into some pseudo-philosophical discussion on what constitutes a good set of examples.

### 2014-02-05 paper/workspace/main.tex
Unrolling lamport's 1-bit algorithm.
Included experiments by Filip.
Drew Petri net for Lamport's algorithm in tikz

--- 2014-02-05 paper/workspace/main.tex
@@
 
-\begin{center}
-  \includegraphics[width=10cm]{lamport}
-\end{center}
-
-\begin{verbatim}
-
-* Subnet of Petri net:
-
-\end{verbatim}
-
-\begin{center}
-\begin{tikzpicture}[node distance=1cm]
-  \node[transition, label=left:t1] (t1) {};
-  \node[place, tokens=1, label=left:a, below of=t1] (a) {};
-  \node[transition, label=left:t2, below of=a] (t2) {};
-  \node[place, tokens=0, label=left:b, below of=t2] (b) {};
-  \node[place, tokens=0, label=right:$\neg bit2$, right of=t2] (nbit2) {};
-  \draw (t1) edge[pre, bend left, min distance=2cm, out=225, in=-45] (b);
-  \draw (t1) edge[post] (a);
-  \draw (t2) edge[pre] (nbit2);
-  \draw (t2) edge[post] (nbit2);
-  \draw (t2) edge[pre] (a);
-  \draw (t2) edge[post] (b);
-\end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property P: a + b <= 1
-\end{verbatim}
-
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
 \begin{figure}
@@
   \begin{tikzpicture}
+  \draw[box] (-2,8.5) rectangle (1,0);
+  \draw[box] (3,8.5) rectangle (7.5,0);
+
+  \node at (-0.5, 0.5) {First Process};
+  \node at (5, 0.5) {Second Process};
+
+  \node[place, label=left:p3] (p3) at (0,7) {};
+  \node[transition, label=left:u1] (u1) at (0,6) {};
+  \node[place, tokens=1, label=left:p1] (p1) at (0,5) {};
+  \node[transition, label=left:u2] (u2) at (0,4) {};
+  \node[place, label=left:p2] (p2) at (0,3) {};
+  \node[transition, label=left:u3] (u3) at (0,2) {};
+
+  \draw (p3) edge[flow] (u1);
+  \draw (u1) edge[flow] (p1);
+  \draw (p1) edge[flow] (u2);
+  \draw (u2) edge[flow] (p2);
+  \draw (p2) edge[flow] (u3);
+  \draw[rounded corners=4mm] (u3) |- (-1.5,1) -- (-1.5,8) [flow] -| (p3);
+
+  \node[transition, label=above:v1] (v1) at (4,7) {};
+  \node[place, label=right:q3] (q3) at (4,6) {};
+  \node[transition, label=right:v2] (v2) at (4,5) {};
+  \node[place, label=right:q4] (q4) at (4,4) {};
+  \node[transition, label=left:v3] (v3) at (4,3) {};
+
+  \node[place, label=right:q2] (q2) at (5.5,7) {};
+  \node[transition, label=right:v4] (v4) at (5.5,6) {};
+  \node[place, label=right:q5] (q5) at (5.5,5) {};
+  \node[transition, label=right:v5] (v5) at (5.5,4) {};
+  \node[place, tokens=1, label=right:q1] (q1) at (5.5,3) {};
+  \node[transition, label=right:v6] (v6) at (5.5,2) {};
+
+  \draw (q2) edge[flow] (v4);
+  \draw (v4) edge[flow] (q5);
+  \draw (q5) edge[flow] (v5);
+  \draw (v5) edge[flow] (q1);
+  \draw (q1) edge[flow] (v6);
+  \draw[rounded corners=4mm] (v6) |- (7,1) -- (7,8) [flow] -| (q2);
+
+  \draw (q2) edge[flow] (v1);
+  \draw (v1) edge[flow] (q3);
+  \draw (q3) edge[flow] (v2);
+  \draw (v2) edge[flow] (q4);
+  \draw (q4) edge[flow] (v3);
+  \draw (v3) edge[flow] (q1);
+
+  \node[place, label=above:bit1] (bit1) at (2,6) {};
+  \node[place, label=below:$\neg$bit1] (nbit1) at (2,4) {};
+  \node[place, label=below:$\neg$bit2] (nbit2) at (2,2) {};
+
+  \draw (bit1) edge[flow] (u1);
+  \draw (u1) edge[flow] (nbit1);
+  \draw (nbit1) edge[flow] (u2);
+  \draw (u2) edge[flow] (bit1);
+  \draw (v1) edge[flow, <->] (bit1);
+  \draw (nbit1) edge[flow, <->, bend left=18] (v4);
+  \draw (nbit1) edge[flow, <->] (v3);
+  \draw (nbit2) edge[flow, <->] (u3);
+  \draw (v2) edge[flow] (nbit2);
+  \draw (v5) edge[flow, bend left=30] ([xshift=-0.2mm, yshift=1mm] nbit2.east);
+  \draw (nbit2) edge[flow] (v6.west);
+\end{tikzpicture}
+\end{center}
+\caption{Petri net for Lamport's 1-bit algorithm}
+\label{fig:lamport-petri-net}
+\end{figure}
+
+\begin{verbatim}
+
+* Subnet of Petri net:
+
+\end{verbatim}
+
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}
+  \node[transition, label=left:t1] (t1) at (0,5) {};
+  \node[place, tokens=1, label=left:a] (a) at (0,4)  {};
+  \node[transition, label=left:t2] (t2) at (0,3)  {};
+  \node[place, tokens=0, label=left:b] (b) at (0,2)  {};
+  \node[place, tokens=0, label=below:$\neg$bit2] (nbit2) at (2,3) {};
+  \draw (t1) edge[flow] (a);
+  \draw (a) edge[flow] (t2);
+  \draw (t2) edge[flow] (b);
+  \draw (nbit2) edge[flow, <->] (t2);
+  \draw[rounded corners=4mm] (b) |- (-1.5,1) -- (-1.5,6) [flow] -| (t1);
+\end{tikzpicture}
+\end{center}
+\caption{Subnet of Petri net for Lamport's 1-bit algorithm}
+\label{fig:lamport-petri-net-subnet}
+\end{figure}
+
+\begin{verbatim}
+* Property P: a + b <= 1
+\end{verbatim}
+
+\begin{verbatim}
+* Method Safety:
+
+  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
+\end{verbatim}
+
+\begin{figure}
+\begin{center}
+  \begin{tikzpicture}[
+      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
+    ]
     \node[state] (begin) {BEGIN};
@@
 \begin{center}
-\begin{tikzpicture}
+  \begin{tikzpicture}[
+      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
+    ]
   \node[state] (begin) {BEGIN};
@@
 \begin{center}
-  \begin{tikzpicture}
+  \begin{tikzpicture}[
+      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
+    ]
     \node[state] (begin) {BEGIN};
@@
 \begin{figure}
-\begin{tikzpicture}
+  \begin{tikzpicture}[
+      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
+    ]
   \node[state] (begin) {BEGIN};

### 2014-02-05 paper/workspace/main.tex
something

--- 2014-02-05 paper/workspace/main.tex
@@
 
+  - Property: p3 + q5 <= 1
+
   - Petri net:
@@
 
+\begin{figure}
 \begin{center}
-  \includegraphics[width=10cm]{lamport}
+  \begin{tikzpicture}
+  \draw[box] (-2,8.5) rectangle (1,0);
+  \draw[box] (3,8.5) rectangle (7.5,0);
+
+  \node at (-0.5, 0.5) {First Process};
+  \node at (5, 0.5) {Second Process};
+
+  \node[place, label=left:p3] (p3) at (0,7) {};
+  \node[transition, label=left:u1] (u1) at (0,6) {};
+  \node[place, tokens=1, label=left:p1] (p1) at (0,5) {};
+  \node[transition, label=left:u2] (u2) at (0,4) {};
+  \node[place, label=left:p2] (p2) at (0,3) {};
+  \node[transition, label=left:u3] (u3) at (0,2) {};
+
+  \draw (p3) edge[flow] (u1);
+  \draw (u1) edge[flow] (p1);
+  \draw (p1) edge[flow] (u2);
+  \draw (u2) edge[flow] (p2);
+  \draw (p2) edge[flow] (u3);
+  \draw[rounded corners=4mm] (u3) |- (-1.5,1) -- (-1.5,8) [flow] -| (p3);
+
+  \node[transition, label=above:v1] (v1) at (4,7) {};
+  \node[place, label=right:q3] (q3) at (4,6) {};
+  \node[transition, label=right:v2] (v2) at (4,5) {};
+  \node[place, label=right:q4] (q4) at (4,4) {};
+  \node[transition, label=left:v3] (v3) at (4,3) {};
+
+  \node[place, label=right:q2] (q2) at (5.5,7) {};
+  \node[transition, label=right:v4] (v4) at (5.5,6) {};
+  \node[place, label=right:q5] (q5) at (5.5,5) {};
+  \node[transition, label=right:v5] (v5) at (5.5,4) {};
+  \node[place, tokens=1, label=right:q1] (q1) at (5.5,3) {};
+  \node[transition, label=right:v6] (v6) at (5.5,2) {};
+
+  \draw (q2) edge[flow] (v4);
+  \draw (v4) edge[flow] (q5);
+  \draw (q5) edge[flow] (v5);
+  \draw (v5) edge[flow] (q1);
+  \draw (q1) edge[flow] (v6);
+  \draw[rounded corners=4mm] (v6) |- (7,1) -- (7,8) [flow] -| (q2);
+
+  \draw (q2) edge[flow] (v1);
+  \draw (v1) edge[flow] (q3);
+  \draw (q3) edge[flow] (v2);
+  \draw (v2) edge[flow] (q4);
+  \draw (q4) edge[flow] (v3);
+  \draw (v3) edge[flow] (q1);
+
+  \node[place, label=above:bit1] (bit1) at (2,6) {};
+  \node[place, label=below:$\neg$bit1] (nbit1) at (2,4) {};
+  \node[place, label=below:$\neg$bit2] (nbit2) at (2,2) {};
+
+  \draw (bit1) edge[flow] (u1);
+  \draw (u1) edge[flow] (nbit1);
+  \draw (nbit1) edge[flow] (u2);
+  \draw (u2) edge[flow] (bit1);
+  \draw (v1) edge[flow, <->] (bit1);
+  \draw (nbit1) edge[flow, <->, bend left=18] (v4);
+  \draw (nbit1) edge[flow, <->] (v3);
+  \draw (nbit2) edge[flow, <->] (u3);
+  \draw (v2) edge[flow] (nbit2);
+  \draw (v5) edge[flow, bend left=30] ([xshift=-0.2mm, yshift=1mm] nbit2.east);
+  \draw (nbit2) edge[flow] (v6.west);
+\end{tikzpicture}
 \end{center}
+\caption{Petri net for Lamport's 1-bit algorithm}
+\label{fig:lamport-petri-net}
+\end{figure}
 
@@
 
-    b     = 0 - t1 + t2
-    nbit2 = 0
+    p1    = 1 + s3 + 
+    p2    = 0 + 
+    p3    = 
+    bit1  = 
+    nbit1 = 
+    q1    = 
+    q2    = 
+    q3    = 
+    q4    = 
+    q5    = 
+    nbit2 =
 

### 2014-02-05 paper/workspace/sec_garbage.tex

--- 2014-02-01 paper/workspace/main-example-filip.tex
@@
-\documentclass{article}
-
-\usepackage{amsmath}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{listings}
-\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-\usepackage{graphicx}
-
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
-	snakes,automata,backgrounds,petri,calc,positioning}
-
-\newcommand{\bfc}{{\sc Bfc}}
-\newcommand{\mist}{{\sc Mist}}
-\newcommand{\iic}{{\sc IIC}}
-\newcommand{\pnerf}{{\sc Pnerf}}
-\newcommand{\zthree}{{\sc Z3}}
-\newcommand{\ttt}[1]{\texttt{#1}}
-
-\tikzstyle{place}=[circle,thick,draw=black!75,fill=white!20,
-  	minimum size=6mm]
-\tikzstyle{red place}=[place,draw=red!75,fill=red!20]
-\tikzstyle{blue place}=[place,draw=blue!75,fill=blue!20]
-\tikzstyle{transition}=[rectangle,thick,draw=black!75,
-  	fill=black!20,minimum size=4mm]
-
-\tikzstyle{state}=[draw, ellipse, aspect=2]
-\tikzstyle{action}=[draw, rectangle, align=center]
-\tikzstyle{decision}=[draw, diamond, aspect=2, align=center]
-\tikzstyle{print}=[draw, trapezium, trapezium left angle=70, trapezium right angle=-70]
-\tikzstyle{every edge}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]
-
-\begin{document}
-
-\title{Report on a Safety Checker for Petri nets}
-\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
-% \institute{Technische Universit\"at M\"unchen}
-\date{Mon Nov 25 14:24:10 CET 2013}
-
-\section{Preliminaries}
-
-\begin{verbatim}
-* Petri net:
- N = (S, T, E, M0), where
- S : places
- T : transitions
- E : edges
- M0: initial marking
-
-* Property:
- P, linear arithmetic formula
-\end{verbatim}
 \iffalse
-\begin{verbatim}
- Examples:
-  x + y < 0
-  x + y < 0 \\and x + z > 0
-  x + y < 0 \\or  x + z > 0
-
- ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
-  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
-\end{verbatim}
-\fi
-
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on a modified Petri net from [Esparza1997].
-
-\begin{verbatim}
-* Code
-
-TODO
-
-* Property: z is never marked
-
-* Petri net:
-
-\end{verbatim}
-
-\begin{center}
-  \includegraphics[width=6cm]{example-filip}
-\end{center}
-
-\begin{verbatim}
-* Property P: z = 0
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-    \node[print, right=of satc] (yes) {N satisfies P};
-    \node[print, below=of satc] (dontknow) {Don't know};
-    \node[state, right=of yes] (end1) {END};
-    \node[state, below=of dontknow] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (yes);
-    \draw (yes) edge (end1);
-    \draw (satc) edge node[right]{YES} (dontknow);
-    \draw (dontknow) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Example:
-  - State constraints C:
-
-    Place equations:
-
-    x = 1 - t
-    ^   ^   ^
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |  # of tokens taken from x
-    |   |
-    |   initial number of tokens in x
-    |
-    number of tokens in x
-
-    y = 0
-    z = 0 + t
-    w = 0 + t
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-    Non-negativity conditions:
-
-    x >= 0
-    y >= 0
-    z >= 0
-    w >= 0
-    t >= 0
-
-  - Negated Property \neg P:
-
-    z >= 1
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-  \end{verbatim}
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land \\
-  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-\begin{center}
-\begin{tikzpicture}
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                    	$\delta:=\Delta(A_\theta)$\\
-                                    	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {N satisfies P};
-  \node[print, right=of satctheta] (dontknow) {Don't know};
-  \node[state, right=of yes] (end1) {END};
-  \node[state, right=of dontknow] (end2) {END};
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (yes);
-  \draw (yes) edge (end1);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (dontknow);
-  \draw (dontknow) edge (end2);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example
-  - State constraints C and \neg P are the same as in the example for method Safety.
-
-  - Assignment A:
-    x = 0
-    y = 0
-    z = 1
-    w = 1
-    t = 1
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-              t
-         -----------
-    x => y \/ z \/ w
-    ^    ^    |    |
-    |    |    |    w \in S
-    |    |    |
-    |    |    z \in S
-    |    |
-    |    y \in S
-    |
-    x \in S
-
-    y => y \/ z \/ w
-    z => true
-    w => true
-
-    - At least one of the initially marked places belongs to S:
-    x
-
-    - None of the marked places in A belongs to S:
-    ~z /\ ~w
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-  - Assignment A_\theta:
-    x = true
-    y = true
-    z = false
-    w = false
-
-  - Refinement constraint \delta:
-    x + y >= 1
-    ^   ^
-    |   |
-    -----
-      |
-      S = {x, y}, therefore \delta excludes assignment A in the next iteration
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land \\
-  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
-  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
-\end{align*}
-
-\iffalse
-
-\newpage
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A.
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
-      $I := \text{Inv}(N, A)$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (satc) edge node[right]{YES} (inv);
-    \draw (noinv) edge (end1);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example:
-
-  - With weakened property P \/ (\neg \delta)
-
-  - Negated property is \neg P /\ \delta
-
-  - Dual state constraints C':
-    0 >= (- cs1) + id1 + idle1 + (- nid1)
-    0 >= (- id1) + (- idle1) + nid1 + req1
-    0 >= cs1 + (- req1)
-    0 >= after_you2 + (- req2)
-    0 >= (- after_you2) + await2 + id2
-    0 >= (- await2) + idle2
-    0 >= cs2 + (- req2)
-    0 >= (- cs2) + id2 + idle2
-    0 >= (- id2) + (- idle2) + req2
-
-    id1 + id2 + idle1 + idle2 < 2 * target_1 + trap_1
-
-    id1        >= trap_1
-    cs1        >= target_1
-    idle1      >= 0
-    req1       >= trap_1
-    nid1       >= 0
-    id2        >= trap_1
-    cs2        >= target_1
-    idle2      >= 0
-    req2       >= trap_1
-    after_you2 >= trap_1
-    await2     >= 0
-    target_1   >= 0
-    trap_1     >= 0
-
-  - Model A:
-
-    id1        = 1
-    cs1        = 1
-    idle1      = 0
-    req1       = 1
-    nid1       = 0
-    id2        = 1
-    cs2        = 1
-    idle2      = 0
-    req2       = 1
-    await2     = 0
-    after_you2 = 1
-    target_1   = 1
-    trap_1     = 1
-
-  - Invariant:
-
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
-
-  - Invariant with trap constraint \delta added:
-
-    id1 + id2 + req1 + req2 + after_you2             >= 1
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
-
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    P             : Property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N = (S, T, E, M0)
-   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- P = ( cs1 + cs2 <= 1 )
-
-TODO: Unroll pseudocode?
-
-* Subprocedure Inv:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C'
-
-  Output:
-    I             : Invariant
-
-TODO: Unroll pseudocode?
-\end{verbatim}
-
-\newpage
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=. 
-We present the method Invariant with Minimization by example on a
-simplified mutual exclusion algorithm by Delzanno2001.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (model1) {$A:=\text{Model}(C')$};
-    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A)$};
-    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
-    \node[action, right=of satcm] (model2) {$A:=\text{Model}(C' \cup C_M)$};
-    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A)$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (noinv) edge (end1);
-    \draw (satc) edge node[right]{YES} (model1);
-    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
-    \draw (cm) edge (satcm);
-    \draw (satcm) edge node[above]{YES} (model2);
-    \draw (model2) |- (edgein);
-    \draw (satcm) edge node[right]{NO} (inv);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A: 
-    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example:
-  - Dual state constraints C' and satisfying assignment A for C' as in example for method Invariant
-
-  - Minimization constraints:
-
-    nc  ~= 0 => m_nc = 1
-    |           --------
-    |              |
-    |           place appears in invariant
-    |
-    place coefficient in the invariant
-
-    nc   = 0 => m_nc  = 0
-    |           ---------
-    |              |
-    |           place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    m   ~= 0 => m_m   = 1
-    m    = 0 => m_m   = 0
-    cs1 ~= 0 => m_cs1 = 1
-    cs1  = 0 => m_cs1 = 0
-    cs2 ~= 0 => m_cs2 = 1
-    cs2  = 0 => m_cs2 = 0
-
-    m_nc + m_m + m_cs1 + m_cs2 < 3
-    --------------------------   |
-                |                |
-                |    number of places appearing in current invariant for A
-                |    = #{s | A(s) ~= 0} = #{m, cs1, cs2}
-                |
-                |
-  number of places appearing in new invariant
-\end{verbatim}
-
-\begin{verbatim}
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Assignment
-  Output:
-    C_M           : Minimization Constraints
-
-Input:
- N   = (S, T, E, M0)
-     = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- A   = ( nc=0, m=1, cs1=1, cs2=1, target1=1 )
-
-TODO: Unroll pseudocode?
-
-\end{verbatim}
 
@@
 \fi
-
-\end{document}

### 2014-02-05 paper/workspace/sec_preliminaries.tex

--- 2014-02-01 paper/workspace/main-example-filip.tex
@@
-\documentclass{article}
-
-\usepackage{amsmath}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{listings}
-\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-\usepackage{graphicx}
-
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
-	snakes,automata,backgrounds,petri,calc,positioning}
-
-\newcommand{\bfc}{{\sc Bfc}}
-\newcommand{\mist}{{\sc Mist}}
-\newcommand{\iic}{{\sc IIC}}
-\newcommand{\pnerf}{{\sc Pnerf}}
-\newcommand{\zthree}{{\sc Z3}}
-\newcommand{\ttt}[1]{\texttt{#1}}
-
-\tikzstyle{place}=[circle,thick,draw=black!75,fill=white!20,
-  	minimum size=6mm]
-\tikzstyle{red place}=[place,draw=red!75,fill=red!20]
-\tikzstyle{blue place}=[place,draw=blue!75,fill=blue!20]
-\tikzstyle{transition}=[rectangle,thick,draw=black!75,
-  	fill=black!20,minimum size=4mm]
-
-\tikzstyle{state}=[draw, ellipse, aspect=2]
-\tikzstyle{action}=[draw, rectangle, align=center]
-\tikzstyle{decision}=[draw, diamond, aspect=2, align=center]
-\tikzstyle{print}=[draw, trapezium, trapezium left angle=70, trapezium right angle=-70]
-\tikzstyle{every edge}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]
-
-\begin{document}
-
-\title{Report on a Safety Checker for Petri nets}
-\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
-% \institute{Technische Universit\"at M\"unchen}
-\date{Mon Nov 25 14:24:10 CET 2013}
-
 \section{Preliminaries}
@@
 \end{verbatim}
+
 \iffalse
@@
 \fi
-
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on a modified Petri net from [Esparza1997].
-
-\begin{verbatim}
-* Code
-
-TODO
-
-* Property: z is never marked
-
-* Petri net:
-
-\end{verbatim}
-
-\begin{center}
-  \includegraphics[width=6cm]{example-filip}
-\end{center}
-
-\begin{verbatim}
-* Property P: z = 0
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-    \node[print, right=of satc] (yes) {N satisfies P};
-    \node[print, below=of satc] (dontknow) {Don't know};
-    \node[state, right=of yes] (end1) {END};
-    \node[state, below=of dontknow] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (yes);
-    \draw (yes) edge (end1);
-    \draw (satc) edge node[right]{YES} (dontknow);
-    \draw (dontknow) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Example:
-  - State constraints C:
-
-    Place equations:
-
-    x = 1 - t
-    ^   ^   ^
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |  # of tokens taken from x
-    |   |
-    |   initial number of tokens in x
-    |
-    number of tokens in x
-
-    y = 0
-    z = 0 + t
-    w = 0 + t
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-    Non-negativity conditions:
-
-    x >= 0
-    y >= 0
-    z >= 0
-    w >= 0
-    t >= 0
-
-  - Negated Property \neg P:
-
-    z >= 1
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-  \end{verbatim}
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land \\
-  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-\begin{center}
-\begin{tikzpicture}
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                    	$\delta:=\Delta(A_\theta)$\\
-                                    	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {N satisfies P};
-  \node[print, right=of satctheta] (dontknow) {Don't know};
-  \node[state, right=of yes] (end1) {END};
-  \node[state, right=of dontknow] (end2) {END};
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (yes);
-  \draw (yes) edge (end1);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (dontknow);
-  \draw (dontknow) edge (end2);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example
-  - State constraints C and \neg P are the same as in the example for method Safety.
-
-  - Assignment A:
-    x = 0
-    y = 0
-    z = 1
-    w = 1
-    t = 1
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-              t
-         -----------
-    x => y \/ z \/ w
-    ^    ^    |    |
-    |    |    |    w \in S
-    |    |    |
-    |    |    z \in S
-    |    |
-    |    y \in S
-    |
-    x \in S
-
-    y => y \/ z \/ w
-    z => true
-    w => true
-
-    - At least one of the initially marked places belongs to S:
-    x
-
-    - None of the marked places in A belongs to S:
-    ~z /\ ~w
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-  - Assignment A_\theta:
-    x = true
-    y = true
-    z = false
-    w = false
-
-  - Refinement constraint \delta:
-    x + y >= 1
-    ^   ^
-    |   |
-    -----
-      |
-      S = {x, y}, therefore \delta excludes assignment A in the next iteration
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land \\
-  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
-  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
-\end{align*}
-
-\iffalse
-
-\newpage
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A.
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
-      $I := \text{Inv}(N, A)$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (satc) edge node[right]{YES} (inv);
-    \draw (noinv) edge (end1);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example:
-
-  - With weakened property P \/ (\neg \delta)
-
-  - Negated property is \neg P /\ \delta
-
-  - Dual state constraints C':
-    0 >= (- cs1) + id1 + idle1 + (- nid1)
-    0 >= (- id1) + (- idle1) + nid1 + req1
-    0 >= cs1 + (- req1)
-    0 >= after_you2 + (- req2)
-    0 >= (- after_you2) + await2 + id2
-    0 >= (- await2) + idle2
-    0 >= cs2 + (- req2)
-    0 >= (- cs2) + id2 + idle2
-    0 >= (- id2) + (- idle2) + req2
-
-    id1 + id2 + idle1 + idle2 < 2 * target_1 + trap_1
-
-    id1        >= trap_1
-    cs1        >= target_1
-    idle1      >= 0
-    req1       >= trap_1
-    nid1       >= 0
-    id2        >= trap_1
-    cs2        >= target_1
-    idle2      >= 0
-    req2       >= trap_1
-    after_you2 >= trap_1
-    await2     >= 0
-    target_1   >= 0
-    trap_1     >= 0
-
-  - Model A:
-
-    id1        = 1
-    cs1        = 1
-    idle1      = 0
-    req1       = 1
-    nid1       = 0
-    id2        = 1
-    cs2        = 1
-    idle2      = 0
-    req2       = 1
-    await2     = 0
-    after_you2 = 1
-    target_1   = 1
-    trap_1     = 1
-
-  - Invariant:
-
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
-
-  - Invariant with trap constraint \delta added:
-
-    id1 + id2 + req1 + req2 + after_you2             >= 1
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
-
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    P             : Property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N = (S, T, E, M0)
-   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- P = ( cs1 + cs2 <= 1 )
-
-TODO: Unroll pseudocode?
-
-* Subprocedure Inv:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C'
-
-  Output:
-    I             : Invariant
-
-TODO: Unroll pseudocode?
-\end{verbatim}
-
-\newpage
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=. 
-We present the method Invariant with Minimization by example on a
-simplified mutual exclusion algorithm by Delzanno2001.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (model1) {$A:=\text{Model}(C')$};
-    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A)$};
-    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
-    \node[action, right=of satcm] (model2) {$A:=\text{Model}(C' \cup C_M)$};
-    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A)$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (noinv) edge (end1);
-    \draw (satc) edge node[right]{YES} (model1);
-    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
-    \draw (cm) edge (satcm);
-    \draw (satcm) edge node[above]{YES} (model2);
-    \draw (model2) |- (edgein);
-    \draw (satcm) edge node[right]{NO} (inv);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A: 
-    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example:
-  - Dual state constraints C' and satisfying assignment A for C' as in example for method Invariant
-
-  - Minimization constraints:
-
-    nc  ~= 0 => m_nc = 1
-    |           --------
-    |              |
-    |           place appears in invariant
-    |
-    place coefficient in the invariant
-
-    nc   = 0 => m_nc  = 0
-    |           ---------
-    |              |
-    |           place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    m   ~= 0 => m_m   = 1
-    m    = 0 => m_m   = 0
-    cs1 ~= 0 => m_cs1 = 1
-    cs1  = 0 => m_cs1 = 0
-    cs2 ~= 0 => m_cs2 = 1
-    cs2  = 0 => m_cs2 = 0
-
-    m_nc + m_m + m_cs1 + m_cs2 < 3
-    --------------------------   |
-                |                |
-                |    number of places appearing in current invariant for A
-                |    = #{s | A(s) ~= 0} = #{m, cs1, cs2}
-                |
-                |
-  number of places appearing in new invariant
-\end{verbatim}
-
-\begin{verbatim}
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Assignment
-  Output:
-    C_M           : Minimization Constraints
-
-Input:
- N   = (S, T, E, M0)
-     = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- A   = ( nc=0, m=1, cs1=1, cs2=1, target1=1 )
-
-TODO: Unroll pseudocode?
-
-\end{verbatim}
-
-\newpage
-\section{Benchmarks of \bfc}
-
-We experimented on the benchmarks \ttt{cprover-PN} and
-\ttt{cprover\_software\_analysis} available at \\
-\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
-The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
-Table~\ref{bfc-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{7cm} | r | r | r | r | } %
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 23 &  0 & 54 &  0 \\
-    Invariant \& Invariant with Minimization    & 23 &  0 & 54 &  0 \\
-    Safety by Refinement                        & 23 &  0 & 54 &  0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \bfc}
-\label{bfc-experiments}
-\end{table}
-
-\section{Benchmarks of \mist}
-
-We experimented on the benchmarks available at \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
-and \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
-The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
-Table~\ref{mist-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{6cm} | r | r | r | r | }
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 14 &  0 & 13 &  0 \\
-    Invariant \& Invariant with Minimization    & 14 &  0 & 13 &  0 \\
-    Safety by Refinement                        & 20 &  0 &  7 &  0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \mist}
-\label{mist-experiments}
-\end{table}
-
-\section{Benchmarks of \iic}
-
-We experimented on the benchmarks available at \\
-\ttt{https://svn.rse.mpi-sws.org/RSE/IC3forWSTS/code/trunk/examples/soter}.
-The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
-Table~\ref{iic-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{6cm} | r | r | r | r | }
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 2 & 0 & 36 & 0 \\
-    Invariant \& Invariant with Minimization    & 2 & 0 & 26 & 0 \\
-    Safety by Refinement                        & 2 & 0 & 36 & 0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \iic}
-\label{iic-experiments}
-\end{table}
-
-We constructed the following two invariants by applying the method
-Invariant with Minimization.
-
-\begin{itemize}
-\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
-\item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
-\end{itemize}
-
-\iic\ constructs the following two invariants.
-
-\begin{itemize}
-\item \verb=firewall__no_pred_called_with_zero__depth_1__fixed.spec=: \verb=(l739 < 1)=
-\item \verb=firewall__no_pred_called_with_zero__depth_2__fixed.spec=: \verb=(l3849 < 1)=
-\end{itemize}
-
-\fi
-
-\end{document}

### 2014-02-05 tests/lamports-alg/input-petri-net.pl
In Lamport input, bit --> b.

@@ -1,30 +1,30 @@
 place(p1).
 place(p2).
 place(p3).
-place(bit1).
-place(nbit1).
+place(b1).
+place(nb1).
 
 place(q1).
 place(q2).
 place(q3).
 place(q4).
 place(q5).
-place(nbit2).
+place(nb2).
 
-transition(s1, [p1,nbit1], [p2,bit1]).
-transition(s2, [p2,nbit2], [p3,nbit2]).
-transition(s3, [p3,bit1], [p1,nbit1]).
+transition(s1, [p1,nb1], [p2,b1]).
+transition(s2, [p2,nb2], [p3,nb2]).
+transition(s3, [p3,b1], [p1,nb1]).
 
-transition(t1, [q1, nbit2], [q2]).
-transition(t2, [q2, bit1], [q3, bit1]).
-transition(t3, [q3], [q4, nbit2]).
-transition(t4, [q4, nbit1], [q1, nbit1]).
-transition(t5, [q2, nbit1], [q5, nbit1]).
-transition(t6, [q5], [q1, nbit2]).
+transition(t1, [q1, nb2], [q2]).
+transition(t2, [q2, b1], [q3, b1]).
+transition(t3, [q3], [q4, nb2]).
+transition(t4, [q4, nb1], [q1, nb1]).
+transition(t5, [q2, nb1], [q5, nb1]).
+transition(t6, [q5], [q1, nb2]).
 
 init(p1).
 init(q1).
-init(nbit1).
-init(nbit2).
+init(nb1).
+init(nb2).
 
-target(1, [([p3,q5],2)]).
+target(1, [([p3, q5], 2)]).

### 2014-02-05 paper/workspace/main.tex
Separated figure lamport network.
Added results for methods safety and safety by refinement

--- 2014-02-05 paper/workspace/main.tex
@@
 
-\begin{center}
-  \includegraphics[width=10cm]{lamport}
-\end{center}
+\input{fig_lamport_net}
 
@@
 
-    a     = 1 + t1 - t2
+    p1    = 1 - s1      + s3
     ^       ^   ^         ^
@@
     |       |   |         |
-    |       |   |   # of tokens taken from a
+    |       |   |        # of tokens given to p1
     |       |   |    
-    |       |  # of tokens given to a
+    |       |  # of tokens taken from p1
     |       |
-    |      initial number of tokens in a
+    |      initial number of tokens in p1
     |
-    number of tokens in a
+    number of tokens in p1
 
-    b     = 0 - t1 + t2
-    nbit2 = 0
+    p2    = 0 + s1 - s2
+    p3    = 0      + s2 - s3
+    bit1  = 0 + s1      - s3
+    nbit1 = 1 - s1      + s3
+    q1    = 1 +              - t1           + t4      + t6
+    q2    = 0 +              + t1 - t2           - t5
+    q3    = 0 +                   + t2 - t3
+    q4    = 0 +                        + t3 - t4
+    q5    = 0 +                                  + t5 - t6
+    nbit2 = 1 +              - t1      + t3           + t6
 
@@
 
-    a     >= 0
-    b     >= 0
+    p1    >= 0
+    p2    >= 0
+    p3    >= 0
+    bit1  >= 0
+    nbit1 >= 0
+    q1    >= 0
+    q2    >= 0
+    q3    >= 0
+    q4    >= 0
+    q5    >= 0
     nbit2 >= 0
@@
 
-    a + b >= 2
+    p3 + q5 >= 2
 \end{verbatim}
@@
 
-  - Property P:
-
-    b = 0
-
-  - Negated Property \neg P:
-
-    b >= 1
-
   - Assignment A:
-    a     = 0
-    b     = 1
+    p1    = 0
+    p2    = 0
+    p3    = 1
+    bit1  = 1
+    nbit1 = 0
+    q1    = 0
+    q2    = 0
+    q3    = 0
+    q4    = 0
+    q5    = 0
     nbit2 = 0
-    t1    = 0
-    t2    = 1
 
@@
     - Trap implications:
-               t2
+                s1
              ----------
-    a     => b \/ nbit2
+    p1    => p2 \/ bit1
     ^           ^      ^
     |           |      |
-    |           |     nbit2 \in S
+    |           |     bit1 \in S
     |           |
-    |          b \in S
+    |          p2 \in S
     |
-    a \in S
+    p1 \in S
 
-    nbit2 => b \/ nbit2
-    b     => a
+    p2    => p3 \/ nbit2
+    p3    => p1 \/ nbit1
+    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
+    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
+    q1    => q2
+    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
+    q3    => q4 \/ nbit2
+    q4    => q1 \/ nbit1
+    q5    => q1 \/ nbit2
+    nbit2 => q2 /\ (p3 \/ nbit2)
 
     - At least one of the initially marked places belongs to S:
-    a
+    p1 \/ q1 \/ nbit1 \/ nbit2
 
     - None of the marked places in A belongs to S:
-    ~b
+    ~p3 /\ ~q5 /\ ~bit1
 
   - Assignment A_\theta:
-    a     = true
-    b     = false
+    p1    = false
+    p2    = true
+    p3    = false
+    bit1  = false
+    nbit1 = true
+    q1    = false
+    q2    = true
+    q3    = true
+    q4    = false
+    q5    = false
     nbit2 = true
@@
   - Refinement constraint \delta:
-    a + b >= 1
-    ^   ^
-    |   |
-    -----
+    p2 + q3 + q5 + nbit1 + nbit2 >= 1
+    ^    ^    ^    ^       ^
+    |    |    |    |       |
+    ------------------------
       |
-      S = {a, b}, therefore \delta excludes assignment A in the next iteration
+      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
+      assignment A in the next iteration
 \end{verbatim}

### 2014-02-05 paper/workspace/fig_dia_method_invariant_w_minimization.tex
Section for introduction, figures for methods.

--- 2014-02-05 paper/workspace/main.tex
@@
-\documentclass{llncs}
-
-\newcommand{\isDraft}{true}
-
-\input{preamble}
-
-\ifthenelse{\equal{\isDraft}{true}}{
-  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-  \pagestyle{plain}
-}{}
-
-\begin{document}
-
-\nocite{DOsualdoSAS13}
-
-\ifthenelse{\equal{\isDraft}{false}}{
-  \title{An SMT-based Approach to Coverability Analysis}
-  \author{
-    Javier Esparza\inst{1}
-    \and
-    Rupak Majumdar\inst{2}
-    \and
-    Philipp Meyer\inst{1}
-    \and
-    Filip Niksic\inst{2}
-    \and
-    Rusl\'{a}n Ledesma-Garza\inst{1}
-  }
-  \institute{
-    Technische Universit\"at M\"unchen
-    \and
-    MPI-SWS, Kaiserslautern and Saarbr\"ucken
-  }
-
-  \maketitle
-
-  \begin{abstract}
-    Model checkers based on Petri net coverability have been used
-    successfully in recent years to verify safety properties of concurrent
-    shared-memory or asynchronous message-passing software. The bottleneck
-    in the verification process is the iterative computation of the
-    (backward) reachable or (forward) coverable set, and most recent
-    research has focused on non-trivial heuristics to speed it up.
-
-    In this paper, we revisit a constraint approach to the problem based on
-    classical Petri net analysis techniques: marking equation, place
-    invariants, and traps. Given a marking M to be covered, the approach
-    constructs constraints corresponding to increasingly precise
-    overapproximations of the reachable markings covering M. If one of
-    these constraints is unsatisfiable, then M is not coverable.
-
-    Previous work on this approach suffered from the absence of efficient
-    decision procedures for linear arithmetic, which nowadays exist as a
-    part of SMT solvers. We show how to utilize an SMT solver to implement
-    the constraint approach, and additionally, to generate a minimized
-    inductive invariant from a safety proof.
-
-    We empirically evaluate our procedure on a large set of existing Petri
-    net benchmarks arising out of software verification. Surprisingly, even
-    though our technique is incomplete, we show that it can quickly
-    discharge all safe instances, usually competitive with, and occasionally
-    much faster than, state-of-the-art techniques. Additionally, the
-    inductive invariants computed are usually orders of magnitude smaller
-    than those produced by existing solvers.
-  \end{abstract}
-}{
-}
-
-\section{Introduction}
-
-This paper makes the following contributions.
-\begin{itemize}
-\item An application of \zthree\ to the constraint approach and its
-  corresponding evaluation.
-\item A method that given a network and a property constructs a
-  minimal invariant.
-\end{itemize}
-
-\input{sec_illustration}
-\input{sec_preliminaries}
-
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
-
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-    \node[print, right=of satc] (yes) {N satisfies P};
-    \node[print, below=of satc] (dontknow) {Don't know};
-    \node[state, right=of yes] (end1) {END};
-    \node[state, below=of dontknow] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (yes);
-    \draw (yes) edge (end1);
-    \draw (satc) edge node[right]{YES} (dontknow);
-    \draw (dontknow) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Safety}
-\label{fig:method-safety-pseudocode}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\If {SAT$(C \cup \{\neg P\})$}
-  \State \Return 'The petri net may not satisfy the property'
-\Else
-  \State \Return 'The petri net satisfies the property!'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Safety}
-\label{fig:method-safety-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land
-    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - Code:
-
-  Process 1:                     |      Process 2:
-                                 |
-      bit1 := false              |      bit2 := false
-      while true do              |      while true do
-  p1:   bit1 := true             |  q1:   bit2 := true
-  p2:   while bit2 do skip od    |  q2:   if bit1 then
-  p3:   <critical section>       |  q3:     bit2 := true
-        bit1 := false            |  q4:     while bit1 do skip od
-      od                         |          goto q1
-                                 |        fi
-                                 |  q5:   <critical section>
-                                 |        bit2 := false
-                                 |      od
-
-  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.
-
-  - Property: p3 + q5 <= 1
-
-  - Petri net:
-
-\end{verbatim}
-
-\input{fig_lamport_net}
-
-\newpage
-
-\begin{verbatim}
-  - State constraints C:
-
-    Place equations:
-
-    p1    = 1 - s1      + s3
-    ^       ^   ^         ^
-    |       |   |         |
-    |       |   |         |
-    |       |   |         |
-    |       |   |        # of tokens given to p1
-    |       |   |    
-    |       |  # of tokens taken from p1
-    |       |
-    |      initial number of tokens in p1
-    |
-    number of tokens in p1
-
-    p2    = 0 + s1 - s2
-    p3    = 0      + s2 - s3
-    bit1  = 0 + s1      - s3
-    nbit1 = 1 - s1      + s3
-    q1    = 1 +              - t1           + t4      + t6
-    q2    = 0 +              + t1 - t2           - t5
-    q3    = 0 +                   + t2 - t3
-    q4    = 0 +                        + t3 - t4
-    q5    = 0 +                                  + t5 - t6
-    nbit2 = 1 +              - t1      + t3           + t6
-
-    Non-negativity conditions:
-
-    p1    >= 0
-    p2    >= 0
-    p3    >= 0
-    bit1  >= 0
-    nbit1 >= 0
-    q1    >= 0
-    q2    >= 0
-    q3    >= 0
-    q4    >= 0
-    q5    >= 0
-    nbit2 >= 0
-
-  - Negated Property \neg P:
-
-    p3 + q5 >= 2
-\end{verbatim}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                    	$\delta:=\Delta(A_\theta)$\\
-                                    	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {N satisfies P};
-  \node[print, right=of satctheta] (dontknow) {Don't know};
-  \node[state, right=of yes] (end1) {END};
-  \node[state, right=of dontknow] (end2) {END};
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (yes);
-  \draw (yes) edge (end1);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (dontknow);
-  \draw (dontknow) edge (end2);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Safety by Refinement}
-\label{fig:method-safety-by-refinement-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\While {SAT$(C \cup \{\neg P\})$}
-  \State $A := $Model$(C \cup \{\neg P\})$
-  \State $C_\theta := $TrapConditions$(N, A)$
-  \If {SAT$(C_\theta)$}
-    \State $A_\theta := $Model$(C_\theta)$
-    \State $\delta := \Delta(A_\theta)$
-    \State $C := C \cup \{\delta\}$
-  \Else
-    \State \Return 'The petri net may not satisfy the property'
-  \EndIf
-\EndWhile
-\State \Return 'The petri net satisfies the property!'
-\end{algorithmic}
-\caption{Pseudocode for Method Safety by Refinement}
-\label{fig:method-safety-by-refinement-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land
-    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
-    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - State constraints C same as in the example for method Safety.
-
-  - Assignment A:
-    p1    = 0
-    p2    = 0
-    p3    = 1
-    bit1  = 1
-    nbit1 = 0
-    q1    = 0
-    q2    = 0
-    q3    = 0
-    q4    = 0
-    q5    = 0
-    nbit2 = 0
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-                s1
-             ----------
-    p1    => p2 \/ bit1
-    ^           ^      ^
-    |           |      |
-    |           |     bit1 \in S
-    |           |
-    |          p2 \in S
-    |
-    p1 \in S
-
-    p2    => p3 \/ nbit2
-    p3    => p1 \/ nbit1
-    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
-    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
-    q1    => q2
-    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
-    q3    => q4 \/ nbit2
-    q4    => q1 \/ nbit1
-    q5    => q1 \/ nbit2
-    nbit2 => q2 /\ (p3 \/ nbit2)
-
-    - At least one of the initially marked places belongs to S:
-    p1 \/ q1 \/ nbit1 \/ nbit2
-
-    - None of the marked places in A belongs to S:
-    ~p3 /\ ~q5 /\ ~bit1
-
-  - Assignment A_\theta:
-    p1    = false
-    p2    = true
-    p3    = false
-    bit1  = false
-    nbit1 = true
-    q1    = false
-    q2    = true
-    q3    = true
-    q4    = false
-    q5    = false
-    nbit2 = true
-
-  - Refinement constraint \delta:
-    p2 + q3 + q5 + nbit1 + nbit2 >= 1
-    ^    ^    ^    ^       ^
-    |    |    |    |       |
-    ------------------------
-      |
-      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
-      assignment A in the next iteration
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A'.
-\end{verbatim}
-
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
-      $I := \text{Inv}(N, A')$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (satc) edge node[right]{YES} (inv);
-    \draw (noinv) edge (end1);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Invariant}
-\label{fig:method-invariant-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C' := \mathcal C'(N, \neg P)$
-\If {SAT$(C')$}
-  \State $A' := $Model$(C')$
-  \State $I := $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant}
-\label{fig:method-invariant-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
-              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N      = (S, T, E, M0)
-        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
-\neg P  = ( a + b >= 2 )
-
-Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
-                      \sum_{(t, s) \in E} s
-                    - \sum_{(s, t) \in E} s \right) \right) \land
-     \left( \sum_{s \in S} M_0(s) \cdot s <
-       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
-     & \left( \bigwedge_{s \in S} \left ( s \ge 
-       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
-     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
-\end{align*}
-
-\begin{verbatim}
-* Subprocedure Inv:
-
-  Input:
-    N = (S, T, E, M0) : Petri net
-    A'                : Satisfying assignment for C'
-
-  Output:
-    I                 : Invariant
-
-  Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
-                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  2 * target_1
-
-    a        >= target_1
-    b        >= target_1
-    nbit2    >= 0
-    target_1 >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-
-  - Invariant:
-
-    a + b + nbit2 <= 1
-
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A'
-\end{verbatim}
-
 \begin{figure}
@@
 \end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C' := \mathcal C'(N, \neg P)$
-\If {SAT$(C')$}
-  \State $A' := $Model$(C')$
-  \State $C_M := $MinConstraints$(N, A')$
-  \While {SAT$(C' \cup C_M)$}
-    \State $A' := $Model$(C' \cup C_M)$
-    \State $C_M := $MinConstraints$(N, A')$
-  \EndWhile
-  \State $I := $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant with Minimization}
-\label{fig:method-invariant-with-minimization-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A': 
-    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant
-
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    A'                 : Satisfying assignment for C'
-  Output:
-    C_M                : Minimization Constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
-      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
-    \right) \right) \land
-    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
-
-  - Minimization constraints C_M:
-
-    a     > 0 => b_a     = 1
-    |            -----------
-    |               |
-    |            place appears in invariant
-    |
-    place coefficient in the invariant
-
-    a     = 0 => b_a     = 0
-    |           ------------
-    |               |
-    |            place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    b     > 0 => b_b     = 1
-    b     = 0 => b_b     = 0
-    nbit2 > 0 => b_nbit2 = 1
-    nbit2 = 0 => b_nbit2 = 0
-
-    b_a + b_b + b_nbit2 < 3
-    -------------------   |
-                |         |
-                |    number of places appearing in current invariant for A'
-                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
-                |
-  number of places appearing in new invariant
-
-  - Model A' for C' \cup C_M:
-
-    a        = 1
-    b        = 1
-    nbit2    = 0
-    target_1 = 1
-
-  - Minimized Invariant:
-
-    a + b <= 1
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant by Refinement}
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety by Refinement.
-
-* Method Invariant by Refinement
-\end{verbatim}
-
-\begin{figure}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
-                                        $D:=\{\}$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                                        $C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                                $\delta:=\Delta(A_\theta)$\\
-                                                $C:=C \cup \{\delta\}$\\
-                                                $D:=D \cup \{\delta\}$};
-  \node[print, right=of satctheta] (noinv2) {No Invariant};
-  \node[state, below=of noinv2] (end3) {END};
-
-  \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, \neg P \land D)$};
-  \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
-  \node[action, below=of satcprime] (inv) {$A':=\text{Model}(C')$\\
-                                        $I:=D \land \text{Inv}(N, A')$};
-  \node[print, below=of inv] (printinv) {Invariant: $I$};
-  \node[print, right=of satcprime] (noinv) {No Invariant};
-  \node[state, below=of noinv] (end1) {END};
-  \node[state, below=of printinv] (end2) {END};
-
-  \draw (cprime) edge (satcprime);
-  \draw (satcprime) edge node[above]{NO} (noinv);
-  \draw (satcprime) edge node[right]{YES} (inv);
-  \draw (inv) edge (printinv);
-  \draw (noinv) edge (end1);
-  \draw (printinv) edge (end2);
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (cprime);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (noinv2);
-  \draw (noinv2) edge (end3);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\caption{Diagram for Method Invariant by Refinement}
-\label{fig:method-invariant-by-refinement-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\State $D := $true
-\While {SAT$(C \cup \{\neg P\})$}
-  \State $A := $Model$(C \cup \{\neg P\})$
-  \State $C_\theta := $TrapConditions$(N, A)$
-  \If {SAT$(C_\theta)$}
-    \State $A_\theta := $Model$(C_\theta)$
-    \State $\delta := \Delta(A_\theta)$
-    \State $C := C \cup \{\delta\}$
-    \State $D := D \land \delta$
-  \Else
-    \State \Return 'Failed at finding an invariant'
-  \EndIf
-\EndWhile
-\State $C' := \mathcal C'(N, \neg P \land D)$
-\If {SAT$(C')$}
-  \State $I := D \land $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant by Refinement}
-\label{fig:method-invariant-by-refinement-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Subprocedures TrapConditions and \Delta are the same as
-  in Method Safety by Refinement
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Trap constraints as in example from section Method Safety by Refinement.
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  target_1 + trap_1
-
-    a        >= trap_1
-    b        >= target_1
-    nbit2    >= trap_1
-    target_1 >= 0
-    trap_1   >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-    trap_1   = 1
-
-  - Invariant:
-
-    a     + nbit2 >= 1
-    a + b + nbit2 <= 1
-\end{verbatim}
-
-\input{sec_experiments}
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}

### 2014-02-05 paper/workspace/fig_dia_method_safety_by_refinement.tex
Section for introduction, figures for methods.

--- 2014-02-05 paper/workspace/main.tex
@@
-\documentclass{llncs}
-
-\newcommand{\isDraft}{true}
-
-\input{preamble}
-
-\ifthenelse{\equal{\isDraft}{true}}{
-  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-  \pagestyle{plain}
-}{}
-
-\begin{document}
-
-\nocite{DOsualdoSAS13}
-
-\ifthenelse{\equal{\isDraft}{false}}{
-  \title{An SMT-based Approach to Coverability Analysis}
-  \author{
-    Javier Esparza\inst{1}
-    \and
-    Rupak Majumdar\inst{2}
-    \and
-    Philipp Meyer\inst{1}
-    \and
-    Filip Niksic\inst{2}
-    \and
-    Rusl\'{a}n Ledesma-Garza\inst{1}
-  }
-  \institute{
-    Technische Universit\"at M\"unchen
-    \and
-    MPI-SWS, Kaiserslautern and Saarbr\"ucken
-  }
-
-  \maketitle
-
-  \begin{abstract}
-    Model checkers based on Petri net coverability have been used
-    successfully in recent years to verify safety properties of concurrent
-    shared-memory or asynchronous message-passing software. The bottleneck
-    in the verification process is the iterative computation of the
-    (backward) reachable or (forward) coverable set, and most recent
-    research has focused on non-trivial heuristics to speed it up.
-
-    In this paper, we revisit a constraint approach to the problem based on
-    classical Petri net analysis techniques: marking equation, place
-    invariants, and traps. Given a marking M to be covered, the approach
-    constructs constraints corresponding to increasingly precise
-    overapproximations of the reachable markings covering M. If one of
-    these constraints is unsatisfiable, then M is not coverable.
-
-    Previous work on this approach suffered from the absence of efficient
-    decision procedures for linear arithmetic, which nowadays exist as a
-    part of SMT solvers. We show how to utilize an SMT solver to implement
-    the constraint approach, and additionally, to generate a minimized
-    inductive invariant from a safety proof.
-
-    We empirically evaluate our procedure on a large set of existing Petri
-    net benchmarks arising out of software verification. Surprisingly, even
-    though our technique is incomplete, we show that it can quickly
-    discharge all safe instances, usually competitive with, and occasionally
-    much faster than, state-of-the-art techniques. Additionally, the
-    inductive invariants computed are usually orders of magnitude smaller
-    than those produced by existing solvers.
-  \end{abstract}
-}{
-}
-
-\section{Introduction}
-
-This paper makes the following contributions.
-\begin{itemize}
-\item An application of \zthree\ to the constraint approach and its
-  corresponding evaluation.
-\item A method that given a network and a property constructs a
-  minimal invariant.
-\end{itemize}
-
-\input{sec_illustration}
-\input{sec_preliminaries}
-
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
-
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-    \node[print, right=of satc] (yes) {N satisfies P};
-    \node[print, below=of satc] (dontknow) {Don't know};
-    \node[state, right=of yes] (end1) {END};
-    \node[state, below=of dontknow] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (yes);
-    \draw (yes) edge (end1);
-    \draw (satc) edge node[right]{YES} (dontknow);
-    \draw (dontknow) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Safety}
-\label{fig:method-safety-pseudocode}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\If {SAT$(C \cup \{\neg P\})$}
-  \State \Return 'The petri net may not satisfy the property'
-\Else
-  \State \Return 'The petri net satisfies the property!'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Safety}
-\label{fig:method-safety-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land
-    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - Code:
-
-  Process 1:                     |      Process 2:
-                                 |
-      bit1 := false              |      bit2 := false
-      while true do              |      while true do
-  p1:   bit1 := true             |  q1:   bit2 := true
-  p2:   while bit2 do skip od    |  q2:   if bit1 then
-  p3:   <critical section>       |  q3:     bit2 := true
-        bit1 := false            |  q4:     while bit1 do skip od
-      od                         |          goto q1
-                                 |        fi
-                                 |  q5:   <critical section>
-                                 |        bit2 := false
-                                 |      od
-
-  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.
-
-  - Property: p3 + q5 <= 1
-
-  - Petri net:
-
-\end{verbatim}
-
-\input{fig_lamport_net}
-
-\newpage
-
-\begin{verbatim}
-  - State constraints C:
-
-    Place equations:
-
-    p1    = 1 - s1      + s3
-    ^       ^   ^         ^
-    |       |   |         |
-    |       |   |         |
-    |       |   |         |
-    |       |   |        # of tokens given to p1
-    |       |   |    
-    |       |  # of tokens taken from p1
-    |       |
-    |      initial number of tokens in p1
-    |
-    number of tokens in p1
-
-    p2    = 0 + s1 - s2
-    p3    = 0      + s2 - s3
-    bit1  = 0 + s1      - s3
-    nbit1 = 1 - s1      + s3
-    q1    = 1 +              - t1           + t4      + t6
-    q2    = 0 +              + t1 - t2           - t5
-    q3    = 0 +                   + t2 - t3
-    q4    = 0 +                        + t3 - t4
-    q5    = 0 +                                  + t5 - t6
-    nbit2 = 1 +              - t1      + t3           + t6
-
-    Non-negativity conditions:
-
-    p1    >= 0
-    p2    >= 0
-    p3    >= 0
-    bit1  >= 0
-    nbit1 >= 0
-    q1    >= 0
-    q2    >= 0
-    q3    >= 0
-    q4    >= 0
-    q5    >= 0
-    nbit2 >= 0
-
-  - Negated Property \neg P:
-
-    p3 + q5 >= 2
-\end{verbatim}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-
 \begin{figure}
@@
 \end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\While {SAT$(C \cup \{\neg P\})$}
-  \State $A := $Model$(C \cup \{\neg P\})$
-  \State $C_\theta := $TrapConditions$(N, A)$
-  \If {SAT$(C_\theta)$}
-    \State $A_\theta := $Model$(C_\theta)$
-    \State $\delta := \Delta(A_\theta)$
-    \State $C := C \cup \{\delta\}$
-  \Else
-    \State \Return 'The petri net may not satisfy the property'
-  \EndIf
-\EndWhile
-\State \Return 'The petri net satisfies the property!'
-\end{algorithmic}
-\caption{Pseudocode for Method Safety by Refinement}
-\label{fig:method-safety-by-refinement-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land
-    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
-    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - State constraints C same as in the example for method Safety.
-
-  - Assignment A:
-    p1    = 0
-    p2    = 0
-    p3    = 1
-    bit1  = 1
-    nbit1 = 0
-    q1    = 0
-    q2    = 0
-    q3    = 0
-    q4    = 0
-    q5    = 0
-    nbit2 = 0
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-                s1
-             ----------
-    p1    => p2 \/ bit1
-    ^           ^      ^
-    |           |      |
-    |           |     bit1 \in S
-    |           |
-    |          p2 \in S
-    |
-    p1 \in S
-
-    p2    => p3 \/ nbit2
-    p3    => p1 \/ nbit1
-    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
-    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
-    q1    => q2
-    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
-    q3    => q4 \/ nbit2
-    q4    => q1 \/ nbit1
-    q5    => q1 \/ nbit2
-    nbit2 => q2 /\ (p3 \/ nbit2)
-
-    - At least one of the initially marked places belongs to S:
-    p1 \/ q1 \/ nbit1 \/ nbit2
-
-    - None of the marked places in A belongs to S:
-    ~p3 /\ ~q5 /\ ~bit1
-
-  - Assignment A_\theta:
-    p1    = false
-    p2    = true
-    p3    = false
-    bit1  = false
-    nbit1 = true
-    q1    = false
-    q2    = true
-    q3    = true
-    q4    = false
-    q5    = false
-    nbit2 = true
-
-  - Refinement constraint \delta:
-    p2 + q3 + q5 + nbit1 + nbit2 >= 1
-    ^    ^    ^    ^       ^
-    |    |    |    |       |
-    ------------------------
-      |
-      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
-      assignment A in the next iteration
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A'.
-\end{verbatim}
-
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
-      $I := \text{Inv}(N, A')$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (satc) edge node[right]{YES} (inv);
-    \draw (noinv) edge (end1);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Invariant}
-\label{fig:method-invariant-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C' := \mathcal C'(N, \neg P)$
-\If {SAT$(C')$}
-  \State $A' := $Model$(C')$
-  \State $I := $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant}
-\label{fig:method-invariant-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
-              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N      = (S, T, E, M0)
-        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
-\neg P  = ( a + b >= 2 )
-
-Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
-                      \sum_{(t, s) \in E} s
-                    - \sum_{(s, t) \in E} s \right) \right) \land
-     \left( \sum_{s \in S} M_0(s) \cdot s <
-       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
-     & \left( \bigwedge_{s \in S} \left ( s \ge 
-       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
-     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
-\end{align*}
-
-\begin{verbatim}
-* Subprocedure Inv:
-
-  Input:
-    N = (S, T, E, M0) : Petri net
-    A'                : Satisfying assignment for C'
-
-  Output:
-    I                 : Invariant
-
-  Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
-                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  2 * target_1
-
-    a        >= target_1
-    b        >= target_1
-    nbit2    >= 0
-    target_1 >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-
-  - Invariant:
-
-    a + b + nbit2 <= 1
-
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A'
-\end{verbatim}
-
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (model1) {$A':=\text{Model}(C')$};
-    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A')$};
-    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
-    \node[action, right=of satcm] (model2) {$A':=\text{Model}(C' \cup C_M)$};
-    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A')$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (noinv) edge (end1);
-    \draw (satc) edge node[right]{YES} (model1);
-    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
-    \draw (cm) edge (satcm);
-    \draw (satcm) edge node[above]{YES} (model2);
-    \draw (model2) |- (edgein);
-    \draw (satcm) edge node[right]{NO} (inv);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Invariant with Minimization}
-\label{fig:method-invariant-with-minimization-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C' := \mathcal C'(N, \neg P)$
-\If {SAT$(C')$}
-  \State $A' := $Model$(C')$
-  \State $C_M := $MinConstraints$(N, A')$
-  \While {SAT$(C' \cup C_M)$}
-    \State $A' := $Model$(C' \cup C_M)$
-    \State $C_M := $MinConstraints$(N, A')$
-  \EndWhile
-  \State $I := $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant with Minimization}
-\label{fig:method-invariant-with-minimization-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A': 
-    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant
-
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    A'                 : Satisfying assignment for C'
-  Output:
-    C_M                : Minimization Constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
-      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
-    \right) \right) \land
-    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
-
-  - Minimization constraints C_M:
-
-    a     > 0 => b_a     = 1
-    |            -----------
-    |               |
-    |            place appears in invariant
-    |
-    place coefficient in the invariant
-
-    a     = 0 => b_a     = 0
-    |           ------------
-    |               |
-    |            place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    b     > 0 => b_b     = 1
-    b     = 0 => b_b     = 0
-    nbit2 > 0 => b_nbit2 = 1
-    nbit2 = 0 => b_nbit2 = 0
-
-    b_a + b_b + b_nbit2 < 3
-    -------------------   |
-                |         |
-                |    number of places appearing in current invariant for A'
-                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
-                |
-  number of places appearing in new invariant
-
-  - Model A' for C' \cup C_M:
-
-    a        = 1
-    b        = 1
-    nbit2    = 0
-    target_1 = 1
-
-  - Minimized Invariant:
-
-    a + b <= 1
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant by Refinement}
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety by Refinement.
-
-* Method Invariant by Refinement
-\end{verbatim}
-
-\begin{figure}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
-                                        $D:=\{\}$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                                        $C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                                $\delta:=\Delta(A_\theta)$\\
-                                                $C:=C \cup \{\delta\}$\\
-                                                $D:=D \cup \{\delta\}$};
-  \node[print, right=of satctheta] (noinv2) {No Invariant};
-  \node[state, below=of noinv2] (end3) {END};
-
-  \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, \neg P \land D)$};
-  \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
-  \node[action, below=of satcprime] (inv) {$A':=\text{Model}(C')$\\
-                                        $I:=D \land \text{Inv}(N, A')$};
-  \node[print, below=of inv] (printinv) {Invariant: $I$};
-  \node[print, right=of satcprime] (noinv) {No Invariant};
-  \node[state, below=of noinv] (end1) {END};
-  \node[state, below=of printinv] (end2) {END};
-
-  \draw (cprime) edge (satcprime);
-  \draw (satcprime) edge node[above]{NO} (noinv);
-  \draw (satcprime) edge node[right]{YES} (inv);
-  \draw (inv) edge (printinv);
-  \draw (noinv) edge (end1);
-  \draw (printinv) edge (end2);
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (cprime);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (noinv2);
-  \draw (noinv2) edge (end3);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\caption{Diagram for Method Invariant by Refinement}
-\label{fig:method-invariant-by-refinement-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\State $D := $true
-\While {SAT$(C \cup \{\neg P\})$}
-  \State $A := $Model$(C \cup \{\neg P\})$
-  \State $C_\theta := $TrapConditions$(N, A)$
-  \If {SAT$(C_\theta)$}
-    \State $A_\theta := $Model$(C_\theta)$
-    \State $\delta := \Delta(A_\theta)$
-    \State $C := C \cup \{\delta\}$
-    \State $D := D \land \delta$
-  \Else
-    \State \Return 'Failed at finding an invariant'
-  \EndIf
-\EndWhile
-\State $C' := \mathcal C'(N, \neg P \land D)$
-\If {SAT$(C')$}
-  \State $I := D \land $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant by Refinement}
-\label{fig:method-invariant-by-refinement-pseudocode}
-\end{figure}
-
-\begin{verbatim}
-* Subprocedures TrapConditions and \Delta are the same as
-  in Method Safety by Refinement
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Trap constraints as in example from section Method Safety by Refinement.
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  target_1 + trap_1
-
-    a        >= trap_1
-    b        >= target_1
-    nbit2    >= trap_1
-    target_1 >= 0
-    trap_1   >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-    trap_1   = 1
-
-  - Invariant:
-
-    a     + nbit2 >= 1
-    a + b + nbit2 <= 1
-\end{verbatim}
-
-\input{sec_experiments}
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}

### 2014-02-05 paper/workspace/main.tex
Section for introduction, figures for methods.

@@
 
-\section{Introduction}
-
-This paper makes the following contributions.
-\begin{itemize}
-\item An application of \zthree\ to the constraint approach and its
-  corresponding evaluation.
-\item A method that given a network and a property constructs a
-  minimal invariant.
-\end{itemize}
-
+\input{sec_introduction}
 \input{sec_illustration}
@@
 
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-    \node[print, right=of satc] (yes) {N satisfies P};
-    \node[print, below=of satc] (dontknow) {Don't know};
-    \node[state, right=of yes] (end1) {END};
-    \node[state, below=of dontknow] (end2) {END};
+\input{fig_dia_method_safety}
 
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (yes);
-    \draw (yes) edge (end1);
-    \draw (satc) edge node[right]{YES} (dontknow);
-    \draw (dontknow) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Safety}
-\label{fig:method-safety-pseudocode}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\If {SAT$(C \cup \{\neg P\})$}
-  \State \Return 'The petri net may not satisfy the property'
-\Else
-  \State \Return 'The petri net satisfies the property!'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Safety}
-\label{fig:method-safety-pseudocode}
-\end{figure}
+\input{fig_method_safety}
 
@@
 
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                    	$\delta:=\Delta(A_\theta)$\\
-                                    	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {N satisfies P};
-  \node[print, right=of satctheta] (dontknow) {Don't know};
-  \node[state, right=of yes] (end1) {END};
-  \node[state, right=of dontknow] (end2) {END};
+\input{fig_dia_method_safety_by_refinement}
 
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (yes);
-  \draw (yes) edge (end1);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (dontknow);
-  \draw (dontknow) edge (end2);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Safety by Refinement}
-\label{fig:method-safety-by-refinement-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\While {SAT$(C \cup \{\neg P\})$}
-  \State $A := $Model$(C \cup \{\neg P\})$
-  \State $C_\theta := $TrapConditions$(N, A)$
-  \If {SAT$(C_\theta)$}
-    \State $A_\theta := $Model$(C_\theta)$
-    \State $\delta := \Delta(A_\theta)$
-    \State $C := C \cup \{\delta\}$
-  \Else
-    \State \Return 'The petri net may not satisfy the property'
-  \EndIf
-\EndWhile
-\State \Return 'The petri net satisfies the property!'
-\end{algorithmic}
-\caption{Pseudocode for Method Safety by Refinement}
-\label{fig:method-safety-by-refinement-pseudocode}
-\end{figure}
+\input{fig_method_safety_by_refinement}
 
@@
 
-The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
+The method Invariant constructs an invariant \verb=I= for given Petri
+net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
 
@@
 
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
-      $I := \text{Inv}(N, A')$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
+\input{fig_dia_method_invariant}
 
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (satc) edge node[right]{YES} (inv);
-    \draw (noinv) edge (end1);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Invariant}
-\label{fig:method-invariant-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C' := \mathcal C'(N, \neg P)$
-\If {SAT$(C')$}
-  \State $A' := $Model$(C')$
-  \State $I := $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant}
-\label{fig:method-invariant-pseudocode}
-\end{figure}
+\input{fig_method_invariant}
 
@@
 
-\begin{figure}
-\begin{center}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (model1) {$A':=\text{Model}(C')$};
-    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A')$};
-    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
-    \node[action, right=of satcm] (model2) {$A':=\text{Model}(C' \cup C_M)$};
-    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A')$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
+\input{fig_dia_method_invariant_w_minimization}
 
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (noinv) edge (end1);
-    \draw (satc) edge node[right]{YES} (model1);
-    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
-    \draw (cm) edge (satcm);
-    \draw (satcm) edge node[above]{YES} (model2);
-    \draw (model2) |- (edgein);
-    \draw (satcm) edge node[right]{NO} (inv);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-\caption{Diagram for Method Invariant with Minimization}
-\label{fig:method-invariant-with-minimization-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C' := \mathcal C'(N, \neg P)$
-\If {SAT$(C')$}
-  \State $A' := $Model$(C')$
-  \State $C_M := $MinConstraints$(N, A')$
-  \While {SAT$(C' \cup C_M)$}
-    \State $A' := $Model$(C' \cup C_M)$
-    \State $C_M := $MinConstraints$(N, A')$
-  \EndWhile
-  \State $I := $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant with Minimization}
-\label{fig:method-invariant-with-minimization-pseudocode}
-\end{figure}
+\input{fig_method_invariant_w_minimization}
 
@@
 
-\begin{figure}
-  \begin{tikzpicture}[
-      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
-    ]
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
-                                        $D:=\{\}$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                                        $C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                                $\delta:=\Delta(A_\theta)$\\
-                                                $C:=C \cup \{\delta\}$\\
-                                                $D:=D \cup \{\delta\}$};
-  \node[print, right=of satctheta] (noinv2) {No Invariant};
-  \node[state, below=of noinv2] (end3) {END};
+\input{fig_dia_method_invariant_by_refinement}
 
-  \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, \neg P \land D)$};
-  \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
-  \node[action, below=of satcprime] (inv) {$A':=\text{Model}(C')$\\
-                                        $I:=D \land \text{Inv}(N, A')$};
-  \node[print, below=of inv] (printinv) {Invariant: $I$};
-  \node[print, right=of satcprime] (noinv) {No Invariant};
-  \node[state, below=of noinv] (end1) {END};
-  \node[state, below=of printinv] (end2) {END};
-
-  \draw (cprime) edge (satcprime);
-  \draw (satcprime) edge node[above]{NO} (noinv);
-  \draw (satcprime) edge node[right]{YES} (inv);
-  \draw (inv) edge (printinv);
-  \draw (noinv) edge (end1);
-  \draw (printinv) edge (end2);
-
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (cprime);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (noinv2);
-  \draw (noinv2) edge (end3);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\caption{Diagram for Method Invariant by Refinement}
-\label{fig:method-invariant-by-refinement-diagram}
-\end{figure}
-
-\begin{figure}
-\begin{algorithmic}[1]
-\State $C := \mathcal C(N)$
-\State $D := $true
-\While {SAT$(C \cup \{\neg P\})$}
-  \State $A := $Model$(C \cup \{\neg P\})$
-  \State $C_\theta := $TrapConditions$(N, A)$
-  \If {SAT$(C_\theta)$}
-    \State $A_\theta := $Model$(C_\theta)$
-    \State $\delta := \Delta(A_\theta)$
-    \State $C := C \cup \{\delta\}$
-    \State $D := D \land \delta$
-  \Else
-    \State \Return 'Failed at finding an invariant'
-  \EndIf
-\EndWhile
-\State $C' := \mathcal C'(N, \neg P \land D)$
-\If {SAT$(C')$}
-  \State $I := D \land $Inv$(N, A')$
-  \State \Return 'Invariant I for the petri net: $I$'
-\Else
-  \State \Return 'Failed at finding an invariant'
-\EndIf
-\end{algorithmic}
-\caption{Pseudocode for Method Invariant by Refinement}
-\label{fig:method-invariant-by-refinement-pseudocode}
-\end{figure}
+\input{fig_method_invariant_by_refinement}
 

### 2014-02-05 paper/workspace/main.tex
Added results for methods invariant and invariant with minimization

--- 2014-02-05 paper/workspace/main.tex
@@
 
-  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.
+  - Property 1: Process 1 and Process 2 are never in their respective critical section at the same time.
 
-  - Property: p3 + q5 <= 1
+  - Property 1: p3 + q5 <= 1
+
+  - Property 2: Process 1 is in at most one state at a time
+
+  - Property 2: p1 + p2 + p3 <= 1
 
@@
 
-  - Negated Property \neg P:
+  - Using property 2
 
-    p3 + q5 >= 2
+  - Negated property 2 \neg P:
+
+    p1 + p2 + p3 >= 2
 \end{verbatim}
@@
   - State constraints C same as in the example for method Safety.
+
+  - Using property 1
+
+  - Negated property 1 \neg P:
+
+    p3 + q5 >= 2
 
@@
 
+  - Using property 2
+
   - Dual state constraints C':
 
-    0        >=    a  - b
-    0        >= (- a) + b
+    0 >= - p1 + p2      + bit1 - nbit2
+    0 >=      - p2 + p3
+    0 >= + p1      - p3 - bit1 + nbit2
 
-    a        <  2 * target_1
+    0 >= - q1 + q2                - nbit2
+    0 >=      - q2 + q3
+    0 >=           - q3 + q4      + nbit2
+    0 >= + q1           - q4
+    0 >=      - q2           + q5
+    0 >= + q1                - q5 + nbit2
 
-    a        >= target_1
-    b        >= target_1
+    p1 + q1 + nb1 + nb2 < 2 * target_1
+
+    p1    >= target_1
+    p2    >= target_1
+    p3    >= target_1
+    bit1  >= 0
+    nbit1 >= 0
+    q1    >= 0
+    q2    >= 0
+    q3    >= 0
+    q4    >= 0
+    q5    >= 0
     nbit2 >= 0
+
     target_1 >= 0
@@
 
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
+    p1       = 3
+    p2       = 2
+    p3       = 2
+    bit1     = 1
+    nbit1    = 0
+    q1       = 0
+    q2       = 0
+    q3       = 0
+    q4       = 0
+    q5       = 0
+    nbit2    = 0
+    target_1 = 2
 
@@
 
-    a + b + nbit2 <= 1
+    3 * p1 + 2 * p2 + 2 * p3 + bit1 <= 3
 
@@
 
-    a     > 0 => b_a     = 1
+    p1     > 0 => b_p1   = 1
     |            -----------
     |               |
-    |            place appears in invariant
+    |            place p1 appears in invariant
     |
-    place coefficient in the invariant
+    place p1 coefficient in the invariant
 
-    a     = 0 => b_a     = 0
+    p1     = 0 => b_p1   = 0
     |           ------------
     |               |
-    |            place does not appear in invariant
+    |            place p1 does not appear in invariant
     |
-    place coefficient in the invariant
+    place p1 coefficient in the invariant
 
-    b     > 0 => b_b     = 1
-    b     = 0 => b_b     = 0
+    p2    > 0 => b_p2    = 1
+    p2    = 0 => b_p2    = 0
+    p3    > 0 => b_p3    = 1
+    p3    = 0 => b_p3    = 0
+    bit1  > 0 => b_bit1  = 1
+    bit1  = 0 => b_bit1  = 0
+    nbit1 > 0 => b_nbit1 = 1
+    nbit1 = 0 => b_nbit1 = 0
+
+    q1    > 0 => b_q1    = 1
+    q1    = 0 => b_q1    = 0
+    q2    > 0 => b_q2    = 1
+    q2    = 0 => b_q2    = 0
+    q3    > 0 => b_q3    = 1
+    q3    = 0 => b_q3    = 0
     nbit2 > 0 => b_nbit2 = 1
@@
 
-    b_a + b_b + b_nbit2 < 3
-    -------------------   |
+    b_p1 + b_p2 + b_p3 + b_bit1 + b_nbit1 + b_q1 + b_q2 + b_q3 + b_q4 + b_q5
+         + b_nbit2 < 4
+    --------------   |
                 |    |
                 |    number of places appearing in current invariant for A'
-                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
+                |    = #{s | A'(s) > 0} = #{p1, p2, p3, bit1}
                 |
@@
 
-    a        = 1
-    b        = 1
+    p1       = 1
+    p2       = 1
+    p3       = 1
+    bit1     = 0
+    nbit1    = 0
+    q1       = 0
+    q2       = 0
+    q3       = 0
+    q4       = 0
+    q5       = 0
     nbit2    = 0
@@
 
-    a + b <= 1
+    p1 + p2 + p3 <= 1
 \end{verbatim}
@@
 
+  - Using property 1
+
   - Trap constraints as in example from section Method Safety by Refinement.

### 2014-02-05 paper/workspace/fig_dia_method_invariant.tex

--- 2014-02-05 paper/workspace/main.tex
@@
-\documentclass{llncs}
+\begin{figure}
+  \begin{center}
+    \begin{tikzpicture}[
+      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
+      ]
+      \node[state] (begin) {BEGIN};
+      \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
+      \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
+      \node[print, right=of satc] (noinv) {No Invariant};
+      \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
+        $I := \text{Inv}(N, A')$};
+      \node[print, below=of inv] (printinv) {Invariant: $I$};
+      \node[state, right=of noinv] (end1) {END};
+      \node[state, below=of printinv] (end2) {END};
 
-\newcommand{\isDraft}{true}
-
-\input{preamble}
-
-\ifthenelse{\equal{\isDraft}{true}}{
-  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-  \pagestyle{plain}
-}{}
-
-\begin{document}
-
-\nocite{DOsualdoSAS13}
-
-\ifthenelse{\equal{\isDraft}{false}}{
-  \title{An SMT-based Approach to Coverability Analysis}
-  \author{
-    Javier Esparza\inst{1}
-    \and
-    Rupak Majumdar\inst{2}
-    \and
-    Philipp Meyer\inst{1}
-    \and
-    Filip Niksic\inst{2}
-    \and
-    Rusl\'{a}n Ledesma-Garza\inst{1}
-  }
-  \institute{
-    Technische Universit\"at M\"unchen
-    \and
-    MPI-SWS, Kaiserslautern and Saarbr\"ucken
-  }
-
-  \maketitle
-
-  \begin{abstract}
-    Model checkers based on Petri net coverability have been used
-    successfully in recent years to verify safety properties of concurrent
-    shared-memory or asynchronous message-passing software. The bottleneck
-    in the verification process is the iterative computation of the
-    (backward) reachable or (forward) coverable set, and most recent
-    research has focused on non-trivial heuristics to speed it up.
-
-    In this paper, we revisit a constraint approach to the problem based on
-    classical Petri net analysis techniques: marking equation, place
-    invariants, and traps. Given a marking M to be covered, the approach
-    constructs constraints corresponding to increasingly precise
-    overapproximations of the reachable markings covering M. If one of
-    these constraints is unsatisfiable, then M is not coverable.
-
-    Previous work on this approach suffered from the absence of efficient
-    decision procedures for linear arithmetic, which nowadays exist as a
-    part of SMT solvers. We show how to utilize an SMT solver to implement
-    the constraint approach, and additionally, to generate a minimized
-    inductive invariant from a safety proof.
-
-    We empirically evaluate our procedure on a large set of existing Petri
-    net benchmarks arising out of software verification. Surprisingly, even
-    though our technique is incomplete, we show that it can quickly
-    discharge all safe instances, usually competitive with, and occasionally
-    much faster than, state-of-the-art techniques. Additionally, the
-    inductive invariants computed are usually orders of magnitude smaller
-    than those produced by existing solvers.
-  \end{abstract}
-}{
-}
-
-\input{sec_introduction}
-\input{sec_illustration}
-\input{sec_preliminaries}
-
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
-
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\input{fig_dia_method_safety}
-
-\input{fig_method_safety}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land
-    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - Code:
-
-  Process 1:                     |      Process 2:
-                                 |
-      bit1 := false              |      bit2 := false
-      while true do              |      while true do
-  p1:   bit1 := true             |  q1:   bit2 := true
-  p2:   while bit2 do skip od    |  q2:   if bit1 then
-  p3:   <critical section>       |  q3:     bit2 := true
-        bit1 := false            |  q4:     while bit1 do skip od
-      od                         |          goto q1
-                                 |        fi
-                                 |  q5:   <critical section>
-                                 |        bit2 := false
-                                 |      od
-
-  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.
-
-  - Property: p3 + q5 <= 1
-
-  - Petri net:
-
-\end{verbatim}
-
-\input{fig_lamport_net}
-
-\newpage
-
-\begin{verbatim}
-  - State constraints C:
-
-    Place equations:
-
-    p1    = 1 - s1      + s3
-    ^       ^   ^         ^
-    |       |   |         |
-    |       |   |         |
-    |       |   |         |
-    |       |   |        # of tokens given to p1
-    |       |   |    
-    |       |  # of tokens taken from p1
-    |       |
-    |      initial number of tokens in p1
-    |
-    number of tokens in p1
-
-    p2    = 0 + s1 - s2
-    p3    = 0      + s2 - s3
-    bit1  = 0 + s1      - s3
-    nbit1 = 1 - s1      + s3
-    q1    = 1 +              - t1           + t4      + t6
-    q2    = 0 +              + t1 - t2           - t5
-    q3    = 0 +                   + t2 - t3
-    q4    = 0 +                        + t3 - t4
-    q5    = 0 +                                  + t5 - t6
-    nbit2 = 1 +              - t1      + t3           + t6
-
-    Non-negativity conditions:
-
-    p1    >= 0
-    p2    >= 0
-    p3    >= 0
-    bit1  >= 0
-    nbit1 >= 0
-    q1    >= 0
-    q2    >= 0
-    q3    >= 0
-    q4    >= 0
-    q5    >= 0
-    nbit2 >= 0
-
-  - Negated Property \neg P:
-
-    p3 + q5 >= 2
-\end{verbatim}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-
-\input{fig_dia_method_safety_by_refinement}
-
-\input{fig_method_safety_by_refinement}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land
-    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
-    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - State constraints C same as in the example for method Safety.
-
-  - Assignment A:
-    p1    = 0
-    p2    = 0
-    p3    = 1
-    bit1  = 1
-    nbit1 = 0
-    q1    = 0
-    q2    = 0
-    q3    = 0
-    q4    = 0
-    q5    = 0
-    nbit2 = 0
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-                s1
-             ----------
-    p1    => p2 \/ bit1
-    ^           ^      ^
-    |           |      |
-    |           |     bit1 \in S
-    |           |
-    |          p2 \in S
-    |
-    p1 \in S
-
-    p2    => p3 \/ nbit2
-    p3    => p1 \/ nbit1
-    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
-    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
-    q1    => q2
-    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
-    q3    => q4 \/ nbit2
-    q4    => q1 \/ nbit1
-    q5    => q1 \/ nbit2
-    nbit2 => q2 /\ (p3 \/ nbit2)
-
-    - At least one of the initially marked places belongs to S:
-    p1 \/ q1 \/ nbit1 \/ nbit2
-
-    - None of the marked places in A belongs to S:
-    ~p3 /\ ~q5 /\ ~bit1
-
-  - Assignment A_\theta:
-    p1    = false
-    p2    = true
-    p3    = false
-    bit1  = false
-    nbit1 = true
-    q1    = false
-    q2    = true
-    q3    = true
-    q4    = false
-    q5    = false
-    nbit2 = true
-
-  - Refinement constraint \delta:
-    p2 + q3 + q5 + nbit1 + nbit2 >= 1
-    ^    ^    ^    ^       ^
-    |    |    |    |       |
-    ------------------------
-      |
-      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
-      assignment A in the next iteration
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri
-net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A'.
-\end{verbatim}
-
-\input{fig_dia_method_invariant}
-
-\input{fig_method_invariant}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
-              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N      = (S, T, E, M0)
-        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
-\neg P  = ( a + b >= 2 )
-
-Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
-                      \sum_{(t, s) \in E} s
-                    - \sum_{(s, t) \in E} s \right) \right) \land
-     \left( \sum_{s \in S} M_0(s) \cdot s <
-       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
-     & \left( \bigwedge_{s \in S} \left ( s \ge 
-       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
-     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
-\end{align*}
-
-\begin{verbatim}
-* Subprocedure Inv:
-
-  Input:
-    N = (S, T, E, M0) : Petri net
-    A'                : Satisfying assignment for C'
-
-  Output:
-    I                 : Invariant
-
-  Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
-                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  2 * target_1
-
-    a        >= target_1
-    b        >= target_1
-    nbit2    >= 0
-    target_1 >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-
-  - Invariant:
-
-    a + b + nbit2 <= 1
-
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A'
-\end{verbatim}
-
-\input{fig_dia_method_invariant_w_minimization}
-
-\input{fig_method_invariant_w_minimization}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A': 
-    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant
-
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    A'                 : Satisfying assignment for C'
-  Output:
-    C_M                : Minimization Constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
-      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
-    \right) \right) \land
-    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
-
-  - Minimization constraints C_M:
-
-    a     > 0 => b_a     = 1
-    |            -----------
-    |               |
-    |            place appears in invariant
-    |
-    place coefficient in the invariant
-
-    a     = 0 => b_a     = 0
-    |           ------------
-    |               |
-    |            place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    b     > 0 => b_b     = 1
-    b     = 0 => b_b     = 0
-    nbit2 > 0 => b_nbit2 = 1
-    nbit2 = 0 => b_nbit2 = 0
-
-    b_a + b_b + b_nbit2 < 3
-    -------------------   |
-                |         |
-                |    number of places appearing in current invariant for A'
-                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
-                |
-  number of places appearing in new invariant
-
-  - Model A' for C' \cup C_M:
-
-    a        = 1
-    b        = 1
-    nbit2    = 0
-    target_1 = 1
-
-  - Minimized Invariant:
-
-    a + b <= 1
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant by Refinement}
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety by Refinement.
-
-* Method Invariant by Refinement
-\end{verbatim}
-
-\input{fig_dia_method_invariant_by_refinement}
-
-\input{fig_method_invariant_by_refinement}
-
-\begin{verbatim}
-* Subprocedures TrapConditions and \Delta are the same as
-  in Method Safety by Refinement
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Trap constraints as in example from section Method Safety by Refinement.
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  target_1 + trap_1
-
-    a        >= trap_1
-    b        >= target_1
-    nbit2    >= trap_1
-    target_1 >= 0
-    trap_1   >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-    trap_1   = 1
-
-  - Invariant:
-
-    a     + nbit2 >= 1
-    a + b + nbit2 <= 1
-\end{verbatim}
-
-\input{sec_experiments}
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}
+      \draw (begin) edge (c);
+      \draw (c) edge (satc);
+      \draw (satc) edge node[above]{NO} (noinv);
+      \draw (satc) edge node[right]{YES} (inv);
+      \draw (noinv) edge (end1);
+      \draw (inv) edge (printinv);
+      \draw (printinv) edge (end2);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Diagram for Method Invariant}
+  \label{fig:method-invariant-diagram}
+\end{figure}

### 2014-02-05 paper/workspace/fig_dia_method_invariant_by_refinement.tex

--- 2014-02-05 paper/workspace/main.tex
@@
-\documentclass{llncs}
+\begin{figure}
+  \begin{tikzpicture}[
+    every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
+    ]
+    \node[state] (begin) {BEGIN};
+    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
+      $D:=\{\}$};
+    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
+    \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
+      $C_{\theta}:=\text{TrapConditions}(N, A)$};
+    \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
+    \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
+      $\delta:=\Delta(A_\theta)$\\
+      $C:=C \cup \{\delta\}$\\
+      $D:=D \cup \{\delta\}$};
+    \node[print, right=of satctheta] (noinv2) {No Invariant};
+    \node[state, below=of noinv2] (end3) {END};
 
-\newcommand{\isDraft}{true}
+    \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, \neg P \land D)$};
+    \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
+    \node[action, below=of satcprime] (inv) {$A':=\text{Model}(C')$\\
+      $I:=D \land \text{Inv}(N, A')$};
+    \node[print, below=of inv] (printinv) {Invariant: $I$};
+    \node[print, right=of satcprime] (noinv) {No Invariant};
+    \node[state, below=of noinv] (end1) {END};
+    \node[state, below=of printinv] (end2) {END};
 
-\input{preamble}
+    \draw (cprime) edge (satcprime);
+    \draw (satcprime) edge node[above]{NO} (noinv);
+    \draw (satcprime) edge node[right]{YES} (inv);
+    \draw (inv) edge (printinv);
+    \draw (noinv) edge (end1);
+    \draw (printinv) edge (end2);
 
-\ifthenelse{\equal{\isDraft}{true}}{
-  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-  \pagestyle{plain}
-}{}
-
-\begin{document}
-
-\nocite{DOsualdoSAS13}
-
-\ifthenelse{\equal{\isDraft}{false}}{
-  \title{An SMT-based Approach to Coverability Analysis}
-  \author{
-    Javier Esparza\inst{1}
-    \and
-    Rupak Majumdar\inst{2}
-    \and
-    Philipp Meyer\inst{1}
-    \and
-    Filip Niksic\inst{2}
-    \and
-    Rusl\'{a}n Ledesma-Garza\inst{1}
-  }
-  \institute{
-    Technische Universit\"at M\"unchen
-    \and
-    MPI-SWS, Kaiserslautern and Saarbr\"ucken
-  }
-
-  \maketitle
-
-  \begin{abstract}
-    Model checkers based on Petri net coverability have been used
-    successfully in recent years to verify safety properties of concurrent
-    shared-memory or asynchronous message-passing software. The bottleneck
-    in the verification process is the iterative computation of the
-    (backward) reachable or (forward) coverable set, and most recent
-    research has focused on non-trivial heuristics to speed it up.
-
-    In this paper, we revisit a constraint approach to the problem based on
-    classical Petri net analysis techniques: marking equation, place
-    invariants, and traps. Given a marking M to be covered, the approach
-    constructs constraints corresponding to increasingly precise
-    overapproximations of the reachable markings covering M. If one of
-    these constraints is unsatisfiable, then M is not coverable.
-
-    Previous work on this approach suffered from the absence of efficient
-    decision procedures for linear arithmetic, which nowadays exist as a
-    part of SMT solvers. We show how to utilize an SMT solver to implement
-    the constraint approach, and additionally, to generate a minimized
-    inductive invariant from a safety proof.
-
-    We empirically evaluate our procedure on a large set of existing Petri
-    net benchmarks arising out of software verification. Surprisingly, even
-    though our technique is incomplete, we show that it can quickly
-    discharge all safe instances, usually competitive with, and occasionally
-    much faster than, state-of-the-art techniques. Additionally, the
-    inductive invariants computed are usually orders of magnitude smaller
-    than those produced by existing solvers.
-  \end{abstract}
-}{
-}
-
-\input{sec_introduction}
-\input{sec_illustration}
-\input{sec_preliminaries}
-
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
-
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\input{fig_dia_method_safety}
-
-\input{fig_method_safety}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land
-    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - Code:
-
-  Process 1:                     |      Process 2:
-                                 |
-      bit1 := false              |      bit2 := false
-      while true do              |      while true do
-  p1:   bit1 := true             |  q1:   bit2 := true
-  p2:   while bit2 do skip od    |  q2:   if bit1 then
-  p3:   <critical section>       |  q3:     bit2 := true
-        bit1 := false            |  q4:     while bit1 do skip od
-      od                         |          goto q1
-                                 |        fi
-                                 |  q5:   <critical section>
-                                 |        bit2 := false
-                                 |      od
-
-  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.
-
-  - Property: p3 + q5 <= 1
-
-  - Petri net:
-
-\end{verbatim}
-
-\input{fig_lamport_net}
-
-\newpage
-
-\begin{verbatim}
-  - State constraints C:
-
-    Place equations:
-
-    p1    = 1 - s1      + s3
-    ^       ^   ^         ^
-    |       |   |         |
-    |       |   |         |
-    |       |   |         |
-    |       |   |        # of tokens given to p1
-    |       |   |    
-    |       |  # of tokens taken from p1
-    |       |
-    |      initial number of tokens in p1
-    |
-    number of tokens in p1
-
-    p2    = 0 + s1 - s2
-    p3    = 0      + s2 - s3
-    bit1  = 0 + s1      - s3
-    nbit1 = 1 - s1      + s3
-    q1    = 1 +              - t1           + t4      + t6
-    q2    = 0 +              + t1 - t2           - t5
-    q3    = 0 +                   + t2 - t3
-    q4    = 0 +                        + t3 - t4
-    q5    = 0 +                                  + t5 - t6
-    nbit2 = 1 +              - t1      + t3           + t6
-
-    Non-negativity conditions:
-
-    p1    >= 0
-    p2    >= 0
-    p3    >= 0
-    bit1  >= 0
-    nbit1 >= 0
-    q1    >= 0
-    q2    >= 0
-    q3    >= 0
-    q4    >= 0
-    q5    >= 0
-    nbit2 >= 0
-
-  - Negated Property \neg P:
-
-    p3 + q5 >= 2
-\end{verbatim}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-
-\input{fig_dia_method_safety_by_refinement}
-
-\input{fig_method_safety_by_refinement}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land
-    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
-    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - State constraints C same as in the example for method Safety.
-
-  - Assignment A:
-    p1    = 0
-    p2    = 0
-    p3    = 1
-    bit1  = 1
-    nbit1 = 0
-    q1    = 0
-    q2    = 0
-    q3    = 0
-    q4    = 0
-    q5    = 0
-    nbit2 = 0
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-                s1
-             ----------
-    p1    => p2 \/ bit1
-    ^           ^      ^
-    |           |      |
-    |           |     bit1 \in S
-    |           |
-    |          p2 \in S
-    |
-    p1 \in S
-
-    p2    => p3 \/ nbit2
-    p3    => p1 \/ nbit1
-    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
-    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
-    q1    => q2
-    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
-    q3    => q4 \/ nbit2
-    q4    => q1 \/ nbit1
-    q5    => q1 \/ nbit2
-    nbit2 => q2 /\ (p3 \/ nbit2)
-
-    - At least one of the initially marked places belongs to S:
-    p1 \/ q1 \/ nbit1 \/ nbit2
-
-    - None of the marked places in A belongs to S:
-    ~p3 /\ ~q5 /\ ~bit1
-
-  - Assignment A_\theta:
-    p1    = false
-    p2    = true
-    p3    = false
-    bit1  = false
-    nbit1 = true
-    q1    = false
-    q2    = true
-    q3    = true
-    q4    = false
-    q5    = false
-    nbit2 = true
-
-  - Refinement constraint \delta:
-    p2 + q3 + q5 + nbit1 + nbit2 >= 1
-    ^    ^    ^    ^       ^
-    |    |    |    |       |
-    ------------------------
-      |
-      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
-      assignment A in the next iteration
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri
-net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A'.
-\end{verbatim}
-
-\input{fig_dia_method_invariant}
-
-\input{fig_method_invariant}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
-              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N      = (S, T, E, M0)
-        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
-\neg P  = ( a + b >= 2 )
-
-Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
-                      \sum_{(t, s) \in E} s
-                    - \sum_{(s, t) \in E} s \right) \right) \land
-     \left( \sum_{s \in S} M_0(s) \cdot s <
-       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
-     & \left( \bigwedge_{s \in S} \left ( s \ge 
-       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
-     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
-\end{align*}
-
-\begin{verbatim}
-* Subprocedure Inv:
-
-  Input:
-    N = (S, T, E, M0) : Petri net
-    A'                : Satisfying assignment for C'
-
-  Output:
-    I                 : Invariant
-
-  Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
-                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  2 * target_1
-
-    a        >= target_1
-    b        >= target_1
-    nbit2    >= 0
-    target_1 >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-
-  - Invariant:
-
-    a + b + nbit2 <= 1
-
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A'
-\end{verbatim}
-
-\input{fig_dia_method_invariant_w_minimization}
-
-\input{fig_method_invariant_w_minimization}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A': 
-    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant
-
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    A'                 : Satisfying assignment for C'
-  Output:
-    C_M                : Minimization Constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
-      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
-    \right) \right) \land
-    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
-
-  - Minimization constraints C_M:
-
-    a     > 0 => b_a     = 1
-    |            -----------
-    |               |
-    |            place appears in invariant
-    |
-    place coefficient in the invariant
-
-    a     = 0 => b_a     = 0
-    |           ------------
-    |               |
-    |            place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    b     > 0 => b_b     = 1
-    b     = 0 => b_b     = 0
-    nbit2 > 0 => b_nbit2 = 1
-    nbit2 = 0 => b_nbit2 = 0
-
-    b_a + b_b + b_nbit2 < 3
-    -------------------   |
-                |         |
-                |    number of places appearing in current invariant for A'
-                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
-                |
-  number of places appearing in new invariant
-
-  - Model A' for C' \cup C_M:
-
-    a        = 1
-    b        = 1
-    nbit2    = 0
-    target_1 = 1
-
-  - Minimized Invariant:
-
-    a + b <= 1
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant by Refinement}
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety by Refinement.
-
-* Method Invariant by Refinement
-\end{verbatim}
-
-\input{fig_dia_method_invariant_by_refinement}
-
-\input{fig_method_invariant_by_refinement}
-
-\begin{verbatim}
-* Subprocedures TrapConditions and \Delta are the same as
-  in Method Safety by Refinement
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Trap constraints as in example from section Method Safety by Refinement.
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  target_1 + trap_1
-
-    a        >= trap_1
-    b        >= target_1
-    nbit2    >= trap_1
-    target_1 >= 0
-    trap_1   >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-    trap_1   = 1
-
-  - Invariant:
-
-    a     + nbit2 >= 1
-    a + b + nbit2 <= 1
-\end{verbatim}
-
-\input{sec_experiments}
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}
+    \draw (begin) edge (c);
+    \draw (c) edge coordinate[pos=.5] (edgein) (satc);
+    \draw (satc) edge node[above]{NO} (cprime);
+    \draw (satc) edge node[right]{YES} (modelc);
+    \draw (modelc) edge (satctheta);
+    \draw (satctheta) edge node[above]{NO} (noinv2);
+    \draw (noinv2) edge (end3);
+    \draw (satctheta) edge node[right]{YES} (modelctheta);
+    \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
+    -| ([xshift=-2cm] modelctheta.west) |- (edgein);
+  \end{tikzpicture}
+  \caption{Diagram for Method Invariant by Refinement}
+  \label{fig:method-invariant-by-refinement-diagram}
+\end{figure}

### 2014-02-05 paper/workspace/fig_dia_method_safety.tex

--- 2014-02-05 paper/workspace/main.tex
@@
-\documentclass{llncs}
+\begin{figure}
+  \begin{center}
+    \begin{tikzpicture}[
+      every path/.style={draw, ->, >=stealth, shorten >=2pt, shorten <=2pt}
+      ]
+      \node[state] (begin) {BEGIN};
+      \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
+      \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
+      \node[print, right=of satc] (yes) {N satisfies P};
+      \node[print, below=of satc] (dontknow) {Don't know};
+      \node[state, right=of yes] (end1) {END};
+      \node[state, below=of dontknow] (end2) {END};
 
-\newcommand{\isDraft}{true}
-
-\input{preamble}
-
-\ifthenelse{\equal{\isDraft}{true}}{
-  \usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-  \pagestyle{plain}
-}{}
-
-\begin{document}
-
-\nocite{DOsualdoSAS13}
-
-\ifthenelse{\equal{\isDraft}{false}}{
-  \title{An SMT-based Approach to Coverability Analysis}
-  \author{
-    Javier Esparza\inst{1}
-    \and
-    Rupak Majumdar\inst{2}
-    \and
-    Philipp Meyer\inst{1}
-    \and
-    Filip Niksic\inst{2}
-    \and
-    Rusl\'{a}n Ledesma-Garza\inst{1}
-  }
-  \institute{
-    Technische Universit\"at M\"unchen
-    \and
-    MPI-SWS, Kaiserslautern and Saarbr\"ucken
-  }
-
-  \maketitle
-
-  \begin{abstract}
-    Model checkers based on Petri net coverability have been used
-    successfully in recent years to verify safety properties of concurrent
-    shared-memory or asynchronous message-passing software. The bottleneck
-    in the verification process is the iterative computation of the
-    (backward) reachable or (forward) coverable set, and most recent
-    research has focused on non-trivial heuristics to speed it up.
-
-    In this paper, we revisit a constraint approach to the problem based on
-    classical Petri net analysis techniques: marking equation, place
-    invariants, and traps. Given a marking M to be covered, the approach
-    constructs constraints corresponding to increasingly precise
-    overapproximations of the reachable markings covering M. If one of
-    these constraints is unsatisfiable, then M is not coverable.
-
-    Previous work on this approach suffered from the absence of efficient
-    decision procedures for linear arithmetic, which nowadays exist as a
-    part of SMT solvers. We show how to utilize an SMT solver to implement
-    the constraint approach, and additionally, to generate a minimized
-    inductive invariant from a safety proof.
-
-    We empirically evaluate our procedure on a large set of existing Petri
-    net benchmarks arising out of software verification. Surprisingly, even
-    though our technique is incomplete, we show that it can quickly
-    discharge all safe instances, usually competitive with, and occasionally
-    much faster than, state-of-the-art techniques. Additionally, the
-    inductive invariants computed are usually orders of magnitude smaller
-    than those produced by existing solvers.
-  \end{abstract}
-}{
-}
-
-\input{sec_introduction}
-\input{sec_illustration}
-\input{sec_preliminaries}
-
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
-
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\input{fig_dia_method_safety}
-
-\input{fig_method_safety}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land
-    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - Code:
-
-  Process 1:                     |      Process 2:
-                                 |
-      bit1 := false              |      bit2 := false
-      while true do              |      while true do
-  p1:   bit1 := true             |  q1:   bit2 := true
-  p2:   while bit2 do skip od    |  q2:   if bit1 then
-  p3:   <critical section>       |  q3:     bit2 := true
-        bit1 := false            |  q4:     while bit1 do skip od
-      od                         |          goto q1
-                                 |        fi
-                                 |  q5:   <critical section>
-                                 |        bit2 := false
-                                 |      od
-
-  - Property: Process 1 and Process 2 are never in their respective critical section at the same time.
-
-  - Property: p3 + q5 <= 1
-
-  - Petri net:
-
-\end{verbatim}
-
-\input{fig_lamport_net}
-
-\newpage
-
-\begin{verbatim}
-  - State constraints C:
-
-    Place equations:
-
-    p1    = 1 - s1      + s3
-    ^       ^   ^         ^
-    |       |   |         |
-    |       |   |         |
-    |       |   |         |
-    |       |   |        # of tokens given to p1
-    |       |   |    
-    |       |  # of tokens taken from p1
-    |       |
-    |      initial number of tokens in p1
-    |
-    number of tokens in p1
-
-    p2    = 0 + s1 - s2
-    p3    = 0      + s2 - s3
-    bit1  = 0 + s1      - s3
-    nbit1 = 1 - s1      + s3
-    q1    = 1 +              - t1           + t4      + t6
-    q2    = 0 +              + t1 - t2           - t5
-    q3    = 0 +                   + t2 - t3
-    q4    = 0 +                        + t3 - t4
-    q5    = 0 +                                  + t5 - t6
-    nbit2 = 1 +              - t1      + t3           + t6
-
-    Non-negativity conditions:
-
-    p1    >= 0
-    p2    >= 0
-    p3    >= 0
-    bit1  >= 0
-    nbit1 >= 0
-    q1    >= 0
-    q2    >= 0
-    q3    >= 0
-    q4    >= 0
-    q5    >= 0
-    nbit2 >= 0
-
-  - Negated Property \neg P:
-
-    p3 + q5 >= 2
-\end{verbatim}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-
-\input{fig_dia_method_safety_by_refinement}
-
-\input{fig_method_safety_by_refinement}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land
-    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
-    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - State constraints C same as in the example for method Safety.
-
-  - Assignment A:
-    p1    = 0
-    p2    = 0
-    p3    = 1
-    bit1  = 1
-    nbit1 = 0
-    q1    = 0
-    q2    = 0
-    q3    = 0
-    q4    = 0
-    q5    = 0
-    nbit2 = 0
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-                s1
-             ----------
-    p1    => p2 \/ bit1
-    ^           ^      ^
-    |           |      |
-    |           |     bit1 \in S
-    |           |
-    |          p2 \in S
-    |
-    p1 \in S
-
-    p2    => p3 \/ nbit2
-    p3    => p1 \/ nbit1
-    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
-    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
-    q1    => q2
-    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
-    q3    => q4 \/ nbit2
-    q4    => q1 \/ nbit1
-    q5    => q1 \/ nbit2
-    nbit2 => q2 /\ (p3 \/ nbit2)
-
-    - At least one of the initially marked places belongs to S:
-    p1 \/ q1 \/ nbit1 \/ nbit2
-
-    - None of the marked places in A belongs to S:
-    ~p3 /\ ~q5 /\ ~bit1
-
-  - Assignment A_\theta:
-    p1    = false
-    p2    = true
-    p3    = false
-    bit1  = false
-    nbit1 = true
-    q1    = false
-    q2    = true
-    q3    = true
-    q4    = false
-    q5    = false
-    nbit2 = true
-
-  - Refinement constraint \delta:
-    p2 + q3 + q5 + nbit1 + nbit2 >= 1
-    ^    ^    ^    ^       ^
-    |    |    |    |       |
-    ------------------------
-      |
-      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
-      assignment A in the next iteration
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri
-net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A'.
-\end{verbatim}
-
-\input{fig_dia_method_invariant}
-
-\input{fig_method_invariant}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
-              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N      = (S, T, E, M0)
-        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
-\neg P  = ( a + b >= 2 )
-
-Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
-                      \sum_{(t, s) \in E} s
-                    - \sum_{(s, t) \in E} s \right) \right) \land
-     \left( \sum_{s \in S} M_0(s) \cdot s <
-       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
-     & \left( \bigwedge_{s \in S} \left ( s \ge 
-       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
-     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
-\end{align*}
-
-\begin{verbatim}
-* Subprocedure Inv:
-
-  Input:
-    N = (S, T, E, M0) : Petri net
-    A'                : Satisfying assignment for C'
-
-  Output:
-    I                 : Invariant
-
-  Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
-                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  2 * target_1
-
-    a        >= target_1
-    b        >= target_1
-    nbit2    >= 0
-    target_1 >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-
-  - Invariant:
-
-    a + b + nbit2 <= 1
-
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A'
-\end{verbatim}
-
-\input{fig_dia_method_invariant_w_minimization}
-
-\input{fig_method_invariant_w_minimization}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A': 
-    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant
-
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    A'                 : Satisfying assignment for C'
-  Output:
-    C_M                : Minimization Constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
-      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
-    \right) \right) \land
-    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
-
-  - Minimization constraints C_M:
-
-    a     > 0 => b_a     = 1
-    |            -----------
-    |               |
-    |            place appears in invariant
-    |
-    place coefficient in the invariant
-
-    a     = 0 => b_a     = 0
-    |           ------------
-    |               |
-    |            place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    b     > 0 => b_b     = 1
-    b     = 0 => b_b     = 0
-    nbit2 > 0 => b_nbit2 = 1
-    nbit2 = 0 => b_nbit2 = 0
-
-    b_a + b_b + b_nbit2 < 3
-    -------------------   |
-                |         |
-                |    number of places appearing in current invariant for A'
-                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
-                |
-  number of places appearing in new invariant
-
-  - Model A' for C' \cup C_M:
-
-    a        = 1
-    b        = 1
-    nbit2    = 0
-    target_1 = 1
-
-  - Minimized Invariant:
-
-    a + b <= 1
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant by Refinement}
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety by Refinement.
-
-* Method Invariant by Refinement
-\end{verbatim}
-
-\input{fig_dia_method_invariant_by_refinement}
-
-\input{fig_method_invariant_by_refinement}
-
-\begin{verbatim}
-* Subprocedures TrapConditions and \Delta are the same as
-  in Method Safety by Refinement
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Trap constraints as in example from section Method Safety by Refinement.
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  target_1 + trap_1
-
-    a        >= trap_1
-    b        >= target_1
-    nbit2    >= trap_1
-    target_1 >= 0
-    trap_1   >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-    trap_1   = 1
-
-  - Invariant:
-
-    a     + nbit2 >= 1
-    a + b + nbit2 <= 1
-\end{verbatim}
-
-\input{sec_experiments}
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}
+      \draw (begin) edge (c);
+      \draw (c) edge (satc);
+      \draw (satc) edge node[above]{NO} (yes);
+      \draw (yes) edge (end1);
+      \draw (satc) edge node[right]{YES} (dontknow);
+      \draw (dontknow) edge (end2);
+    \end{tikzpicture}
+  \end{center}
+  \caption{Diagram for Method Safety}
+  \label{fig:method-safety-pseudocode}
+\end{figure}

### 2014-02-05 paper/workspace/main.tex

--- 2014-02-05 paper/workspace/main.tex
@@
 * Example:
+
+  - Using property 1
 

### 2014-02-05 paper/workspace/sec_introduction.tex

\section{Introduction}

This paper makes the following contributions.
\begin{itemize}
\item An application of \zthree\ to the constraint approach and its
  corresponding evaluation.
\item A method that given a network and a property constructs a
  minimal invariant.
\end{itemize}

### 2014-02-05 paper/workspace/main.tex
Added results for method invariant by refinement

@@
     q4    = 0
-    q5    = 0
+    q5    = 1
     nbit2 = 0
@@
   - Refinement constraint \delta:
-    p2 + q3 + q5 + nbit1 + nbit2 >= 1
+    p2 + q2 + q3 + nbit1 + nbit2 >= 1
     ^    ^    ^    ^       ^
@@
       |
-      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
+      S = {p2, q2, q3, nbit1, nbit2}, therefore \delta excludes
       assignment A in the next iteration
@@
     q3    = 0 => b_q3    = 0
+    q4    > 0 => b_q4    = 1
+    q4    = 0 => b_q4    = 0
+    q5    > 0 => b_q5    = 1
+    q5    = 0 => b_q5    = 0
     nbit2 > 0 => b_nbit2 = 1
@@
     target_1 = 1
+
+    b_p1     = 1
+    b_p2     = 1
+    b_p3     = 1
+    b_bit1   = 0
+    b_nbit1  = 0
+    b_q1     = 0
+    b_q2     = 0
+    b_q3     = 0
+    b_q4     = 0
+    b_q5     = 0
+    b_nbit2  = 0
 
@@
 
-    0        >=    a  - b
-    0        >= (- a) + b
+    0 >= - p1 + p2      + bit1 - nbit2
+    0 >=      - p2 + p3
+    0 >= + p1      - p3 - bit1 + nbit2
 
-    a        <  target_1 + trap_1
+    0 >= - q1 + q2                - nbit2
+    0 >=      - q2 + q3
+    0 >=           - q3 + q4      + nbit2
+    0 >= + q1           - q4
+    0 >=      - q2           + q5
+    0 >= + q1                - q5 + nbit2
 
-    a        >= trap_1
-    b        >= target_1
+    p1 + q1 + nb1 + nb2 < 2 * target_1 + trap_1
+
+    p1    >= 0
+    p2    >= trap_1
+    p3    >= target_1
+    bit1  >= 0
+    nbit1 >= trap_1
+    q1    >= 0
+    q2    >= trap_1
+    q3    >= trap_1
+    q4    >= 0
+    q5    >= target_1
     nbit2 >= trap_1
+
     target_1 >= 0
@@
 
-    a        = 1
-    b        = 1
+    p1       = 0
+    p2       = 1
+    p3       = 1
+    bit1     = 0
+    nbit1    = 1
+    q1       = 0
+    q2       = 1
+    q3       = 1
+    q4       = 0
+    q5       = 1
     nbit2    = 1
@@
 
-    a     + nbit2 >= 1
-    a + b + nbit2 <= 1
+    p2        q2 + q3      + nbit1 + nbit2 >= 1
+    p2 + p3 + q2 + q3 + q5 + nbit1 + nbit2 <= 2
+
 \end{verbatim}

### 2014-02-05 paper/workspace/main.tex
Merge?

@@
 
-\newcommand{\isDraft}{true}
+\newcommand{\isDraft}{false}
 
@@
 
-\input{fig_dia_method_safety}
+\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_safety}}{}
 
@@
 
-\input{fig_dia_method_safety_by_refinement}
+\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_safety_by_refinement}}{}
 
@@
 
-\input{fig_dia_method_invariant}
+\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_invariant}}{}
 
@@
 
-\input{fig_dia_method_invariant_w_minimization}
+\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_invariant_w_minimization}}{}
 
@@
 
-\input{fig_dia_method_invariant_by_refinement}
+\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_invariant_by_refinement}}{}
 

### 2014-02-05 paper/workspace/sec_preliminaries.tex
Outline for preliminaries.

@@
 \begin{verbatim}
-* Petri net:
+* Petri net: definition
  N = (S, T, E, M0), where
@@
 
-* Property:
- P, linear arithmetic formula
+* Marking: definition
+
+* Visual representation of Petri Nets: description, example
+
+* Property P: safety property expressed over linear arithmetic formulas.
+
+* Linear arithmetic formula: examples
+
+* N \models P: definition
+
+* Whatever you believe should be defined for the paper.
 \end{verbatim}

### 2014-02-05 paper/workspace/sec_preliminaries.tex
Updated preliminaries with definitions

--- 2014-02-05 paper/workspace/sec_preliminaries.tex
@@
 
-\begin{verbatim}
-* Petri net:
- N = (S, T, E, M0), where
-     S : places
-     T : transitions
-     E : edges
-     M0: initial marking
+A \emph{Petri net} is a tuple $(P, T, F, M_0)$, where $P$ is the set of
+\emph{places}, $T$ is the set of \emph{transitions},
+$F \subseteq P\times T \cup T\times P$ is the \emph{flow relation}
+and $M_0$ is the initial marking.
+We identify $F$ with its characteristic function
+$P\times T \cup T\times P \to \{0, 1\}$.
 
-* Property:
- P, linear arithmetic formula
-\end{verbatim}
+For $x\in P\cup T$, the \emph{pre-set} is
+$\pre{x}=\{y\in P\cup T\mid (y,x) \in F\}$
+and the \emph{post-set} is $\post{x}=\{y\in P\cup T\mid (x,y) \in F\}$.
+The pre- and post-set of a subset of $P \cup T$ are the union of
+the pre and post-sets of its elements.
+
+A \emph{marking} of a Petri net is a function $M\colon P \to \mathbb{N}$,
+which describes the number of tokens $M(p)$ that the marking puts in
+each place $p$.
+
+Petri nets are represented graphically as follows: places and transitions
+are represented as circles and boxes, respectively. An element $(x,y)$
+of the flow relation is represented by an arc leading from $x$ to $y$.
+A token on a place $p$ is represented by a black dot in the circle
+corresponding to $p$.
+
+A transition $t \in T$ is \emph{enabled at $M$} iff
+$\forall p \in \pre{t} \colon M(p) \ge F(p, t)$.
+If $t$ is enabled at $M$, then $t$ may \emph{fire} or \emph{occur},
+yielding a new marking $M'$ (denoted as $M \xrightarrow{t} M'$),
+where $M'(p) = M(p) + F(t,p) - F(p,t)$.
+
+A sequence of transitions, $\sigma = t_1 t_2 \ldots t_r$ is an
+$\emph{occurence sequence}$ of $N$ iff there exist markings
+$M_1, \ldots, M_r$ such that $M_0 \xrightarrow{t_1} M_1
+\xrightarrow{t_2} M_2 \ldots \xrightarrow{t_r} M_r$. The marking
+$M_r$ is said to be \emph{reachable} from $M_0$ by the occurence
+of $\sigma$ (denoted $M \xrightarrow{\sigma} M_r$).
+
+A property $P$ is a safety property expressed over linear arithmetic
+formulas. The property $P$ holds on a marking $M$ iff $M \models P$.
+Examples of properties are $s_1 \le 2$, $s_1 + s_2 \ge 1$ and
+$((s_1 \le 1) \land (s_2 \ge 1)) \lor (s_3 \le 1)$.
+
+A Petri net $P$ satisfies a property $P$ (denoted by $N \models P$)
+iff for all reachable markings $M_0 \xrightarrow{\sigma} M$
+$M \models P$ holds.
 

### 2014-02-05 paper/workspace/main.tex

--- 2014-02-05 paper/workspace/main.tex
@@
   \author{
-    Javier Esparza\inst{1}
+    Javier~Esparza\inst{1}
     \and
-    Rupak Majumdar\inst{2}
+    Rupak~Majumdar\inst{2}
     \and
-    Philipp Meyer\inst{1}
+    Philipp~Meyer\inst{1}
     \and
-    Filip Niksic\inst{2}
+    Filip~Niksic\inst{2}
     \and
-    Rusl\'{a}n Ledesma-Garza\inst{1}
+    Rusl\'{a}n~Ledesma-Garza\inst{1}
   }
@@
     \and
-    MPI-SWS, Kaiserslautern and Saarbr\"ucken
+    MPI-SWS,~Kaiserslautern~and~Saarbr\"ucken
   }
@@
 \input{sec_preliminaries}
+\newpage
+\input{sec_constraint_approach}
+\newpage
+\input{sec_invariants}
+\newpage
+\input{sec_experiments}
+\input{sec_rel_work}
+\input{sec_conclusions}
 
-\newpage
-\section{Method Safety}
-
-The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
-
-\begin{verbatim}
-* Method Safety:
-
-  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
-\end{verbatim}
-
-\input{fig_dia_method_safety}
-
-\input{fig_method_safety}
-
-\begin{verbatim}
-* Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Place equation:
-
-  For a given place s the place equation is
-
-  # of tokens in s = initial number of tokens of place s
-                     + # times each input transition of s fires
-                     - # times each output transition of s fires
-
-* Non-negativity conditions:
-
-  # of tokens in place s           >= 0
-  # of times transition t is fired >= 0
-
-* Subprocedure \mathcal{C}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-  Output:
-    C : State constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left(
-    s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land
-    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
-    \left( \bigwedge_{t \in T} t \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - Code:
-
-  Process 1:                     |      Process 2:
-                                 |
-      bit1 := false              |      bit2 := false
-      while true do              |      while true do
-  p1:   bit1 := true             |  q1:   bit2 := true
-  p2:   while bit2 do skip od    |  q2:   if bit1 then
-  p3:   <critical section>       |  q3:     bit2 := true
-        bit1 := false            |  q4:     while bit1 do skip od
-      od                         |          goto q1
-                                 |        fi
-                                 |  q5:   <critical section>
-                                 |        bit2 := false
-                                 |      od
-
-  - Property 1: Process 1 and Process 2 are never in their respective critical section at the same time.
-
-  - Property 1: p3 + q5 <= 1
-
-  - Property 2: Process 1 is in at most one state at a time
-
-  - Property 2: p1 + p2 + p3 <= 1
-
-  - Petri net:
-
-\end{verbatim}
-
-\input{fig_lamport_net}
-
-\newpage
-
-\begin{verbatim}
-  - State constraints C:
-
-    Place equations:
-
-    p1    = 1 - s1      + s3
-    ^       ^   ^         ^
-    |       |   |         |
-    |       |   |         |
-    |       |   |         |
-    |       |   |        # of tokens given to p1
-    |       |   |    
-    |       |  # of tokens taken from p1
-    |       |
-    |      initial number of tokens in p1
-    |
-    number of tokens in p1
-
-    p2    = 0 + s1 - s2
-    p3    = 0      + s2 - s3
-    bit1  = 0 + s1      - s3
-    nbit1 = 1 - s1      + s3
-    q1    = 1 +              - t1           + t4      + t6
-    q2    = 0 +              + t1 - t2           - t5
-    q3    = 0 +                   + t2 - t3
-    q4    = 0 +                        + t3 - t4
-    q5    = 0 +                                  + t5 - t6
-    nbit2 = 1 +              - t1      + t3           + t6
-
-    Non-negativity conditions:
-
-    p1    >= 0
-    p2    >= 0
-    p3    >= 0
-    bit1  >= 0
-    nbit1 >= 0
-    q1    >= 0
-    q2    >= 0
-    q3    >= 0
-    q4    >= 0
-    q5    >= 0
-    nbit2 >= 0
-
-  - Using property 2
-
-  - Negated property 2 \neg P:
-
-    p1 + p2 + p3 >= 2
-\end{verbatim}
-
-\newpage
-
-\section{Method Safety by Refinement}
-
-The method Safety by Refinement applies trap conditions to check that a given Petri net \verb=N= never violates a property \verb=P=.
-
-\begin{verbatim}
-* Method Safety by Refinement:
-
-  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
-  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
-\end{verbatim}
-
-\input{fig_dia_method_safety_by_refinement}
-
-\input{fig_method_safety_by_refinement}
-
-\begin{verbatim}
-* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
-  1. S is a trap in the net N
-  2. S is marked in the initial marking M0
-  3. S is unmarked in the assignment A
-
-* Property of A_\theta: for each place s, A_\theta(s) iff s \in S
-
-* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
-  1. A ^ \delta is unsat (\delta excludes A)
-  2. N |= \delta is sat  (\delta is a property of N)
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land
-    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
-    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example
-
-  - State constraints C same as in the example for method Safety.
-
-  - Using property 1
-
-  - Negated property 1 \neg P:
-
-    p3 + q5 >= 2
-
-  - Assignment A:
-    p1    = 0
-    p2    = 0
-    p3    = 1
-    bit1  = 1
-    nbit1 = 0
-    q1    = 0
-    q2    = 0
-    q3    = 0
-    q4    = 0
-    q5    = 0
-    nbit2 = 0
-
-  - Trap conditions C_\theta:
-
-    - Trap implications:
-                s1
-             ----------
-    p1    => p2 \/ bit1
-    ^           ^      ^
-    |           |      |
-    |           |     bit1 \in S
-    |           |
-    |          p2 \in S
-    |
-    p1 \in S
-
-    p2    => p3 \/ nbit2
-    p3    => p1 \/ nbit1
-    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
-    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
-    q1    => q2
-    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
-    q3    => q4 \/ nbit2
-    q4    => q1 \/ nbit1
-    q5    => q1 \/ nbit2
-    nbit2 => q2 /\ (p3 \/ nbit2)
-
-    - At least one of the initially marked places belongs to S:
-    p1 \/ q1 \/ nbit1 \/ nbit2
-
-    - None of the marked places in A belongs to S:
-    ~p3 /\ ~q5 /\ ~bit1
-
-  - Assignment A_\theta:
-    p1    = false
-    p2    = true
-    p3    = false
-    bit1  = false
-    nbit1 = true
-    q1    = false
-    q2    = true
-    q3    = true
-    q4    = false
-    q5    = false
-    nbit2 = true
-
-  - Refinement constraint \delta:
-    p2 + q3 + q5 + nbit1 + nbit2 >= 1
-    ^    ^    ^    ^       ^
-    |    |    |    |       |
-    ------------------------
-      |
-      S = {p2, q3, q5, nbit1, nbit2}, therefore \delta excludes
-      assignment A in the next iteration
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Trap implication:
-
-  place s \in S =>    /\       \    /
-                     /  \       \  /     place p \in S
-                    /    \       \/
-                   t \in s*   p \in t*
-
-  "if s is in trap S then for each output transition t at least one successor p is in trap S"
-
-* Refinement constraint \delta:
-
-  \Sigma s >= 1  
-  A(s)
-
-  "At least one place in S is always marked"
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri
-net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A'.
-\end{verbatim}
-
-\input{fig_dia_method_invariant}
-
-\input{fig_method_invariant}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
-              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N      = (S, T, E, M0)
-        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
-\neg P  = ( a + b >= 2 )
-
-Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
-                      \sum_{(t, s) \in E} s
-                    - \sum_{(s, t) \in E} s \right) \right) \land
-     \left( \sum_{s \in S} M_0(s) \cdot s <
-       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
-     & \left( \bigwedge_{s \in S} \left ( s \ge 
-       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
-     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
-\end{align*}
-
-\begin{verbatim}
-* Subprocedure Inv:
-
-  Input:
-    N = (S, T, E, M0) : Petri net
-    A'                : Satisfying assignment for C'
-
-  Output:
-    I                 : Invariant
-
-  Pseudocode:
-\end{verbatim}
-
-\begin{align*}
-  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
-                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Using property 2
-
-  - Dual state constraints C':
-
-    0 >= - p1 + p2      + bit1 - nbit2
-    0 >=      - p2 + p3
-    0 >= + p1      - p3 - bit1 + nbit2
-
-    0 >= - q1 + q2                - nbit2
-    0 >=      - q2 + q3
-    0 >=           - q3 + q4      + nbit2
-    0 >= + q1           - q4
-    0 >=      - q2           + q5
-    0 >= + q1                - q5 + nbit2
-
-    p1 + q1 + nb1 + nb2 < 2 * target_1
-
-    p1    >= target_1
-    p2    >= target_1
-    p3    >= target_1
-    bit1  >= 0
-    nbit1 >= 0
-    q1    >= 0
-    q2    >= 0
-    q3    >= 0
-    q4    >= 0
-    q5    >= 0
-    nbit2 >= 0
-
-    target_1 >= 0
-
-  - Model A':
-
-    p1       = 3
-    p2       = 2
-    p3       = 2
-    bit1     = 1
-    nbit1    = 0
-    q1       = 0
-    q2       = 0
-    q3       = 0
-    q4       = 0
-    q5       = 0
-    nbit2    = 0
-    target_1 = 2
-
-  - Invariant:
-
-    3 * p1 + 2 * p2 + 2 * p3 + bit1 <= 3
-
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A'
-\end{verbatim}
-
-\input{fig_dia_method_invariant_w_minimization}
-
-\input{fig_method_invariant_w_minimization}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A': 
-    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant
-
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    N = (S, T, E, M0)  : Petri net
-    A'                 : Satisfying assignment for C'
-  Output:
-    C_M                : Minimization Constraints
-
-  Pseudocode:
-
-\end{verbatim}
-
-\begin{align*}
-  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
-      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
-    \right) \right) \land
-    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
-\end{align*}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant
-
-  - Minimization constraints C_M:
-
-    p1     > 0 => b_p1   = 1
-    |            -----------
-    |               |
-    |            place p1 appears in invariant
-    |
-    place p1 coefficient in the invariant
-
-    p1     = 0 => b_p1   = 0
-    |           ------------
-    |               |
-    |            place p1 does not appear in invariant
-    |
-    place p1 coefficient in the invariant
-
-    p2    > 0 => b_p2    = 1
-    p2    = 0 => b_p2    = 0
-    p3    > 0 => b_p3    = 1
-    p3    = 0 => b_p3    = 0
-    bit1  > 0 => b_bit1  = 1
-    bit1  = 0 => b_bit1  = 0
-    nbit1 > 0 => b_nbit1 = 1
-    nbit1 = 0 => b_nbit1 = 0
-
-    q1    > 0 => b_q1    = 1
-    q1    = 0 => b_q1    = 0
-    q2    > 0 => b_q2    = 1
-    q2    = 0 => b_q2    = 0
-    q3    > 0 => b_q3    = 1
-    q3    = 0 => b_q3    = 0
-    nbit2 > 0 => b_nbit2 = 1
-    nbit2 = 0 => b_nbit2 = 0
-
-    b_p1 + b_p2 + b_p3 + b_bit1 + b_nbit1 + b_q1 + b_q2 + b_q3 + b_q4 + b_q5
-         + b_nbit2 < 4
-    --------------   |
-                |    |
-                |    number of places appearing in current invariant for A'
-                |    = #{s | A'(s) > 0} = #{p1, p2, p3, bit1}
-                |
-  number of places appearing in new invariant
-
-  - Model A' for C' \cup C_M:
-
-    p1       = 1
-    p2       = 1
-    p3       = 1
-    bit1     = 0
-    nbit1    = 0
-    q1       = 0
-    q2       = 0
-    q3       = 0
-    q4       = 0
-    q5       = 0
-    nbit2    = 0
-    target_1 = 1
-
-  - Minimized Invariant:
-
-    p1 + p2 + p3 <= 1
-\end{verbatim}
-
-\newpage
-
-\section{Method Invariant by Refinement}
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety by Refinement.
-
-* Method Invariant by Refinement
-\end{verbatim}
-
-\input{fig_dia_method_invariant_by_refinement}
-
-\input{fig_method_invariant_by_refinement}
-
-\begin{verbatim}
-* Subprocedures TrapConditions and \Delta are the same as
-  in Method Safety by Refinement
-
-* Subprocedures \mathcal{C'} and Inv are the same as
-  in Method Invariant
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-* Example:
-
-  - Using property 1
-
-  - Trap constraints as in example from section Method Safety by Refinement.
-
-  - Dual state constraints C':
-
-    0        >=    a  - b
-    0        >= (- a) + b
-
-    a        <  target_1 + trap_1
-
-    a        >= trap_1
-    b        >= target_1
-    nbit2    >= trap_1
-    target_1 >= 0
-    trap_1   >= 0
-
-  - Model A':
-
-    a        = 1
-    b        = 1
-    nbit2    = 1
-    target_1 = 1
-    trap_1   = 1
-
-  - Invariant:
-
-    a     + nbit2 >= 1
-    a + b + nbit2 <= 1
-\end{verbatim}
-
-\input{sec_experiments}
 \bibliographystyle{abbrv}

### 2014-02-05 paper/workspace/sec_constraint_approach.tex

--- 2014-02-01 paper/workspace/main-example-filip.tex
@@
-\documentclass{article}
-
-\usepackage{amsmath}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{listings}
-\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-\usepackage{graphicx}
-
-\usepackage{tikz}
-\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
-	snakes,automata,backgrounds,petri,calc,positioning}
-
-\newcommand{\bfc}{{\sc Bfc}}
-\newcommand{\mist}{{\sc Mist}}
-\newcommand{\iic}{{\sc IIC}}
-\newcommand{\pnerf}{{\sc Pnerf}}
-\newcommand{\zthree}{{\sc Z3}}
-\newcommand{\ttt}[1]{\texttt{#1}}
-
-\tikzstyle{place}=[circle,thick,draw=black!75,fill=white!20,
-  	minimum size=6mm]
-\tikzstyle{red place}=[place,draw=red!75,fill=red!20]
-\tikzstyle{blue place}=[place,draw=blue!75,fill=blue!20]
-\tikzstyle{transition}=[rectangle,thick,draw=black!75,
-  	fill=black!20,minimum size=4mm]
-
-\tikzstyle{state}=[draw, ellipse, aspect=2]
-\tikzstyle{action}=[draw, rectangle, align=center]
-\tikzstyle{decision}=[draw, diamond, aspect=2, align=center]
-\tikzstyle{print}=[draw, trapezium, trapezium left angle=70, trapezium right angle=-70]
-\tikzstyle{every edge}=[draw, ->, >=stealth, shorten >=2pt, shorten <=2pt]
-
-\begin{document}
-
-\title{Report on a Safety Checker for Petri nets}
-\author{Philipp Meyer \and Rusl\'{a}n Ledesma-Garza}
-% \institute{Technische Universit\"at M\"unchen}
-\date{Mon Nov 25 14:24:10 CET 2013}
-
-\section{Preliminaries}
-
-\begin{verbatim}
-* Petri net:
- N = (S, T, E, M0), where
- S : places
- T : transitions
- E : edges
- M0: initial marking
-
-* Property:
- P, linear arithmetic formula
-\end{verbatim}
-\iffalse
-\begin{verbatim}
- Examples:
-  x + y < 0
-  x + y < 0 \\and x + z > 0
-  x + y < 0 \\or  x + z > 0
-
- ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
-  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
-\end{verbatim}
-\fi
-
-\newpage
 \section{Method Safety}
@@
 The method Safety checks that a given Petri net \verb=N= never violates a property \verb=P=.
-We present the method Safety by example on a modified Petri net from [Esparza1997].
+We present the method Safety by example on Lamport's 1-bit algorithm [Esparza1997].
 
-\begin{verbatim}
-* Code
-
-TODO
-
-* Property: z is never marked
-
-* Petri net:
-
-\end{verbatim}
-
-\begin{center}
-  \includegraphics[width=6cm]{example-filip}
-\end{center}
-
-\begin{verbatim}
-* Property P: z = 0
-\end{verbatim}
-
-\newpage
 \begin{verbatim}
@@
 
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-    \node[print, right=of satc] (yes) {N satisfies P};
-    \node[print, below=of satc] (dontknow) {Don't know};
-    \node[state, right=of yes] (end1) {END};
-    \node[state, below=of dontknow] (end2) {END};
+\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_safety}}{}
 
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (yes);
-    \draw (yes) edge (end1);
-    \draw (satc) edge node[right]{YES} (dontknow);
-    \draw (dontknow) edge (end2);
-  \end{tikzpicture}
-\end{center}
+\input{fig_method_safety}
 
@@
 * Property of state constraints C: If C U {\neg P} is unsat then N |= P.
-
-* Example:
-  - State constraints C:
-
-    Place equations:
-
-    x = 1 - t
-    ^   ^   ^
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |   |
-    |   |  # of tokens taken from x
-    |   |
-    |   initial number of tokens in x
-    |
-    number of tokens in x
-
-    y = 0
-    z = 0 + t
-    w = 0 + t
-\end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-    Non-negativity conditions:
-
-    x >= 0
-    y >= 0
-    z >= 0
-    w >= 0
-    t >= 0
-
-  - Negated Property \neg P:
-
-    z >= 1
 
@@
 \end{verbatim}
+
 \begin{align*}
@@
     s = M_0(s) + \sum_{(t, s) \in E} t - \sum_{(s, t) \in E} t
-  \right) \right) \land \\
-  & \left( \bigwedge_{s \in S} s \ge 0 \right) \land
+  \right) \right) \land
+    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
     \left( \bigwedge_{t \in T} t \ge 0 \right)
 \end{align*}
+
+\newpage
+
+\begin{verbatim}
+* Example
+
+  - Code:
+
+  Process 1:                     |      Process 2:
+                                 |
+      bit1 := false              |      bit2 := false
+      while true do              |      while true do
+  p1:   bit1 := true             |  q1:   bit2 := true
+  p2:   while bit2 do skip od    |  q2:   if bit1 then
+  p3:   <critical section>       |  q3:     bit2 := true
+        bit1 := false            |  q4:     while bit1 do skip od
+      od                         |          goto q1
+                                 |        fi
+                                 |  q5:   <critical section>
+                                 |        bit2 := false
+                                 |      od
+
+  - Property 1: Process 1 and Process 2 are never in their respective critical section at the same time.
+
+  - Property 1: p3 + q5 <= 1
+
+  - Property 2: Process 1 is in at most one state at a time
+
+  - Property 2: p1 + p2 + p3 <= 1
+
+  - Petri net:
+
+\end{verbatim}
+
+\input{fig_lamport_net}
+
+\newpage
+
+\begin{verbatim}
+  - State constraints C:
+
+    Place equations:
+
+    p1    = 1 - s1      + s3
+    ^       ^   ^         ^
+    |       |   |         |
+    |       |   |         |
+    |       |   |         |
+    |       |   |        # of tokens given to p1
+    |       |   |    
+    |       |  # of tokens taken from p1
+    |       |
+    |      initial number of tokens in p1
+    |
+    number of tokens in p1
+
+    p2    = 0 + s1 - s2
+    p3    = 0      + s2 - s3
+    bit1  = 0 + s1      - s3
+    nbit1 = 1 - s1      + s3
+    q1    = 1 +              - t1           + t4      + t6
+    q2    = 0 +              + t1 - t2           - t5
+    q3    = 0 +                   + t2 - t3
+    q4    = 0 +                        + t3 - t4
+    q5    = 0 +                                  + t5 - t6
+    nbit2 = 1 +              - t1      + t3           + t6
+
+    Non-negativity conditions:
+
+    p1    >= 0
+    p2    >= 0
+    p3    >= 0
+    bit1  >= 0
+    nbit1 >= 0
+    q1    >= 0
+    q2    >= 0
+    q3    >= 0
+    q4    >= 0
+    q5    >= 0
+    nbit2 >= 0
+
+  - Using property 2
+
+  - Negated property 2 \neg P:
+
+    p1 + p2 + p3 >= 2
+\end{verbatim}
 
@@
 \end{verbatim}
-\begin{center}
-\begin{tikzpicture}
-  \node[state] (begin) {BEGIN};
-  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$};
-  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
-  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
-                               	$C_{\theta}:=\text{TrapConditions}(N, A)$};
-  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
-  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
-                                    	$\delta:=\Delta(A_\theta)$\\
-                                    	$C:=C \cup \{\delta\}$};
-  \node[print, right=of satc] (yes) {N satisfies P};
-  \node[print, right=of satctheta] (dontknow) {Don't know};
-  \node[state, right=of yes] (end1) {END};
-  \node[state, right=of dontknow] (end2) {END};
 
-  \draw (begin) edge (c);
-  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
-  \draw (satc) edge node[above]{NO} (yes);
-  \draw (yes) edge (end1);
-  \draw (satc) edge node[right]{YES} (modelc);
-  \draw (modelc) edge (satctheta);
-  \draw (satctheta) edge node[above]{NO} (dontknow);
-  \draw (dontknow) edge (end2);
-  \draw (satctheta) edge node[right]{YES} (modelctheta);
-  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
-  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
-\end{tikzpicture}
-\end{center}
+\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_safety_by_refinement}}{}
+
+\input{fig_method_safety_by_refinement}
 
@@
 \newpage
+
+\begin{verbatim}
+* Subprocedure TrapConditions:
+
+  Input:
+    (S, T, E, M0) : Petri net
+    A             : Satisfying assignment for C \cup { ~P }
+  Output:
+    C_\theta      : Trap conditions
+
+  Pseudocode:
+
+\end{verbatim}
+\begin{align*}
+  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
+      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
+    \right) \right) \land
+    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
+    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
+\end{align*}
+\begin{verbatim}
+
+* Subprocedure \Delta:
+
+  Input:
+    A_\theta      : Satisfying assignment for C_\theta
+  Output:
+    \delta        : Refinement constraint \delta
+
+  Pseudocode:
+
+\end{verbatim}
+\begin{align*}
+  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)
+\end{align*}
+
+\newpage
+
 \begin{verbatim}
 * Example
-  - State constraints C and \neg P are the same as in the example for method Safety.
+
+  - State constraints C same as in the example for method Safety.
+
+  - Using property 1
+
+  - Negated property 1 \neg P:
+
+    p3 + q5 >= 2
 
   - Assignment A:
-    x = 0
-    y = 0
-    z = 1
-    w = 1
-    t = 1
+    p1    = 0
+    p2    = 0
+    p3    = 1
+    bit1  = 1
+    nbit1 = 0
+    q1    = 0
+    q2    = 0
+    q3    = 0
+    q4    = 0
+    q5    = 1
+    nbit2 = 0
 
@@
     - Trap implications:
-              t
-         -----------
-    x => y \/ z \/ w
-    ^    ^    |    |
-    |    |    |    w \in S
+                s1
+             ----------
+    p1    => p2 \/ bit1
+    ^           ^      ^
     |           |      |
-    |    |    z \in S
+    |           |     bit1 \in S
     |           |
-    |    y \in S
+    |          p2 \in S
     |
-    x \in S
+    p1 \in S
 
-    y => y \/ z \/ w
-    z => true
-    w => true
+    p2    => p3 \/ nbit2
+    p3    => p1 \/ nbit1
+    bit1  => (p1 \/ nbit1) /\ (q3 \/ bit1)
+    nbit1 => (p2 \/ bit1) /\ (q1 \/ nbit1) /\ (q5 \/ nbit1)
+    q1    => q2
+    q2    => (q3 \/ bit1) /\ (q5 \/ nbit1)
+    q3    => q4 \/ nbit2
+    q4    => q1 \/ nbit1
+    q5    => q1 \/ nbit2
+    nbit2 => q2 /\ (p3 \/ nbit2)
 
     - At least one of the initially marked places belongs to S:
-    x
+    p1 \/ q1 \/ nbit1 \/ nbit2
 
     - None of the marked places in A belongs to S:
-    ~z /\ ~w
+    ~p3 /\ ~q5 /\ ~bit1
+
+  - Assignment A_\theta:
+    p1    = false
+    p2    = true
+    p3    = false
+    bit1  = false
+    nbit1 = true
+    q1    = false
+    q2    = true
+    q3    = true
+    q4    = false
+    q5    = false
+    nbit2 = true
+
+  - Refinement constraint \delta:
+    p2 + q2 + q3 + nbit1 + nbit2 >= 1
+    ^    ^    ^    ^       ^
+    |    |    |    |       |
+    ------------------------
+      |
+      S = {p2, q2, q3, nbit1, nbit2}, therefore \delta excludes
+      assignment A in the next iteration
 \end{verbatim}
@@
 \begin{verbatim}
-  - Assignment A_\theta:
-    x = true
-    y = true
-    z = false
-    w = false
-
-  - Refinement constraint \delta:
-    x + y >= 1
-    ^   ^
-    |   |
-    -----
-      |
-      S = {x, y}, therefore \delta excludes assignment A in the next iteration
 
@@
   "At least one place in S is always marked"
-
 \end{verbatim}
-
-\newpage
-
-\begin{verbatim}
-
-* Subprocedure TrapConditions:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C \cup { ~P }
-  Output:
-    C_\theta      : Trap conditions
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  C_\theta(S, T, E, M_0) :=& \left( \bigwedge_{s \in S} \left( s \Rightarrow
-      \bigwedge_{(s, t) \in E} \bigvee_{(t, p) \in E} p
-    \right) \right) \land \\
-  & \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land \\
-  & \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right) \\
-\end{align*}
-\begin{verbatim}
-
-* Subprocedure \Delta:
-
-  Input:
-    A_\theta      : Satisfying assignment for C_\theta
-  Output:
-    \delta        : Refinement constraint \delta
-
-  Pseudocode:
-
-\end{verbatim}
-\begin{align*}
-  & \delta(A_\theta) := \left( \sum_{A_\theta(s)} s \ge 0 \right)\\
-\end{align*}
-
-\iffalse
-
-\newpage
-\section{Method Invariant}
-
-The method Invariant constructs an invariant \verb=I= for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant
-
-  Subprocedure \mathcal{C'} constructs dual state constraints C' corresponding to N and P.
-  Subprocedure Model returns assignment A such that A satisfies C'.
-  Subprocedure Inv constructs invariant I corresponding to N and A.
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
-      $I := \text{Inv}(N, A)$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (satc) edge node[right]{YES} (inv);
-    \draw (noinv) edge (end1);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of dual state constraints C':  If C' is sat then N |= P.
-
-* Property of invariant I:
-  - I is reachable: For each reachable marking M, I(M) is valid
-  - I is safe:      For markings that violate the property, I(M) is unsat
-  - I is inductive: For each marking M, if I(M) is valid and M -> M1 then I(M1) is valid
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example:
-
-  - With weakened property P \/ (\neg \delta)
-
-  - Negated property is \neg P /\ \delta
-
-  - Dual state constraints C':
-    0 >= (- cs1) + id1 + idle1 + (- nid1)
-    0 >= (- id1) + (- idle1) + nid1 + req1
-    0 >= cs1 + (- req1)
-    0 >= after_you2 + (- req2)
-    0 >= (- after_you2) + await2 + id2
-    0 >= (- await2) + idle2
-    0 >= cs2 + (- req2)
-    0 >= (- cs2) + id2 + idle2
-    0 >= (- id2) + (- idle2) + req2
-
-    id1 + id2 + idle1 + idle2 < 2 * target_1 + trap_1
-
-    id1        >= trap_1
-    cs1        >= target_1
-    idle1      >= 0
-    req1       >= trap_1
-    nid1       >= 0
-    id2        >= trap_1
-    cs2        >= target_1
-    idle2      >= 0
-    req2       >= trap_1
-    after_you2 >= trap_1
-    await2     >= 0
-    target_1   >= 0
-    trap_1     >= 0
-
-  - Model A:
-
-    id1        = 1
-    cs1        = 1
-    idle1      = 0
-    req1       = 1
-    nid1       = 0
-    id2        = 1
-    cs2        = 1
-    idle2      = 0
-    req2       = 1
-    await2     = 0
-    after_you2 = 1
-    target_1   = 1
-    trap_1     = 1
-
-  - Invariant:
-
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
-
-  - Invariant with trap constraint \delta added:
-
-    id1 + id2 + req1 + req2 + after_you2             >= 1
-    id1 + id2 + req1 + req2 + after_you2 + cs1 + cs2 <= 2
-
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Derivation of invariant I for Petri net N that satisfies property P:
-
-  M                                    : marking          ~ places
-  C                                    : incidence matrix ~ rows correspond to places, columns correspond to transitions, relates places to transitions
-  X                                    : firing vector    ~ transitions
-
-  The following constraints C1 are unsat.
-  M = M0 + C*X                         : place equations
-  M >= 0                               : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-  AM >= b                              : property P negated
-
-  Substitute M to obtain constraints C2.
-  A(M0 + CX) >= b                      : property P negated
-  M0 + CX >= 0                         : non-negativity conditions for places
-  X >= 0                               : non-negativity conditions for transitions
-
-  Rewrite each system to obtain constraints C3.
-  (-A*C)      (A*M0-b)                 : property P negated
-  (  -C)*X <= (  M0  )                 : non-negativity conditions for places
-  (  -I)      (   0  )                 : non-negativity conditions for transitions
-
-  Apply Farkas' Lemma to obtain constraints C4.
-  yT*(-A*C)                            : 
-     (  -C) = 0                        : 
-     (  -I)                            :
-
-  yT*(A*M0-b)                          : 
-     (  M0  ) < 0                      : 
-     (   0  )                          :
-
-  y >= 0                               :
-
-  The constraints C4 are sat iff the following constraints C5 are sat.
-  y1 * A * C + y2 * C + y3 = 0         : 
-  y1 * (A*M - b) + y2 * M0 < 0         : 
-  y1 >= 0                              : 
-  y2 >= 0                              : 
-  y3 >= 0                              :
-
-  The constraints C5 are sat iff the following constraints C6 are sat.
-  (y1 * A + y2) * C  <= 0              : 
-  (y1 * A + y2) * M0 < y1 * b          : 
-  y1 >= 0                              : 
-  y2 >= 0                              :
-
-  The constraints C6 are sat iff the following constraints C' are sat.
-  \lambda * C  <= 0                    : inductivity constraint
-  \lambda * M0 < y1 * b                : safety constraint
-  \lambda >= y1 * A                    : property constraint
-  y1 >= 0                              : non-negativity constraint
-
-  For \lambda satisfying C' the invariant is the following.
-  I(M) = (\lambda * M <= \lambda * M0) : invariant
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Subprocedure \mathcal{C'}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    P             : Property
-  Output:
-    C'            : Dual state constraints
-
-Input:
- N = (S, T, E, M0)
-   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- P = ( cs1 + cs2 <= 1 )
-
-TODO: Unroll pseudocode?
-
-* Subprocedure Inv:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Satisfying assignment for C'
-
-  Output:
-    I             : Invariant
-
-TODO: Unroll pseudocode?
-\end{verbatim}
-
-\newpage
-\section{Method Invariant with Minimization}
-
-The method Invariant with Minimization constructs an invariant \verb=I=
-that uses a minimal number of places for given Petri net \verb=N= and
-\verb=P= when \verb=N= never violates \verb=P=. 
-We present the method Invariant with Minimization by example on a
-simplified mutual exclusion algorithm by Delzanno2001.
-
-\begin{verbatim}
-* Code, property, and Petri net: same as in section Method Safety.
-\end{verbatim}
-
-\begin{verbatim}
-* Method Invariant with Minimization
-
-  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A
-\end{verbatim}
-
-\begin{center}
-  \begin{tikzpicture}
-    \node[state] (begin) {BEGIN};
-    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
-    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
-    \node[print, right=of satc] (noinv) {No Invariant};
-    \node[action, below=of satc] (model1) {$A:=\text{Model}(C')$};
-    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A)$};
-    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
-    \node[action, right=of satcm] (model2) {$A:=\text{Model}(C' \cup C_M)$};
-    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A)$};
-    \node[print, below=of inv] (printinv) {Invariant: $I$};
-    \node[state, right=of noinv] (end1) {END};
-    \node[state, below=of printinv] (end2) {END};
-
-    \draw (begin) edge (c);
-    \draw (c) edge (satc);
-    \draw (satc) edge node[above]{NO} (noinv);
-    \draw (noinv) edge (end1);
-    \draw (satc) edge node[right]{YES} (model1);
-    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
-    \draw (cm) edge (satcm);
-    \draw (satcm) edge node[above]{YES} (model2);
-    \draw (model2) |- (edgein);
-    \draw (satcm) edge node[right]{NO} (inv);
-    \draw (inv) edge (printinv);
-    \draw (printinv) edge (end2);
-  \end{tikzpicture}
-\end{center}
-
-\begin{verbatim}
-* Property of minimization constraints C_M generated from A: 
-    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
-\end{verbatim}
-
-\newpage
-\begin{verbatim}
-* Example:
-  - Dual state constraints C' and satisfying assignment A for C' as in example for method Invariant
-
-  - Minimization constraints:
-
-    nc  ~= 0 => m_nc = 1
-    |           --------
-    |              |
-    |           place appears in invariant
-    |
-    place coefficient in the invariant
-
-    nc   = 0 => m_nc  = 0
-    |           ---------
-    |              |
-    |           place does not appear in invariant
-    |
-    place coefficient in the invariant
-
-    m   ~= 0 => m_m   = 1
-    m    = 0 => m_m   = 0
-    cs1 ~= 0 => m_cs1 = 1
-    cs1  = 0 => m_cs1 = 0
-    cs2 ~= 0 => m_cs2 = 1
-    cs2  = 0 => m_cs2 = 0
-
-    m_nc + m_m + m_cs1 + m_cs2 < 3
-    --------------------------   |
-                |                |
-                |    number of places appearing in current invariant for A
-                |    = #{s | A(s) ~= 0} = #{m, cs1, cs2}
-                |
-                |
-  number of places appearing in new invariant
-\end{verbatim}
-
-\begin{verbatim}
-* Subprocedure \text{MinConstraints}:
-
-  Input:
-    (S, T, E, M0) : Petri net
-    A             : Assignment
-  Output:
-    C_M           : Minimization Constraints
-
-Input:
- N   = (S, T, E, M0)
-     = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
- A   = ( nc=0, m=1, cs1=1, cs2=1, target1=1 )
-
-TODO: Unroll pseudocode?
-
-\end{verbatim}
-
-\newpage
-\section{Benchmarks of \bfc}
-
-We experimented on the benchmarks \ttt{cprover-PN} and
-\ttt{cprover\_software\_analysis} available at \\
-\ttt{http://www.mpi-sws.org/\~{}jkloos/iic-experiments/cprover.zip}.
-The set of benchmarks consists 77 benchmarks, 26 positive and 51 negative.
-Table~\ref{bfc-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{7cm} | r | r | r | r | } %
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 23 &  0 & 54 &  0 \\
-    Invariant \& Invariant with Minimization    & 23 &  0 & 54 &  0 \\
-    Safety by Refinement                        & 23 &  0 & 54 &  0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \bfc}
-\label{bfc-experiments}
-\end{table}
-
-\section{Benchmarks of \mist}
-
-We experimented on the benchmarks available at \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/boundedPN}
-and \\
-\ttt{https://github.com/pierreganty/mist/tree/master/examples/PN}.
-The set of benchmarks consists 27 benchmarks, 23 positive and 4 negative.
-Table~\ref{mist-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{6cm} | r | r | r | r | }
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 14 &  0 & 13 &  0 \\
-    Invariant \& Invariant with Minimization    & 14 &  0 & 13 &  0 \\
-    Safety by Refinement                        & 20 &  0 &  7 &  0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \mist}
-\label{mist-experiments}
-\end{table}
-
-\section{Benchmarks of \iic}
-
-We experimented on the benchmarks available at \\
-\ttt{https://svn.rse.mpi-sws.org/RSE/IC3forWSTS/code/trunk/examples/soter}.
-The set of benchmarks consists 38 benchmarks, 28 positive and 10 negative.
-Table~\ref{iic-experiments} shows the corresponding experimental results.
-
-\begin{table}[h]
-\begin{center}
-  \begin{tabular}{ | p{6cm} | r | r | r | r | }
-    \hline
-    Method & pos & neg & don't know & TO \\
-    \hline
-    Safety (over real numbers or naturals)      & 2 & 0 & 36 & 0 \\
-    Invariant \& Invariant with Minimization    & 2 & 0 & 26 & 0 \\
-    Safety by Refinement                        & 2 & 0 & 36 & 0 \\
-    \hline
-  \end{tabular}
-\end{center}
-\caption{Experimental results for the benchmarks of \iic}
-\label{iic-experiments}
-\end{table}
-
-We constructed the following two invariants by applying the method
-Invariant with Minimization.
-
-\begin{itemize}
-\item \verb?firewall__no_pred_called_with_zero__depth_1__fixed?: \verb?(assert (<= l739 0))?
-\item \verb?firewall__no_pred_called_with_zero__depth_2__fixed?: \verb?(assert (<= l3849 0))?
-\end{itemize}
-
-\iic\ constructs the following two invariants.
-
-\begin{itemize}
-\item \verb=firewall__no_pred_called_with_zero__depth_1__fixed.spec=: \verb=(l739 < 1)=
-\item \verb=firewall__no_pred_called_with_zero__depth_2__fixed.spec=: \verb=(l3849 < 1)=
-\end{itemize}
-
-\fi
-
-\end{document}

### 2014-02-05 paper/workspace/sec_constraint_approach.tex
Added third property

@@
 
+  - Property 3: Variable bit1 is either true or false
+
+  - Property 3: bit1 + nbit1 = 1, alternatively written as
+                bit1 + nbit1 <= 1 /\ bit1 + nbit1 >= 1
+
   - Petri net:
@@
 
-  - Using property 2
+  - Using property 1, with negated property \neg P:
 
-  - Negated property 2 \neg P:
+    p3 + q5 >= 2
+
+    The constraints are satisfiable, therefore we can not
+    prove safety with this method.
+
+  - Using property 2, with negated property \neg P:
 
     p1 + p2 + p3 >= 2
+
+    The constraints are unsatisfiable, therefore we
+    prove safety with this method.
+
+  - Using property 3, with negated property \neg P:
+
+    nbit1 + nbit1 <= 0 \/ nbit1 + nbit1 >= 2
+
+    The constraints are unsatisfiable, therefore we
+    prove safety with this method.
 \end{verbatim}
@@
 
-  - Using property 1
-
-  - Negated property 1 \neg P:
+  - Using property 1, with negated property 1 \neg P:
 

### 2014-02-05 paper/workspace/sec_preliminaries.tex
Updated preliminaries

@@
 
-A Petri net $P$ satisfies a property $P$ (denoted by $N \models P$)
+A Petri net $N$ satisfies a property $P$ (denoted by $N \models P$)
 iff for all reachable markings $M_0 \xrightarrow{\sigma} M$
@@
 
-\iffalse
-\begin{verbatim}
- Examples:
-  x + y < 0
-  x + y < 0 \\and x + z > 0
-  x + y < 0 \\or  x + z > 0
+An invariant $I$ of a Petri net $N$ is a property such that $N$ satisfies $I$.
+The invariant $I$ is inductive iff for all markings
+$M$, if $M \models I$ and $M \xrightarrow{t} M'$ for some $t$, then
+$M' \models I$.
 
- ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
-  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
-\end{verbatim}
-\fi
+A trap is a set of places $S \subseteq P$ such that $\post{S} \subseteq \pre{S}$.
+If a trap $S$ is marked in $M_0$, i.e. $\sum_{p \in S} M_0(p) > 0$, then it is also marked in all reachable markings.

### 2014-02-06 paper/example-latex-threads/POPL2011-proceedings/experiments.tex
Example

\section{Experimental results} \label{sec-experiments}

In this section, we describe a proof-of-concept implementation of our
proposed algorithm as an extension of the model checker ARMC 
\cite{RybalPodelskiPADL07}.

\paragraph{Tool description}
The verifier we built takes as input a number of functions (written in
the C language) representing threads that should execute concurrently.
The input file also contains the description of an initial state and
a number of assertions to be proven correct.
Our tool uses a frontend based on the CIL infrastructure
\cite{NeculaCC02} to translate a C program to its corresponding
multi-threaded transition system that is formalized in
Section~\ref{sec-prelim}.
The main component of our tool is an implementation of our algorithm
done using SICStus Prolog \cite{sicstus}.

An important design decision in our implementation concerns the 
treatment of control-location and data variables.
Even if both control-location variables and data variables can be
handled uniformly by our algorithm, we found that different
abstraction domains and refinement for the two domains can lead to
significant improvement.
In our implementation, the \algRefine procedure first splits the
constraints into data variable constraints and control-location
constraints. 
The splitting procedure preserves satisfiability/unsatisfiability of
the original constraint since there is no atomic formula in the
program transitions that relates both control variables and data
variables.
If the data constraints are satisfiable, the algorithm proceeds as in
Figure~\ref{fig-alg-refine}.
If the data constraints are unsatisfiable, our implementation relies
on a specialized refinement procedure (described in
\cite{GuptaATVA10}) that takes advantage of the simpler form of
control counterexamples.
For these counterexamples, control variables range over a finite
domain and no atomic formula from the program transitions involves
different control variables.

\input{table-vs-no-bias}

\paragraph{Benchmark programs}

We tested our prototype implementation using a collection of programs
that have intricate correctness proofs for their safety assertions.
The first four programs shown in Table~\ref{table-vs-no-bias} are
derived from two buggy examples highlighted as figures in
\cite{LuASPLOS08}, together with their fixes from the \Mozilla CVS
repository.
The property to verify is that two operations performed by different
threads are executed in the correct order.
The next three examples model the stopping procedure of a Windows NT
Bluetooth driver \cite{QadeerPLDI04}.
\BluetoothTwo contains two threads, one worker thread and another thread
to model the stopping procedure of the driver.
%\BluetoothThree includes an additional stopper thread and illustrates
%a bug in the driver model. 
%(The bug cannot be triggered for the two-thread model.)
\BluetoothTwoFixed and \BluetoothThreeFixed are the fixed versions of
the model with two and respectively three threads.
\Scull\cite{CorbetBook05} is a Linux character device driver that
implements access to a global memory area.
The property to verify is that read and write operations are performed
in critical section.

We also include some examples which are not particularly favorable to
a modular reasoning approach.
These examples are algorithms that establish mutual exclusion and mainly
deal with global variables (no local computation is included in the
critical region).
The mutual exclusion property of the na\"ive version of the Bakery
algorithm \cite{MannSafetyBook95} holds only when assuming
assignments are performed atomically. 
(Our verifier was able to confirm the bug present in the code without
such atomicity assumption.)
\Bakery\cite{LamportCACM74} is the complete version of the Bakery
algorithm, while \Lamport\cite{LamportTCS87} is an algorithm with an
optimized path in the absence of memory contention.
\QRCU\cite{McKenneyLWN07} is an algorithm implementing the
Read-copy-update synchronization algorithm.
It is an alternative to a readers-writer lock having wait-free read
operations.

\paragraph{Performance of our tool}

To explain our experimental results, we first articulate a working
hypothesis.
This hypothesis suggests that, when verifying a program that does not
have a modular proof, the algorithm with preference for modular
solutions (denoted as \emph{verification with bias}) is expected to
pay a penalty by insisting to search for modular solutions that do
not exist.
On the other hand, for cases where a modular proof does exist, the
non-biased verification could fail to find a modular proof and instead
return a more detailed non-modular proof.
Therefore, the hypothesis suggests that in these cases the biased
verification is expected to succeed faster compared to the non-biased
verification.

We report statistical data for each of the programs in
Table~\ref{table-vs-no-bias}.
We show the number of lines of code (LOC) and whether a modular proof
exists for a program (see Column 3).
Our implementation has two modes. 
Column 4 shows the verification results, when using our algorithm with
a preference for modular solutions.
The last column of the table shows the verification results for the 
non-biased implementation of our algorithm.
The results demonstrate that our approach to verification of
multi-threaded programs is feasible and that the constraint solving
procedure with bias is able to produce modular proofs more
often than the non-biased verification.
Furthermore, without the bias, the verification procedure times-out
for Scull and QRCU examples showing the benefits of modular proofs.

As another experiment, we tested some of our smaller examples
using two state-of-the-art model checkers for sequential C programs,
Blast \cite{BlastLazy02} and ARMC\cite{RybalchenkoARMC}.
For each of the tested programs (Fig2-fixed, Fig4-fixed, Dekker,
Peterson, and Lamport), we instrumented the program counter as
explicit program variables ($\pcone$ and $\pctwo$) and obtained a
sequential model of the multi-threaded examples.
Both Blast and ARMC eagerly consider all interleavings and obtained
timeouts after 30 minutes for both Fig4-fixed and Lamport.
Comparatively, our tool exploits the thread structure of these
programs and obtains conclusive verification results fast.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/POPL2011-proceedings/fig-ex-cex.tex
Example

\begin{figure*}[p]
  \small
  \centering
  \hrule
%-----------------------------------------------------------------------------
\begin{minipage}[t]{0.13\linewidth} 
  \vspace{12pt}
  \centering
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto, node distance=1.1cm,
                    semithick]

  \node[stateBox,node distance=4cm] (m1) {$m_1$};
  \node[stateBox, below of=m1] (m2) {$m_2$};
  \path (m1) edge node[left]{$\rel_{0}$} (m2);
  \node[stateBox, below of=m2] (m3) {$m_3$};
  \path (m2) edge node[left]{$\rel_1$} (m3);
  \node[stateBox, errorState, below of=m3] (m4) {$m_4$};
  \path (m3) edge node[left]{$e_2$} (m4);

  \node[stateBox,right of=m2, node distance=1.2cm] (n1) {$n_1$};
  \node[stateBox, below of=n1] (n2) {$n_2$};
  \path (n1) edge node[right]{$\rel_2$} (n2);
  \node[stateBox,errorState, below of=n2] (n3) {$n_3$};
  \path (n2) edge node[right]{$e_1$} (n3);
  \path (2mm, -16mm) edge[envTransition] (10mm,-29mm);
  \path (10mm,-16mm) edge[envTransition]( 2mm,-29mm);
  \end{tikzpicture}~\vspace{12pt}\\
\textbf{(a)}
\end{minipage}
%---------------------------------------------------------------------------
\begin{minipage}[t]{0.38\linewidth}
\centering
\begin{equation*}
  \begin{array}[t]{ll}
  \SymbInit \limplies \unkOf{m_1}(\Vars)           \\
  \unkOf{m_1}(\Vars) \land  \rel_{0} \limplies \unkOf{m_2}(\Vars')  \\
  \unkOf{m_2}(\Vars) \land  \rel_1 \limplies \unkOf{m_3}(\Vars')  \\
  \unkOf{m_2}(\Vars) \land  \rel_1 \limplies
  \unkOf{e_1}(\Vars,\Vars')  \\
  \unkOf{m_3}(\Vars) \land \unkOf{e_2}(\Vars,\Vars') \land \relLocEqOf{1} \limplies\unkOf{m_4}(\Vars')
  \vspace{10pt}\\
  \SymbInit \limplies \unkOf{n_1}(\Vars) \\
  \unkOf{n_1}(\Vars) \land \rel_{2} \limplies \unkOf{n_2}(\Vars') \\
  \unkOf{n_1}(\Vars) \land \rel_{2} \limplies \unkOf{e_2}(\Vars,\Vars') \\
  \unkOf{n_2}(\Vars) \land \unkOf{e_1}(\Vars,\Vars')  \land  \relLocEqOf{2} \limplies \unkOf{n_3}(\Vars')
  \vspace{10pt}\\
  \unkOf{m_4}(\Vars) \land \unkOf{n_3}(\Vars) \land  \SymbError \limplies \lfalse 
\end{array}
\end{equation*}
\textbf{(b)}
\end{minipage}
\hfill
%----------------------------------------------------------------------------
\begin{minipage}[t]{0.46\linewidth}
\centering
\begin{equation*}
  \begin{array}[t]{ll}
  \SymbInit \limplies \unkOf{m_1}(\GlobVars,\LocVarsOf{1})           \\
  \unkOf{m_1}(\GlobVars,\LocVarsOf{1}) \land  \rel_{0} \limplies \unkOf{m_2}(\GlobVars',\LocVarsOf{1}')  \\
  \unkOf{m_2}(\GlobVars,\LocVarsOf{1}) \land  \rel_1 \limplies \unkOf{m_3}(\GlobVars',\LocVarsOf{1}')  \\
  \unkOf{m_2}(\GlobVars,\LocVarsOf{1}) \land  \rel_1 \limplies
  \unkOf{e_1}(\GlobVars,\GlobVars')  \\
  \unkOf{m_3}(\GlobVars,\LocVarsOf{1}) \land \unkOf{e_2}(\GlobVars,\GlobVars')
  \land \relLocEqOf{1} \limplies \unkOf{m_4}(\GlobVars',\LocVarsOf{1}') 
  \vspace{10pt}\\
  \SymbInit \limplies \unkOf{n_1}(\GlobVars,\LocVarsOf{2}) \\
  \unkOf{n_1}(\GlobVars,\LocVarsOf{2}) \land \rel_{2} \limplies \unkOf{n_2}(\GlobVars',\LocVarsOf{2}') \\
  \unkOf{n_1}(\GlobVars,\LocVarsOf{2}) \land \rel_{2} \limplies \unkOf{e_2}(\GlobVars,\GlobVars') \\
  \unkOf{n_2}(\GlobVars,\LocVarsOf{2}) \land \unkOf{e_1}(\GlobVars,\GlobVars')  \land\relLocEqOf{2} \limplies 
  \unkOf{n_3}(\GlobVars',\LocVarsOf{2}')\vspace{10pt}\\

  \unkOf{m_4}(\GlobVars,\LocVarsOf{1})\land\unkOf{n_3}(\GlobVars,\LocVarsOf{2})\land\SymbError\limplies\lfalse 
\end{array}
\end{equation*}
\textbf{(c)}
\end{minipage}
\hfill 
\vspace{1ex}
%----------------------------------------------------------------------------------------------
\hrule
\begin{minipage}[t]{\linewidth} 
\centering
\vspace{1pt}
  \begin{tikzpicture}[<-,>=stealth',shorten >=1pt,auto, node distance=9mm,
                    semithick]
 \node[stateBox] (err) {$1:\lfalse$};

%----------------------------
 \node [stateBox,xshift=-83mm,yshift=2mm,above left of=err] (m47) {$2:\unkOf{m_4}(\Vars^1)$};\path (m47) edge (err);
 \node [stateBox,xshift=-34mm,yshift=3mm, above left of=m47] (m36) {$3:\unkOf{m_3}(\Vars^2)$};\path (m36) edge (m47);
 \node [stateBox, above of=m36 ] (m25) {$4:\unkOf{m_2}(\Vars^3)$}; \path (m25) edge (m36);
 \node [stateBox, above of=m25 ] (m14) {$5:\unkOf{m_1}(\Vars^4)$}; \path (m14) edge (m25);

 \node [draw,xshift=13.5mm,yshift=3mm,above right of=m36] (t1) {$8:\rel_1(\Vars^3,\Vars^2)$};\path (t1) edge (m36);
 \node [draw,xshift=13.5mm,yshift=3mm,above right of=m25] (te1) {$7:\rel_{0}(\Vars^4,\Vars^3)$};\path (te1)edge(m25);
 \node [draw, above of=m14]  (i1)  {$6:\SymbInit(\Vars^4)$};    \path (i1) edge (m14);

 \node [stateBox,above of=m47] (e2) {$9:\unkOf{e_2}(\Vars^2,\Vars^1)$}; \path (e2) edge (m47);
 \node [stateBox, above of=e2 ] (n16) {$10:\unkOf{n_1}(\Vars^2)$}; \path (n16) edge (e2);

 \node [draw,xshift=17mm,yshift=3mm,above right of=e2] (et2) {$12:\rel_2(\Vars^2,\Vars^1)$}; \path (et2) edge (e2);
 \node [draw, above of=n16] (ei2){$11:\SymbInit(\Vars^2)$};  \path (ei2) edge (n16);

 \node [draw,xshift=17mm,yshift=3mm, above right of=m47,text centered] (eq2) 
 {$13:\relLocEqOf{1}(\Vars^2,\Vars^1)$}; \path (eq2) edge (m47);

%---------------------------

 \node [stateBox,above of=err] (n37) {$14:\unkOf{n_3}(\Vars^1)$};\path (n37) edge (err);
 \node [stateBox,xshift=-36mm,yshift=3mm,above left of=n37 ] (n23) {$15:\unkOf{n_2}(\Vars^5)$};\path (n23)edge(n37);
 \node [stateBox,     above of=n23 ] (n12) {$16:\unkOf{n_1}(\Vars^6)$}; \path (n12) edge (n23);

 \node [draw,xshift=14.5mm,yshift=2.5mm,above right of=n23](t2){$18:\rel_2(\Vars^6,\Vars^5)$}; \path (t2) edge (n23);
 \node [draw, above of=n12] (i2)  {$17:\SymbInit(\Vars^6)$}; \path (i2) edge (n12);

 \node [stateBox,above of=n37] (e1)  {$19:\unkOf{e_1}(\Vars^5,\Vars^1)$};\path (e1) edge (n37);
 \node [stateBox, above of=e1 ] (m23) {$20:\unkOf{m_2}(\Vars^5)$}; \path (m23) edge (e1);
 \node [stateBox, above of=m23] (m11) {$21:\unkOf{m_1}(\Vars^7)$}; \path (m11) edge (m23);

 \node[draw,xshift=17mm,yshift=3mm,above right of=e1] (et1) {$24:\rel_1(\Vars^5,\Vars^1)$};\path (et1) edge (e1);
 \node[draw,xshift=17mm,yshift=3mm,above right of=m23] (ete1) {$23:\rel_0(\Vars^7,\Vars^5)$};\path (ete1) edge (m23);
 \node[draw, above of=m11]  (ei1)  {$22:\SymbInit(\Vars^7)$};\path (ei1) edge (m11);

 \node [draw,xshift=17mm,yshift=3mm,above right of=n37,text centered ] (eq1) 
 {$25:\relLocEqOf{2}(\Vars^5,\Vars^1)$}; \path (eq1) edge (n37);

%---------------------------
  \node [draw, xshift=22mm, yshift=-1mm, above of=err] (errn) {$26:\SymbError(\Vars^1)$};\path (errn) edge (err);
  \end{tikzpicture}\vspace{-10pt}\\
\textbf{(d)}
\end{minipage}
\hrule
%-------------------------------------------------------------------------------
% Pred map
\begin{minipage}[t]{\linewidth} 
\centering
\vspace{1pt}
  \begin{tikzpicture}[<-,>=stealth',shorten >=1pt,auto, node distance=9mm,
                    semithick]
 \node[stateBox] (err) {$1:\linfalse$};

%----------------------------
 \node [stateBox,xshift=-83mm,yshift=2mm,above left of=err] (m47) {$2:\linfalse$};\path (m47) edge (err);
 \node [stateBox,xshift=-34mm,yshift=3mm, above left of=m47] (m36) {$3:1\leq\local^2$};\path (m36) edge (m47);
 \node [stateBox, above of=m36 ] (m25) {$4:1\leq\local^3$}; \path (m25) edge (m36);
 \node [stateBox, above of=m25 ] (m14) {$5:\lintrue$}; \path (m14) edge (m25);

 \node [draw,xshift=13.5mm,yshift=3mm,above right of=m36] (t1) {$8:$$\local^3\leq\local^2$};\path (t1) edge (m36);
 \node [draw,xshift=13.5mm,yshift=3mm,above right of=m25] (te1) {$7:1\leq\local^3$};\path (te1)edge(m25);
 \node [draw, above of=m14]  (i1)  {$6:\lintrue$};    \path (i1) edge (m14);

 \node [stateBox,above of=m47] (e2) {$9:\local^2\leq0$}; \path (e2) edge (m47);
 \node [stateBox, above of=e2 ] (n16) {$10:\local^2\leq0$}; \path (n16) edge (e2);

 \node [draw,xshift=16mm,yshift=3mm,above right of=e2] (et2) {$12:\lintrue$}; \path (et2) edge (e2);
 \node [draw, above of=n16] (ei2){$11:\local^2\leq0$};  \path (ei2) edge (n16);

 \node [draw, xshift=16mm,yshift=2mm, above right of=m47,text centered] (eq2) {$13:\lintrue$}; \path (eq2) edge (m47);

%---------------------------
 \node [stateBox,above of=err] (n37) {$14:\lintrue$};\path (n37) edge (err);
 \node [stateBox,xshift=-36mm,yshift=3mm,above left of=n37 ] (n23) {$15:\lintrue$};\path (n23)edge(n37);
 \node [stateBox,     above of=n23 ] (n12) {$16:\lintrue$}; \path (n12) edge (n23);

 \node [draw,xshift=14.5mm,yshift=2.5mm,above right of=n23](t2){$18:\lintrue$}; \path (t2) edge (n23);
 \node [draw, above of=n12] (i2)  {$17:\lintrue$}; \path (i2) edge (n12);

 \node [stateBox,above of=n37] (e1)  {$19:\lintrue$};\path (e1) edge (n37);
 \node [stateBox, above of=e1 ] (m23) {$20:\lintrue$}; \path (m23) edge (e1);
 \node [stateBox, above of=m23] (m11) {$21:\lintrue$}; \path (m11) edge (m23);

 \node[draw,xshift=18mm,yshift=3mm,above right of=e1] (et1) {$24:\lintrue$};\path (et1) edge (e1);
 \node[draw,xshift=18mm,yshift=3mm,above right of=m23] (ete1) {$23:\lintrue$};\path (ete1) edge (m23);
 \node[draw, above of=m11]  (ei1)  {$22:\lintrue$};\path (ei1) edge (m11);

 \node [draw,xshift=19mm,yshift=2.5mm,above right of=n37,text centered ] (eq1) 
 {$25:\lintrue$}; \path (eq1) edge (n37);

%---------------------------
  \node [draw, xshift=22mm, yshift=-1mm, above of=err] (errn) {$26:\lintrue$};\path (errn) edge (err);
  \end{tikzpicture}\vspace{-10pt}\\
\textbf{(e)}
\end{minipage}
%-------------------------------------------------------------------------------
% 
\hrule
\begin{minipage}[t]{\linewidth} 
\centering
\vspace{1pt}
  \begin{tikzpicture}[<-,>=stealth',shorten >=1pt,auto, node distance=9mm,
                    semithick]
 \node [stateBox] (m47) {$2:\unkOf{m_4}(\GlobVars^1,\LocVarsOf{1}^1)$};
 \node [stateBox,xshift=-70mm,yshift=3mm, above left of=m47] (m36) {$3:\unkOf{m_3}(\GlobVars^2,\LocVarsOf{1}^2)$};
 \path (m36) edge (m47);
 \node [stateBox, above of=m36 ] (m25) {$4:\unkOf{m_2}(\GlobVars^3,\LocVarsOf{1}^3)$}; \path (m25) edge (m36);
 \node [stateBox, above of=m25 ] (m14) {$5:\unkOf{m_1}(\GlobVars^4,\LocVarsOf{1}^4)$}; \path (m14) edge (m25);

 \node [draw,xshift=29mm,yshift=2mm,above right of=m36] (t1)
 {$8:\rel_1(\GlobVars^3,\LocVarsOf{1}^3,\LocVarsOf{2}^{\textsc{IV}},
            \GlobVars^2,\LocVarsOf{1}^2,\LocVarsOf{2}^{\textsc{V}})$};
 \path (t1) edge (m36);
 \node [draw,xshift=29mm,yshift=2mm,above right of=m25] (te1) 
 {$7:\rel_{0}(\GlobVars^4,\LocVarsOf{1}^4,\LocVarsOf{2}^{\textsc{II}},
              \GlobVars^3,\LocVarsOf{1}^3,\LocVarsOf{2}^{\textsc{III}})$};
 \path (te1) edge (m25);
 \node [draw, above of=m14] (i1)
 {$6:\SymbInit(\GlobVars^4,\LocVarsOf{1}^4,\LocVarsOf{2}^{\textsc{I}})$};
 \path (i1) edge (m14);

 \node [stateBox,above of=m47] (e2) {$9:\unkOf{e_2}(\GlobVars^2,\GlobVars^1)$}; \path (e2) edge (m47);
 \node [stateBox, above of=e2 ] (n16) {$10:\unkOf{n_1}(\GlobVars^2,\LocVarsOf{2}^2)$}; \path (n16) edge (e2);

 \node [draw,xshift=30mm,yshift=2mm,above right of=e2] (et2) 
 {$12:\rel_2(\GlobVars^2,\LocVarsOf{1}^{\textsc{VI}},\LocVarsOf{2}^2,
             \GlobVars^2,\LocVarsOf{1}^{\textsc{VII}},\LocVarsOf{2}^1)$};
 \path (et2) edge (e2);
 \node [draw, above of=n16] (ei2)
 {$11:\SymbInit(\GlobVars^2,\LocVarsOf{1}^{\textsc{VIII}},\LocVarsOf{2}^2)$};
 \path (ei2) edge (n16);

 \node [draw, xshift=18mm,yshift=2mm, above right of=m47,text centered] (eq2) 
 {$13:\dots$}; \path (eq2) edge (m47);
%---------------
  \node [xshift=20mm, yshift=5mm, below right of=m47] (errt) {};\path (m47) edge (errt);
  \node [xshift=20mm, yshift=5mm, below right of=errt] (err) {};\path (errt) edge[-,dashed] (err);
  \end{tikzpicture}\vspace{-7pt}\\
\textbf{(f)}
\end{minipage}
\caption{
(a) Reachability trees constructed by \aret computation.
(b) Corresponding Horn clauses generated using \algMkHornClauses.
(c) Horn clauses generated with preference for modular solutions.
(d) A tree representation of the clauses from (b) as generated by \algMkHornTree.
Each node shows its $\cFormula$ attribute.
The superscript of the node name identifies the set of
variables appearing in the attribute of the node.
(e) The $\cInterPol$ map generated by \algSolveLinearHornClauses from the clauses in (b).
(f) The $\cFormula$ map generated by \algMkHornClauses from the clauses in (c).
}
\label{fig-ex-cex}
\end{figure*}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/POPL2011-proceedings/proof-rule-symb.tex
Example

\section{Proof rule, fixpoints, and abstraction}\label{sec-proof-rule}
\input{fig-eq-abst-mod-reach}

In this section we develop the foundations for our verification
algorithm.
We present a compositional proof rule and then derive a corresponding
characterization in terms of least fixpoints and their approximations.
We present the ability of our proof rule to facilitate modular
reasoning, when admitted by the program, without losing the ability
for global reasoning otherwise.

\paragraph{Proof rule}

Figure~\ref{fig-eq-abst-mod-reach} presents a proof rule \ProofRule
for compositional verification of program safety.
The proof rule is inspired by the existing proof rules for
compositional safety reasoning, see
e.g.~\cite{OwickiAI76,JonesTOPLAS83,CohenFMSD09,HenzingerPLDI04}.
Our formulation of \ProofRule directly leads to a pre-fixpoint
characterization, thus, providing a basis for the proof rule
automation using abstraction and refinement techniques.

\ProofRule relies on \emph{thread reachability} assertions $\Reach_1,
\dots, \Reach_\Num$ that keep track of program states reached
by threads $1, \dots, \Num\ $ together with their respective
\emph{environment transitions} $\Env_1, \dots, \Env_\Num\ $.
The environment transition of each thread keeps track of modifications
of program states by other threads.
The auxiliary assertions used in our proof rule can refer to all
program variables, that is, they are not restricted to a combination
of global variables and local variables of a particular thread.

If the provided auxiliary assertions satisfy all premises of the proof
rule, i.e., \ruleInit, \dots, \ruleCheck, then the program is safe.
The premise \ruleInit requires that each thread reachability
over-approximates the initial program states.
\ruleTransStep ensures that the thread reachability of each thread is
invariant under the application of the thread transitions.
In addition, \ruleEnvStep requires invariance under the environment
transitions of the thread. 
The conjunct $\relLocEqOf{i}$ in \ruleEnvStep selects the subset of
the environment transition that does not modify the local variables of
the thread.
Given a thread $j\ $, the premise \ruleTransEnv collects transitions
that start from states in the thread reachability of other threads and
combines them into the environment transition for~$j\ $.
Finally, \ruleCheck checks that there is no error state that appears
in each thread reachability set.

The proof rule \ProofRule can be directly used to prove program safety
following a two-step procedure. 
First, we need to identify candidate assertions for the thread
reachability and environment transitions.
Second, we need to check that these candidate assertions satisfy the
premises of proof rule.
The correctness of the conclusion is formalized by the following
theorems.

%
\begin{theorem}[Soundness]
  \label{thm-rule-sound}
  % 
The proof rule \ProofRule is sound.\,\eofClaim
%
\end{theorem}
\includeProof{
\begin{proof}
%  See Appendix~\ref{app-proofs}.
% 
Let $\Reach_1, \dots, \Reach_\Num$ and $\Env_1, \dots, \Env_\Num$
satisfy the premises \ruleInit, \dots, \ruleCheck.
We show that the program is safe.
To prove safety, for each reachable state $s\models\SymbReach$ we
prove that $s\models\Reach_1\land\dots\land\Reach_\Num$ by induction
over the length $k$ of a shortest computation segment $s_1, \dots,
s_k$ such that $s_1\models \SymbInit$ and~$s_k=s\ $.

For the base case $k=1$, the inclusion holds due to the premise
\ruleInit.
For the induction step, we assume that the above statement holds for
states reachable in $k \geq 1$ steps and prove the statement for their
immediate successors.
That is, let $s_k\models\SymbReach$ and hence $s_k \models
\Reach_1\land\dots\land\Reach_\Num\ $.
If $s_k$ does not have any successor, i.e., $\neg (\exists s_{k+1} :
(s_k, s_{k+1}) \models \relProg)\ $, then there are no more states to
consider.
Otherwise, we choose a successor state $s_{k+1}$ of $s_k$ that is
reached by taking a transition in a thread~$i\ $, i.e., $(s_k,
s_{k+1})\models\relTIdOf{i}\ $.
From \ruleTransStep follows that~$s_{k+1}\models\Reach_i\ $.

To show that $s_{k+1}\models\Reach_j$ for each
$j\in\ThreadIds\setminus\set{i}$ we rely on the premises \ruleTransEnv
and \ruleEnvStep.
By induction hypothesis $s_k\models\Reach_i$ and due to \ruleTransEnv
we have $(s_k, s_{k+1})\models\Env_j\ $.
Now $s_{k+1}\in\Reach_j$ follows from \ruleEnvStep.
\end{proof}
}
%
\begin{theorem}[Relative completeness]\label{thm-rule-complete}
%
The proof rule \ProofRule is complete relative to first-order
reasoning.\eofClaim
%
\end{theorem}
%
\includeProof{
\begin{proof} 
Let $\program$ be safe.
We define $\Reach_1 = \dots = \Reach_\Num = \SymbReach$ and 
$\Env_1 = \dots = \Env_\Num = \SymbReach \land \relProg\ $.
Then, the premises \ruleInit, \dots, \ruleCheck are immediately
satisfied.
\end{proof}
}

\paragraph{Modular and  global proofs}

Reasoning about multi-threaded program is more complex than reasoning
about sequential programs since thread interaction needs to be taken
into account.
Some programs admit modular reasoning that deals with each thread in
isolation, i.e., assertions used in the proof only refer to the global
variables and the local variables of one thread at a time.

The proof rule \ProofRule facilitates \emph{modular} reasoning about
multi-threaded programs. 
If a program has a modular safety proof, then the following
\emph{modular} assertions satisfy the proof rule premises:
%
\begin{equation*}
  \begin{array}[t]{l@{\defeq}l@{\qquad}c@{}}
    \Reach_i & \exists \Vars\setminus (\GlobVars \cup \LocVarsOf{i}) :
    \SymbReach\ , & \text{for }
    i \in \ThreadIds\\[\jot]
    \Env_i & \exists (\Vars\cup \Vars')\setminus (\GlobVars \cup
    \GlobVars') : \SymbReach\land \relProg\ . & \text{for }
    i  \in \ThreadIds
  \end{array}
\end{equation*}
%

\ProofRule is \emph{not} restricted to modular proofs. 
Since the assertions used in \ProofRule can refer to each of the program
variables, non-modular proofs can be directly used.
In fact, the proof of Theorem~\ref{thm-rule-complete} relies on
non-modular assertions, since $\SymbReach$ may refer to local
variables of different threads.

In Section~\ref{sec-algo-safety} we will present our algorithm that
can discover modular assertions for \ProofRule if the program admits
modular proofs, and delivers non-modular assertions otherwise.

\paragraph{Fixpoints}

The proof rule \ProofRule in Figure~\ref{fig-eq-abst-mod-reach}
directly leads to a fixpoint-based characterization, which defines our
algorithm in Section~\ref{sec-algo-safety}.

From the premises \ruleTransStep, \ruleEnvStep, and \ruleTransEnv we
obtain a function $\eqPreFunc$ on $\Num$-tuples of assertions over the
program variables and $\Num$-tuples of assertions over the unprimed
and primed program variables such that
%
\begin{equation}\label{eq-pre-func}
  \begin{split}
    & \eqPreFunc(S_1, \dots, S_\Num, T_1, \dots, T_\Num) \defeq \\[\jot]
    &\;\,\,  (
    \begin{array}[t]{@{}l@{\; \dots,\; }l}
      \post(\relTIdOf{1} \lor T_1\land \rho^=_1, S_1), 
      &
      \post(\relTIdOf{\Num} \lor T_\Num\land \rho^=_\Num, S_\Num),\\[2\jot]
      \Lor_{i\in\ThreadIds\setminus\set{1}}S_i \land \relTIdOf{i},
      &
      \Lor_{i\in\ThreadIds\setminus\set{\Num}}S_i \land \relTIdOf{i} )\ .
    \end{array}
\end{split}
\end{equation} 
%
We formalize the relation between $\eqPreFunc$ and \ProofRule as
follows.
%
\begin{lemma}\label{lem-rule-fix}
  %
      Each pre-fixpoint of $\eqPreFunc$ satisfies the premises
\emph{\ruleTransStep,} \emph{\ruleEnvStep,} and \emph{\ruleTransEnv}
of \ProofRule.
That is, if
%
\begin{equation*}
  \eqPreFunc(\Reach_1, \dots, \Reach_\Num, \Env_1, \dots, \Env_\Num) \limplies
  (\Reach_1, \dots, \Reach_\Num, \Env_1, \dots, \Env_\Num)
\end{equation*}
%
then $\Reach_1, \dots, \Reach_\Num, \Env_1, \dots, \Env_\Num$
satisfies \emph{\ruleTransStep,} \emph{\ruleEnvStep,} and
\emph{\ruleTransEnv}.\eofClaim
%
\end{lemma}
%
We define a distinguished tuple $\eqPreBot$:
%
\begin{equation}\label{eq-pre-bot}
  \eqPreBot \defeq (\SymbInit,\dots,\SymbInit, \underbrace{\lfalse, \dots,
    \lfalse}_{\Num \text{ times}})\ .
\end{equation}
%
Then, each pre-fixpoint of $\eqPreFunc$ that is greater than
$\eqPreBot$ satisfies the premise \ruleInit.
By choosing a pre-fixpoint $(\Reach_1, \dots, \Reach_\Num, \Env_1,
\dots, \Env_\Num)$ above $\eqPreBot$ such that $\Reach_1 \land \dots,
\Reach_\Num \land \SymbError \limplies \lfalse$ we will satisfy all
premises of the proof rule \ProofRule, and hence prove the program
safety.

\paragraph{Fixpoint abstraction}

Computing pre-fixpoints of $\eqPreFunc$ that satisfy \ruleInit and
\ruleCheck is a difficult task.
We automate this computation using the framework of abstract
interpretation~\cite{lattice77}, which uses over-approximation to
strike a balance between reasoning precision and efficiency.
To implement required over-approximation functions, we will use a
collection of abstraction functions $\abstTIdOf{i}$ and
$\abstTransTIdOf{i}{j}$, where $i \neq j \in\ThreadIds$, that
over-approximate sets and binary relations over programs states,
respectively.

We define a function $\eqFunc$ that over-approximates $\eqPreFunc$
using given abstraction functions:
%
\begin{equation}\label{eq-func}
  \begin{split}
    & \eqFunc(S_1, \dots, S_\Num, T_1, \dots, T_\Num) \defeq \\[\jot]
    &\qquad  (
    \begin{array}[t]{@{}l}
      \abst_1(\post(\relTIdOf{1}, S_1)) \lor 
      \abst_1(\post(T_1 \land \relLocEqOf{1}, S_1))\ , \\[\jot]
      \qquad\qquad\cdots\\
      \abst_\Num(\post(\relTIdOf{\Num}, S_\Num) \lor 
      \abst_\Num(\post(T_\Num \land \relLocEqOf{\Num}, S_\Num))\ ,
    \\[3\jot]
    \Lor_{i\in\ThreadIds\setminus\set{1}}\abstTransTIdOf{i}{1}(S_i \land \relTIdOf{i})\ ,\\[2\jot]
    \qquad\qquad\cdots \\[\jot]
    \Lor_{i\in\ThreadIds\setminus\set{\Num}}\abstTransTIdOf{i}{\Num}(S_i \land \relTIdOf{i} ))\ .
  \end{array}
\end{split}
\end{equation} 
%
Let $\eqBot$ be an over-approximation of $\eqPreFunc$ such that
%
\begin{equation}\label{eq-bot}
  \eqBot \defeq (\abst_1(\SymbInit),\dots,\abst_\Num(\SymbInit), \underbrace{\lfalse, \dots,
    \lfalse}_{\Num \text{ times}})\ .
\end{equation}
%
The least pre-fixpoint of $\eqFunc$ above $\eqBot$ can be used to
prove program safety by applying the following theorem, and is the key
outcome of our algorithm in Section~\ref{sec-algo-safety}.
%
\begin{theorem}[Abstract fixpoint checking]\label{thm-abst-fix}
  % 
    If the least pre-fixpoint of $\eqFunc$ above $\eqBot$, say
$(\Reach_1, \dots, \Reach_\Num, \Env_1, \dots, \Env_\Num)$, satisfies
the premise \emph{\ruleCheck} then the program is safe.\eofClaim
  %
\end{theorem}
%
\includeProof{
\begin{proof} The theorem follows directly from the soundness of the
proof rule \ProofRule, Lemma~\ref{lem-rule-fix}, and
over-approximations introduced by the applied abstraction functions.
\end{proof}
}

The choice of the abstract domains, i.e., the range sets of the
abstraction functions, determines if the least fixpoint of $\eqFunc$
yields a modular proof.
Our abstraction discovery algorithm in Section~\ref{sec-refinement}
automatically chooses the abstraction such that modular proofs are
preferred.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/POPL2011-proceedings/refinement.tex
Example

\section{Abstraction refinement}
\label{sec-refinement}

\input{fig-alg-refine}
\input{fig-alg-mkHornClauses}
\input{fig-alg-mkTMHornClauses}
\input{fig-ex-cex-prog}

\paragraph{Procedure \algRefine}
In Figure~\ref{fig-alg-refine}, we present the procedure $\algRefine$
that takes as argument an error tuple and, if possible,
refines the abstraction functions to include predicates that witness
the fact that the error state is unreachable.
The procedure $\algRefine$ generates a set of Horn clauses
corresponding to the error tuple (lines~1--4).
Next, the \algRefine~algorithm invokes a solving procedure for Horn
clauses (lines~5--7).
Lastly, the procedure \algRefine updates the abstraction functions
using the solution of Horn clauses at lines 8--12.
We consider the solution $\Solution$ and add the atomic predicates
that appear in Sol(\unkOf{S}(\Vars) to the set of predicates
$\artPreds_i\ $. 
The index $i$ is chosen to be that of the thread where $S$ originated
from.
Similarly, the procedure updates the transition abstraction functions
at line~12.
Here, we only assume that \algSolveHornClauses returns a correct
solution to the set of Horn clauses received as argument.
In Section~\ref{sec-horn-solving}, we present a solving algorithm for
recursion-free Horn clauses over the linear arithmetic domain.

\paragraph{Function \algMkHornClauses} 
The generation of the Horn clauses is started from lines~1--4 of
Figure~\ref{fig-alg-refine}.
One clause requires that the solutions corresponding to the abstract
states from the error tuple do not intersect $\SymbError\ $:
$\set{\unkOf{S_1}(\Vars) \land \dots \land \unkOf{S_{\Num}}(\Vars)
  \land \SymbError \rightarrow \lfalse}\ $.
The other clauses are generated by invoking $\algMkHornClauses(S_i)$
for $i\in\ThreadIds\ $.
The function $\algMkHornClauses$ generates Horn clauses for
transitions considered during \aret computation as follows.
If the abstract state $\StateDst$ was produced by following a local
transition, i.e., $\artParent(\StateDst)=(\StateSrc,\rel)$ and
$\artParent(\rel)=\bot\ $, then one Horn clause corresponds to the
application of the local transition at line~12:
$\set{\unkOf{\StateSrc}(\Vars)\land \rel \rightarrow
  \unkOf{\StateDst}(\Vars')}\ $.
Additional Horn clauses are generated recursively for the parent
state $\StateSrc$ at line~5.
If the abstract state $\StateDst$ was produced by following an
environment transition, i.e., $\artParent(\StateDst)=(\StateSrc,\rel)$
and $\artParent(\rel)=(\StateOrig, \OrigRel)\ $, then two Horn clauses 
correspond to the generation of the environment transition (line~9)
and to the application of the environment transition (line~10).
Finally, if $\artParent(\StateDst)=\bot\ $, then one Horn clause 
constrains the solution of $\unkOf{\StateDst}(\Vars)$ at line~14:
$\set{\SymbInit \rightarrow\unkOf{\StateDst}(\Vars)}\ $.
Note that solutions for unknown states are expressed in terms of all
program variables $\Vars$, while solutions for unknown transitions are
expressed in terms of $\Vars$ and $\Vars'\ $.
Consequently, these solutions may lead to non-modular proofs even for
a set of Horn clauses that has modular solutions.

\paragraph{Discovery of  Modular Predicates}
We present modifications to our abstraction refinement algorithm that
guarantee the discovery of modular solutions whenever they exist.
With these modifications, solutions for unknown states originating in
thread $i$ can only be expressed in terms of $\GlobVars, \LocVarsOf{i}$
rather than the whole set of program variables $\Vars\ $.
Solutions for unknown transitions are restricted to the set of global
variables $\GlobVars, \GlobVars'\ $.
To implement these changes, we change line 1 from the \algRefine
procedure as follows:
%
\begin{center}
  \begin{minipage}[t]{.04\columnwidth}
    1
  \end{minipage}
  \begin{minipage}[t]{.94\columnwidth}
    \tabT $\HornClauses$ \algAssgn $\set{\unkOf{S_1}(\GlobVars, \LocVarsOf{1})\land \dots
    \land \unkOf{S_\Num}(\GlobVars, \LocVarsOf{\Num}) \land \SymbError}$
  \end{minipage}
\end{center}
% 
We also replace lines~9--14 from \algMkHornClauses with the
fragment shown in Figure~\ref{fig-alg-TMMkHornClauses}.
The rest of the function \algMkHornClauses is unchanged.
If the resulting Horn clauses have no solution, i.e.,
\algSolveHornClauses throws an \algUnsolvable exception, then it may
still be possible that a non-modular solution exists.
In this case, we invoke the abstraction refinement once again, this
time generating Horn clauses using the unmodified function
\algMkHornClauses from Figure~\ref{fig-alg-MkHornClauses}.

\paragraph{Example}
We illustrate the generation of Horn clauses using an expanded
version of $\TakeLockBit$ shown in Figure~\ref{fig-ex-cex-prog}.
This example contains an additional variable $\local$ local to the
first thread.
The initial symbolic state of the program $\SymbInit$ constrains both
$\local$ and $\lock$ to the value $0\ $. 
The transition relation of the first thread is extended with $\rel_0$,
which increments $\local$ by $1$ and assumes that the incremented value
is greater than or equal to $1\ $.
Similar to the example from Section~\ref{sec-illustration},
$\SymbError$ encodes the violation of the mutual exclusion property.
We show in Figure~\ref{fig-ex-cex}(a) the reachability trees as
computed by the \aret computation.
The error tuple consists of $m_4$ and $n_3\ $, i.e., $m_4
\land n_3 \land \SymbError$ is satisfiable.

From this error tuple, $\algMkHornClauses$ generates Horn clauses
following the procedure from Figure~\ref{fig-alg-MkHornClauses}.
These Horn clauses are shown in Figure~\ref{fig-ex-cex}(b).
The Horn clauses have unknown states 
$\unkOf{m_1}(\Vars)$, $\unkOf{m_2}(\Vars)$, $\unkOf{m_3}(\Vars)$, 
$\unkOf{m_4}(\Vars)$, $\unkOf{n_1}(\Vars)$, $\unkOf{n_2}(\Vars)\ $, and 
$\unkOf{n_3}(\Vars)\ $. 
The unknown transitions are $\unkOf{e_1}(\Vars,\Vars')$ and
$\unkOf{e_2}(\Vars,\Vars')\ $.

Comparatively, we show in Figure~\ref{fig-ex-cex}(c) the Horn clauses
generated with preference for modular solutions.
The solutions for the unknown states of thread 1 can refer only to
$(\GlobVars,\LocVarsOf{1})\ $, while the unknown states of thread 2
are restricted to $(\GlobVars,\LocVarsOf{1})\ $.
The unknown transitions are $\unkOf{e_1}(\GlobVars,\GlobVars')$ and
$\unkOf{e_2}(\GlobVars,\GlobVars')\ $.

\begin{theorem}[Progress of abstraction refinement]
\label{thm-refinement-progress}
%
The procedure \algRefine guarantees progress of abstraction
refinement, i.e., the same set of Horn clauses is never discovered
twice.\eofClaim
%
\end{theorem}
%
\iffalse
\includeProof{
\begin{proof} 
Let us consider an error tuple $(S_1, \dots, S_N)$ obtained
during \aret computation and let $\HornClauses$ be the set of Horn
clauses that are generated.
We assume here that the procedure \algSolveHornClauses returns a
solution that satisfies the generated Horn clauses. 
The predicates that appear in these solutions are added to
$\artPreds_i$ and $\artTPredsOf{i}{j}$ as in lines~8--12 of the
\algRefine procedure. 
Then, in the next \aret computation the abstract states will be at
least as precise as these solutions. 
Consequently, we can guarantees that, if abstract states corresponding
to $S_1, \dots, S_N$ are still reachable, then the error condition no
longer holds, i.e., $S_1 \land \dots S_N \land \SymbError$ is unsatisfiable.

Since each iteration of the loop from the \algMain function refines
the abstraction functions used in the previous iterations, we can
guarantee that all the previously discovered sets of Horn clauses remain
excluded in subsequent \aret computations.
\end{proof}
}
\fi

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/abstract.tex
Example?

%
Automated verification of multi-threaded programs requires explicit
identification of the interplay between interacting threads, so-called
environment transitions, to enable scalable, compositional reasoning.  
Once the environment transitions are identified, we can prove program
properties by considering each program thread in isolation, as the
environment transitions keep track of the interleaving with other
threads.
Finding adequate environment transitions that are sufficiently precise
to yield conclusive results and yet do not overwhelm the verifier with
unnecessary details about the interleaving with other threads is a
major challenge.
In this paper we propose a method for safety verification of
multi-threaded programs that applies (transition) predicate
abstraction-based discovery of environment transitions, exposing a
minimal amount of information about the thread interleaving.
The crux of our method is an abstraction refinement procedure that
uses recursion-free Horn clauses to declaratively state abstraction
refinement queries.
Then, the queries are resolved by a corresponding constraint solving
algorithm.
We present preliminary experimental results for mutual exclusion
protocols and multi-threaded device drivers.

\iffalse % popl submission abstract
%
In order to scale, verification of multi-threaded programs requires
definition of auxiliary variables that reveal some details about the
local state of the concurrent threads and construction of
rely/guarantee assertions that describe the thread interaction in
terms of shared and auxiliary variables.
%
In this paper we present an automatic method for the discovery of
auxiliary variables and relies/guarantees.
Our method is based on a generalization of counterexample guided
abstraction refinement to the multi-threaded setting.
The crux of the generalization is a notion of tree-like
counterexamples that explicitly account for thread interleaving.
By solving the corresponding tree-like path constraints, our method
reveals additional local facts through auxiliary variables and
iteratively strengthens relies/guarantees.
Furthermore, the tree-like path constraints can avoid the introduction
of auxiliary variables when they are not necessary.
%Furthermore, the tree-like path constraints enable the discovery of
%modular relies/guarantees (whose auxiliary variables do not reveal any
%local facts) whenever a modular proof exists.
We devised a solver for the tree-like constraints that is based on
Craig interpolation and used it in a prototype implementation of our
method.
We present preliminary experimental results for mutual exclusion
protocols and concurrent device drivers.
\fi

\iffalse % PLDI abstract
Automated verification of multi-threaded programs requires explicit
identification of the interplay between interacting threads, so-called
environment assumptions, to enable scalable reasoning.  
Once identified, these assumptions can be used for reasoning with one
program thread at a time, which is possible by using the respective
environment assumption to model the interleaving with other threads.
Finding adequate assumptions that are sufficiently precise to yield
conclusive results and yet keep track only of necessary facts about
the execution environment in order to scale well is a major challenge.
In this paper we propose a constraint-based technique for the
inference of such assumptions.
Our technique automatically steers towards an optimal
precision/efficiency trade-off between the extremes of efficient, but
incomplete thread-modular reasoning and complete, but prohibitively
expensive consideration of all interleavings.
For this task, we pinpoint a declarative formulation of modular
verification that allows one to express requisite environment
assumptions using constraints and admits algorithmic solutions based
on abstract fixpoint checking.
We describe an application of our environment assumption inference for
the verification of reachability and termination properties of
multi-threaded programs, and present our experience with its
implementation as well as evaluation in practice. 
\fi

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/algo-symb.tex
Example?

\section{Thread reachability and environment transitions}
\label{sec-algo-safety}

In this section, we present our rely-guarantee based verification
algorithm for proving safety properties of multi-threaded programs.
The algorithm is based on Theorem~\ref{thm-abst-fix} and consists of
three  main steps.
The first step computes for each thread a tree that is decorated by
abstract states and environment transitions, so-called \aret, and
analyses the discovered abstract states.
If an intersection with the error states of the program is found, then
the second step generates a set of corresponding Horn clauses, see
Section~\ref{sec-refinement}.
At the third step, we solve the constraint defined by the conjunction
of the generated Horn clauses and use the solutions to the refine the
abstraction functions used for the \aret computation, see
Section~\ref{sec-horn-solving}.

\input{fig-alg-main}

\paragraph{Function \algMain}

The main function of our algorithm \algMain is shown in
Figure~\ref{fig-alg-main}.
\algMain takes as input the multi-threaded program $\program\ $.
The \algRepeat loop iterates through the three main steps of the
algorithm.
First, we construct the abstraction functions $\abst_i$ and
$\abstTrans_{\envFromTo{i}{j}}$ at lines~4--6 from a given set of
(transition) predicates, which is empty initially.
Next, the \aret computation is performed in line~7 using these
abstraction functions.
In lines~8--9, the abstract states in the computed \aret's are
analyzed wrt.\ the safety property.
In case of a positive outcome of this check, \algMain constructs and
returns a safety proof in lines~17--18.
If the safety check fails, then \algRefine is executed on the
violating abstract states.
If \algRefine terminates normally, and hence succeeds in eliminating
the violation by refining the abstraction functions, then \algMain
continues with the next iteration of the \algRepeat loop.
In case an \algUnsolvable exception is thrown, \algCexPath from
Figure~\ref{fig-alg-cex-path} constructs a counterexample path that we
report to the user.

\paragraph{Procedure \algAbstReachEnvSolve}
\input{fig-alg-abst-reach}

See Figure~\ref{fig-alg-abst-reach} for the procedure
\algAbstReachEnvSolve that implements \aret computation using the
abstraction functions $\abst_i$ and $\abstTrans_{\envFromTo{i}{j}}\ $.
We use $\artParent$ and $\artTId$ to maintain information about the
constructed trees, and initialize them with the empty function $\bot$
in line~1.
$\artReachOf{i}$ and $\EnvTransOf{i}$ keep track of abstract states
and environment transitions for a thread $i\in\ThreadIds\ $. 
$\EnvTransOf{i}$ is initialized to an empty set in line~4, while
$\artReachOf{i}$ contains the abstraction of the initial program
states computed for the thread~$i\ $.
The \aret computation is performed iteratively in the \algRepeat loop,
see lines~5--27.

% =========== compute states

The first part of the loop (see lines~7--17) implements a standard, 
least fixpoint computation over reachable states.
At line~7, the algorithm picks an already reachable states
$S\in\artReachOf{i}$ in order to compute its abstract successors. 
After computing at lines~10--11 one successor of $S\ $, line~12 implements
a fixpoint check, which succeeds if $S'$ contains program states that
have not been reached yet.
The new states reachable in thread $i$ are stored in
$\artReachOf{i}\ $.
At line~14, the function $\artParent$ is updated to keep track of the
child-parent relation between abstract states, while $\artTId$ maps
the new reachable state to its parent thread.

% ======= compute environment transitions

The second part of the loop (see lines~$19$--$26$) performs a least
fixpoint computation over environment transitions. 
Each time a transition from a thread $i$ is picked at line~7, the
abstraction of its effect computed in line~20 is propagated to each
other thread~$j\ $.
Note however, that the propagation only happens for environment
transitions that are not subsumed by the existing ones, which is
checked in line~21.
Additional environment transitions are recorded in line~22.
Environment transitions are taken into consideration when computing
abstract state reachability, see line~9.

Upon termination, which is guaranteed by the finiteness of our
abstract domains, the function \algAbstReachEnvSolve computes sets of
abstract states $\artReachOf{1}, \dots, \artReachOf{\Num}$ and sets of
environment transitions $\EnvTransOf{1}, \dots, \EnvTransOf{\Num}\ $.

\input{fig-alg-mkPath}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/fig-ex-mk-tree.tex
Example?

\begin{figure*}[t]
  \centering
\begin{minipage}[t]{1.0\linewidth}
%  \begin{tikzpicture}[grow'=up, ->]

  \begin{tikzpicture}[grow'=right, ->]

    \tikzset{level distance=60pt,sibling distance=3pt}
    \tikzset{execute at begin node=\strut}

    \iffalse
    edge from parent/.style=
    {draw, edge from parent path={(\tikzparentnode.north)
                                  -- +(0,+3pt)
                                  -| (\tikzchildnode)}}
    \fi

    \Tree 
    [.\node [draw] {\small $\lfalse$};
      [.\node [stateBox] {\small $\unkOf{m_4}(\Vars^1)$};
        [.\node [stateBox] {\small $\unkOf{m_3}(\Vars^2)$};
          [.\node [stateBox] {\small $\unkOf{m_2}(\Vars^6)$};
            [.\node [stateBox] {\small $\unkOf{m_1}(\Vars^8)$};
            \node [draw] {\small $\StatesInit(\Vars^8)$};
            ]
            [.\node[draw] {\small $\rel_0(\Vars^8,\Vars^6)$};]
          ]
          [.\node [draw] {\small $\rel_1(\Vars^6,\Vars^2)$};]
        ]
        [.\node [draw]
        {\small $\set{
            \begin{array}[t]{@{}l@{}}
              \pcone^2 = \pcone^1\ , \\ 
              \local^2 = \local^1}
          \end{array}
          $};
        ]
        [.\node [draw] {\small $\unkOf{e_2}(\Vars^2,\Vars^1)$};
          [.\node [stateBox] {\small $\unkOf{n_1}(\Vars^2)$};
          \node [draw] {\small $\StatesInit(\Vars^2)$};
          ]
          [.\node [draw] {\small $\rel_2(\Vars^2,\Vars^1)$};]
        ]
        ]
      [.\node [stateBox] {\small $\unkOf{n_3}(\Vars^1)$};
        [.\node [stateBox] {\small $\unkOf{n_2}(\Vars^{11})$};
          [.\node [stateBox] {\small $\unkOf{n_1}(\Vars^{15})$};
          \node [draw] {\small $\StatesInit(\Vars^{15})$};
          ]
          [.\node [draw] {\small $\rel_2(\Vars^{15},\Vars^{11})$};]
        ]
        [.\node [draw] {\small $\set{\pctwo^{11}=\pctwo^{1}}$};]
        [.\node [stateBox] {\small $\unkOf{e_1}(\Vars^{11},\Vars^{1})$};
          [.\node [stateBox] {\small $\unkOf{m_2}(\Vars^{11})$};]
          [.\node [draw] {\small $\rel_1(\Vars^{11},\Vars^1)$};]
        ]
      ]
      [.\node [draw] {\small $\StatesError(\Vars^1)$};]
     ]

  \iffalse

  \node [stateBox, above of=m36 ] (m25) {$\unkOf{m_2}(\Vars^3)$}; \path (m25) edge (m36);

  \node [draw,xshift=1cm,yshift=2.5mm,above right of=m25]  (te1) {$\rel_{0}(\Vars^4,\Vars^3)$};\path (te1) edge(m25);
  \node [draw, above of=m14]  (i1)  {$\StatesInit(\Vars^4)$};    \path (i1) edge (m14);

  \node [stateBox,xshift=24mm,yshift=2mm,above right of=err] (n37) {$\unkOf{n_3}(\Vars^1)$};\path (n37) edge (err);

  \node [draw,xshift=1cm,yshift=2.5mm,above right of=n23] (t2)  {$\rel_2(\Vars^6,\Vars^5)$}; \path (t2) edge (n23);
  \node [draw, above of=n12] (i2)  {$\StatesInit(\Vars^6)$}; \path (i2) edge (n12);

  \node [stateBox,xshift=14mm,yshift=2mm,above right of=m47] (e2) {$\unkOf{e_2}(\Vars^2,\Vars^1)$};
  \path (e2) edge (m47);

 \node [draw, above of=n16] (ei2){$\StatesInit(\Vars^2)$};  \path (ei2) edge (n16);

 \node [stateBox, above of=m23] (m11) {$\unkOf{m_1}(\Vars^6)$}; \path (m11) edge (m23);

  \node[draw,xshift=1cm,yshift=2.5mm,above right of=e1] (et1) {$\rel_1(\Vars^5,\Vars^1)$};\path (et1) edge (e1);
 \node[draw, above of=m11]  (ei1)  {$\StatesInit(\Vars^6)$};\path (ei1) edge (m11);

%----------------------------

 \fi
  \end{tikzpicture}\\
\end{minipage}
\\
\begin{minipage}[t]{1.0\linewidth}
  \begin{tikzpicture}[<-,>=stealth',shorten >=1pt,auto, node distance=9mm,
                    semithick]
  \node[stateBox] (err) {$\lfalse$}; 
  \node [draw, xshift=8cm, above of=err] (errn) {$\StatesError(\Vars^1)$};\path (errn) edge (err);

  \node [stateBox,xshift=-3.6cm,yshift=2mm,above left of=err] (m47) {$\unkOf{m_4}(\Vars^1)$}; \path (m47) edge (err);
  \node [stateBox,xshift=-17mm,yshift=2mm, above left of=m47] (m36) {$\unkOf{m_3}(\Vars^2)$};\path (m36) edge (m47);
  \node [stateBox, above of=m36 ] (m25) {$\unkOf{m_2}(\Vars^3)$}; \path (m25) edge (m36);
  \node [stateBox, above of=m25 ] (m14) {$\unkOf{m_1}(\Vars^4)$}; \path (m14) edge (m25);

  \node [draw,xshift=1cm,yshift=3mm,above right of=m36] (t1) {$\rel_1(\Vars^3,\Vars^2)$}; \path (t1) edge (m36);
  \node [draw,xshift=1cm,yshift=2.5mm,above right of=m25]  (te1) {$\rel_{0}(\Vars^4,\Vars^3)$};\path (te1) edge(m25);
  \node [draw, above of=m14]  (i1)  {$\StatesInit(\Vars^4)$};    \path (i1) edge (m14);

  \node [stateBox,xshift=24mm,yshift=2mm,above right of=err] (n37) {$\unkOf{n_3}(\Vars^1)$};\path (n37) edge (err);
  \node [stateBox,xshift=-12mm,yshift=2mm,above left of=n37 ] (n23) {$\unkOf{n_2}(\Vars^5)$};\path (n23) edge (n37);
  \node [stateBox,     above of=n23 ] (n12) {$\unkOf{n_1}(\Vars^6)$}; \path (n12) edge (n23);

  \node [draw,xshift=1cm,yshift=2.5mm,above right of=n23] (t2)  {$\rel_2(\Vars^6,\Vars^5)$}; \path (t2) edge (n23);
  \node [draw, above of=n12] (i2)  {$\StatesInit(\Vars^6)$}; \path (i2) edge (n12);

  \node [stateBox,xshift=14mm,yshift=2mm,above right of=m47] (e2) {$\unkOf{e_2}(\Vars^2,\Vars^1)$};
  \path (e2) edge (m47);
  \node [stateBox, above of=e2 ] (n16) {$\unkOf{n_1}(\Vars^2)$}; \path (n16) edge (e2);

  \node [draw,xshift=1cm,yshift=2.5mm,above right of=e2] (et2) {$\rel_2(\Vars^2,\Vars^1)$}; \path (et2) edge (e2);
  \node [draw, above of=n16] (ei2){$\StatesInit(\Vars^2)$};  \path (ei2) edge (n16);

  \node [stateBox,xshift=14mm,yshift=2.5mm,above right of=n37] (e1)  {$\unkOf{e_1}(\Vars^5,\Vars^1)$};
  \path (e1) edge (n37);
  \node [stateBox, above of=e1 ] (m23) {$\unkOf{m_2}(\Vars^5)$}; \path (m23) edge (e1);
  \node [stateBox, above of=m23] (m11) {$\unkOf{m_1}(\Vars^6)$}; \path (m11) edge (m23);

  \node[draw,xshift=1cm,yshift=2.5mm,above right of=e1] (et1) {$\rel_1(\Vars^5,\Vars^1)$};\path (et1) edge (e1);
  \node[draw,xshift=1cm,yshift=2.5mm,above right of=m23] (ete1) {$\rel_0(\Vars^6,\Vars^5)$};\path (ete1) edge (m23);
  \node[draw, above of=m11]  (ei1)  {$\StatesInit(\Vars^6)$};\path (ei1) edge (m11);

%----------------------------

  \node [draw, yshift=1mm, above of=m47,text centered] (eq2) 
  {    
    \begin{array}[t]{@{}l@{}}
      \pcone^{1}=\pcone^{2}\\ 
      \local^{1}=\local^{2}\vspace{-1ex}
    \end{array}

  }; \path (eq2) edge (m47);

  \node [draw,yshift=-1mm,above of=n37,text centered ] (eq1) 
  {$\pctwo^{5}=\pctwo^{1}$}; \path (eq1) edge (n37);

  \end{tikzpicture}\\
\end{minipage}

\end{figure*}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/horn-solving.tex
Example?

\section{Solving Horn clauses over linear inequalities} \label{sec-horn-solving}

As presented in the previous section, \algRefine calls the function
\algSolveHornClauses.
In this section we present a function \algSolveLinearHornClauses that
can be used as an implementation of \algSolveHornClauses takes as
input a set of clauses $\HornClauses$ over linear inequalities that is
recursion-free.

To simplify the presentation of the algorithm, we make two additional
assumptions on~$\HornClauses$.
First, we assume that for each pair of clauses $(\dots\!\,) \limplies
\HornBody(\FreeVars)$ and $(\dots\!\,) \limplies
\HornBody'(\FreeVars')$ from $\HornClauses$ we have~$\HornBody \neq
\HornBody'$ and~$\HornBody \neq (\leq)\neq \HornBody'$.
Second, we assume that $\HornClauses$ contains a clause $(\dots\!\,)
\limplies \lfalse\ $.

The additional assumptions are satisfied by the clauses generated in
Section~\ref{sec-refinement}.
In case \algSolveLinearHornClauses is applied on a set of
recursion-free Horn clauses over linear arithmetic that violates the
two assumptions above, we can apply a certain renaming of relation
symbols and introduction of additional clauses to meet the
assumptions.

%-----------------------------------------------------------------------------------

\input{fig-ex-cex}

\paragraph{Function \algMkHornTree}

The function \algMkHornTree generates a tree representation for a set
of Horn clauses and is shown in Figure~\ref{fig-alg-mkTree}.
For every relation appearing in the Horn clauses, the algorithm
generates a corresponding tree node.
The children of a node are maintained in a function $\cChildren$ as follows.
Nodes that correspond to linear arithmetic relations have no children,
see lines~11--12.
A node that corresponds to an unknown relation with a relation symbol
$b$ has as children those nodes that represent relation symbols that
depend on~$b\ $.
The $\cFormula$ attribute of the tree nodes is initialized to a linear
arithmetic constraint for leaves of the tree in line~11, and to an
unknown relation for internal tree nodes in line~13.

\paragraph{Function \algSolveLinearHornClauses}
See Figure~\ref{fig-alg-solveHorn} for the pseudocode of the procedure
$\algSolveLinearHornClauses\ $.
This procedure creates a tree representation for $\HornClauses$ at line~2.
At line~3, we build a set containing all the $\cFormula$ attributes of
leaf nodes and store this set in $\cLeaves\ $ .
The input set of Horn clauses is satisfiable if and only if
$\Land \cLeaves$ is unsatisfiable.
If $\Land \cLeaves$ is unsatisfiable, the test at lines~4--5 succeeds
and returns a proof of unsatisfiability in the form of weights for
each linear inequality. 
This test can be implemented using some linear arithmetic constraint
solver.
If the constraint solver fails to find a proof, an exception
$\algUnsolvable$ is thrown at line~12.

At line~7, $\algSolveLinearHornClauses$ calls the procedure
$\algBuildPred\ $, which is presented in Figure~\ref{fig-alg-annotPred}.
This procedure recursively traverses the input tree in postorder.
If this procedure is invoked for a leaf node $\cnode\ $, it directly
computes the value of $\cInterPolOf{\cnode}$ as a linear combination
of atomic formulas with weights given by the $\proofVec$ function (see
line~2).
If this procedure is invoked for an internal node $\cnode\ $, the
attribute $\cInterPol$ of $\cnode$'s children is computed using a
recursive call at line~5.
After completing the recursive call, $\cInterPolOf{\cnode}$ is
calculated by adding the values of the $\cInterPol$ attributes of
$\cnode$'s children.

Since there may be multiple nodes in the tree corresponding to the
same unknown relation, the algorithm has to account for the
$\cInterPol$ attributes of all these nodes.
Therefore, at lines~8--9 we compute solutions for each
$\HornBody(\FreeVars)$ in $\Unknowns$ by taking conjunction of
$\cInterPol$ of each node of the tree that is labeled with
$\HornBody(u)$ for some~$u\ $.

%The algorithm $\algSolveLinearHornClauses$ is closely related to the
%interpolation generation method described in \cite{McMillanTACAS04}.

%-----------------------------------------------------------------------------------

\begin{theorem}
\algSolveLinearHornClauses computes a solution for a set of Horn
clauses  $\HornClauses$ if and only if the conjunction of the clauses
in $\HornClauses$ is satisfiable.
\end{theorem}
\iffalse
\todo{proof sketch}
\includeProof{
\begin{proof}
  Given that each relational variable appears only once as the head of
  a Horn clause in $\HornClauses\ $, we denote by $body(b(w))$ the left
  hand side of this particular clause.
  Given the recursion-free structure of the set of clauses
  $\HornClauses\ $, we can replace each relational variable $b_i(w_i)$
  that appears in a body of a clause with $body(b_i(w_i))$ (with
  proper renamings).
  This replacement preserves the solutions of $\HornClauses\ $.
  After replacement, the set of Horn clauses is reduced to one Horn
  clause such that all the terms that appear in the body of this
  clause are known predicates.

  From the hypothesis that $\HornClauses$ has a solution, we can
  deduce that the conjunction of the terms in the body of the clause
  is unsatisfiable.
  From the fact that the solution of $\HornClauses$ is in the
  conjunctive linear inequality domain, we can guarantee that a proof
  of the unsatisfiability is found using a linear arithmetic solver.
  Therefore the test at lines~4--5 from \algSolveLinearHornClauses
  succeeds and returns a function $\proofVec\ $.
  From the $\proofVec$ function, \algBuildPred computes solutions for
  each relational variable which satisfy the input set of Horn
  clauses.
\end{proof}
}
\fi

\paragraph{Example}
We illustrate the solving procedure using the same example from the
previous section. 
Given the Horn clauses from Figure~\ref{fig-ex-cex}(b), \algMkHornTree
constructs a tree that is shown in Figure~\ref{fig-ex-cex}(d).
This tree contains nodes which we label for convenience with
identifiers from $1$ to $26\ $.
In Figure~\ref{fig-ex-cex}(d) we show the $\cFormula$ map of the tree.
A witness of the unsatisfiability of $\Land \cLeaves$ is given by the
following atomic formulas:
\begin{equation*}  
  \begin{array}[t]{l}
    \begin{array}[t]{@{}c@{}}
      (\local^{3} \geq 1 )\\
      {\downin}\\
      {\cFormulaOf{7}}    
    \end{array}
    \land 
    \begin{array}[t]{@{}c@{}}
      (\local^{3} = \local^{2})\\
      {\downin} \\ 
      {\cFormulaOf{8}}
    \end{array}
    \land 
    \begin{array}[t]{@{}c@{}}
      (\local^{2}=0)\\
      {\downin}\\ 
      {\cFormulaOf{11}}
    \end{array}
\end{array}\end{equation*}
Our solver treats each linear equality as a conjunction of two linear
inequalities.
The equality $\local^{2}=0$ is split in two inequalities
$\local^{2}\leq0\land -\local^{2}\leq0\ $.
The proof of unsatisfiability is:
%
\begin{equation*}  \begin{array}[t]{l}
(1 \leq \local^{3} )+
(\local^{3} \leq \local^{2} ) +
( \local^{2} \leq 0) = (\linfalse).
\end{array}\end{equation*}
%
This is encoded in the $\proofVec$ map with values of $1$ at locations
corresponding to the three atomic formulas above and values of $0$ for
all the other atomic formula.
Next, we show in Figure~\ref{fig-ex-cex}(e) the values for the
$\cInterPol$ map as computed by $\algBuildPred\ $.
The final solution of the Horn clauses is built by a conjunction of
the $\cInterPol$ attributes for nodes with the same unknown label. 
The resulting solution $\Solution$ is shown below.
%
\begin{equation*} 
  \begin{array}[t]{l@{\;\;=\;\;}l@{\;\;=\;\;}l}
    \SolutionOf{\unkOf{m_1}(\Vars)} & \SolutionOf{\unkOf{n_2}(\Vars)} &
    (\lintrue)\\[\jot]
    \SolutionOf{\unkOf{e_1}(\Vars,\Vars')}& & (\lintrue) \\[\jot]
    \SolutionOf{\unkOf{m_2}(\Vars)} & \SolutionOf{\unkOf{m_3}(\Vars)} &
    (1 \leq \local) \\[\jot]
    \SolutionOf{\unkOf{n_1}(\Vars)} & \SolutionOf{\unkOf{e_2}(\Vars,\Vars')} & 
    (\local \leq 0) 
  \end{array}
\end{equation*}

\input{fig-alg-solveHorn}
\input{fig-alg-mkTree}
\input{fig-alg-annotPred}

\paragraph{Solving the clauses shown in Figure~\ref{fig-ex-cex}(c)}
Given the Horn clauses shown in this figure, \algMkHornTree returns a
tree representation with a similar $\cChildren$ map structure but with
different $\cFormula$ attributes.
The part of the tree that contributes to the proof of unsatisfiability
is shown in Figure~\ref{fig-ex-cex}(f).
The variable $\local^2$ does not appear in the subtree of the node $9$
since $\cFormulaOf{9}=\unkOf{e_2}(\GlobVars^2,\GlobVars^1)\ $. 
Part of this subtree is the node 11.
Let us name the variable at this node as $\local^{\textsc{VIII}}\ $.
The proof of unsatisfiability shown above does no longer hold, since
the following formula is satisfiable:
\[
 (1\leq\local^{3} ) \land (\local^{3} = \local^{2}) \land (\local^{\textsc{VIII}}=0)
\]
However, the conjunction of the elements from the $\cLeaves$ set is
still unsatisfiable, indicating that a modular solution exists.
We find that the following atoms contribute to a proof of unsatisfiability:
\begin{equation*}
  \begin{array}[t]{@{}c@{}}
      \lock^2=1\\
      \downin\\
      {\cFormulaOf{8}}    
    \end{array}
  \land 
  \begin{array}[t]{@{}c@{}}
      \lock^2=0\\
      \downin\\
      {\cFormulaOf{12}}
  \end{array}
\end{equation*}
After splitting the equalities in equivalent inequalities, our
algorithm computes the following solution:
\begin{equation*}  
  \begin{array}[t]{r@{\;}lr@{\;}l}
    \cInterPolOf{8} &= (1\leq\lock^{2} ) & \cInterPolOf{12} &= (\lock^{2}\leq 0 ) \\
    \cInterPolOf{3} &= (1\leq\lock^{2} ) & \cInterPolOf{9} &= (\lock^{2} \leq 0 ) \\
    \cInterPolOf{2} &= (\linfalse)  
  \end{array} 
\end{equation*}
%
From this $\cInterPol$ map, our algorithm derives a solution
$\Solution$ in lines 8--9 and succeeds in computing modular
predicates.

% \todo{ PC vs. Data.
% Each constraint originating from the initial states, the transition
% relations and the error states can be split in to two parts.
% The first constraint is over program counter variables and the other
% constraint relates the data variables. 
% The splitting procedure preserves unsatisfiability of the original
% constraint since there in no atomic formula that relates both program
% counter variables and data variables.
% The original constraint system is unsatisfiable if and only if either
% of the two parts is unsatisfiable.
% For solving the program counter constraints, \algRefinePC at line 13
% implements an algorithm proposed in our earlier work
% \cite{GuptaATVA10}.
% }

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/illustration.tex
Example?

\section{Illustration}\label{sec-illustration}

In this section we illustrate our algorithm using two multi-threaded
examples.
The first example does not have a modular proof, hence our algorithm
reasons about relationship between the local variables of different
threads.
For the second example, our algorithm succeeds in finding a modular
proof by applying an abstraction refinement procedure that guarantees
the discovery of a modular abstraction whenever it exists.

\iffalse

Firstly, the algorithm generates thread reachability assertions
in the form of trees of reachable states for each thread.
If a reachable state is found to violate the mutual exclusion
property, then our algorithm generates a set of Horn clauses that
encodes how the abstract error state was reached.
If the set of Horn clauses is satisfiable, our solving procedure
returns a solution which is used for abstraction refinement.

The first multi-threaded program does not have a thread-modular proof
and hence its verification leads to environment transitions that
reveal some facts about the local state of the threads. 
The second program can be proven correct using a thread-modular proof,
however, non-modular proofs that (unnecessarily) involve local
variables can also be used to prove correctness of this program. 
\fi

\subsection{Example \TakeLockBit}\label{subsec-takebit}

See Figure~\ref{fig-take-lock} for the program \TakeLockBit that
consists of two threads.
The threads attempt to access a critical section, and synchronize their
accesses using a global variable~$\lock\ $.
We assume that initially the lock is not taken, i.e., $\lock = 0\ $,
and that the locking statement \verb|take_lock| waits until the lock
is released and then assigns the value of its second parameter
to~$\lock\ $, thus taking the lock.
We write $\Vars = (\lock, \pcone, \pctwo)$ for the program variables,
where $\pcone$ and $\pctwo$ are local program counter variables of the
first and second thread, respectively.

% ===================== Program description ===============

We start by representing the program using assertions $\SymbInit$ and
$\SymbError$ over program variables that describe the initial and
error states of the program, together with assertions over unprimed
and primed program variables $\rel_1$ and $\rel_2$ that describe the
transition relations for program statements.
%
\begin{equation*}
  \begin{array}[t]{r@{}l@{}}
    \SymbInit = (&\pcone=a \land \pctwo=p \land \lock=0)\ , \\[\jot]
    \SymbError = (&\pcone=b \land \pctwo=q)\ ,\\[\jot]
    \rel_1 = (&\lock=0 \land \lock'=1 \land \pcone=a \land
    \pcone'=b \land  \relLocEqOf{2})\ ,
    \\[\jot]
    \relLocEqOf{1} = (&\pcone=\pcone')\ , \\[\jot]
    \rel_2 = (&\lock=0 \land \lock'=1 \land \pctwo=p \land \pctwo'=q
    \land \relLocEqOf{1})\ , \\[\jot]
    \relLocEqOf{2} = (&\pctwo=\pctwo')\ .
 \end{array}
\end{equation*} 
%
The auxiliary assertions $\relLocEqOf{1}$ and $\relLocEqOf{2}$ state
that the local variable of the first and second thread, respectively,
is preserved during the transition.

\iffalse
The transition relations $\rel_1$ and $\rel_2$ refer to all program
variables, i.e., the global variable $\lock$ as well as the local
variables $\pcone$ and $\pctwo\ $.

For example, the transition relation $\rel_1$ of the first thread
requires that the local variable of the second thread does not change,
which is expressed by the conjunct $\pctwo = \pctwo'$ in~$\rel_1\ $.
\fi

To verify \TakeLockBit, our algorithm computes a sequence of \aret{}s
(Abstract Reachability and Environment Trees).
Each tree computation amounts to a combination of i) a standard
abstract reachability computation that is performed for each thread
and is called thread reachability, and ii) a construction and
application of environment transitions.
Abstract states represent sets of (concrete) program states, while
environment transitions are binary relations of program states.

\begin{figure}[h!]
\hrule
\centering
\begin{minipage}[t]{.48\columnwidth} 
\begin{verbatim}

// Thread 1 
a: take_lock(lock, 1); 
b: // critical 
\end{verbatim}
\end{minipage}
%
\begin{minipage}[t]{.48\columnwidth} 
\begin{verbatim}

  // Thread 2
  p: take_lock(lock, 1);
  q: // critical

\end{verbatim}
\end{minipage}
\caption{Example program \TakeLockBit.
  Each thread waits until the lock is released, and then assigns the
integer $1$ to~$\lock\ $.
}
\label{fig-take-lock}
\end{figure}

% ===================== Iteration 1 =======================
\paragraph{First \aret computation}

The thread reachability computation for the first thread starts by
computing an abstraction of the initial program states~$\SymbInit\ $.
Here, we use an abstraction function $\abst_1\ $, where the dot
indicates that this function over-approximates sets of program states
(and not sets of pairs of states, as will take place later) and the
index $1$ indicates that this abstraction function is used for the
first thread.
In this example, we assume that the abstraction function only tracks
the value of the program counter of the first thread, i.e.,
$\artPreds_1 = \set{\pcone=a, \pcone=b}\ $, and is computed as
follows: $\abst_1(S) = \Land \set{\dot{p} \in \artPreds_1 \mid \forall
\Vars:S \limplies \dot{p}}\ $.
We obtain the initial abstract state $m_1$ as follows:
%
\begin{equation*}
  m_1  = \abst_1(\SymbInit) = (\pcone = a)\ .
\end{equation*}
%
 Next, we compute an abstract successor of $m_1$ with respect to the
transition~$\rel_1$ using the strongest postcondition operator $\post\
$ that is combined with $\abstTIdOf{1}$:
%
\begin{equation*}
  m_2 = \abst_1(\post(\rel_1, m_1)) = (\pcone = b)\ .
\end{equation*}

\input{fig-ex-takebit-art}

Similarly, we compute the thread reachability for the second thread.
Using predicates over the program counter of the second thread, i.e.,
$\artPreds_2 = \set{\pctwo=p, \pctwo=q}\ $, we compute the following
two abstract states:
%
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}l@{\;=\;}l}
    n_1 &  \abst_2(\SymbInit) &  (\pctwo = p)\ ,\\[\jot]
    n_2 &  \abst_2(\post(\rel_2, n_1)) &  (\pctwo = q)\ .
  \end{array}
\end{equation*}
%
For each thread, we organize the computed abstract states in a tree,
see Figure~\ref{fig-ex-takebit-art}(a).

We stop the \aret computation since we discover that the error states
overlap with the intersection of the abstract state $m_2$ from the
thread reachability of the first thread and $n_2$ from the second
thread, i.e., $m_2 \land n_2 \land \SymbError$ is satisfiable.

\paragraph{First abstraction refinement}

We treat the pair $m_2$ and $n_2$ as a possible evidence that the
error states of the program can be reached.
Yet, we cannot assert that the program is incorrect, since abstraction
was involved when computing $m_2$ and $n_2\ $.

We check if the discovered evidence is spurious by formulating a
constraint that is satisfiable if and only if the abstraction can be
refined to exclude the spuriousness.
For each abstract state involved in the reachability of and including
$m_2$ and $n_2$ we create an unknown predicate that denotes a set of
program states. 
We obtain $\unkOf{m_1}(\Vars)$, $\unkOf{n_1}(\Vars)$,
$\unkOf{m_2}(\Vars)$, and~$\unkOf{n_2}(\Vars)\ $, which correspond to
$m_1$, $n_1$, $m_2$, and~$n_2$, respectively.
Then, we record the relation between the unknown predicates using
constraints in the form of Horn clauses.
For example, since $m_1$ was an abstraction of the initial program
states, we require that $\unkOf{m_1}(\Vars)$ over-approximates
$\SymbInit$ as well, and represent this requirement by a Horn clause
$\SymbInit \limplies \unkOf{m_1}(\Vars)\ $.
As a result, we obtain the following set of clauses.
%
\begin{equation*}
  \HornClauses_1 = \set{
  \begin{array}[t]{@{}l}
    \SymbInit \limplies \unkOf{m_1}(\Vars),\ 
    \unkOf{m_1}(\Vars) \land \rel_1 \limplies \unkOf{m_2}(\Vars'),\\[\jot]
    \SymbInit \limplies \unkOf{n_1}(\Vars),\ 
    \unkOf{n_1}(\Vars) \land \rel_2 \limplies \unkOf{n_2}(\Vars'),\\[\jot]
    \unkOf{m_2}(\Vars) \land \unkOf{n_2}(\Vars) \land \SymbError \limplies 
    \lfalse }
  \end{array}
\end{equation*}
%
The last clause in $\HornClauses_1$ requires that the intersection of
the refined versions of the abstract states $m_2$ and $n_2$ is
disjoint from the error states of the program.

We check if the conjunction of the clauses in $\HornClauses_1$ is
satisfiable using a SAT-based algorithm presented
in~\cite{GuptaATVA10}.
(Section~\ref{sec-horn-solving} presents an algorithm for solving Horn
clauses over linear inequalities.)
We obtain the following satisfying assignment~$\Solution$ that maps
each unknown predicate to an assertion of the program variables.
%
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}ll@{\;=\;}l}
    \SolutionOf{\unkOf{m_1}(\Vars)} & (\pctwo=p) & 
    \SolutionOf{\unkOf{n_1}(\Vars)} & (\pcone=a)\\[\jot]
    \SolutionOf{\unkOf{m_2}(\Vars)} & (\pctwo=p) & 
    \SolutionOf{\unkOf{n_2}(\Vars)} & (\pcone=a)
  \end{array}
\end{equation*} 
%
The existence of $\Solution$ indicates that the discovered evidence is
spurious. 
We use $\Solution$ to refine the abstraction functions and hence
eliminate the source of spuriousness.
We collect the predicates that appear in the solution for abstract
states from the first thread, add them to the sets of predicates
$\artPreds_1$, and perform a similar step for the second thread.
The resulting sets of predicates are shown below.
%
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}l}
  \artPreds_1 & \set{\pcone=a, \pcone=b, \pctwo=p}\\[\jot]
  \artPreds_2 & \set{\pctwo=p, \pctwo=q, \pcone=a}
  \end{array}
\end{equation*}
%
They guarantee that the same spuriousness will not appear during
subsequent \aret computations.

\iffalse
\todo{Drop?}
We observe that the assignments for $\unkOf{m_1}(\Vars)$ and
$\unkOf{m_2}(\Vars)$ from the first thread involve the local variable
$\pctwo$ of the second thread, and the assignments for
$\unkOf{n_1}(\Vars)$ and $\unkOf{n_2}(\Vars)$ involve the variable
$\pcone\ $. 
While unavoidable for this constraint system, such solutions lead to
non modular reasoning in the next \aret computation.
\fi

% ===================== Iteration 2 =======================
\paragraph{Second \aret computation}
We re-start the \aret computation using the previously discovered 
predicates. 
Figure~\ref{fig-ex-takebit-art}(b) shows the two trees computed with
the refined abstraction functions where
%
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}ll@{\;=\;}l}
    m_1 & (\pcone=a \land \pctwo=p)\ , & n_1 & (\pcone=a \land
    \pctwo=p)\ ,
    \\[\jot] 
    m_2 & (\pcone=b \land \pctwo=p)\ , & n_2 & (\pcone=a \land
    \pctwo=q)\ .
  \end{array}
\end{equation*}
%
Due to the first abstraction refinement step, $m_2 \land n_2 \land
\SymbError$ is unsatisfiable.
The thread reachability computation for each thread does not discover
any further abstract states.

The \aret computation proceeds by considering interleaving of the
transitions from one thread with the transitions from the other
thread. 
We account for thread interleaving by constructing and applying
environment transitions.
First, we construct an environment transition $e_1$ that records the
effect of applying $\rel_1$ on $m_1$ in the first thread on the thread
reachability in the second thread.
This effect is over-approximated by using an abstraction
function~$\abstTransTIdOf{1}{2}$.
In this function, the double dot indicates that the function abstracts
binary relations over states (and not sets of states).
The index $\envFromTo{1}{2}$ indicates that this function is applied
to abstract effect of the first thread on the second thread.
Initially, we use the empty set of transition predicates (over pairs
of states) $\artTPreds_{\envFromTo{1}{2}}=\emptyset\ $ to
define~$\abstTransTIdOf{1}{2}$.
The environment transition $e_1$ is defined as
%
\begin{equation*}
  e_1 = \abstTrans_{\envFromTo{1}{2}}(m_1 \land \rel_1) = \ltrue\ ,
\end{equation*}
%
and it non-deterministically updates the program variables (since
$\ltrue$ does not impose any restrictions on the successor states of
the transition).

Next, we add $e_1$ to the transitions of the second thread.
Then, its thread reachability computation uses $e_1$ during the
abstract successor computation, and creates an abstract state $n_3$ by
applying $e_1$ on~$n_2$ as follows:
%
\begin{equation*}
  n_3 = \abstTIdOf{2}(\post(e_1 \land \relLocEqOf{2}, n_2)) = (\pctwo
  = q)\ .
\end{equation*}
%
The conjunct $\relLocEqOf{2}$ ensures that the local variable of the
second thread is not changed by the environment transition.

Symmetrically, we use a function $\abstTransTIdOf{2}{1}$ to abstract
the effect of applying transitions in the second thread on the thread
reachability of the first thread.
The application of $\rel_2$ on $n_1$ results in the environment
transition $e_2$ such that
%
\begin{equation*}
  e_2 = \abstTransTIdOf{2}{1}(n_1 \land \rel_2) = \ltrue\ .
\end{equation*}
%
We apply $e_2$ to contribute an abstract successor $m_3$ of the
abstract state $m_2$ to the thread reachability of the first thread:
%
\begin{equation*}
  m_3 = \abstTIdOf{1}(\post(e_2 \land \relLocEqOf{1}, m_2)) = (\pcone=b)\ .
\end{equation*}
%

\iffalse
We denote by $\EnvTransOf{j}$ the set of all environment
transitions that are to be applied in thread $j\ $. 
For example, after the previous step, $\EnvTransOf{1} = \set{e_2}$ and
$\EnvTransOf{2} = \set{e_1}\ $.
The effect of the environment transitions $e_1$ and, respectively,
$e_2$ is to havoc the values of all program variables except $\pctwo$
and respectively $\pcone$ (local variables in the thread that applies
the environment transition).
%
\begin{equation*}
  \begin{array}[t]{r@{\arrSpace}l@{\arrSpace}l}
    m_3 & = \abst_1(\post(e_2 \land \pcone=\pcone', m_2)) & =(\pcone=b)\\[\jot]
    n_3 & = \abst_2(\post(e_1 \land \pctwo=\pctwo', n_2)) & =(\pctwo=q)\\[\jot]
  \end{array}
\end{equation*}
%
\fi

We observe that the intersection of the abstract states $m_3$ and
$n_3$ contains a non-empty set of error states, i.e., $m_3 \land n_3
\land \SymbError$ is satisfiable, thus delivering a possible evidence
for incorrectness.

% Note that the branches that give the tree-shape of the counterexample
% originate either from the error check or from the application of
% environment transitions.

\paragraph{Second abstraction refinement}

Similarly to the first abstraction refinement step, we construct a set
of Horn clauses $\HornClauses_2$ to check if the discovered evidence
is spurious.
We consider predicates $\unkOf{m_1}(\Vars)$, $\unkOf{n_1}(\Vars)$,
$\unkOf{m_2}(\Vars)$, $\unkOf{n_2}(\Vars)$, $\unkOf{m_3}(\Vars)$, and
$\unkOf{n_3}(\Vars)$ that represent unknown sets of program states,
together with $\unkOf{e_1}(\Vars,\Vars')$ and
$\unkOf{e_2}(\Vars,\Vars')$ that represent unknown binary relations
over program states.
%
\begin{equation*}
  \HornClauses_2 = \set{
    \begin{array}[t]{@{}l}
      \SymbInit \limplies \unkOf{m_1}(\Vars),\ 
      \unkOf{m_1}(\Vars) \land \rel_1 \limplies \unkOf{m_2}(\Vars'), \\[\jot]
      \SymbInit \limplies \unkOf{n_1}(\Vars),\ 
      \unkOf{n_1}(\Vars) \land \rel_2 \limplies \unkOf{n_2}(\Vars'), \\[\jot]
      \unkOf{m_1}(\Vars) \land \rel_1 \limplies \unkOf{e_1}(\Vars,\Vars'),
      \\[\jot]
      \unkOf{n_2}(\Vars) \land \unkOf{e_1}(\Vars,\Vars') \land \relLocEqOf{2}
      \limplies \unkOf{n_3}(\Vars'), \\[\jot]
      \unkOf{n_1}(\Vars) \land \rel_2 \limplies \unkOf{e_2}(\Vars,\Vars'),
      \\[\jot]
      \unkOf{m_2}(\Vars) \land \unkOf{e_2}(\Vars,\Vars') \land \relLocEqOf{1}
      \limplies \unkOf{m_3}(\Vars'),\\[\jot]
      \unkOf{m_3}(\Vars) \land \unkOf{n_3}(\Vars) \land \SymbError \limplies 
      \lfalse }
  \end{array}
\end{equation*}
%
The conjunction of clauses in $\HornClauses_2$ is satisfiable.
We obtain the following satisfying assignment~$\Solution$.
%
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}ll@{\;=\;}l}
    \SolutionOf{\unkOf{m_1}(\Vars)}&\ltrue & 
    \SolutionOf{\unkOf{m_2}(\Vars)}&(\lock=1) \\[\jot]
    \SolutionOf{\unkOf{n_1}(\Vars)}&\ltrue &
    \SolutionOf{\unkOf{n_2}(\Vars)}&(\lock=1)\\[\jot]
    \SolutionOf{\unkOf{m_3}(\Vars)}&\lfalse & 
    \SolutionOf{\unkOf{e_1}(\Vars,\Vars')}&(\lock=0) \\[\jot]
    \SolutionOf{\unkOf{n_3}(\Vars)}&\lfalse &
    \SolutionOf{\unkOf{e_2}(\Vars,\Vars')}&(\lock=0)
  \end{array}
\end{equation*}
%
This solution constrains $\unkOf{m_2}(\Vars)$ and $\unkOf{n_2}(\Vars)$
to states where the lock is held $(\lock=1)\ $, while the environment
transitions $\unkOf{e_1}(\Vars,\Vars')$ and
$\unkOf{e_2}(\Vars,\Vars')$ are applicable only in states for which
the lock is not held by the respective thread $(\lock=0)\ $.

We add the (transition) predicates that appear in the environment
transition $e_2$ of the first thread to the set $\artTPredsOf{2}{1}\ $,
and, symmetrically, we add the predicates from $e_1$ to
$\artTPredsOf{1}{2}$.
For the next \aret computation we have the following set of
predicates:
%
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}l}
    \artPreds_1 & \set{\pcone=a, \pcone=b, \pctwo=p, \lock=1}\ ,\\[\jot]
    \artPreds_2 & \set{\pctwo=p, \pctwo=q, \pcone=a, \lock=1}\ ,\\[\jot]
    \artTPredsOf{1}{2} & \set{\lock=0}\ ,\\[\jot]
    \artTPredsOf{2}{1} & \set{\lock=0}\ .
  \end{array}
\end{equation*}
%

% ===================== Iteration 4 =======================
\paragraph{Last \aret computation}

We perform another \aret computation and a subsequent abstraction
refinement step.
We add the predicate $\lock'=1$ to both $\artTPredsOf{1}{2}$
and~$\artTPredsOf{2}{1}$, and proceed with the final \aret computation.
%
Figure~\ref{fig-ex-takebit-art}(c) shows the resulting trees.
The application of thread transitions produces the following abstract
states and environment transitions:
%
\begin{equation*}
  \begin{array}[t]{l@{\arrSpace}l@{\arrSpace}l@{}}
    m_1 & = \abst_1(\SymbInit) & = (\pcone = a \land \pctwo = p)\ ,  \\[\jot]
    n_1 & = \abst_2(\SymbInit) & = (\pcone = a \land \pctwo = p)\ , \\[\jot] 
    m_2 & = \abst_1(\post(\rel_1, m_1)) & = (\pcone = b \land \pctwo = p 
    \land \lock = 1)\ , \\[\jot] 
    n_2 & = \abst_2(\post(\rel_2, n_1)) & = (\pcone = a \land \pctwo = q 
    \land \lock = 1)\ , \\[\jot] 
    e_1 & = \abstTrans_{\envFromTo{1}{2}}(m_1 \land \rel_1) & = (\lock = 0 
    \land \lock'=1)\ , \\[\jot]
    e_2 & = \abstTrans_{\envFromTo{2}{1}}(n_1 \land \rel_2) & = (\lock = 0 
    \land \lock'=1)\ .
  \end{array}
\end{equation*}
%
The environment transitions $e_1$ and $e_2$ produce the abstract
states $m_3$ and $n_3$ whose intersection does not contain any error
states.
%
\begin{equation*}
  \begin{array}[t]{l@{\arrSpace}l@{\arrSpace}l}
    m_3 &= \abst_1(\post(e_2 \land \relLocEqOf{1}, m_2)) &= (\pcone = a 
    \land \lock = 1) \\[\jot] 
    n_3 &= \abst_2(\post(e_1 \land \relLocEqOf{2}, n_2)) &= (\pctwo = p 
    \land \lock = 1)
  \end{array}
\end{equation*}

Neither thread nor environment transitions can be applied from the
abstract states $m_3$ and $n_3$, while no further abstract states are
found.
Since each pair of abstract states from different threads yields an
intersection that is disjoint from the error states, we conclude that
\TakeLockBit is safe.
The labeling of the computed trees can be directly used to construct a
safety proof for \TakeLockBit, as Sections~\ref{sec-proof-rule}
and~\ref{sec-algo-safety} will show.

\iffalse
We denote by $\artReachOf{1}$ and $\artReachOf{2}$ the set of all
abstract reachable states found in this last \aret computation.
\fi

\subsection{Example \TakeLockThreadId} \label{subsec-takethreadid}

\begin{figure}[t]
\small
\hrule
\begin{minipage}[t]{.48\columnwidth} 
\begin{verbatim}

    // Thread 1 
    a: take_lock(lock, 1); 
    b: // critical 
\end{verbatim}
\end{minipage}
\hfill
\begin{minipage}[t]{.48\columnwidth} 
\begin{verbatim}

    // Thread 2
    p: take_lock(lock, 2);
    q: // critical

\end{verbatim}
\end{minipage}
  \caption{Example program $\TakeLockThreadId\ $.}
  \label{fig-take-lock-2}
\end{figure}

Our second example \TakeLockThreadId, shown in
Figure~\ref{fig-take-lock-2}, is a variation of \TakeLockBit.
\TakeLockThreadId uses an integer variable $\lock$ (instead of a
single bit) to record which thread holds the lock.
Due to this additional information recorded in the global variable,
the example \TakeLockThreadId has a modular proof, which does not
refer to any local variables.
We show how our algorithm discovers such a proof by only admitting
modular predicates in the abstraction refinement step.

\TakeLockThreadId differs from \TakeLockBit in its transition relation
$\rho_2$:
%
% ===================== Program description ===============
\begin{equation*}
  \rel_2 = (\lock=0 \land \lock'=2 \land \pctwo=p \land \pctwo'=q
  \land \relLocEqOf{1})\ .
\end{equation*}

% ===================== Iteration 1 =======================
\paragraph{First \aret computation}

Similarly to \TakeLockBit, we discover that $m_2 \land n_2 \land
\SymbError$ is satisfiable, and compute the following set of Horn
clauses:
%
\begin{equation*}
  \HornClauses_3 = \set{
  \begin{array}[t]{@{}l}
     \SymbInit \limplies \unkOf{m_1}(\Vars),\
     \unkOf{m_1}(\Vars) \land \rel_1 \limplies \unkOf{m_2}(\Vars'),\\[\jot]
     \SymbInit \limplies \unkOf{n_1}(\Vars),\ 
     \unkOf{n_1}(\Vars) \land \rel_2 \limplies \unkOf{n_2}(\Vars'),\\[\jot]
     \unkOf{m_2}(\Vars) \land \unkOf{n_2}(\Vars) \land \SymbError \limplies 
     \lfalse }\ .
  \end{array}
\end{equation*}
%
One possible satisfying assignment $\Solution$ is:
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}ll@{\;=\;}l@{}}
    \SolutionOf{\unkOf{m_1}(\Vars)} & (\pctwo=p)\ , & 
    \SolutionOf{\unkOf{n_1}(\Vars)} & (\pcone=a)\ , \\[\jot]
    \SolutionOf{\unkOf{m_2}(\Vars)} & (\pctwo=p)\ , & 
    \SolutionOf{\unkOf{n_2}(\Vars)} & (\pcone=a)\ .
  \end{array}
\end{equation*} 
%
This assignment uses a predicate $\pctwo=p$ over the local variable of
the second thread as a solution for the abstract state
$\unkOf{m_1}(\Vars)$ in the thread reachability of the first thread.
By collecting and using the corresponding predicates, we will discover
a non-modular proof.

To avoid the drawbacks of non-modular proofs, our algorithm does not
use $\HornClauses_3$ and attempts to find modular predicates for
abstraction refinement instead.
We express the preference for modular predicates declaratively, using
a set of Horn clauses in which the unknown predicates are restricted
to the desired variables, as described in
Sections~\ref{sec-refinement}.
For the abstract states in the first thread, we require that the
corresponding solutions are over the global variable $\lock$ and the
local variable $\pcone$ of the first thread, i.e., we have the unknown
predicates $\unkOf{m_1}(\lock, \pcone)$ and~$\unkOf{m_2} (\lock,
\pcone)$.
Similarly, for the second thread we obtain $\unkOf{n_1}(\lock,
\pctwo)$ and~$\unkOf{n_2}(\lock, \pctwo)$.
Instead of~$\HornClauses_3$,  we use a set of Horn clauses $\HornClauses_4$ shown below:
%
\begin{equation*}
  \HornClauses_4 = \set{
    \begin{array}[t]{@{}l@{}}
       \SymbInit \limplies \unkOf{m_1}(\lock, 
       \pcone),\\[\jot]
       \unkOf{m_1}(\lock, \pcone) \land \rel_1 \limplies \unkOf{m_2}(\lock', \pcone'),\\[\jot]
       \SymbInit \limplies \unkOf{n_1}(\lock, \pctwo),\\[\jot]
       \unkOf{n_1}(\lock, \pctwo) \land \rel_2 \limplies \unkOf{n_2}(\lock', \pctwo'),\\[\jot]
       \unkOf{m_2}(\lock, \pcone) \land \unkOf{n_2}(\lock, \pctwo) \land \SymbError \limplies 
       \lfalse }\ .
  \end{array}
\end{equation*}
%
The conjunction of clauses from $\HornClauses_4$ can be satisfied by
an assignment $\Solution$ such that
%
\begin{equation*}
  \begin{array}[t]{r@{\;=\;}l}
    \SolutionOf{\unkOf{m_1}(\lock, \pcone)} &\ltrue\ , \\[\jot]
    \SolutionOf{\unkOf{m_2}(\lock, \pcone)} &(\lock=1)\ , \\[\jot]
    \SolutionOf{\unkOf{n_1}(\lock, \pctwo)} &\ltrue\ ,\\[\jot]
    \SolutionOf{\unkOf{n_2}(\lock, \pctwo)} &(\lock=2)\ ,
  \end{array}
\end{equation*}
%
which contains only modular predicates.

\input{fig-ex-take-threadid-art-3}

\paragraph{Last \aret computation}

We present the last \aret computation for \TakeLockBit, which uses on
the following (transition) predicates collected so far:
%
\begin{equation*}
  \begin{array}[t]{l@{\arrSpace}l}
    \artPreds_1 &= \set{\pcone=a, \pcone=b, \lock=1}\ ,\\[\jot]
    \artPreds_2 &= \set{\pctwo=p, \pctwo=q, \lock=2}\ ,\\[\jot]
    \artTPreds_{\envFromTo{2}{1}} &= \set{\lock=0}\ ,\\[\jot]
    \artTPreds_{\envFromTo{1}{2}} &= \set{\lock=0}\ .\\[\jot]
  \end{array}
\end{equation*}
%
Figure~\ref{fig-ex-take-threadid-art-3} shows the resulting abstract
reachability and environment trees constructed as follows:
%
\begin{equation*}
  \begin{array}[t]{l@{\;=\;}l@{\;=\;}l}
    m_1 & \abst_1(\SymbInit) &  (\pcone=a) \\[\jot] 
    n_1 & \abst_2(\SymbInit) &  (\pctwo=p) \\[\jot]
    m_2 & \abst_1(\post(\rel_1, m_1)) & (\pcone=b \land \lock = 1)\\[\jot]
    n_2 & \abst_2(\post(\rel_2, n_1)) & (\pctwo=q \land \lock = 2)\\[\jot]
    e_1 & \abstTransTIdOf{1}{2}(m_1 \land \rel_1) & (\lock = 0) \\[\jot]
    e_2 & \abstTransTIdOf{2}{1}(n_1 \land \rel_2) & (\lock = 0)  \\[\jot]
  \end{array}
\end{equation*}
%
The \aret construction is completed, since
\begin{equation*}
\begin{array}[t]{l@{\;\limplies\;}ll@{\;=\;}l@{}}
    \abst_1(\post(e_2 \land \relLocEqOf{1}, m_1)) & m_1\ ,
    & \abst_1(\post(e_2, m_2)) & \lfalse\ ,\\[\jot] 
    \abst_2(\post(e_1 \land \relLocEqOf{2}, n_1)) & n_1\ ,
    &
    \abst_2(\post(e_1, n_2)) & \lfalse\ .
  \end{array}
\end{equation*}
%
By inspecting pairs of abstract states from different trees we
conclude that \TakeLockThreadId is safe.

Furthermore no predicate in $\artPreds_1$ refers to the local variable
of the second thread, the symmetric condition holds for
$\artPreds_2$, and the predicates in $\artTPredsOf{1}{2}$ as well as
$\artTPredsOf{2}{1}$ do not refer to any local variables.
Thus, from the trees in Figure~\ref{fig-ex-take-threadid-art-3} we can
construct a modular safety proof.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/intro.tex
Example?

\section{Introduction}

The ubiquitous availability of parallel computing infrastructures
facilitated by the advent of multicore architectures requires a shift
towards multi-threaded programming to take full advantage of the
available computing resources.
Writing correct multi-threaded software is a difficult task, as the
programmer needs to keep track of a very large number of possible
interactions between the program threads.
Automated program analysis and verification tools can support
programmer in dealing with this challenge by systematically and
exhaustively exploring program behaviours and checking their
correctness.

Direct treatment of all possible thread interleavings by reasoning
about the program globally is a prohibitively expensive task, even for
small programs. 
By applying rely-guarantee techniques, see
e.g.~\cite{RelyGuarantee,OwickiAI76}, such global reasoning can be
avoided by considering each program thread in isolation, using
environment transitions to summarize the effect of executing other
threads, and applying them on the thread at hand.
The success of such an approach depends on the ability to
automatically discover environment transitions that are precise enough
to deliver a conclusive analysis/verification outcome, and yet do not
keep track of unnecessary details in order to avoid sub-optimal
efficiency.

%\todo{Modular proofs play an important role in rely-guarantee
%approaches.}

In this paper we present a method that automates rely-guarantee
reasoning for verifying safety of multi-threaded programs.
Our method relies on an automated discovery of environment transitions
using (transition) predicate abstraction~\cite{GrafSaidi,TransPred05}.
It performs a predicate abstraction-based reachability computation for
each thread and interleaves it with the construction of environment
transitions that over-approximate the effect of executing thread
transitions using transition predicates.
The success of our method crucially depends on an abstraction
refinement procedure that discovers (transition) predicates.
The refinement procedure attempts to minimize the amount of details
that are exposed by the environment transitions, in order to avoid
unnecessary details about thread interaction.

The crux of our refinement approach is in using a declarative
formulation of the abstraction refinement algorithm that can deal with
the thread reachability, environment transitions, and their mutual
dependencies.
We use Horn clauses to describe constraints on the desired
(transition) predicates, and solve these constraints using a general
algorithm for recursion-free Horn clauses.
Our formalization can accommodate additional requirements that express
the preference for modular predicates that do not refer to the local
variables of environment threads, together with the preference for
modular transition predicates that only deal with global variables and
their primed versions.

We implemented the proposed method in a verification tool for
multi-threaded programs and applied it on a range of benchmarks,
which includes fragments of open source software, ticket-based mutual
exclusion protocols, and multi-threaded Linux device drivers.
%We observed that some of the discovered environment assumptions are
%thread-modular, while some others involve reasoning about thread-local
%variables.
The results of the experimental evaluation indicate that our
declarative abstraction refinement approach can be effective in
finding adequate environment transitions for the verification of
multi-threaded programs.

This paper makes the following contributions:
\begin{enumerate}
\item the automatic, rely-guarantee based method for verifying
multi-threaded programs using (transition) predicate abstraction;
%together with a corresponding abstraction refinement algorithm;
\item the novel formulation of abstraction refinement schemes using
Horn clauses, and its application for the (transition) abstraction
discovery for multi-threaded programs;
\item the algorithm for solving recursion-free Horn clauses over
linear arithmetic constraints;
\item the prototype implementation and its evaluation.
\end{enumerate}

The rest of the paper is organized as follows. 
First, we illustrate our method in Section~\ref{sec-illustration}.
In Section~\ref{sec-prelim} we present necessary definitions.
Section~\ref{sec-proof-rule} presents a proof rule that provides a
basis for our method, and shows how the proof rule can be automated
using the connection to fixpoints and abstraction techniques.
We present the main algorithm in Section~\ref{sec-algo-safety}.
Section~\ref{sec-refinement} focuses on the abstraction refinement using
Horn clauses, while Section~\ref{sec-horn-solving} presents a
constraint solving algorithm for Horn clauses over linear
inequalities.
We discuss the experimental evaluation in
Section~\ref{sec-experiments}.
Related work is presented in Section~\ref{sec-related}.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/popl213af-gupta.tex
Example?

\documentclass{sigplanconf}
\sloppy

\usepackage{graphicx,pgf,tikz,tikz-qtree}
\usepackage{macros,ar-alg}
\usepackage{amsmath,amssymb,xspace}
%\usepackage{mathabx} % provides \smalltriangleright
\usepackage{color} % graphicx scaprs ams symbols and we don't use it anyway
%\usepackage{extarrows}
\usetikzlibrary{arrows,automata}
\usepackage{ifthen}

\newcommand{\isTechReport}{true} % true or false
\newcommand{\includeProof}[1]{
  \ifthenelse{\equal{\isTechReport}{true}}{
    #1
  }{
  }
}

\begin{document}

\conferenceinfo{POPL'11,} {January 26--28, 2011, Austin, Texas, USA.}
\CopyrightYear{2011}
\copyrightdata{978-1-4503-0490-0/11/01}

\titlebanner{}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

%\title{Predicate Abstraction and Refinement with Horn Clauses for Verifying Multi-Threaded Programs}
\title{Predicate Abstraction and Refinement for Verifying Multi-Threaded Programs}
%\title{Constraint Based Iterative Inference of Auxiliary Variables 
%  and Relies/Guarantees}
%\title{Constrained Environment Inference for\\ 
% Verification of Multi-Threaded Programs}
%\title{Constraint Based Inference of Auxiliary Variables 
%  and Relies/Guarantees for Verifying Multi-Threaded Programs}
% \subtitle{Draft. Please don't distribute.}

\authorinfo{Ashutosh Gupta \and Corneliu Popeea \and Andrey Rybalchenko}
           {Institut f\"ur Informatik, Technische Universit\"at M\"unchen\\Germany}
           {\{guptaa,popeea,rybal\}@in.tum.de}
%\authorinfo{Corneliu Popeea}{\TUM}{}
%\authorinfo{Andrey Rybalchenko}{\TUM}{}

\maketitle

\category{D.2.4}{Software Engineering}{Software/Program Verification}
\category{F.3.1}{Logics and Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}
\terms
Algorithms, Languages, Reliability, Verification.
\keywords
Multi-threaded programs, safety, proof rule,
modular reasoning, environment transitions, (transition) predicate
abstraction, abstraction refinement, Horn clauses.

% TBD in final version
\vspace{-0.3cm}
\acks

The first author was supported by the DFG Graduiertenkolleg 1480 (PUMA).
We thank Byron Cook, Rusl\'an Ledesma Garza, and Peter O'Hearn for
comments and suggestions.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.
%%%%%%% \bibliography{biblio}

\begin{thebibliography}{34}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Bar-David and Taubenfeld(2003)]{BarDavidDISC03}
Y.~Bar-David and G.~Taubenfeld.
\newblock Automatic discovery of mutual exclusion algorithms.
\newblock In \emph{DISC}, pages 136--150, 2003.

\bibitem[Basler et~al.(2009)Basler, Mazzucchi, Wahl, and Kroening]{BaslerCAV09}
G.~Basler, M.~Mazzucchi, T.~Wahl, and D.~Kroening.
\newblock Symbolic counter abstraction for concurrent software.
\newblock In \emph{CAV}, pages 64--78, 2009.

\bibitem[Basler et~al.(2010)Basler, Hague, Kroening, Ong, Wahl, and
  Zhao]{BaslerTACAS10}
G.~Basler, M.~Hague, D.~Kroening, C.-H.~L. Ong, T.~Wahl, and H.~Zhao.
\newblock Boom: Taking boolean program model checking one step further.
\newblock In \emph{TACAS}, pages 145--149, 2010.

\bibitem[Clarke et~al.(2000)Clarke, Grumberg, Jha, Lu, and Veith]{ClarkeCEGAR}
E.~M. Clarke, O.~Grumberg, S.~Jha, Y.~Lu, and H.~Veith.
\newblock Counterexample-guided abstraction refinement.
\newblock In \emph{CAV}, pages 154--169, 2000.

\bibitem[Cohen and Namjoshi(2009)]{CohenFMSD09}
A.~Cohen and K.~S. Namjoshi.
\newblock Local proofs for global safety properties.
\newblock \emph{FMSD}, 34\penalty0 (2):\penalty0 104--125, 2009.

\bibitem[Corbet et~al.(2005)Corbet, Rubini, and Kroah-Hartman]{CorbetBook05}
J.~Corbet, A.~Rubini, and G.~Kroah-Hartman.
\newblock \emph{Linux Device Drivers, 3rd Edition}.
\newblock O'Reilly Media, Inc., 2005.

\bibitem[Cousot and Cousot(1977)]{lattice77}
P.~Cousot and R.~Cousot.
\newblock Abstract interpretation: A unified lattice model for static analysis
  of programs by construction or approximation of fixpoints.
\newblock In \emph{POPL}, pages 238--252, 1977.

\bibitem[Flanagan and Godefroid(2005)]{FlanaganPOPL05}
C.~Flanagan and P.~Godefroid.
\newblock Dynamic partial-order reduction for model checking software.
\newblock In \emph{POPL}, pages 110--121, 2005.

\bibitem[Flanagan and Qadeer(2003)]{ThreadModularFlanaganQadeer}
C.~Flanagan and S.~Qadeer.
\newblock Thread-modular model checking.
\newblock In \emph{SPIN}, pages 213--224, 2003.

\bibitem[Flanagan et~al.(2002)Flanagan, Freund, and Qadeer]{CalvinESOP}
C.~Flanagan, S.~N. Freund, and S.~Qadeer.
\newblock Thread-modular verification for shared-memory programs.
\newblock In \emph{ESOP}, pages 262--277, 2002.

\bibitem[Godefroid(1994)]{GodefroidPhd94}
P.~Godefroid.
\newblock \emph{Partial-Order Methods for the Verification of Concurrent
  Systems - An Approach to the State-Explosion Problem}.
\newblock PhD thesis, University of Liege, Computer Science Department, 1994.

\bibitem[Graf and Sa\"{\i}di(1997)]{GrafSaidi}
S.~Graf and H.~Sa\"{\i}di.
\newblock Construction of abstract state graphs with {PVS}.
\newblock In \emph{CAV}, pages 72--83, 1997.

\bibitem[Gupta et~al.(2010)Gupta, Popeea, and Rybalchenko]{GuptaATVA10}
A.~Gupta, C.~Popeea, and A.~Rybalchenko.
\newblock Non-monotonic refinement of control abstraction for concurrent
  programs.
\newblock In \emph{ATVA}, pages 188--202, 2010.

\bibitem[Henzinger et~al.(2002)Henzinger, Jhala, Majumdar, and
  Sutre]{BlastLazy02}
T.~A. Henzinger, R.~Jhala, R.~Majumdar, and G.~Sutre.
\newblock Lazy abstraction.
\newblock In \emph{POPL}, pages 58--70, 2002.

\bibitem[Henzinger et~al.(2004)Henzinger, Jhala, and Majumdar]{HenzingerPLDI04}
T.~A. Henzinger, R.~Jhala, and R.~Majumdar.
\newblock Race checking by context inference.
\newblock In \emph{PLDI}, pages 1--13, 2004.

\bibitem[Jones(1983{\natexlab{a}})]{JonesTOPLAS83}
C.~B. Jones.
\newblock Tentative steps toward a development method for interfering programs.
\newblock \emph{ACM Trans. Program. Lang. Syst.}, 5\penalty0 (4):\penalty0
  596--619, 1983{\natexlab{a}}.

\bibitem[Jones(1983{\natexlab{b}})]{RelyGuarantee}
C.~B. Jones.
\newblock Specification and design of (parallel) programs.
\newblock In \emph{IFIP Congress}, pages 321--332, 1983{\natexlab{b}}.

\bibitem[Lamport(1974)]{LamportCACM74}
L.~Lamport.
\newblock A new solution of {D}ijkstra's concurrent programming problem.
\newblock \emph{Commun. ACM}, 17\penalty0 (8):\penalty0 453--455, 1974.

\vfill\eject

\bibitem[Lamport(1987)]{LamportTCS87}
L.~Lamport.
\newblock A fast mutual exclusion algorithm.
\newblock \emph{ACM Trans. Comput. Syst.}, 5\penalty0 (1):\penalty0 1--11,
  1987.

\bibitem[Lu et~al.(2008)Lu, Park, Seo, and Zhou]{LuASPLOS08}
S.~Lu, S.~Park, E.~Seo, and Y.~Zhou.
\newblock Learning from mistakes: a comprehensive study on real world
  concurrency bug characteristics.
\newblock In \emph{ASPLOS}, pages 329--339, 2008.

\bibitem[Malkis et~al.(2006)Malkis, Podelski, and Rybalchenko]{MalkisICTAC06}
A.~Malkis, A.~Podelski, and A.~Rybalchenko.
\newblock Thread-modular verification is cartesian abstract interpretation.
\newblock In \emph{ICTAC}, pages 183--197, 2006.

\bibitem[Manna and Pnueli(1995)]{MannSafetyBook95}
Z.~Manna and A.~Pnueli.
\newblock \emph{Temporal verification of reactive systems: safety}.
\newblock Springer-Verlag, 1995.

\bibitem[McKenney(2007)]{McKenneyLWN07}
P.~McKenney.
\newblock Using {Promela} and {Spin} to verify parallel algorithms.
\newblock \emph{LWN.net weekly edition}, 2007.

\bibitem[Musuvathi et~al.(2008)Musuvathi, Qadeer, Ball, Basler, Nainar, and
  Neamtiu]{MusuvathiOSDI08}
M.~Musuvathi, S.~Qadeer, T.~Ball, G.~Basler, P.~A. Nainar, and I.~Neamtiu.
\newblock Finding and reproducing {Heisenbugs} in concurrent programs.
\newblock In \emph{OSDI}, pages 267--280, 2008.

\bibitem[Necula et~al.(2002)Necula, McPeak, Rahul, and Weimer]{NeculaCC02}
G.~C. Necula, S.~McPeak, S.~P. Rahul, and W.~Weimer.
\newblock {CIL}: Intermediate language and tools for analysis and
  transformation of {C} programs.
\newblock In \emph{CC}, pages 213--228, 2002.

\bibitem[Owicki and Gries(1976)]{OwickiAI76}
S.~S. Owicki and D.~Gries.
\newblock An axiomatic proof technique for parallel programs {I}.
\newblock \emph{Acta Inf.}, 6:\penalty0 319--340, 1976.

\bibitem[Pnueli et~al.(2002)Pnueli, Xu, and Zuck]{PnueliCAV02}
A.~Pnueli, J.~Xu, and L.~D. Zuck.
\newblock Liveness with (0, 1, infty)-counter abstraction.
\newblock In \emph{CAV}, pages 107--122, 2002.

\bibitem[Podelski and Rybalchenko(2005)]{TransPred05}
A.~Podelski and A.~Rybalchenko.
\newblock Transition predicate abstraction and fair termination.
\newblock In \emph{POPL}, pages 132--144, 2005.

\bibitem[Podelski and Rybalchenko(2007)]{RybalPodelskiPADL07}
A.~Podelski and A.~Rybalchenko.
\newblock {ARMC}: The logical choice for software model checking with
  abstraction refinement.
\newblock In \emph{PADL}, pages 245--259, 2007.

\bibitem[Qadeer and Wu(2004)]{QadeerPLDI04}
S.~Qadeer and D.~Wu.
\newblock {KISS}: keep it simple and sequential.
\newblock In \emph{PLDI}, pages 14--24, 2004.

\bibitem[Rybalchenko()]{RybalchenkoARMC}
A.~Rybalchenko.
\newblock {The ARMC tool}.
\newblock \emph{Available from http://www7.in.tum.de/\~{}rybal/armc/}.

\bibitem[Szymanski(1988)]{SzymanskiICS88}
B.~K. Szymanski.
\newblock A simple solution to {L}amport's concurrent programming problem with
  linear wait.
\newblock In \emph{ICS}, pages 621--626, 1988.

\bibitem[{The Intelligent Systems Laboratory}(2001)]{sicstus}
{The Intelligent Systems Laboratory}.
\newblock \emph{SICStus Prolog User's Manual}.
\newblock Swedish Institute of Computer Science, 2001.
\newblock Release 3.8.7.

\bibitem[Wang et~al.(2008)Wang, Yang, Kahlon, and Gupta]{WangTACAS08}
C.~Wang, Z.~Yang, V.~Kahlon, and A.~Gupta.
\newblock Peephole partial order reduction.
\newblock In \emph{TACAS}, pages 382--396, 2008.

\end{thebibliography}

\iffalse
\appendix
% \input{auxiliaries}
% \input{redundancy}
% \input{algorithms}
% \input{tables}

\clearpage
\input{fig-alg-main}
\input{fig-alg-abst-reach}
\input{fig-alg-cex-path}
\input{fig-alg-refine}
\input{fig-alg-mkHornClauses}
\input{fig-alg-solveHorn}
\fi
%\clearpage
%\input{fig-ex-mk-tree}

\end{document}

### 2014-02-06 paper/example-latex-threads/prelim-symb.tex
Example?

\section{Preliminaries}\label{sec-prelim}

In this section we briefly describe multi-threaded programs, their
computations and correctness. 
We also introduce auxiliary definitions that we apply for reasoning
about programs.

\paragraph{Programs}

% = \program_1 \parcomp \dots \parcomp \program_\Num

We consider a \emph{multi-threaded} program $\program$ that consists
of $\Num\geq 1$ concurrent threads.
Let $\ThreadIds$ be the set~$\set{1, \dots, \Num}\ $.
We assume that the \emph{program} variables $\Vars = (\GlobVars,
\LocVarsOf{1}, \dots, \LocVarsOf{\Num})$ are partitioned into
\emph{global} variables $\GlobVars$ that are shared by all threads,
and \emph{local} variables $\LocVarsOf{1}, \dots, \LocVarsOf{\Num}$
that are only accessible by the threads $1, \dots, \Num\ $,
respectively.

The set of \emph{global states} $\Glob$ consists of the valuations of
global variables, and the sets of \emph{local states} $\Loc_1, \dots,
\Loc_\Num$ consist of the valuations of the local variables of
respective threads.
By taking the product of the global and local state spaces, we obtain
the set of \emph{program states} $\States = \Glob \times \Loc_1 \times
\dots \times \Loc_\Num\ $.
We represent sets of program states using assertions over program
variables. 
Binary relations between sets of program states are represented using
assertions over unprimed and primed variables.
Let $\models$ denote the satisfaction relation between (pairs) of
states and assertions.

The set of \emph{initial} program states is denoted by $\SymbInit\ $,
and the set of \emph{error} states is denoted by~$\SymbError\ $.
For each thread $i\in\ThreadIds$ we have a finite set of transition
relations~$\Trans_i\ $, which are abbreviated as \emph{transitions}.
Each transition $\rel \in \Trans_i$ can change the values of the
global variables and the local variables of the thread~$i\ $. 
Let $\relLocEqOf{i}$ be a constraint requiring that the local
variables of the thread $i$ do not change, i.e., $\relLocEqOf{i} =
(\LocVarsOf{i} = \LocVarsOf{i}')\ $.
Then, $\rel \in\Trans_i$ has the form
%
\begin{equation*}
  \rel^{\text{update}}(\GlobVars, \LocVarsOf{i}, \GlobVars', \LocVarsOf{i}') \land
  \Land_{j\in\ThreadIds\setminus{\set{i}}} \relLocEqOf{j}\ ,
\end{equation*}
%
where the first conjunct represents the update of the variables in the
scope of the thread $i$ and the remaining conjuncts ensure that the
local variables of other threads do not change.
%\todo{add notation for update and eq part?}
We write $\relTIdOf{i}$ for the union of the transitions of the thread
$i\ $, i.e., $\relTIdOf{i} = \Lor \Trans_i\ $.
The transition relation of the program is $\relProg =
\relTIdOf{1} \lor \dots \lor \relTIdOf{\Num}\ $.

\paragraph{Computations}

A \emph{computation} of $\program$ is a sequence of program states
$s_1, s_2, \dots$ such that $s_1$ is an initial state, i.e., $s_1
\models \SymbInit\ $, and each pair of consecutive states $s_i$ and
$s_{i+1}$ in the sequence is connected by some transition $\rel$ from
a program thread, i.e., $(s_i, s_{i+1})\models \rel\ $.
A \emph{path} is a sequence of transitions.
We write $\emptypath$ for the \emph{empty} sequence.

Let $\substXtoY{\FreshVars}{\FreeVars}$ be a substitution function
such that $\varphi\substXtoY{\FreshVars}{\FreeVars}$ replaces
$\FreeVars$ by $\FreshVars$ in~$\varphi\ $.
Let $\comp$ be the \emph{relational composition} function for binary
relations given by assertions over unprimed and primed variables such
that for assertions $\varphi$ and $\psi$ we have
%
$
\varphi \comp \psi \defeq
\exists \Vars'' : \varphi\substXtoY{\Vars''}{\Vars'} \land \psi\substXtoY{\Vars''}{\Vars} \ .
$
%
Then, a \emph{path relation} is a relational composition of transition
relations along the path, i.e., for $\path = \rho_1 \cdots \rho_n$ we
have $\rho_\path = \rho_1 \comp \dots \comp \rho_n\ $. 
A path $\path$ is \emph{feasible} if its path relation is not empty,
i.e.,~$\exists \Vars, \Vars' : \rho_\path\ $.

A program state is \emph{reachable} if it appears in some computation.
Let $\SymbReach$ denote the set of reachable states.
The program is \emph{safe} if \emph{none} of its error states is
reachable, i.e., $\SymbReach \land \SymbError \limplies \lfalse\ $.

\paragraph{Auxiliary definitions}

We define a \emph{successor} function $\post$ such that for a
binary relation over states $\rel$ and a set of states $\varphi$ we
have
%
\begin{equation*}
  \post(\rel, \varphi) \defeq 
  \exists \Vars'': \varphi\substXtoY{\Vars''}{\Vars} \land \rel\substXtoY{\Vars''}{\Vars}\substXtoY{\Vars}{\Vars'}\ .
\end{equation*}
%
We also extend the logical implication to tuples of equal length, i.e.,
%
\begin{equation*}
  (\varphi_1, \dots, \varphi_n) \limplies (\psi_1, \dots, \psi_n) \defeq
  \varphi_1 \limplies \phi_1 \land \dots \land
  \varphi_n \limplies \phi_n\ ,
\end{equation*}
%
where each implication is implicitly universally quantified over the
free variables occurring in it.
From now on, we assume that tuples of assertions are partially ordered
by the above extension of~$\limplies\ $.

A \emph{Horn clause} 
%
$%\begin{equation*}
  \HornBody_1(\FreeVars_1)\land \dots\land \HornBody_n(\FreeVars_n)
  \rightarrow \HornBody(\FreeVars)\ 
$%\end{equation*} 
% 
consists of relation symbols $\HornBody_1$,\dots, $\HornBody_n$,
$\HornBody$, and vectors of variables $\FreeVars_1$,\dots,
$\FreeVars_n$,~$\FreeVars$.
\iffalse
The left-hand side of the implication is called the \emph{body} of the
clause and right-hand side is called the \emph{head}.
\fi 
For the algorithm \algSolveLinearHornClauses in
Section~\ref{sec-horn-solving} we only consider Horn clauses over linear
arithmetic.
We say that $\HornBody$ \emph{depends} on the relation symbols
$\set{\HornBody_i \mid i\in\betweenOf{1}{n}\land \HornBody_i \neq
(\leq)}\ $.
A set of Horn clauses is \emph{recursion-free} if the transitive
closure of the corresponding dependency relation is well founded.

\iffalse
We reason about program safety by computing approximations of
reachable states and relations between reachable states. 
The following auxiliary notions provide fundamental building blocks
for the development of algorithms for program
verification~\cite{Floyd,lattice77,CousotAPCT84}.

The set of reachable program states is the least fixpoint of $\post$
applied on the program transition relation that is above the initial
program states, i.e.,
%
\begin{equation*}
  \SymbReach \lequivalent \lfp(\lambda
  \varphi.\post(
  \relProg, \varphi),
  \SymbInit)\ .
\end{equation*}
%

\todo{Move/drop this?}
Unfortunately, the precise least fixpoint computation is not practical
for programs with very large or infinite state spaces.
To obtain a practical solution, we can apply abstraction to achieve
approximate, yet sound results.
We formalize abstraction as over-approximation following the framework
of abstract interpretation~\cite{lattice77,CousotAPCT84}.
For this purpose, we will use abstraction functions $\abst$ and
$\abstTrans$ that over-approximate sets and binary relations over
programs states, respectively.

\fi

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-02-06 paper/example-latex-threads/related.tex
Example?

\section{Related work}\label{sec-related}

The main inspiration for our work draws from the rely-guarantee
reasoning method \cite{JonesTOPLAS83, RelyGuarantee} and automatic
abstraction refinement approach to verification \cite{ClarkeCEGAR}.

The seminal work on rely-guarantee reasoning \cite{JonesTOPLAS83,
  RelyGuarantee} initially offered an approach to reason about
multi-threaded programs by making explicit the interference between
threads.
Subsequently, rely-guarantee reasoning was used to tackle the problem
of state explosion in verification of multi-threaded programs.
Rely-guarantee reasoning was mechanized and firstly implemented in the 
Calvin model checker \cite{CalvinESOP} for Java shared-memory programs.
Calvin reduces the verification of the multi-threaded program to the
verification of several sequential programs with the help of a
programmer specified environment assumption.
In \cite{ThreadModularFlanaganQadeer}, thread-modular model checking
was proposed to infer automatically environment assumptions that
propagate only global variable changes to other threads. 
The algorithm has low complexity, polynomial in the number of threads,
but is incomplete and fails to discover environment assumptions that
refer to the local states of a thread.
Thread-modular verification is formalized by \cite{MalkisICTAC06} in
the framework of abstract interpretation as Cartesian product of sets
of states.

% we - weakest environment and refinement by strengthening
% RR - strongest environment and refinement by weakening

The method of \cite{HenzingerPLDI04} uses a richer abstraction scheme
that computes contextual thread reachability, where the context in
which a thread executes includes information on both global and local
states of threads.
The context (or environment) is computed using bisimilarity quotients
in steps that are interleaved with abstract reachability computations.
The verification starts with the strongest possible environment
assumption and, by refinement, the environment is weakened until
it over-approximates the transitions of the other threads.
In contrast, our approach refines iteratively the environment based on
over-approximation, starting with the weakest environment and
strengthening it at every iteration. 
% More importantly, our algorithm integrates the computation of
% environment assumptions with the abstract state reachability.
For abstraction refinement, a counterexample from
\cite{HenzingerPLDI04} is reduced to a concrete sequential path by
replacing environment transitions with their corresponding local
transitions.

The approach of \cite{CohenFMSD09} presents another
solution to overcome the incompleteness of local reasoning.
Guided by counterexamples, it refines the abstraction by exposing 
a local variable of a thread as a global variable. 
This refinement recovers the completeness of reasoning, but is
applicable to finite-state systems and may compute an unnecessarily
precise abstraction.
In contrast, our refinement procedure relies on interpolation and
includes predicates on local variables as needed during verification.

Another approach to overcome the state explosion problem of monolithic
reasoning over multi-threaded programs is to translate the
multi-threaded program to a sequential program assuming a bound on the
number of context switches.
This scheme was initially proposed and implemented in KISS
\cite{QadeerPLDI04}, a multi-threaded checker for C programs, and later
evolved to handle and reproduce even difficult to find Heisenbugs
\cite{MusuvathiOSDI08}.
Monolithic reasoning can be greatly facilitated by using techniques
evolved from partial-order reduction \cite{GodefroidPhd94}, like
dynamic partial-order reduction \cite{FlanaganPOPL05} or peephole
partial order reduction \cite{WangTACAS08}.
Yet another technique to fight state explosion is to factor out
redundancy due to thread replication as proposed in counter
abstraction \cite{PnueliCAV02} and implemented in the 
model checker Boom \cite{BaslerCAV09,BaslerTACAS10}.
% shape analysis for concurrent programs: 
% GotsmanPLDI07, RGSep-Action-Inference VMCAI10, HongseokYang-RCU-unpublished
We view these techniques as paramount in obtaining practical
multi-threaded verifiers, but orthogonal to our proposal for automatic
environment inference.

% Future work:
% - distribute the model checking algorithm: as in DArmc-VMCAI'11. Even more natural for
% compositional model checking algorithms like ours (as also suggested in Cohen-EC'09).
% - prove that multi-threaded programs eventually do something good Cook-POPL'07

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 tests/lamports-alg/input-petri-net.pl
Updated property in example

@@ -27,4 +27,4 @@ init(q1).
 init(nb1).
 init(nb2).
 
-target(1, [([p3, q5], 2)]).
+target(1, [([p3], 1), ([q5], 1)]).

### 2014-03-17 paper/workspace/appendix.tex
Added submitted version of paper

\section{Proof of Theorem \ref{thm:dual}}

We use the following version of Farkas' Lemma:

\begin{theorem}[Farkas' lemma (variant) \cite{Schrijver86}]
\label{lem:farkas}
Let $A$ be a matrix and let $b$ be a vector.
Then the system $Ax \le b$ of linear inequalities has a solution $x$,
if and only if $yb \ge 0$ for each row vector $y \ge 0$ with $yA = 0$.
\end{theorem}

\begin{reftheorem}{thm:dual}
The system $\mathcal{S}'$ of constraints given by
\begin{align*}
  \lambda C    & \le 0            && \text{inductivity constraint} \\
  \lambda m_0  & <   Y_1 b + Y_2 1 && \text{safety constraint} \\
  \lambda      & \ge Y_1 A + Y_2 D && \text{property constraint} \\
  Y_1, Y_2     & \ge 0     && \text{non-negativity constraint} 
\end{align*}
is feasible over the rational numbers if{}f the system $\mathcal{S}$ is infeasible.
\end{reftheorem}
\begin{proof}
Recall that the system $\mathcal{S}$ is given by:
\begin{align*}
  \mathcal{C}&(P, T, F, m_0)  && \text{state constraints} \\
  A M & \ge b                 && \text{negation of property $P$} \\
  D M & \ge 1                 && \text{trap constraints}
\end{align*}
Eliminating $M$ we get
\begin{align*}
A (m_0 + C X) & \ge b && \text{property $P$ negated} \\
D (m_0 + C X) & \ge 1 && \text{trap constraints} \\
   m_0 + C X  & \ge 0 && \text{non-negativity conditions for places} \\
           X  & \ge 0 && \text{non-negativity conditions for transitions}
\end{align*}
which can be rewritten in matrix form as
%
\begin{align*}
\begin{pmatrix}
  -  A  C \\
  -  D  C \\
  -     C \\
  -     I
\end{pmatrix} X & \le
\begin{pmatrix}
  A m_0 - b \\
  D m_0 - 1 \\
    m_0     \\
     0
\end{pmatrix}
\end{align*}
%
Lemma \ref{lem:farkas} yields the set $\mathcal{S}_1$ given by  
%
\begin{align*}
  Y^T
\begin{pmatrix}
  -  A  C \\
  -  D  C \\
  -     C \\
  -     I
\end{pmatrix} & = 0  &
  Y^T
\begin{pmatrix}
  A m_0 - b \\
  D m_0 - 1 \\
    m_0     \\
     0
\end{pmatrix} & < 0  &
Y & \ge 0
\end{align*}
So, by the Lemma, $\mathcal{S}_1$ is satisfiable if and only if $\mathcal{S}$ is unsatisfiable. 
Splitting the vector $Y$ into four segments $Y := (Y_1|Y_2|Y_3|Y_4)$ 
we obtain that $\mathcal{S}_1$ is equisatisfiable with 
\begin{align*}
  Y_1 A C         + Y_2 D C         + Y_3 C  + Y_4    & = 0 \\
  Y_1 (A m_0 - b) + Y_2 (D m_0 - 1) + Y_3 m_0         & < 0 \\
  Y_i                               & \ge 0 \ \ i=1,2,3,4
\end{align*}
and, since $Y_4\ge 0$ is only requiring that $Y_1 D C + Y_2 A C + Y_3 C \leq 0$ holds,
also with 
\begin{align*}
  (Y_1 A + Y_2 D + Y_3) C    & \le 0          \\
  (Y_1 A + Y_2 D + Y_3) m_0  & < Y_1 b + Y_2 1 \\
  Y_1,Y_2,Y_3 & \ge 0 
\end{align*}
Finally, defining $\lambda := Y_1 A + Y_2 D + Y_3$, we obtain that
$\mathcal{S}_1$ is equisatisfiable with $\mathcal{S}'$:
\begin{align*}
  \lambda C    & \le 0             \\
  \lambda m_0  & <   Y_1 b + Y_2 1  \\
  \lambda      & \ge Y_1 A + Y_2 D  \\
  Y_1, Y_2        & \ge 0    
\end{align*}
\end{proof}

### 2014-03-17 paper/workspace/fig_safety_by_refinement.tex
Added submitted version of paper

\begin{figure}[t]
  \begin{algorithmic}[1]
    \State $C := \mathcal C(N)$
    \While {SAT$(C \cup \{\neg P\})$}
    \State $A := $Model$(C \cup \{\neg P\})$
    \State $C_\theta := $TrapConditions$(N, A)$
    \If {SAT$(C_\theta)$}
    \State $A_\theta := $Model$(C_\theta)$
    \State $\delta := \Delta(A_\theta)$
    \State $C := C \cup \{\delta\}$
    \Else
    \State \Return 'The petri net may not satisfy the property'
    \EndIf
    \EndWhile
    \State \Return 'The petri net satisfies the property!'
  \end{algorithmic}
  \caption{Pseudocode for Function Safety by Refinement}
  \label{fig_function_safety_by_refinement_pseudocode}
\end{figure}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 paper/workspace/main.tex
Added submitted version of paper

@@
 \ifthenelse{\equal{\isDraft}{false}}{
-  \title{An SMT-based Approach to Coverability Analysis}
+  \title{An SMT-based Approach to\\
+         Coverability Analysis}
   \author{
     Javier~Esparza\inst{1}
+    \and
+    Rusl\'{a}n~Ledesma-Garza\inst{1}
     \and
@@
     Filip~Niksic\inst{2}
-    \and
-    Rusl\'{a}n~Ledesma-Garza\inst{1}
   }
@@
     \and
-    MPI-SWS,~Kaiserslautern~and~Saarbr\"ucken
+    Max Planck Institute for Software Systems (MPI-SWS)
   }
@@
     successfully in recent years to verify safety properties of concurrent
-    shared-memory or asynchronous message-passing software. The bottleneck
-    in the verification process is the iterative computation of the
-    (backward) reachable or (forward) coverable set, and most recent
-    research has focused on non-trivial heuristics to speed it up.
-
-    In this paper, we revisit a constraint approach to the problem based on
-    classical Petri net analysis techniques: marking equation, place
-    invariants, and traps. Given a marking M to be covered, the approach
-    constructs constraints corresponding to increasingly precise
-    overapproximations of the reachable markings covering M. If one of
-    these constraints is unsatisfiable, then M is not coverable.
-
-    Previous work on this approach suffered from the absence of efficient
-    decision procedures for linear arithmetic, which nowadays exist as a
-    part of SMT solvers. We show how to utilize an SMT solver to implement
+    shared-memory or asynchronous message-passing software. We revisit 
+    a constraint approach to the problem based on
+    classical Petri net analysis techniques. We show how to 
+    utilize an SMT solver to implement
     the constraint approach, and additionally, to generate a minimized
-    inductive invariant from a safety proof.
-
-    We empirically evaluate our procedure on a large set of existing Petri
-    net benchmarks arising out of software verification. Surprisingly, even
-    though our technique is incomplete, we show that it can quickly
-    discharge all safe instances, usually competitive with, and occasionally
-    much faster than, state-of-the-art techniques. Additionally, the
+    inductive invariant from a safety proof. We empirically evaluate 
+    our procedure on a large set of existing Petri
+    net benchmarks. Even though our technique is incomplete, 
+    it can quickly discharge all safe instances. Additionally, the
     inductive invariants computed are usually orders of magnitude smaller
@@
 \input{sec_introduction}
-\input{sec_illustration}
+\input{sec_rel_work}
 \input{sec_preliminaries}
-\newpage
-\input{sec_constraint_approach}
-\newpage
-\input{sec_invariants}
-\newpage
+\input{sec_safety}
+\input{sec_safety_by_refinement}
+%\input{sec_invariant}
+\input{sec_invariant_by_refinement}
+\input{sec_invariant_w_minimization}
 \input{sec_experiments}
-\input{sec_rel_work}
-\input{sec_conclusions}
+% \input{sec_conclusions}
 
@@
 
+\newpage
+\appendix
+\input{appendix}
+
 \end{document}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2014-03-17 paper/workspace/sec_experiments.tex
Added submitted version of paper

@@
-\section{Experiments}
+\section{Experimental Evaluation}
+\label{sec_experiments}
 
-To evaluate the described methods, we implemented them in a tool called
-\emph{Petrinizer}. Petrinizer is essentially a set of Bash and Prolog scripts
-that mediate between the input, specified in the input format of
-MIST2\footnote{\url{https://github.com/pierreganty/mist}}, and the SMT
-solver Z3\footnote{\url{http://z3.codeplex.com/}} \cite{DeMouraTACAS08}. With
-the choice of Bash and Prolog over languages with bindings for Z3 API, we
-restricted Z3's full power. Indeed, by using Z3 as an external tool, we
-were not able to fully exploit the incremental nature of its linear arithmetic
-solver (TODO: cite appropriate papers). However, even the non-optimal
-implementation turned out to be robust and scalable, with the main advantage of
-being able to tackle problem instances that are out of reach for
-state-of-the-art coverability checkers.
-
-The evaluation of Petrinizer had three main goals. First, we wanted to compare
-its performance against state-of-the-art tools like MIST2, BFC
-and IIC (cite!). Second, as the method it
-implements is incomplete, we wanted to measure its rate of success on safe problem instances. As a
-subgoal, we wanted to investigate the usefulness and necessity of traps, as
-well as to what extent real arithmetic suffices over integer arithmetic in
-proving safety. And last, since the language of linear arithmetic, used by
-Petrinizer, allows for a more succint representation of formulas than the
-language used by IIC, we wanted to compare sizes of invariants the two tools produce.
-
-Inputs that were used in tests come from various sources (TODO: write a
-sentence to justify the variety).
-One source is the collection of Petri nets from the literature that is part of the MIST2 toolkit.
-The second source are Petri nets originating from the analysis of concurrent C
-programs that were used in the evaluation of BFC \cite{KaiserCONCUR12}. Then
-there are inputs originating from the provenance analysis of a medical and a
-bug-tracking system \cite{MajumdarSAS13}. Finally, we generated a set of Petri
-nets from Erlang programs, using an Erlang verification tool Soter
-\cite{DOsualdoSAS13}. (TODO: Acknowledge Emanuele D'Osualdo for help.)
-
-All experiments were performed on the identical machines, equipped with
-Intel Xeon 2.66 GHz CPUs and 48 GB of memory, running Linux 3.2.48.1 in 64-bit
-mode. Execution time was limited to 100,000 seconds (27 hours, 46 minutes and
-40 seconds), and memory to 2 GB.
-
-\subsection{Performance}
-
-Explain what we observe:
-\begin{itemize}
-  \item On small examples IIC mostly outperforms Petrinizer. Probably due
-    to a slower parser, setting up and invoking Z3.
-  \item On examples of a similar size, coming from the same source,
-    unlike with IIC, resource consumption is fairly uniform. (Medical examples.)
-  \item Petrinizer is able to handle some huge examples that are out of
-    scope for IIC.
-  \item There were no cases when IIC was able to finish, but Petrinizer ran
-    out of resources. As a matter of fact, Petrinizer finishes in all cases,
-    whereas other tools somethimes fail.
-\end{itemize}
-
-\emph{Invariant sizes.} Again, explain how we measure it. For IIC, we measure
-the number of non-null literals, the number of clauses, total length of
-the invariant in characters. For Petrinizer, we measure the number of
-non-null coefficients in the linear expressions and the total length of the
-invariant in characters. Argue why it makes sense to compare any of these
-quantities. Point out the difference in size with and without the invariant
+We have implemented our algorithms in a tool called
+\emph{Pe\-tri\-ni\-zer}. 
+Petrinizer is implemented as a combination of Bash and Prolog scripts,
+and uses Z3 \cite{MouraB08} as a backend SMT solver.
+It takes as input coverability problem instances encoded in the MIST
+input format\footnote{\url{https://github.com/pierreganty/mist}}, and it
+runs one of the selected methods. 
+We implemented all possible combinations of
+methods: with and without trap refinement, with real and integer arithmetic,
+with and without invariant construction, with and without invariant
 minimization.
 
-\emph{Rate of success.} Point out that Petrinizer mostly successfully proves
-safety. Specifically, it proves safety in all of the safe Soter examples.
-Point out the cases where traps were actually useful, and potentially progress
-into some pseudo-philosophical discussion on what constitutes a good set of examples.
+The evaluation of Petrinizer had two main goals. First, as the underlying
+methods are incomplete, we wanted to measure their success rate on standard
+benchmark sets. 
+As a subgoal, we wanted to investigate
+the usefulness and necessity of traps, the benefit of using integer arithmetic
+over real arithmetic, and the sizes of the constructed invariants.
+The second goal was to measure Petrinizer's performance and to compare it
+with state-of-the-art tools: IIC \cite{KloosMNP13}, BFC\footnote{The most recent version of BFC at the time of
+writing the paper was 2.0. However, we noticed it sometimes reports
+inconsistent results, so we used version 1.0 instead. The tool can be
+obtained at \url{http://www.cprover.org/bfc/}.} \cite{KaiserKWCONCUR12}, and
+MIST\footnote{MIST consists of several methods, most of them based on
+  EEC \cite{GeeraertsRB06}. We used the abstraction
+  refinement method that tries to minimize the number of places in the net \cite{GantyFI08}.}.
+
+\paragraph{Benchmarks.} For the inputs used in the experiments, we collected a
+large number of coverability problem instances originating from various sources. The
+collection contains 178 examples, out of which 115 are safe, and is
+organized into five example suites. The first suite is a collection of
+Petri net examples from the MIST toolkit. This suite contains a
+mixture of 29 examples, both safe and unsafe. It contains both real-world and
+artificially created examples. The second suite consists of 46 Petri nets
+that were used in the evaluation of BFC \cite{KaiserKWCONCUR12}. They
+originate from the analysis of concurrent C programs, and they are mostly
+unsafe. The third and the forth suite come from the
+provenance analysis of messages in a medical and a bug-tracking system
+\cite{MajumdarSAS13}. The medical suite contains 12 safe
+examples, and the bug-tracking suite contains 41 examples, all
+safe except for one. The final, fifth suite contains 50 examples that
+come from the analysis of Erlang programs \cite{DOsualdoSAS13}. We generated
+them ourselves using an Erlang verification tool Soter \cite{DOsualdoSAS13},
+from the example programs found on Soter's
+website\footnote{http://mjolnir.cs.ox.ac.uk/soter/}. Out of 50 examples in this
+suite, 38 are safe. This suite also contains the largest example in the
+collection, with 66,950 places and 213,635 transitions.
+For our evaluation, only the 115 safe instances are interesting.
+
+\begin{table}[t]
+  \centering
+  \caption{Safe examples that were successfully proved safe.}
+  \label{table:rate-of-success}
+  \scriptsize
+  \begin{tabular}{|l|r|r|r|r|r|r|r|}
+    \hline
+    \multicolumn{1}{|c|}{Suite} & \multicolumn{1}{|c|}{Safety}
+    & \multicolumn{1}{|c|}{Ref./reals} & \multicolumn{1}{|c|}{Ref./ints}
+    & \multicolumn{1}{|c|}{IIC} & \multicolumn{1}{|c|}{BFC}
+    & \multicolumn{1}{|c|}{MIST} & \multicolumn{1}{|c|}{Total} \\
+    \hline
+    MIST        & 14 & 20 & 20 & {\bf 23} & 21 & 19 & 23 \\
+    BFC          & {\bf 2} & {\bf 2} & {\bf 2} & {\bf 2} & {\bf 2} & {\bf 2} & 2 \\
+    Medical      & 4 & 4 & 4 & 9 & {\bf 12} & 10 & 12 \\
+    Bug-tracking & {\bf 32} & {\bf 32} & {\bf 32} & 0 & - & 0 & 40 \\
+    Erlang       & 32 & 36 & {\bf 38} & 17 & 26 & 2 & 38 \\
+    \hline
+    Total        & 84 & 94 & {\bf 96} & 51 & 61 & 33 & 115 \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\paragraph{Rate of success on safe examples.} As shown in Table
+\ref{table:rate-of-success}, even with the weakest of the methods---safety based
+on marking equation over reals---Petrinizer is able to prove safety for 84 out
+of 115 examples. Switching to integer arithmetic does not help: The number of
+examples proved safe remains 84. Using refinement via traps, Petrinizer proves
+safety for 94 examples. Switching to integer arithmetic in this case helps:
+Another two examples are proved safe, totaling with 96 out of 115 examples. In
+contrast to these numbers, the most successful existing tool turned out to be BFC, proving safety for
+only 61 examples\footnote{This number comes with a caveat: We were not able to
+run BFC on the bug-tracking suite, as the examples were too big for the tool that translates
+MIST format into BFC format. For the BFC and Erlang suites, which contain
+some similarly large or even larger examples, we initially had the inputs in BFC
+format, and translated them into MIST format, thus avoiding the issue. Translation
+utilities are provided by the authors of BFC.}. The small numbers are due to
+time and memory limits. Even though the methods these tools implement
+are complete, the tools themselves are limited by the time and space they can
+use.
+
+Looking at the results accross different suites, we see that Petrinizer
+performed poorest on the medical suite, proving safety for only 4 out of 12
+examples. On the other hand, on the bug-tracking suite, which was completely
+intractable for other tools, it proved safety for 32 out of 40 examples.
+Furthermore, using traps and integer arithmetic, Petrinizer successfuly proved
+safety for all safe Erlang examples. We find this result particularly
+surprising, as the original verification problems for these examples seem non-trivial.
+
+\begin{figure}[t]
+  \centering
+  \input{fig_invariant_size}
+  \caption{Graph on the left shows a relation of sizes of constructed
+  invariants to the number of places in the corresponding Petri nets. Graph
+  on the right shows comparison in size of invariants produced by Petrinizer and
+  IIC. Axes represent size on a logarithmic scale. Each dot represents one example.}
+  \label{fig:invariant-size}
+\end{figure}
+
+\paragraph{Invariant sizes.} As the invariant construction is a new addition to
+the existing methods, we wanted to inspect sizes of the constructed invariants.
+We took the number of atomic terms appearing in an invariant's linear expressions
+as a measure of its size. When we relate sizes of invariants to number of places
+in the corresponding Petri net (left graph in Fig.~\ref{fig:invariant-size}), we
+see that invariants are most of the times very succinct. As an example, the
+largest invariant had 814 atomic terms, and the corresponding Petri net, coming
+from the Erlang suite, had 4,763 places. For the largest Petri net in the
+collection, the one with 66,950 places, the constructed invariant had 339 atomic
+terms.
+
+We also compared sizes of constructed invariants with
+sizes of invariants produced by IIC \cite{KloosMNP13}. IIC's invariants
+are expressed as CNF formulas over atoms of the form $x<a$,
+for a variable $x$ and a constant $a$. As a measure of size for these formulas,
+we took the number of atoms they contain. As the right graph in
+Fig.~\ref{fig:invariant-size} shows, when compared to IIC's invariants, Petrinizer's invariants
+are never larger, and are often orders of magnitude smaller.
+
+\paragraph{Performance.} To ensure accuracy and fairness, all experiments were
+performed on identical machines, equipped with Intel Xeon 2.66 GHz CPUs and 48 GB of memory, running Linux 3.2.48.1
+in 64-bit mode. Execution time was limited to 100,000 seconds (27
+hours, 46 minutes and 40 seconds), and memory to 2 GB.
+
+Due to dissimilarities between the compared tools, selecting a fair
+measure of time turned out to be tricky. On the one hand, as
+Petrinizer communicates with Z3 via temporary files, it spends a
+considerable amount of time doing I/O operations.  On the other hand,
+as BFC performs both a forward and a backward search, it naturally
+splits the work into two threads, and runs them in parallel on two CPU
+cores. In both cases, the actual elapsed time does not quite
+correspond to the amount of computational effort we wanted to measure.
+Therefore, for the measure of time we selected the \emph{user time},
+as reported by the \emph{time} utility on Linux. User time measures
+the total CPU time spent executing the process and its children. In
+the case of Petrinizer, it excludes the I/O overhead, and in the case
+of BFC, it includes total CPU time spent on both CPU cores.
+
+\begin{figure}[t]
+  \centering
+  \input{fig_performance_various}
+  \caption{Time overhead of trap refinement, invariant
+  construction and invariant minimization. Axes represent time in seconds on a
+  logarithmic scale. Each dot represents
+  execution time on one example. The graph on the left
+  only shows examples for which at least one trap appeared in the refinement.
+  Similarly, the other two graphs only show safe examples.}
+  \label{fig:performance-various}
+\end{figure}
+
+\paragraph{Time overhead of Petrinizer's methods.} Before comparing Petrinizer
+with other tools, we analyze time overhead of trap refinement, invariant construction, and invariant minimization.
+Fig.~\ref{fig:performance-various} summarizes the results.
+The leftmost graph in Fig.~\ref{fig:performance-various} shows that traps
+incur a significant overhead. This is not too surprising as, each time
+a trap is found, the main system has to be updated with a new trap
+constraint and solved again. Thus the actual overhead depends on the number
+of traps that appear during the refinement. In the experiments, there
+were 32 examples for refinement with integer arithmetic where traps
+appeared at least once. The maximal number of traps in a single example was 9.
+
+In the case of invariant construction, as shown on the middle graph in
+Fig.~\ref{fig:performance-various}, the overhead is more
+uniform and predictable. The reason is that constructing the invariant
+involves solving the dual form of the main system as many times as
+there are disjuncts in the property violation constraint. In most
+cases, the property violation constraint has one disjunt. A single example
+with many disjuncts, having 8989 of them, appears on the graph
+as an outlier.
+
+In the case of invariant minimization, as the rightmost
+graph in Fig.~\ref{fig:performance-various} shows, time overhead is
+quite severe.
+The underlying data contains examples of slowdowns of up to 30$\times$.
+At the same time, the added benefit of minimization is negligible: There were only four examples where the invariant
+was reduced, and that was for mere 2-3\%. 
+Therefore, invariant minimization does not seem to pay off for these examples.
+
+\begin{figure}[t]
+  \centering \input{fig_performance_vs_tools}
+  \caption{Comparison of execution time for Petrinizer vs.~IIC,
+  BFC and MIST. Graphs in the top row show comparison in the case without
+  invariant construction, and graphs in the bottom row show comparison in
+  the case with invariant construction. Axes represent time in seconds on a
+  logarithmic scale. Each dot represents execution time on one example.}
+  \label{fig:performance-vs-tools}
+\end{figure}
+
+\paragraph{Comparison with other tools.} Six graphs in
+Fig.~\ref{fig:performance-vs-tools} show the comparison of execution time for Petrinizer vs.~IIC, BFC, and MIST. 
+In the comparison, we used
+the refinement methods, both with and without invariant construction.
+In general, we observe that other tools outperform
+Petrinizer on small examples, an effect that can be explained by the
+overhead of starting Bash and Prolog interpreters and Z3. 
+However, on large examples Petrinizer consistently outperforms other tools.
+Not only does it finish in all cases within the given time and memory
+constraints, it even finishes in under 100
+seconds in all but two cases. The two cases are the
+large example from the Erlang suite, with 66,950 places and 213,635
+transitions and, in the case of invariant construction, the example
+from the MIST suite, with 8989 disjuncts in the property violation constraint.
+
+\paragraph{Conclusions.}
+%
+Marking equations and traps are classical techniques in Petri net theory, but have fallen out of favor
+in recent times in comparison with state-space traversal techniques in combination with abstractions
+or symbolic representations.
+Our experiments demonstrate that, when combined with the power of a modern SMT solver, these techniques
+can be surprisingly effective in finding proofs of correctness (inductive invariants) of common benchmark examples arising
+out of software verification.
+It remains to be seen whether these techniques can be used to check liveness properties of concurrent
+software, which often reduce to Petri net \emph{reachability} problems and are out of scope of other
+state-space traversal techniques.
+
+\paragraph{Acknowledgements.}
+%
+We thank Emanuele D'Osualdo for help with the Soter tool.

### 2014-03-17 paper/workspace/sec_illustration.tex
Added submitted version of paper

\section{An Example: Lamport's 1-bit Algorithm}

We illustrate three constraint-based functions for
invariant construction on Lamport's 1-bit algorithm for mutual
exclusion\cite{LamportJACM86}.
Figure~\ref{fig_lamport_src} shows the instance of the algorithm 
for two processes. Figure~\ref{fig_lamport_net} 
shows the corresponding Petri net.

\input{fig_lamport_src}

\subsection{An Invariant for Process 1}

We consider the property that the program counter of process 1 is
always either at location $p_1$, $p_2$, or $p_3$, i.e. $p_1 + p_2 +
p_3 \leq 1$. For the given network and property we construct an
invariant in the following way.

We obtain the following set of state constraints \setCp\ corresponding
to the initial marking, flow relation, and the property.
The set \setCp\ is such that a solution to the set corresponds to an
invariant.
\\ \mbox{ } \\
\begin{minipage}{\columnwidth}
  \emph{Inductivity constraints} \\[-2em]
  \begin{minipage}[t]{.49\columnwidth}
    \centering
    \begin{alignat*}{25}
      &{}-{}& p_1 &{}+{}& p_2 &     &     &{}+{}& bit_1 &{}-{}& notbit_1 &{}\le{}& 0 \\[-0.4em]
      &     &     &{}-{}& p_2 &{}+{}& p_3 &     &       &     &          &{}\le{}& 0 \\[-0.4em]
      &     & p_1 &     &     &{}-{}& p_3 &{}-{}& bit_1 &{}+{}& notbit_1 &{}\le{}& 0
    \end{alignat*}
  \end{minipage}
  \begin{minipage}[t]{.49\columnwidth}
    \centering
    \begin{alignat*}{25}
      &{}-{}& q_1 &{}+{}& q_2 &     &     &     &     &     &     &{}-{}& notbit_2 &{}\le{}& 0  \\[-0.4em]
      &     &     &{}-{}& q_2 &{}+{}& q_3 &     &     &     &     &     &          &{}\le{}& 0  \\[-0.4em]
      &     &     &     &     &{}-{}& q_3 &{}+{}& q_4 &     &     &{}+{}& notbit_2 &{}\le{}& 0  \\[-0.4em]
      &     & q_1 &     &     &     &     &{}-{}& q_4 &     &     &     &          &{}\le{}& 0  \\[-0.4em]
      &     &     &{}-{}& q_2 &     &     &     &     &{}+{}& q_5 &     &          &{}\le{}& 0  \\[-0.4em]
      &     & q_1 &     &     &     &     &     &     &{}-{}& q_5 &{}+{}& notbit_2 &{}\le{}& 0 
    \end{alignat*}
  \end{minipage} \\
  \emph{Safety constraint}\\[-2em]
  \begin{align*}
    p_1 + q_1 + \neg bit_1 + \neg bit_2 & < 2 * target_1
  \end{align*}
\end{minipage}
\begin{minipage}{\columnwidth}
  \emph{Property constraints} \\[-2em]
  \begin{minipage}[t]{.49\columnwidth}
    \centering
    \begin{align*}
      p_1   & \ge target_1 \\[-0.4em]
      p_2   & \ge target_1 \\[-0.4em]
      p_3   & \ge target_1 \\[-0.4em]
      bit_1 & \ge 0        \\[-0.4em]
      notbit_1 & \ge 0
    \end{align*}
  \end{minipage}
  \begin{minipage}[t]{.49\columnwidth}
    \centering
    \begin{align*}
      q_1   & \ge 0 \\[-0.4em]
      q_2   & \ge 0 \\[-0.4em]
      q_3   & \ge 0 \\[-0.4em]
      q_4   & \ge 0 \\[-0.4em]
      q_5   & \ge 0 \\[-0.4em]
      notbit_2 & \ge 0
    \end{align*}
  \end{minipage}
\end{minipage}
\begin{minipage}{\columnwidth}
  \emph{Non-negativity constraint}\\[-2em]
  \begin{align*}
    target_1 & \ge 0
  \end{align*}
\end{minipage}
\\ \mbox{ } \\
The \emph{inductivity constraints} correspond to the flow relation of the
Petri~net.
The \emph{safety constraint} corresponds to the initial marking and
the property.
The \emph{property constraint} corresponds to the property and the set
of places.
The \emph{non-negativity constraints} corresponds to the property.

For the construction of an invariant, we consider the
following solution \solA\ to the set \setCp.
\\[-2em]
\begin{minipage}[t]{.49\columnwidth}
  \begin{align*}
    p_1      & = 3 \\[-0.4em]
    p_2      & = 2 \\[-0.4em]
    p_3      & = 2 \\[-0.4em]
    bit_1    & = 1 \\[-0.4em]
    notbit_1 & = 0 \\[-0.4em]
  \end{align*}
\end{minipage}
\begin{minipage}[t]{.49\columnwidth}
  \begin{align*}
    q_1      & = 0 \\[-0.4em]
    q_2      & = 0 \\[-0.4em]
    q_3      & = 0 \\[-0.4em]
    q_4      & = 0 \\[-0.4em]
    q_5      & = 0 \\[-0.4em]
    notbit_2 & = 0 \\[-0.4em]
    target_1 & = 2
  \end{align*}
\end{minipage}

Given the solution, we construct the following
invariant \invI.
\begin{align*}
  3 p_1 + 2 p_2 + 2 p_3 + bit_1 & \le 3
\end{align*}
The invariant corresponds to the property that the program
counter of process 1 is always either at location $p_1$, $p_2$, or
$p_3$ and when the counter is at $p_1$ the flag $bit_1$ is not set.
The left-hand-side is a linear combination of the places of the
Petri~net where the coefficients are given by the solution.
The right hand side is the value of the combination at the initial
marking.

\input{fig_lamport_net}

\subsection{Example: minimal invariant for process 1}

\subsection{Example: invariant by refinement for process 1}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 paper/workspace/sec_introduction.tex
Added submitted version of paper

@@
 
-This paper makes the following contributions.
-\begin{itemize}
-\item An application of \zthree\ to the constraint approach and its
-  corresponding evaluation.
-\item A method that given a network and a property constructs a
-  minimal invariant.
-\end{itemize}
+In recent years many papers have proposed and developed techniques for
+the verification of concurrent software
+\cite{KaiserKW10,GantyM12,BouajjaniEmmi12,KaiserKWCONCUR12,DOsualdoSAS13}.
+In particular, model checkers based on Petri net coverability have been successfully
+applied.  Petri nets are a natural model for this problem for two
+resons.  The first and obvious one is that they are a simple and
+natural automata-like model for concurrent systems. The second, more
+specific, is that they can model certain programs with an unbounded
+number of threads or thread creation. In a nutshell, the places of the
+net correspond to program points, and the number of tokens in a place
+models the number of threads that are currently at that point. This
+point, which was first observed in \cite{GS92}, was later revisited in
+\cite{DelzannoRB02}, and, more implicitly, in \cite{KaiserKW10}.
+
+The problem whether at least one thread can reach a given program
+point (modelling some kind of error), naturally reduces to the {\em
+  coverability problem} for Petri nets: given a net $N$ and a marking
+$M$, decide whether some reachable marking of $N$ {\em covers} $M$,
+i.e., puts at least as many tokens as $M$ on each place. While the
+decidability and EXPSPACE-completeness of the coverability problem
+were settled long ago \cite{KarpM69,Rackoff78}, new algorithmic ideas
+have been developed in recent years
+\cite{GeeraertsRB06,GantyFI08,ValmariH12,KaiserKWCONCUR12,KloosMNP13}. 
+The techniques are based on forward or backward state-space exploration, 
+which is accelerated in a number of ways in order to cope with the possibly 
+infinite number of states.
+
+In this paper we revisit an approach to the coverability problem based
+on classical Petri net analysis techniques: the marking equation and
+traps \cite{murata1989petri,EsparzaM00,Reisig13}.
+The marking equation is a system of linear constraints that can be
+easily derived from the net, and whose set of solutions
+overapproximates the set of reachable markings.  The marking equation
+can be supplemented with linear constraints specifying a set of unsafe
+markings, and solved using standard linear or integer programming. If
+the equations are infeasible, then all reachable markings are safe. If
+not, then, as suggested in \cite{EsparzaM00}, one can use a
+counterexample to derive an additional constraint in the shape of a
+{\em trap}: a set of places that, loosely speaking, once marked cannot
+be ``emptied''.
+
+While these techniques are well known, their application has always
+suffered from the absence of efficient decision procedures for 
+linear arithmetic together with Boolean satisfiability.  
+Moreover, and to the best of our knowledge, their
+systematic application to verification of concurrent software has not
+been studied yet. Profiting from recent advances in SMT-solving
+technology, we reimplement the techniques on top of the Z3 SMT solver
+\cite{MouraB08}, and apply them to a large collection of benchmarks.
+
+Our first and surprising finding is that, while the technique is theoretically not
+complete, it is powerful enough to prove safety of 96 out of a total
+of 115 safe benchmarks gathered from current research papers in concurrent software
+verification. In contrast, state-of-the-art tools for coverability
+proved only 61, 51, or 33 of these 115 cases!
+Moreover, and possibly due to the characteristics of the
+application domain, even the simplest version of the technique ---based on the marking equation--- is
+successful in 84 cases. 
+As a second contribution, and inspired
+by work on interpolation, we show that a dual version of the classical
+set of constraints, equivalent in expressive power, can be used not
+only to check safety, but to produce an inductive invariant.
+While some existing solvers based on state-space
+exploration can also produce such invariants, we show that inductive invariants obtained
+through our technique are usually orders
+of magnitude smaller.  Additionally, while we can use the 
+SMT solver iteratively to minimize the invariant, the tool almost always
+provides a minimal one at the first attempt.
+
+% The paper is organized as follows.
+% We introduce the basic notation in Section~\ref{sec_preliminaries}.
+% We recall the marking equation and traps approach in
+% Sections~\ref{sec_method_safety}
+% and~\ref{sec_method_safety_by_refinement}.
+% We introduce three novel methods to constructing invariants in
+% Sections~\ref{sec_method_invariant}
+% % \ref{sec_method_invariant_by_refinement},
+% and
+% \ref{sec_method_invariant_w_minimization}.
+% Section~\ref{sec_experiments} presents (1) our experimental evaluation of
+% the application of SMT solver \zthree\ to the constraint-based
+% approach, and (2) our experimental evaluation of our methods for
+% invariant construction.
+% Section~\ref{sec_rel_work} reviews related
+% work. Section~\ref{sec_conclusions} presents our conclusions.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "main"
+%%% End:

### 2014-03-17 paper/workspace/sec_invariant.tex
Added submitted version of paper

\section{Constructing Invariants from Constraints}
\label{sec_method_invariant}

\subsection{A first approach}

We make a simple observation which, combined with the flexibility of
SMT-solvers, leads to an interesting improvement on the Safety
Method. We propose method \invariant\ with the same power as the
method \safety\ over the rational numbers, but dual to it: the
constraints of method \safety\ are unfeasible over the rational
numbers if{}f the constraints of method \invariant\ are
feasible. The advantage is that the solution delivered by the
method \invariant\ is an inductive linear invariant satisfied by all
reachable markings, but violated by all markings that do not satisfy
the property. So this invariant provides an {\em explanation} of why
the property holds.

We now derive the constraints given by method \invariant.
Consider the matrix form of the negation of property $P$ below.
$$ \neg P = ( A M \ge b ) $$
for some matrix $A$ and vector $b$.
Consider the following primal system of linear inequations $C$.
\begin{align*}
  \mathcal{C}&(S, T, F, M_0)  && \text{constraints given by \safety} \\
  A M & \ge b                 && \text{negation of property $P$}
\end{align*}

We have the following theorem:

\begin{theorem}
\label{thm:dual}
The system $C'$ of constraints given by
\begin{align*}
  \lambda C    & \le 0     && \text{inductivity constraint} \\
  \lambda M_0  & <   Y_1 b && \text{safety constraint} \\
  \lambda      & \ge Y_1 A && \text{property constraint} \\
  Y_1          & \ge 0     && \text{non-negativity constraint} 
\end{align*}
is feasible over the rational numbers if{}f the system $C$ is unfeasible
over teh rational numbers.
\end{theorem}
\begin{proof}
See the Appendix.
\end{proof}

We can now easily interpret the meaning of a solution $\lambda$ of $C'$. Recall that
for every reachable marking $M$ we find a solution of 
$M = M_0 + CX$. Multiplying by $\lambda$, and since $\lambda$ is a solution
of $C'$, we get 
$$\lambda M = \lambda M_0 + \lambda CX \leq \lambda M_0$$
\noindent On the other hand, for every marking $M$ violating the property
we have $ AM \geq b$, and so
$$\lambda M \geq Y_1 AM \geq Y_1 b > \lambda M_0$$

Therefore, the linear constraint
$$I(M) = \lambda M \leq \lambda M_0$$
is an invariant satisfied by all reachable markings $M$, 
but by no marking violating the property.

We can now prove that $I(M)$ is an inductive invariant, that is, if
$I(M)$ holds for some marking $M$ (reachable or not), and $M \xrightarrow{t} M'$
for some transition $t$, then $I(M')$ holds as well.
Indeed, in this case we have $M' = M + C e_t$ where $e_t = (0,\ldots,0,\underbrace{1}_{t},0,\ldots,0)^T$ is the vector whose components are all $0$ but the one
corresponding to $t$, which has value $1$. So we get
$$\lambda M' = \lambda(M + Ce_t) = \lambda M + \lambda Ce_t
\le \lambda M \le \lambda M_0$$
\noindent and so $I(M')$ holds.

\subsection{Example}

Again, we consider Lamport's algorithm and the mutual
exclusion property.  As in this case, we can not prove safety 
without using trap constraints, we can also
not obtain an invariant by simply applying method \invariant.

However, let us assume we wanted to prove that there is no marking
violating the mutual exclusion property and also marking each trap
found by method \safetyref. We can do this by adding
the trap constraints to the negation of the property, which yields
$$(p_3 \ge 1) \land (q_5 \ge 1) \land
(p_2 + q_2 + q_3 + notbit_1 + notbit_2 \ge 1)$$

The constraints constructed by the method \invariant\ then are

\noindent \emph{Inductivity constraints}
$$
\begin{array}[t]{clclclclclcl}
    -  & p_1 &  +  & p_2 &     &     &  +  & bit_1 &  -  & notbit_1 &  \le  & 0 \\
       &     &  -  & p_2 &  +  & p_3 &     &       &     &          &  \le  & 0 \\
       & p_1 &     &     &  -  & p_3 &  -  & bit_1 &  +  & notbit_1 &  \le  & 0
\end{array}\quad
\begin{array}[t]{clclclclclclclclclcl}
    -  & q_1 &  +  & q_2 &     &     &     &     &     &     &  -  & notbit_2 &  \le  & 0  \\
       &     &  -  & q_2 &  +  & q_3 &     &     &     &     &     &          &  \le  & 0  \\
       &     &     &     &  -  & q_3 &  +  & q_4 &     &     &  +  & notbit_2 &  \le  & 0  \\
       & q_1 &     &     &     &     &  -  & q_4 &     &     &     &          &  \le  & 0  \\
       &     &  -  & q_2 &     &     &     &     &  +  & q_5 &     &          &  \le  & 0  \\
       & q_1 &     &     &     &     &     &     &  -  & q_5 &  +  & notbit_2 &  \le  & 0 
\end{array}
$$
\emph{Safety constraint}
$$    
p_1 + q_1 + \neg bit_1 + \neg bit_2 < target_1 + target_2 + trap_1
$$
\emph{Property constraints}
$$
\begin{array}[t]{rcl}
  p_1   & \ge & 0 \\
  p_2   & \ge & trap_1 \\
  p_3   & \ge & target_1 \\
  bit_1 & \ge & 0 \\
  notbit_1 & \ge 0
\end{array}
\qquad
\begin{array}[t]{rcl}
  bit_1 & \ge & 0 \\
  notbit_1 & \ge & trap_1
\end{array}
\qquad
\begin{array}[t]{rcl}
  notbit_2 & \ge & trap_1
\end{array}
\qquad
\begin{array}[t]{rcl}
  q_1   & \ge & 0 \\
  q_2   & \ge & trap_1 \\
  q_3   & \ge & trap_1 \\
  q_4   & \ge & 0 \\
  q_5   & \ge & target_2 \\
\end{array}
$$
\emph{Non-negativity constraints}
$$
\begin{array}[t]{rcl}
  target_1 & \ge & 0 \\
  target_2 & \ge & 0 \\
  trap_1   & \ge & 0
\end{array}
$$

\noindent where $\lambda$ corresponds to the places and $Y_1 = (target_1\ target_2\ trap_1)$
corresponds to the two parts of the property and the added trap.

For the construction of an invariant, we consider the
following solution
$$
\begin{array}[t]{rcl}
  p_1      & = & 1 \\
  p_2      & = & 2 \\
  p_3      & = & 2 \\
  bit_1    & = & 0 \\
  notbit_1 & = & 0
\end{array}
\qquad
\begin{array}[t]{rcl}
  bit_1    & = & 0 \\
  notbit_1 & = & 1
\end{array}
\qquad
\begin{array}[t]{rcl}
  notbit_2 & = & 1
\end{array}
\qquad
\begin{array}[t]{rcl}
  q_1   & = & 0 \\
  q_2   & = & 1 \\
  q_3   & = & 1 \\
  q_4   & = & 0 \\
  q_5   & = & 1 \\
\end{array}
\begin{array}[t]{rcl}
  target_1 & = & 2 \\
  target_2 & = & 1 \\
  trap_1   & = & 1
\end{array}
$$

Given the solution, we construct the following
invariant \invI.
$$
I(M) = (p_1 + 2 p_2 + 2 p_3 + notbit_1 + notbit_2 + q_2 + q_3 + q_5 \le 3)
$$

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 paper/workspace/sec_invariant_by_refinement.tex
Added submitted version of paper

\section{Constructing Invariants from Constraints}
\label{sec_method_invariant_by_refinement}

We now show that one can compute inductive invariants from the method \safetyref.
That is, given a Petri net $N$ and a property $\varphi$, if \safetyref\ (over the rationals)
can prove $N$ satisfies $\varphi$, then in fact we can construct a linear inductive invariant
that contains $m_0$ and does not intersect $\lnot\varphi$. We call the
new method \invariantref.

The key observation is to use a constraint system dual to the constraint system for \safetyref.
%
% We make a simple observation which, combined with the flexibility of
% SMT-solvers, leads to an interesting improvement on the method \safetyref.
% We propose method \invariantref\ with the same power as the
% method \safetyref\ over the rational numbers, but dual to it: the
% constraints of method \safetyref\ are infeasible over the rational
% numbers if{}f the constraints of method \invariantref\ are
% feasible. 
% The advantage is that the solution delivered by the
% method \invariantref\ is an inductive linear invariant satisfied by all
% reachable markings, but violated by all markings that do not satisfy
% the property. 
% So this invariant provides an {\em explanation} of why
% the property holds.
We assume the negation of property $\varphi$ is represented as the constraints:
$$ \neg \varphi\  :: \ A M \ge b $$
for some matrix $A$ and vector $b$.
% consider the matrix form of the trap constraints $C_T$ below
% $$ C_T = ( D M \ge 1 ) $$
% for some matrix $D$.
Consider the following primal system $\mathcal{S}$:
\begin{align*}
  \mathcal{C}&(P, T, F, m_0)  && \text{marking constraints} \\
  A M & \ge b                 && \text{negation of property $\varphi$} \\
  D M & \ge 1                 && \text{trap constraints}
\end{align*}
%
Note that the trap constraints~\eqref{eq:trapcons} are collected in matrix form $D M \geq 1$.
We have the following theorem. (The proof is in Appendix A.)

\begin{theorem}
\label{thm:dual}
The system $\mathcal{S}'$ of constraints given by
\begin{align*}
  \lambda C    & \le 0            && \text{inductivity constraint} \\
  \lambda m_0  & <   Y_1 b + Y_2 1 && \text{safety constraint} \\
  \lambda      & \ge Y_1 A + Y_2 D && \text{property constraint} \\
  Y_1, Y_2     & \ge 0     && \text{non-negativity constraint} 
\end{align*}
is feasible over the rational numbers if{}f the system $\mathcal{S}$ is infeasible
over the rational numbers.
\end{theorem}

So $\mathcal{S}'$ is satisfiable if{}f $\mathcal{S}$ is unsatisfiable. We can
interpret the meaning of a solution $\lambda$. Recall that
for every reachable marking $m$ we find a solution of 
$m = m_0 + CX$. Multiplying by $\lambda$, and since $\lambda$ is a solution
of $\mathcal{S}'$, we get 
$$\lambda m = \lambda m_0 + \lambda CX \leq \lambda m_0\,.$$
Furthermore, any reachable marking will satisfy the trap constraints $DM\geq 1$.
On the other hand, for every marking $m$ violating the property
and satisfying the trap constraints, 
we have $ Am \geq b$ and $ Dm \geq 1$, and so
$$\lambda m \geq (Y_1 A + Y_2 D)m = Y_1 A m + Y_2 D m \geq
Y_1 b + Y_2 1 > \lambda m_0\,.$$
%
Therefore, the linear constraint
$$I(M)\  :: \  DM\geq 1 \land (\lambda M \leq \lambda m_0)$$
is an invariant satisfied by all reachable markings $m$, 
but by no marking violating the property.

We can now prove that $I(M)$ is an inductive invariant, that is, if
$I(m)$ holds for some marking $m$ (reachable or not), and $m \xrightarrow{t} m'$
for some transition $t$, then $I(m')$ holds as well.
Indeed, in this case we have $m' = m + C e_t$ where
$e_t$ is the unit vector which has a $1$ in the $t$th component and $0$ everywhere else.
% $e_t = (0,\ldots,0,\underbrace{1}_{t},0,\ldots,0)^T$ is the vector whose
% components are all $0$ but the one corresponding to $t$, which has value $1$.
So we get $$\lambda m' = \lambda(m + Ce_t) = \lambda m + \lambda Ce_t
\le \lambda m \le \lambda m_0\,,$$
and furthermore, as $m$ marks all traps in the trap constraints, 
$m'$ also marks all traps, so $I(m')$ holds.

% \begin{enumerate}
% \item The constraints given by \safetyref\ are unsat if{}f the
%   constraints by \invariantref\ are satisfiable.
% \item If the constraints by \invariantref\ are sat, then there is some
%   inductive invariant.
% \end{enumerate}

% \\ \mbox{ } \\
% \\ \mbox{ } \\
% \\ \mbox{ } \\
% \\ \mbox{ } \\

% Note we can express a set of trap constraints as a linear equation
% of the form $DM \ge 1$, where $D$ encodes the places in each trap.
% We can combine these with the negation of the property $AM \ge b$
% to obtain a new constraint $A'M \ge b'$, with
% $A' = (D A)^T$ and $b' = (1 b)^T$.
% The new constraint is then equivalent to $DM \ge 1 \land AM \ge b$.

% We can use this constraint instead of the negation of the property
% to obtain a linear constraint $I(M)$ with method Invariant.
% However, to obtain an invariant which only excludes violations of
% the property, we have to add back the trap constraints.
% By this, we obtain the invariant

% $$I'(M) := DM \ge 1 \land I(M)$$

% The invariant $I'(M)$ holds for all reachable markings, as the traps are marked
% initially and therefore also marked in all reachable markings and
% $I(M)$ holds for all reachable markings.

% Let $M$ be a marking $M$ violating the property, i.e. $AM \ge b$.
% If $DM \ge 1$ does not hold, then $I'(M)$ does not hold.
% Otherwise, $DM \ge 1 \land AM \ge b$ holds and by the
% the properties of $I(M)$ also $I(M)$ and $I'(M)$ do not hold.

% The invariant $I'(M)$ is inductive, as traps are inductive and $I(M)$ is inductive.

\subsection{Example}

Consider again Lamport's algorithm and the mutual
exclusion property. Recall that the negation of the property for this example is $p_3 \ge 1 \land q_5 \ge 1$, and the trap constraint is
$$p_2 + q_2 + q_3 + notbit_1 + notbit_2 \ge 1\,.$$
%
Method \invariantref{} constructs the following constraints:

\smallskip
\noindent \emph{Inductivity constraints}
$$
\begin{array}[t]{clclclclclcl}
    -  & p_1 &  +  & p_2 &     &     &  +  & bit_1 &  -  & notbit_1 &  \le  & 0 \\
       &     &  -  & p_2 &  +  & p_3 &     &       &     &          &  \le  & 0 \\
       & p_1 &     &     &  -  & p_3 &  -  & bit_1 &  +  & notbit_1 &  \le  & 0
\end{array}\quad
\begin{array}[t]{clclclclclclclclclcl}
    -  & q_1 &  +  & q_2 &     &     &     &     &     &     &  -  & notbit_2 &  \le  & 0  \\
       &     &  -  & q_2 &  +  & q_3 &     &     &     &     &     &          &  \le  & 0  \\
       &     &     &     &  -  & q_3 &  +  & q_4 &     &     &  +  & notbit_2 &  \le  & 0  \\
       & q_1 &     &     &     &     &  -  & q_4 &     &     &     &          &  \le  & 0  \\
       &     &  -  & q_2 &     &     &     &     &  +  & q_5 &     &          &  \le  & 0  \\
       & q_1 &     &     &     &     &     &     &  -  & q_5 &  +  & notbit_2 &  \le  & 0 
\end{array}
$$
\emph{Safety constraint}
$$    
p_1 + q_1 + \neg bit_1 + \neg bit_2 < target_1 + target_2 + trap_1
$$
\emph{Property constraints}
$$
\begin{array}[t]{rcl}
  p_1   & \ge & 0 \\
  p_2   & \ge & trap_1 \\
  p_3   & \ge & target_1 \\
  bit_1 & \ge & 0 \\
  notbit_1 & \ge & 0
\end{array}
\qquad
\begin{array}[t]{rcl}
  bit_1 & \ge & 0 \\
  notbit_1 & \ge & trap_1
\end{array}
\qquad
\begin{array}[t]{rcl}
  notbit_2 & \ge & trap_1
\end{array}
\qquad
\begin{array}[t]{rcl}
  q_1   & \ge & 0 \\
  q_2   & \ge & trap_1 \\
  q_3   & \ge & trap_1 \\
  q_4   & \ge & 0 \\
  q_5   & \ge & target_2 \\
\end{array}
$$
\emph{Non-negativity constraints}
$$
\begin{array}[t]{rcl}
  target_1, target_2, trap_1  & \ge & 0 
%  target_2 & \ge & 0 \\
% trap_1   & \ge & 0
\end{array}
$$
%
where $\lambda$ corresponds to the places,
$Y_1 = (target_1\ target_2)$ corresponds to the two parts of the property, and
$Y_2 = (trap_1)$ corresponds to the trap.

The SMT solver yields the following solution
$$
\begin{array}[t]{rcl}
  p_1      & = & 1 \\
  p_2      & = & 2 \\
  p_3      & = & 2 \\
  bit_1    & = & 0 \\
  notbit_1 & = & 0
\end{array}
\qquad
\begin{array}[t]{rcl}
  bit_1    & = & 0 \\
  notbit_1 & = & 1
\end{array}
\qquad
\begin{array}[t]{rcl}
  notbit_2 & = & 1
\end{array}
\qquad
\begin{array}[t]{rcl}
  q_1   & = & 0 \\
  q_2   & = & 1 \\
  q_3   & = & 1 \\
  q_4   & = & 0 \\
  q_5   & = & 1 \\
\end{array}
\begin{array}[t]{rcl}
  target_1 & = & 2 \\
  target_2 & = & 1 \\
  trap_1   & = & 1
\end{array}
$$
\noindent which leads to the invariant
\begin{align}\label{eq:invariant}
& (p_2 + q_2 + q_3 + notbit_1 + notbit_2 \ge 1 ) \; \land \nonumber \\
& (p_1 + 2 p_2 + 2 p_3 + notbit_1 + notbit_2 + q_2 + q_3 + q_5 \le 3)\,.
\end{align}
\noindent This ``explains'' why the mutual exclusion property holds: every
reachable marking satisfies this invariant, but no marking putting at least 
a token in $p_3$ and $q_5$ does. 
However, it is not the ``simplest'' explanation; we show next how to simplify invariants.
% The next section shows how to get a simpler one.

\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_invariant_by_refinement}}{}

% \input{fig_method_invariant_by_refinement}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 paper/workspace/sec_invariant_w_minimization.tex
Added submitted version of paper

\subsection{Minimizing Invariants}
\label{sec_invariant_w_minimization}

There may be multiple solutions to the system $\mathcal{S}'$
leading to different invariants. 
We now propose a method \invariantwmin\ to obtain a minimal
invariant, where we define the size of an invariant as the number of 
place variables appearing in it.

Given an assignment $\lambda$ to the dual state constraints,
we add further constraints whose solution yields a
smaller invariant, if one exists. 
We iterate this process to obtain the minimal invariant.

To minimize a given invariant, we define a binary vector $b_\lambda$ which projects the
non-zero components of $\lambda$ to 1 and the other components to 0.
The sum of this vector is then equal to the number of places appearing
in the invariant.
We add constraints of the following form for each place $p$:
\begin{align*}
p &> 0 \implies b_{p} = 1 && \text{$p$ appears in the invariant} \\
p &= 0 \implies b_{p} = 0 && \text{$p$ does not appear in the invariant}
\end{align*}
Finally, we add the constraint $\sum_{p\in P} b_p < K$, where $K$ is the number
of places appearing in the current invariant.
For the invariant~\eqref{eq:invariant} for mutual exclusion, this constraint is
$$
    b_{p_1} + b_{p_2} + b_{p_3} + b_{bit_1} + b_{notbit_1} + b_{notbit_2} +
    b_{q_1} + b_{q_2} + b_{q_3} + b_{q_4} + b_{q_5} < 8\,.
$$
With the addded constraints, we obtain a new solution for our 
example, setting $p_2$, $p_3$, $notbit_1$, $notbit_2$, $q_2$, $q_3$,
$target_1$, $target_2$, and $trap_1$ to $1$, and all other variables
to $0$. The invariant for this solution is
$$
I(M)\ ::\  p_2 + p_3 + notbit_1 + notbit_2 + q_2 + q_3 + q_5 \le 2\,.
$$
which is the minimal invariant. Together with the trap constraint we 
finally get:
\begin{align*}
I'(M)\ ::&\ 
(p_2 + q_2 + notbit_1 + notbit_2 + q_3 \ge 1 ) \; \land \\
&\ (p_2 + p_3 + notbit_1 + notbit_2 + q_2 + q_3 + q_5 \le 2)\,.
\end{align*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%I think this remark can be removed
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Note that with the added trap constraints, the invariant might not be
%minimal as we are not guaranteed to find traps of minimal size
%or a minimal number of traps. However, we might still be able to
%reduce the size of the invariant.

% \input{fig_method_invariant_w_minimization}

\iffalse

\begin{verbatim}
* Property of minimization constraints C_M generated from A': 
    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
\end{verbatim}

\newpage

\begin{verbatim}
* Subprocedures \mathcal{C'} and Inv are the same as in Method Invariant

* Subprocedure \text{MinConstraints}:

  Input:
    N = (S, T, E, M0)  : Petri net
    A'                 : Satisfying assignment for C'
  Output:
    C_M                : Minimization Constraints

  Pseudocode:

\end{verbatim}

\begin{align*}
  \text{MinConstraints}(N, A') =& \left( \bigwedge_{s \in S} \left(
      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
    \right) \right) \land
    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
\end{align*}

\iffalse
\begin{verbatim}
* Example:

  - Dual state constraints C' and satisfying assignment A' for C' as in example for method Invariant with property 2

  - Minimization constraints C_M:
\end{verbatim}

\begin{verbatim}
    p1     > 0 => b_p1   = 1
    |            -----------
    |               |
    |            place p1 appears in invariant
    |
    place p1 coefficient in the invariant

    p1     = 0 => b_p1   = 0
    |           ------------
    |               |
    |            place p1 does not appear in invariant
    |
    place p1 coefficient in the invariant
\end{verbatim}

\begin{align*}
(     p_1    > 0 & \implies b_{p_1}   = 1 ) & &\land &
(     p_1    = 0 & \implies b_{p_1}    = 0 ) \\
(     p_2    > 0 & \implies b_{p_2}    = 1 ) & &\land &
(     p_2    = 0 & \implies b_{p_2}    = 0 ) \\
(     p_3    > 0 & \implies b_{p_3}    = 1 ) & &\land &
(     p_3    = 0 & \implies b_{p_3}    = 0 ) \\
(     bit_1  > 0 & \implies b_{bit_1}  = 1 ) & &\land &
(     bit_1  = 0 & \implies b_{bit_1}  = 0 ) \\
( notbit_1 > 0 & \implies b_{notbit_1} = 1 ) & &\land &
( notbit_1 = 0 & \implies b_{notbit_1} = 0 ) \\
(     q_1    > 0 & \implies b_{q_1}    = 1 ) & &\land &
(     q_1    = 0 & \implies b_{q_1}    = 0 ) \\
(     q_2    > 0 & \implies b_{q_2}    = 1 ) & &\land &
(     q_2    = 0 & \implies b_{q_2}    = 0 ) \\
(     q_3    > 0 & \implies b_{q_3}    = 1 ) & &\land &
(     q_3    = 0 & \implies b_{q_3}    = 0 ) \\
(     q_4    > 0 & \implies b_{q_4}    = 1 ) & &\land &
(     q_4    = 0 & \implies b_{q_4}    = 0 ) \\
(     q_5    > 0 & \implies b_{q_5}    = 1 ) & &\land &
(     q_5    = 0 & \implies b_{q_5}    = 0 ) \\
( notbit_2 > 0 & \implies b_{notbit_2} = 1 ) & &\land &
( notbit_2 = 0 & \implies b_{notbit_2} = 0 )
\end{align*}

\begin{align*}
    b_{p_1} + b_{p_2} + b_{p_3} + b_{bit_1} + b_{notbit_1} +
    b_{q_1} + b_{q_2} + b_{q_3} + b_{q_4} + b_{q_5} + b_{notbit_2} & < 4
\end{align*}

\iffalse
\begin{verbatim}
    b_p1 + b_p2 + b_p3 + b_bit1 + b_nbit1 + b_q1 + b_q2 + b_q3 + b_q4 + b_q5
         + b_nbit2 < 4
    --------------   |
                |    |
                |    number of places appearing in current invariant for A'
                |    = #{s | A'(s) > 0} = #{p1, p2, p3, bit1}
                |
  number of places appearing in new invariant
\end{verbatim}
\fi

\begin{verbatim}
  - Model A' for C' \cup C_M:
\end{verbatim}

\begin{align*}
     p_1      & = 1 &    b_{p_1}     & = 1 \\
     p_2      & = 1 &    b_{p_2}     & = 1 \\
     p_3      & = 1 &    b_{p_3}     & = 1 \\
     bit_1    & = 0 &    b_{bit_1}   & = 0 \\
notbit_1    & = 0 &    b_{notbit_1}  & = 0 \\
     q_1      & = 0 &    b_{q_1}     & = 0 \\
     q_2      & = 0 &    b_{q_2}     & = 0 \\
     q_3      & = 0 &    b_{q_3}     & = 0 \\
     q_4      & = 0 &    b_{q_4}     & = 0 \\
     q_5      & = 0 &    b_{q_5}     & = 0 \\
notbit_2    & = 0 &    b_{notbit_2}  & = 0 \\
     target_1 & = 1 \\
\end{align*}

\begin{verbatim}
  - Minimized Invariant:
\end{verbatim}

\begin{align*}
    p_1 + p_2 + p_3 & \le 1
\end{align*}
\fi

\fi

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 paper/workspace/sec_preliminaries.tex
Added submitted version of paper

--- 2014-02-05 paper/workspace/sec_preliminaries.tex
@@
 \section{Preliminaries}
+\label{sec_preliminaries}
 
-\begin{verbatim}
-* Petri net: definition
- N = (S, T, E, M0), where
-     S : places
-     T : transitions
-     E : edges
-     M0: initial marking
+A \emph{Petri net} is a tuple $(P, T, F, m_0)$, where $P$ is a set of
+\emph{places}, $T$ is a (disjoint) set of \emph{transitions},
+$F : (P\times T) \cup (T\times P) \to \set{0,1}$ is the \emph{flow function},
+and $m_0: P \rightarrow \mathbb{N}$ is the initial marking.
+%
+For $x\in P\cup T$, the \emph{pre-set} is
+$\pre{x}=\{y\in P\cup T\mid F(y,x) =1\}$
+and the \emph{post-set} is $\post{x}=\{y\in P\cup T\mid F(x,y) = 1\}$.
+We extend the pre- and post-set to a subset of $P \cup T$ as the union of
+the pre- and post-sets of its elements.
 
-* Marking: definition
+A \emph{marking} of a Petri net is a function $m\colon P \to \mathbb{N}$,
+which describes the number of tokens $m(p)$ in 
+each place $p\in P$. 
+Assuming an enumeration $p_1, \ldots, p_n$ of $P$, we
+often identify $m$ and the vector $(m(p_1), \ldots, m(p_n))$.
+For a subset $P'\subseteq P$ of places, we write $m(P') = \sum_{p\in P'} m(p)$.
 
-* Visual representation of Petri Nets: description, example
+A transition $t \in T$ is \emph{enabled at $m$} iff
+for all $p \in \pre{t}$, we have $m(p) \ge F(p, t)$.
+A transition $t$ enabled at $m$ may \emph{fire},
+yielding a new marking $m'$ (denoted $m \xrightarrow{t} m'$),
+where $m'(p) = m(p) + F(t,p) - F(p,t)$.
 
-* Property P: safety property expressed over linear arithmetic formulas.
+A sequence of transitions, $\sigma = t_1 t_2 \ldots t_r$ is an
+$\emph{occurrence sequence}$ of $N$ iff there exist markings
+$m_1, \ldots, m_r$ such that $m_0 \xrightarrow{t_1} m_1
+\xrightarrow{t_2} m_2 \ldots \xrightarrow{t_r} m_r$. The marking
+$m_r$ is said to be \emph{reachable} from $m_0$ by the occurrence
+of $\sigma$ (denoted $m_0 \xrightarrow{\sigma} m_r$).
 
-* Linear arithmetic formula: examples
+A \emph{property} $\varphi$ is a linear arithmetic constraint over the free variables $P$. 
+The property $\varphi$ holds on a marking $m$ iff $m \models \varphi$.
+% Examples of properties are $s_1 \le 2$, $s_1 + s_2 \ge 1$ and
+% $((s_1 \le 1) \land (s_2 \ge 1)) \lor (s_3 \le 1)$.
+% 
+A Petri net $N$ satisfies a property $\varphi$ (denoted by $N \models \varphi$)
+iff for all reachable markings $m_0 \xrightarrow{\sigma} m$, we have
+$m \models \varphi$.
+A property $\varphi$ is an \emph{invariant} of $N$ if it holds for every reachable marking.
+A property is inductive if whenever $m\models \varphi$ and $m \xrightarrow{t} m'$ for some $t\in T$
+and marking $m'$, we have $m'\models \varphi$.
 
-* N \models P: definition
+% A trap is a set of places $S \subseteq P$ such that $\post{S} \subseteq \pre{S}$.
+% If a trap $S$ is marked in $m_0$, i.e. $\sum_{p \in S} m_0(p) > 0$, then it is also marked in all reachable markings.
 
-* Whatever you believe should be defined for the paper.
-\end{verbatim}
+Petri nets are represented graphically as follows: places and transitions
+are represented as circles and boxes, respectively. For $x, y \in P \cup T$,
+there is an arc leading from $x$ to $y$ if{}f $F(x,y)=1$.
 
-\iffalse
-\begin{verbatim}
- Examples:
-  x + y < 0
-  x + y < 0 \\and x + z > 0
-  x + y < 0 \\or  x + z > 0
+\input{fig_lamport_src}
+\input{fig_lamport_net}
 
- ~P = (x + y <  0 \\and x + z >  0) \\or  x + z >  0
-  P = (x + y >= 0 \\or  x + z <= 0) \\and x + z <= 0
-\end{verbatim}
-\fi
+Petri nets can be used to model behavior of concurrent software. As an
+example, consider Lamport's 1-bit
+algorithm for mutual exclusion \cite{LamportJACM86}, shown in
+Fig.~\ref{fig_lamport_src}. 
+A Petri net model for the code is shown in Fig.~\ref{fig_lamport_net}.
+The two grey blocks model the control flow of the two
+processes. For instance, the token on place $p_1$ models the current
+position of process 1 at program location $p_1$. The three places in the middle of the
+diagram model the current values of the variables. For instance, a
+token on place $notbit_1$ indicates that the variable ${\it bit}_1$ is currently
+set to ${\it false}$.
+The mutual exclusion property, which states that the two processes cannot be in the
+critical section at the same time, corresponds to the property that
+places $p_3$ and $q_5$ cannot both have a token at the same time.

### 2014-03-17 paper/workspace/sec_rel_work.tex
Added submitted version of paper

\smallskip
\noindent\textbf{Related Work.\ }
%
Our starting point was the work of Esparza and Melzer on
extending the marking equation with trap conditions to gain a stronger method
for proving safety of Petri nets \cite{EsparzaM00}. We combined the
constraint-based approach there with modern SMT solvers.
Their focus on (integer) linear programming tools of the time enforced some limitations. 
First, while traps are naturally encoded using Boolean variables,
\cite{EsparzaM00} encoded traps and the marking equation together into a set of 
linear constraints.
This encoding came at a practical cost: the encoding required (roughly) $n\times m$ constraints
for a Petri net with $n$ places and $m$ transitions, whereas the natural Boolean encoding
requires $m$ constraints.
Moreover, (I)LP solvers were not effective in searching large Boolean state
spaces; our use of modern SAT techniques alleviates this problem. 
%    constraints is cumbersome: it requires $n\times k$ constraints, where $n$ is
%    the number of places, and $k$ is the number of arcs from places to
%    transitions (succintly: $\#\{(p,t)\mid F(p,t)>0\}$. Expressing the trap
%    condition using boolean constraints requires only $m$ constraints, where $m$
%    is the number of transitions. 
Second, (I)LP solvers used by \cite{EsparzaM00} did not handle strict inequalities. 
Because of this, the authors used additional tricks, such as posing the problem that includes a 
strict inequality as a minimization problem, with the goal of minimizing the involved expression, 
and testing if the minimal value equaled zero. 
Unfortunately, this trick led to numerical instabilities. 
All of these concerns vanish by using an SMT solver.

In contrast to other recent techniques for coverability \cite{GantyFI08,KaiserKWCONCUR12,KloosMNP13},
our technique is incomplete.
However, we empirically demonstrate that for safe instances arising out of software verification,
our technique is extremely effective, often beating well-optimized techniques in terms of time or memory usage.
At the same time, our technique applies not only to coverability but also to \emph{reachability}.
It will be interesting to see whether the techniques can effectively verify liveness properties for concurrent
programs, which can be reduced to Petri net reachability questions.

% Main differences:
% \begin{itemize}
%     \item We have extended the techniques to construct inductive invariants,
%     which serve as certificates of safety.
%     \item Their focus on (integer) linear programming tools of the time enforced
%     some limitations. First, expressing the trap condition using linear
%     constraints is cumbersome: it requires $n\times k$ constraints, where $n$ is
%     the number of places, and $k$ is the number of arcs from places to
%     transitions (succintly: $\#\{(p,t)\mid F(p,t)>0\}$. Expressing the trap
%     condition using boolean constraints requires only $m$ constraints, where $m$
%     is the number of transitions. Second, (I)LP solvers of the time could not
%     handle strict inequalities. Because of this, tricks were used, such as
%     posing the problem that includes a strict inequality as a minimization
%     problem, with the goal of minimizing the involved expression, and testing
%     if the minimal value equals zero. Subsequently, this trick led to numerical
%     problems. All of these concerns vanish by using an SMT solver.
%     \item They focused their evaluation on proving safety mostly for mutual
%     exclusion protocols (check!). We have evaluated the methods on a
%     variety of examples that appear in software verification, provenance
%     analysis\ldots
% \end{itemize}
% 
% Other methods and tools \cite{KloosMNP13, KaiserKWCONCUR12, GeeraertsRB06,
% GantyFI08, ValmariH12}\ldots Even though the methods are complete, the tools
% suffer from time and memory limitations. However, they can also give refutations in the form of
% conterexamples.
% 
% Theoretical results \cite{KarpM69, Rackoff78}?

### 2014-03-17 paper/workspace/sec_safety.tex
Added submitted version of paper

\section{Marking Equation}
\label{sec_method_safety}

We now recall a well-known method, which we call \safety, that provides a sufficient
condition for a given Petri net $N$ to satisfy a property $\varphi$ 
by reducing the problem to checking satisfiability of a linear arithmetic formula.
We illustrate the method on Lamport's 1-bit algorithm for mutual
exclusion.

Before going into details, we state several conventions. For a
Petri net $N = (P, T, F, m_0)$, we introduce a vector of
$|P|$ variables $M$, and a vector of $|T|$ variables $X$.
$M$ and $X$ will be used to represent the current
marking, and the number of occurrences
of transitions in the occurrence sequence leading to the current marking.
If a place or a transition is given a specific
name, we use the same name for its associated variable. Given a place
$p$, the intended meaning of a constraint like $p \geq 3$ is 
``at the current marking place $p$ must have at least 3 tokens''. 
Given a transition $t$, the intended meaning of a constraint like $t \leq 2$ is ``in
the occurrence sequence leading to the current marking, transition $t$
must fire at most twice''.

The key idea of the \safety{} method lies in the \emph{marking
  equation}:
$$M = m_0 + C X\,,$$
where the incidence matrix $C$ is a $|P| \times |T|$ matrix given by
$$C(p,t) = F(t,p) - F(p,t)\,.$$
For each place $p$, the marking equation contains a constraint
that formulates a simple token conservation law: the number of tokens in
$p$ at the current marking is equal to the initial number of tokens $m_0(p)$,
plus the number of tokens added by the input transitions of $p$, minus
the number of tokens removed by the output transitions. 
So, for instance, in Lamport's algorithm the constraint for place $notbit_1$
is:
$$notbit_1 = 1 + s_3 + t_5 + t_4 - s_1 - t_4 - t_5 = 1 +s_3
-s_1\,.$$

We equip the marking equation with the non-negativity conditions,
modeling that the number of tokens in a place, or the number of
occurrences of a transition in an occurrence sequence cannot become
negative. All together, we get the following set of \emph{marking constraints}:
\begin{align*}
  \mathcal{C}(P, T, F, m_0) ::&
  \begin{cases}
    M = m_0 + C X  & \text{marking equation} \\
    M \ge 0        & \text{non-negativity conditions for places} \\
    X \ge 0        & \text{non-negativity conditions for transitions} \\
  \end{cases}
\end{align*}

Method \safety{} for checking that a property $\varphi$ is
invariant for a Petri net $N=(P,T,F,m_0)$ consists of checking for
satisfiability of the constraints
\begin{equation}\label{eq:safety}
\mathcal{C}(P,T,F,m_0) \wedge \lnot \varphi(M)\,.
\end{equation}
If the constraints are unsatisfiable, then no reachable marking
violates $\varphi$. To see that this is true, consider the converse:
If there exists an occurrence sequence $m_0
\xrightarrow{\sigma} m$ leading to a marking $m$ that violates the
property, then we can construct a valuation of the variables that assigns $m(p)$ to
$M(p)$ for each place $p$, and the number of occurrences of $t$ in
$\sigma$ to $X(t)$ for each transition $t$. This valuation then
satisfies the constraints.

The method does not work in the other direction: If the constraints
\eqref{eq:safety} are satisfiable, we cannot conclude that the
property $\varphi$ is violated.

As an example, consider the Lamport's algorithm. \safety{}
successfully proves the property ``if process
1 is at location $p_3$, then $\mathit{bit_1}={\it true}$'' by showing
that $\mathcal{C}(P,T,F,m_0) \wedge p_3 \geq 1 \wedge \mathit{bit}_1
\neq 1$ is unsatisfiable. However, if we apply it to the mutual
exclusion property, i.e.~check for satisfiability of
$\mathcal{C}(P,T,F,m_0) \wedge p_3 \geq 1\wedge q_5 \geq 1$, we obtain
a solution, and we cannot conclude that the mutual exclusion property
does not hold.

Note that the marking constraints~\eqref{eq:safety} are interpreted over integer
variables.
As usual in program analysis, one can solve the constraints over rationals to get an approximation
of the method. Solving the constraints over rationals will become
useful in Section \ref{sec_method_invariant_by_refinement}.

\iffalse

\begin{verbatim}
* Method Safety:

  Subprocedure \mathcal{C} constructs state constraints C corresponding to N.
\end{verbatim}

\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_safety}}{}

\input{fig_method_safety}

\begin{verbatim}
* Property of state constraints C: If C U {\neg P} is unsat then N |= P.

* Place equation:

  For a given place s the place equation is

  # of tokens in s = initial number of tokens of place s
                     + # times each input transition of s fires
                     - # times each output transition of s fires

* Non-negativity conditions:

  # of tokens in place s           >= 0
  # of times transition t is fired >= 0

* Subprocedure \mathcal{C}:

  Input:
    N = (S, T, F, M0) : Petri net
  Output:
    C : State constraints

  Pseudocode:

\end{verbatim}

\begin{align*}
  \mathcal{C}(N) =& \left( \bigwedge_{s \in S} \left(
    s = M_0(s) + \sum_{t \in \pre{s}} t - \sum_{t \in \post{s}} t
  \right) \right) \land
    \left( \bigwedge_{s \in S} s \ge 0 \right) \land
    \left( \bigwedge_{t \in T} t \ge 0 \right)
\end{align*}

\newpage

\begin{verbatim}
* Example

  - Property 1: Process 1 and Process 2 are never in their respective critical section at the same time.

  - Property 1: p3 < 1 \/ q5 < 1

  - Petri net:

\end{verbatim}

% \input{fig_lamport_net}

\newpage

\begin{verbatim}

\end{verbatim}

\iffalse
\begin{verbatim}
  - State constraints C:
    p1    = 1 - s1      + s3
    ^       ^   ^         ^
    |       |   |         |
    |       |   |         |
    |       |   |         |
    |       |   |        # of tokens given to p1
    |       |   |    
    |       |  # of tokens taken from p1
    |       |
    |      initial number of tokens in p1
    |
    number of tokens in p1
\end{verbatim}
\fi

\begin{verbatim}
- State constraints C:
\end{verbatim}

\input{fig_setC}

\begin{verbatim}
  - Using property 1

  - Negated property 1 \neg P:
\end{verbatim}

\begin{align*}
     p3 \ge 1 & \land q5 \ge 1
\end{align*}

\fi

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 paper/workspace/sec_safety_by_refinement.tex
Added submitted version of paper

\section{Refining Marking Equations with Traps}
\label{sec_method_safety_by_refinement}

The bottom line of the work of Esparza and Melzer \cite{EsparzaM00} was to
strengthen \safety{} with additional
\emph{trap constraints}. A {\em trap} of a Petri net $N=(P, T, F, m_0)$ is a subset of places
$Q \subseteq P$ satisfying the following condition for every
transition $t \in T$: if $t$ is an output transition of at least one
place of $Q$, then it is also an input transition of at least one
place of $Q$. Equivalently, $Q$ is a trap if its set of output
transitions is included in its set of input transitions, i.e., if
$\post{Q} \subseteq \pre{Q}$. Here we present a variant of Esparza's and
Melzer's method that encodes traps using Boolean constraints. We
call the new method \safetyref.

The method \safetyref\ is based on the following observation about traps.
If $Q$ is a trap and a marking $m$ marks $Q$, i.e., $m(p) > 0$ for some $p\in Q$,
then for each occurrence sequence $\sigma$ and marking $m'$ such that $m \xrightarrow{\sigma} m'$,
we also have $m'(p') > 0$ for some $p'\in Q$.  
Indeed, by the trap property any transition removing tokens from places of
$Q$ also adds at least one token to some place of $Q$. 
So, while $m'(Q)$ can be smaller than $m(Q)$, it can never become $0$. 
In particular, if a trap $Q$ satisfies $m_0(Q) > 0$, then every
reachable marking $m$ satisfies $m(Q) > 0$ as well.

Since the above property must hold for any trap, we can restrict the constrains
from method \safety{} as follows.
We additionally add a Boolean variable for each place $p\in P$.
These variables will be used as indicators for traps.
First, we write a constraint to specify that the Boolean variables encode a trap:
\[
\mathit{trap}(B) ::= \bigwedge_{t\in T}\left[ \bigvee_{p \in \pre{t}} B(p) \Rightarrow \bigvee_{p\in \post{t}} B(p) \right]\,.
\]
Next, we define a predicate $\mathit{mark}(m, B)$ that specifies that a marking marks a trap:
\[
\mathit{mark}(m, B) ::= \bigvee_{p\in P} B(p)\wedge (m(p) > 0)\,.
\]
Finally, we conjoin the following constraint to the constraints~\eqref{eq:safety}: 
\begin{equation}\label{eq:trap}
\forall B: \mathit{trap}(B) \wedge \mathit{mark}(m_0, B) \Rightarrow \mathit{mark}(M, B)\,.
\end{equation}
%
This constraint conceptually enumerates over all subsets of places, and ensures that if
the subset forms a trap, and this trap is marked by the initial marking, then
it is also marked in the current marking. 
Thus, markings violating the trap constraint are eliminated.

While the above constraint provides a refinement of the \safety\ method, it requires the
SMT solver to reason with universally quantified variables.
Instead of directly using universal quantifiers, we use a counterexample-guided
heuristic \cite{EsparzaM00,CEGIS} of adding trap constraints one-at-a-time in
the following way.

If the set of constraints constructed so far (for instance, the set
given by the method \safety) is feasible, the SMT solver delivers a
model that assigns values to each place, corresponding to a potentially reachable marking $m$.  
We search for a trap $P_m$ that violates the trap condition \eqref{eq:trap} for this specific model $m$.
If we find such a trap, then we know that $m$ is unreachable, and we can
add the constraint $\sum_{p \in P_m} M(p) \geq 1$ to exclude all markings that violate
this specific trap condition.

The search for $P_m$ is a pure Boolean satisfiability question.
We ask for an assignment to 
\begin{equation}\label{eq:booltrap}
\mathit{trap}(B) \wedge \mathit{mark}(m_0,B) \wedge \lnot \mathit{mark}(m, B)
\end{equation}
(and notice that for a fixed marking $m$, the predicate $\mathit{mark}(m,B)$ can
be evaluated to a Boolean predicate).
Given a satisfying assignment $b$ for this formula, we add the constraint
\begin{equation}\label{eq:trapcons}
\sum_{p \in P \mid b(p) = 1} M(p) \geq 1
\end{equation} 
to the constraints to rule out solutions that
do not satisfy this trap constraint.
We iteratively add such constraints until either the constraints are unsatisfiable
or the Boolean constraints~\eqref{eq:booltrap} are unsatisfiable (i.e., no traps are found
to invalidate the current solution). 
% How do we find $P_M$? The integration of boolean and arithmetic
% constraints allows to use the SMT-solver. The constraint problem
% delivering a trap that is marked at $M_0$ but not marked at a given
% marking $M$ is constructed as follows. As in the method \safety, we
% introduce a variable $p$ for every place $P$, but this time of {\em
%   boolean type}, with the following intended meaning: if $p$ has the
% value {\it true}, then $p$ belongs to the trap.
% 
% For each transition $t$ we add the constraint
% $$\bigvee_{p \in \pre{t}} p  \Rightarrow \bigvee_{p' \in \post{t}} p'$$
% \noindent which formalizes the trap property. Moreover we add
% %
% $$\bigvee_{p \in P, M_0(p) > 0} p \wedge \bigwedge_{p' \in P, M(p') > 0} \neg p'$$
% 
% \noindent modelling that the trap is marked at $M_0$ but not at $M$.

This yields the method \safetyref.  The method is still not
complete \cite{EsparzaM00}: one can find nets and unreachable markings that 
mark all traps of the net.

Let us apply the algorithm \safetyref{} to Lamport's algorithm and the
mutual exclusion property. 
Recall that the markings violating the property are those satisfying $ p_3 \geq 1$ and  $q_5 \geq 1$. \safety{} yields a
satisfying assignment with $p_3 = bit_1 = q_5 = 1$,
and $p=0$ for all other places $p$, 
which corresponds to a potentially reachable marking $m$. 
We search for a trap marked at $m_0$ but not at $m$. The constraints 
derived from the trap property are:
$$\begin{array}{l}
\begin{array}[t]{lcl}
  p_1 \lor notbit_1 &\implies& p_2 \lor    bit_1 \\
  p_2 \lor notbit_2 &\implies& p_3 \lor notbit_2 \\
  p_3 \lor    bit_1 &\implies& p_1 \lor notbit_1
\end{array} \qquad
\begin{array}[t]{lcl}
  q_1 \lor notbit_2 &\implies& q_2               \\
  q_2 \lor    bit_1 &\implies& q_3 \lor    bit_1 \\
  q_3               &\implies& q_4 \lor notbit_2 \\
  q_4 \lor notbit_1 &\implies& q_1 \lor notbit_1 \\
  q_2 \lor notbit_1 &\implies& q_5 \lor notbit_1 \\
  q_5               &\implies& q_1 \lor notbit_2
\end{array}
\end{array}
$$
and the following constraints model that at least one of the places 
initially marked belongs to the trap, but none of the places marked at 
the satisfying assigment do:
$$p_1 \lor q_1 \lor notbit_1 \lor notbit_2  \qquad \neg p_3 \land \neg q_5 \land
\neg bit_1$$ For this set of constraints we find the satisfying assignment 
that sets $p_2$, $notbit_1$, $notbit_2$, $q_2$, $q_3$ to $\mathit{true}$ and all
other variables to $\mathit{false}$. 
So this set of places is an initially marked trap, and 
so every reachable marking should put at least one token in it. 
Hence we can add the refinement constraint to marking
constraints~\eqref{eq:safety}:
\begin{align*}
  p_2 + q_2 + q_3 + notbit_1 + notbit_2 & \ge 1\,.
\end{align*}
%
On running the SMT solver again, we find the constraints are unsatisfiable, 
proving that the mutual exclusion property holds.

\iffalse

\begin{verbatim}
* Method Safety by Refinement:

  Subprocedure TrapConditions constructs trap conditions C_\theta corresponding to N and A.
  Subprocedure \Delta constructs refinement constraint \delta corresponding to A_\theta.
\end{verbatim}

\ifthenelse{\equal{\isDraft}{true}}{\input{fig_dia_method_safety_by_refinement}}{}

\input{fig_method_safety_by_refinement}

\begin{verbatim}
* Property of trap conditions C_\theta: If C_\theta is sat then there is a set S such that
  1. S is a trap in the net N
  2. S is marked in the initial marking M0
  3. S is unmarked in the assignment A

* Property of A_\theta: for each place s, A_\theta(s) iff s \in S

* Property of refinement constraint \delta: Constraint \delta refines the abstraction, i.e.
  1. A ^ \delta is unsat (\delta excludes A)
  2. N |= \delta is sat  (\delta is a property of N)
\end{verbatim}

\newpage

\begin{verbatim}
* Subprocedure TrapConditions:

  Input:
    N = (S, T, F, M0) : Petri net
    A             : Satisfying assignment for C \cup { ~P }
  Output:
    C_\theta      : Trap conditions

  Pseudocode:

\end{verbatim}
\begin{align*}
  \text{TrapConditions}(N, A) =& \left(
      \bigwedge_{s \in S} \left( s \Rightarrow
      \bigwedge_{t \in \post{s}} \bigvee_{p \in \post{t}} p
    \right) \right) \land
    \left( \bigvee_{s \in S: M_0(s) > 0} s \right) \land
    \left( \bigwedge_{s \in S: A(s) > 0} \neg s \right)
\end{align*}
\begin{verbatim}

* Subprocedure \Delta:

  Input:
    A_\theta      : Satisfying assignment for C_\theta
  Output:
    \delta        : Refinement constraint \delta

  Pseudocode:

\end{verbatim}
\begin{align*}
  & \Delta(A_\theta) = \left( \sum_{A_\theta(s)} s \ge 0 \right)
\end{align*}

\newpage

\begin{verbatim}
* Example

  - State constraints C same as in the example for method Safety.

  - Using property 1

  - Negated property 1 \neg P:
\end{verbatim}

{\Large OK, but how does it continue? Is this enough to prove the property,
or do we need another trap?}

\iffalse
\begin{verbatim}
    p2 + q2 + q3 + nbit1 + nbit2 >= 1
    ^    ^    ^    ^       ^
    |    |    |    |       |
    ------------------------
      |
      S = {p2, q2, q3, nbit1, nbit2}, therefore \delta excludes
      assignment A in the next iteration
\end{verbatim}
\fi

\newpage

\begin{verbatim}

* Trap implication:

  place s \in S =>    /\       \    /
                     /  \       \  /     place p \in S
                    /    \       \/
                   t \in s*   p \in t*

  "if s is in trap S then for each output transition t at least one successor p is in trap S"

* Refinement constraint \delta:

  \Sigma s >= 1  
  A(s)

  "At least one place in S is always marked"
\end{verbatim}

\fi

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2014-03-17 paper/workspace/sync-workspace.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+(cd ../../../../pnerf/paper/workspace && make clean)
+rsync ../../../../pnerf/paper/workspace/* .

### 2014-03-17 mytrash/workspace/peterson-simplified.pl
Removed trash

@@ -1,75 +1,0 @@
-place(s5).
-place(s6).
-place(s7).
-place(s8).
-place(s9).
-place(s10).
-place(s11).
-place(s12).
-place(s13).
-place(s14).
-place(s15).
-place(s16).
-place(s17).
-place(s18).
-place(s19).
-place(s20).
-place(s21).
-place(s22).
-place(s23).
-place(l1).
-place(l2).
-place(l3).
-place(l4).
-place(l5).
-place(l6).
-place(l7).
-place(l8).
-place(l9).
-place(l10).
-place(l11).
-place(l12).
-place(l13).
-place(l14).
-transition(t6, [l1, s5], [s6, l2]).
-transition(t7, [l2, s5], [s7, l3]).
-transition(t8, [l2, s5], [s8, l3]).
-transition(t9, [l3, s5], [s10, l4]).
-transition(t10, [l3, s5], [s12, l4]).
-transition(t11, [l4, s5], [s14, l1]).
-transition(t12, [l5, s5], [s19, l14]).
-transition(t13, [l7, s5], [s17, l13]).
-transition(t14, [l8, s5], [s16, l7]).
-transition(t15, [l8, s5], [s21, l14]).
-transition(t16, [l9, s5], [s23, l10]).
-transition(t17, [l10, s5], [s15, l9]).
-transition(t18, [l5, s6], [s5, l6]).
-transition(t19, [l7, s7], [s5, l8]).
-transition(t20, [l8, s8], [s9]).
-transition(t21, [s9], [s5, l8]).
-transition(t22, [l10, s10], [s11]).
-transition(t23, [s11], [s5, l10]).
-transition(t24, [l7, s12], [s13]).
-transition(t25, [s13], [s5, l7]).
-transition(t26, [l6, s14], [s5, l5]).
-transition(t27, [l11, s15], [s5, l12]).
-transition(t28, [l12, s16], [s5, l13]).
-transition(t29, [l12, s17], [s18]).
-transition(t30, [s18], [s5, l7]).
-transition(t31, [l13, s19], [s20]).
-transition(t32, [s20], [s5, l5]).
-transition(t33, [l13, s21], [s22]).
-transition(t34, [s22], [s5, l8]).
-transition(t35, [l14, s23], [s5, l11]).
-init(l1, 1).
-init(l5, 1).
-init(l8, 1).
-init(l10, 1).
-init(l11, 1).
-init(s5, 1).
[…]

### 2014-03-17 mytrash/workspace/run_with_animation.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-
-function animated_wait {
-  chars=('|' '\' '-' '/')
-  chars_len=${#chars[@]}
-  c=0
-  fout="/tmp/animated_wait.out"
-  line=1
-  echo $*
-  $* &
-  pid=$!
-  trap handle_ctrl_c INT
-  tput civis
-  while (ps -p $pid >/dev/null); do
-    lines_count=`<$fout wc -l`
-    tail -n +$line $fout
-    line=$((lines_count+1))
-    
-    echo -n ${chars[$c]}
-    c=$(((c+1) % chars_len))
-    sleep .2s
-    tput cub1
-  done
-  tput cnorm
-}
-
-function handle_ctrl_c {
-  tput cnorm
-  kill $pid 2>/dev/null
-  wait $pid 2>/dev/null
-  tail -n +$line $fout
-}
-
-animated_wait $*
